### 2018-09-12 Manuscript/results_outline.Rmd
results outline

---
title: "Manuscript Outline"
author: "Spiro Stilianoudakis"
date: "September 11, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

   + Paragraph 1:
      - What are TADs and what role do they play in the human genome?
      - Why do we care about TADs?

   + Paragraph 2:
      - What is in the current literature?

   + Paragraph 3:
      - What are the issues?
      - Why are they issues?

   + Paragraph 4:
      - What are the current methods out there?
      - What drawbacks do they have?

   + Paragraph 5:
      - Defined the problem
      - What have we done?
      - Why have we done it?

# Methods

* The Data
   + Paragraph 1:
      - Where was the data obtained?
   + Paragraph 2:
      - How it was used?
      - How was the response vector created

* Genomic Annotations
   + Paragraph 1: 
      - Where was the annotation data obtained?
      - How were they used?

* Data Pipeline:
   + What was the pipeline used for?
   + Set up structure of pipeline
   + How were techniques evaluated?

* Normalization:
   + What techniques were used/compared?
   + How were they used/how do they worked?

* Class Imbalance:
   + What techniques were used/compared?
   + How were they used/how do they worked?

* Variable Selection:
   + What techniques were used/compared?
   + How were they used/how do they worked?

* Final Model:
   + What model(s) we used to evaluat the pipeline processed data
   + How the model(s) were performed
   + What we compared our models to

# Results