### 2018-09-28 ToDo.Rmd
completed tasks

---
title: "To Do List"
author: "Spiro Stilianoudakis"
date: "September 4, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 09/07/18

####To-Do
+ use width in base pairs as type of feature instead of counts of overlap
   - distance and width in base pairs of overlap as features
+ finish introduction, start on methods

## 09/10/18

####Questions

* Should we use elastic-net and RF as final models or just one or the other?
* In results section should we compare between mourad models visually or verbally or none at all?
   + visually and verbally
   + AUROCs, F1 and MCC
* I've mentioned other methods in intro, it will be hard to compare to the others; how to handle this?
* Hong et al uses TFBS should I also include them?
   + use tfbs encode clustered (in separate model)
* How to handle k562
   +1 unit bin, 1unit flank (10 kb; 20kb)

####To-Do

+ Make a bullet pointed list for manuscript
+ Apply RF, GBM, SVM (later), MLR using our features
+ Finish Methods
+ Finish running code applied to 100 iterations of undersampling
+ Apply pipeline to K562 cell line

## 09/10/18

####Questions

* Should figures be placed in back of paper (in Appendix section)?
* Line 60 of manuscript: Check for accuracy?

####To-Do

+ in table with total features put (*) for removed then state why
+ order figures and create legends
* Finish results outline
* Finish results manuscript

## 09/17/18

####Questions

* clustered vs cell type specific tfbs
   +/home/sequencing/data/ExtData/db_5.00_07-22-2015/grsnp_db/hg19/ENCODE/

####To-Do

* Finish TFBS models
+ Include column for feature table for why a feature was excluded
+ Use tables instead of barplots for performance metrics
+ class imbalance barplot unnecessary, put in words
* clean up class imbalance plot put bar underneath to categorize each technique; put x axis diagonally
* put tad in boundary before tad not in boundary for class imbalance table
+ include venn diagram of selected features
* create a heat map for AIC values from commonly selected features
+ look at chromosome specific models and compare
* look at different resolutions (talk with john)
+ look at other type of transformation (inverse sine transformation)

## 09/19/18

+ split feature list into distance then counts
+ run mlr/lasso without distances to compare with mourad
* find out how to obtain consensus tree from random forest
+ do we need distance?
+ look at models with just counts vs just distance; compare

## 09/19/18

+ compare between mourad model and our pipeline at each step (full data vs balanced; distance vs no distance features)
   - lasso to lasso
   - mlr to mlr

## 09/24/18

+ consider inverse normal distribution as another normalization technique
   - understand conceptually what is happening

## 09/24/18

+ disregard count overlaps and use binary overlap with percent overlap if partial
+ reformat table1 into asking question how bad in unbalanced data
 + question 1 should be how severe is unbalanced data?
    - distance not needed here
* consider standardizing data prior to modeling
* is scaling important
 +just percent
 +just dist
 +both unscaled
 +both scaled