### 2018-10-03 Manuscript/results_outline.Rmd
edits

---
title: "Manuscript Outline"
author: "Spiro Stilianoudakis"
date: "September 11, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

   + Paragraph 1:
      - What are TADs and what role do they play in the human genome?
      - Why do we care about TADs?

   + Paragraph 2:
      - What is in the current literature?

   + Paragraph 3:
      - What are the issues?
      - Why are they issues?

   + Paragraph 4:
      - What are the current methods out there?
      - What drawbacks do they have?

   + Paragraph 5:
      - Defined the problem
      - What have we done?
      - Why have we done it?

# Methods

* The Data
   + Paragraph 1:
      - Where was the data obtained?
   + Paragraph 2:
      - How it was used?
      - How was the response vector created

* Genomic Annotations
   + Paragraph 1: 
      - Where was the annotation data obtained?
      - How were they used?

* Data Pipeline:
   + What was the pipeline used for?
   + Set up structure of pipeline
   + How were techniques evaluated?

* Normalization:
   + What techniques were used/compared?
   + How were they used/how do they worked?

* Class Imbalance:
   + What techniques were used/compared?
   + How were they used/how do they worked?

* Variable Selection:
   + What techniques were used/compared?
   + How were they used/how do they worked?

* Final Model:
   + What model(s) we used to evaluat the pipeline processed data
   + How the model(s) were performed
   + What we compared our models to

# Results

* Summary of data
   + How many TADs
   + How many genomic bins
   + What genomic elements were used

* Results of class imbalance
   + which technique performed best?

* Results for including distance predictors?
   + results for model with just distance
   + results for model with both types
   + compare 
   + results from normalization techniques

* Results from classification algorithms
   + MLR
   + LASSO
   + ENET
   + RF
   + GBM
   + SVM 
   + Naiive Bayes
   + Compare

# Discussion

* State the results that were interesting 
   - Specific to each cell line/tfbs
   - Then results all together
* State how much better models using our pipeline performed compared to Mourad
   - Was it due to the model choice or the use of distance features?
* Limitations/Future
   - Mention the need for HiC experiments using higher resolutions
   - Experimenting with TAD data using other methods
   - Exploring other cell lines