### 2018-12-03 Presentation/Abstract_MSU.Rmd

---
title: Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance
  Settings
output:
  pdf_document: default
  word_document: default
csl: C:/Users/Spiro Stilianoudakis/Documents/TAD_predictive_modeling/Manuscript/styles.ref/genomebiology.csl
---

Spiro C. Stilianoudakis^1^ (stilianoudasc@mymail.vcu.edu), Mikhail G. Dozmorov^1\*^ (mikhail.dozmorov@vcuhealth.org)

^1^ Dept. of Biostatistics, Virginia Commonwealth University, Richmond, VA, 23298, USA

# ABSTRACT

**Background.** Chromosome conformation capture sequencing technologies have shown that the three-dimensional (3D) structure of the genome is folded into distinct compartments, known as topologically associated domains (TADs) - units of coordinated gene expression. The location of TAD boundaries is highly conserved, suggesting the presence of epigenomic marks aiding in TAD formation. The ability to predict which epigenomic features are most associated with the TAD boundaries will allow to better understand the regulatory role of the 3D structure of the genome.

Existing methods for predicting associations between genomic elements tend to ignore key characteristics of the genomic data. Specifically, the number of TAD boundaries is much less than the number of other genomic regions, leading to heavily imbalanced classes. Furthermore, most methods utilize direct overlap as a means to quantify the association, while distance, the measure of spatial relationships, remains unaccounted for. Consequently, distances on a genomic scale vary widely, leaving uncertainty how the heavily right-tailed distribution of distance measures will affect the model’s performance.

**Methods.** We propose a novel data pre-processing pipeline that addresses those shortcomings. Lasso and Elastic Net were used for feature selection, coupled with five classifier methods. A number of classifier performance metrics were assessed, including the F1 measure and Matthew Correlation Coefficient (MCC), and area under the ROC curve (AUROC).

**Results.** Data preprocessing (log2-transformation and standardization) improves the performance of classification algorithms and allows for the ability to more accurately predict which genomic features are most associated with TAD boundaries. In contrast to overlap-based association measures, the distance between genomic elements was the most predictive measure. The random undersampling strategy addresses the class imbalance problem the most effectively in nearly all settings.

**Conclusions.** Current methods used to model the epigenomic features associated with TAD boundaries are insufficiently robust to handle properties of genomic data. Models applied to unprocessed data can have poor predictive performances. Focusing solely on standard performance assessment metrics, such as AUROC, can mask poor performance of the models; thus, the use of more balanced metrics, such as F1 and MCC, is warranted. Our model results in better performances and more accurate identification of important features associated with the formation of TAD boundaries.