### 2019-01-24 Manuscript/Tables/table_legends.Rmd
updated legends

---
title: "Legends for Tables"
author: "Spiro Stilianoudakis"
date: "November 19, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```