### 2019-01-25 Scripts/performances.rmd
added scripts folder used to create figures for results

---
title: "Performances Code"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)

```

# Evaluating Distance Densities
##GM12878
###10kb

```{r}
gm12878_10kb <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/train.rds")

gm12878_10kb_ld <- gm12878_10kb[,c(1,grep("_dist", names(gm12878_10kb)))]

###############
#gm12878_10kb_ld <- gm12878_10kb_ld[,c(1:5)]
###############

gm12878_10kb_d <- gm12878_10kb_ld
gm12878_10kb_d[,-1] <- exp(gm12878_10kb_ld[,-1]) - 1
gm12878_10kb_d$Average_dist <- rowMeans(gm12878_10kb_d[,-1])

gm12878_10kb_ld$Average_dist <- rowMeans(gm12878_10kb_ld[,-1])
names(gm12878_10kb_ld)[-1] <- paste(substr(names(gm12878_10kb_ld)[-1],
                                           1, 
                                           nchar(names(gm12878_10kb_ld)[-1])-4), 
                                    "logdist", 
                                    sep="")

gm12878_10kb_alldist <- cbind.data.frame(gm12878_10kb_d[,-1], gm12878_10kb_ld[,-1])

long10kb <- melt(gm12878_10kb_alldist)
long10kb$variable <- as.character(long10kb$variable)
long10kb$Transformation <- NA
long10kb$Transformation[grep("_logdist", long10kb$variable)] <- rep("Log Distance", length(long10kb$variable[grep("_logdist", long10kb$variable)]))
long10kb$Transformation[grep("_logdist", long10kb$variable, invert = TRUE)] <- rep("Distance", length(long10kb$variable[grep("_logdist", long10kb$variable, invert = TRUE)]))

table(long10kb$Transformation)

distdensity10kb <- ggplot() + 
  geom_density(data=long10kb[intersect(grep("_dist", long10kb$variable), grep("Average", long10kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long10kb[grep("Average_dist", long10kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,2.5e+07) + 
	ylab("")+ 
  scale_x_continuous(limits = c(0,2.5e+07), labels = c("0.0e+00"="0", "5.0e+06"="5", "1e+07"="10", "1.5e+07"="15", "2.0e+07"="20", "2.0e+07"="25"))+
	scale_fill_manual(values="red")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

logdistdensity10kb <- ggplot() + 
  geom_density(data=long10kb[intersect(grep("logdist", long10kb$variable), grep("Average", long10kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long10kb[grep("Average_logdist", long10kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("")+ 
	scale_fill_manual(values="blue")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

#rescaling nontransformed and combining two plots
long10kb$value[long10kb$Transformation=="Distance"] <- long10kb$value[long10kb$Transformation=="Distance"]/1000000
long10kb <- long10kb[grep("Average", long10kb$variable, invert = TRUE),]
long10kb$variable <- factor(long10kb$variable)
long10kb$Transformation <- factor(long10kb$Transformation)

alldensity10kb <- ggplot(long10kb, aes(x=value, group=variable, fill=Transformation)) + 
  geom_density(alpha=0.01) +
  scale_fill_manual(name = "", labels = c("Untransformed", "Log2"), values = c("red", "blue")) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("10 kb")+ 
  ylim(0,.8)+
  ggtitle("GM12878") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

###25kb

```{r}
gm12878_25kb <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/training_testing/train.rds")

gm12878_25kb_ld <- gm12878_25kb[,c(1,grep("_dist", names(gm12878_25kb)))]

###############
#gm12878_25kb_ld <- gm12878_25kb_ld[,c(1:5)]
###############

gm12878_25kb_d <- gm12878_25kb_ld
gm12878_25kb_d[,-1] <- exp(gm12878_25kb_ld[,-1]) - 1
gm12878_25kb_d$Average_dist <- rowMeans(gm12878_25kb_d[,-1])

gm12878_25kb_ld$Average_dist <- rowMeans(gm12878_25kb_ld[,-1])
names(gm12878_25kb_ld)[-1] <- paste(substr(names(gm12878_25kb_ld)[-1],
                                           1, 
                                           nchar(names(gm12878_25kb_ld)[-1])-4), 
                                    "logdist", 
                                    sep="")

gm12878_25kb_alldist <- cbind.data.frame(gm12878_25kb_d[,-1], gm12878_25kb_ld[,-1])

long25kb <- melt(gm12878_25kb_alldist)
long25kb$variable <- as.character(long25kb$variable)
long25kb$Transformation <- NA
long25kb$Transformation[grep("_logdist", long25kb$variable)] <- rep("Log Distance", length(long25kb$variable[grep("_logdist", long25kb$variable)]))
long25kb$Transformation[grep("_logdist", long25kb$variable, invert = TRUE)] <- rep("Distance", length(long25kb$variable[grep("_logdist", long25kb$variable, invert = TRUE)]))

table(long25kb$Transformation)

distdensity25kb <- ggplot() + 
  geom_density(data=long25kb[intersect(grep("_dist", long25kb$variable), grep("Average", long25kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long25kb[grep("Average_dist", long25kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,2.5e+07) + 
	ylab("")+ 
  scale_x_continuous(limits = c(0,2.5e+07), labels = c("0.0e+00"="0", "5.0e+06"="5", "1e+07"="10", "1.5e+07"="15", "2.0e+07"="20", "2.0e+07"="25"))+
	scale_fill_manual(values="red")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

logdistdensity25kb <- ggplot() + 
  geom_density(data=long25kb[intersect(grep("logdist", long25kb$variable), grep("Average", long25kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long25kb[grep("Average_logdist", long25kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) +  
	ylab("")+ 
	scale_fill_manual(values="blue")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

#rescaling nontransformed and combining two plots
long25kb$value[long25kb$Transformation=="Distance"] <- long25kb$value[long25kb$Transformation=="Distance"]/1000000
long25kb <- long25kb[grep("Average", long25kb$variable, invert = TRUE),]
long25kb$variable <- factor(long25kb$variable)
long25kb$Transformation <- factor(long25kb$Transformation)

alldensity25kb <- ggplot(long25kb, aes(x=value, group=variable, fill=Transformation)) + 
  geom_density(alpha=0.01) +
  scale_fill_manual(name = "", labels = c("Untransformed", "Log2"), values = c("red", "blue")) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("25 kb")+ 
  ylim(0,.8)+
  #ggtitle("GM12878") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

###50kb

```{r}
gm12878_50kb <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/training_testing/train.rds")

gm12878_50kb_ld <- gm12878_50kb[,c(1,grep("_dist", names(gm12878_50kb)))]

###############
#gm12878_50kb_ld <- gm12878_50kb_ld[,c(1:5)]
###############

gm12878_50kb_d <- gm12878_50kb_ld
gm12878_50kb_d[,-1] <- exp(gm12878_50kb_ld[,-1]) - 1
gm12878_50kb_d$Average_dist <- rowMeans(gm12878_50kb_d[,-1])

gm12878_50kb_ld$Average_dist <- rowMeans(gm12878_50kb_ld[,-1])
names(gm12878_50kb_ld)[-1] <- paste(substr(names(gm12878_50kb_ld)[-1],
                                           1, 
                                           nchar(names(gm12878_50kb_ld)[-1])-4), 
                                    "logdist", 
                                    sep="")

gm12878_50kb_alldist <- cbind.data.frame(gm12878_50kb_d[,-1], gm12878_50kb_ld[,-1])

long50kb <- melt(gm12878_50kb_alldist)
long50kb$variable <- as.character(long50kb$variable)
long50kb$Transformation <- NA
long50kb$Transformation[grep("_logdist", long50kb$variable)] <- rep("Log Distance", length(long50kb$variable[grep("_logdist", long50kb$variable)]))
long50kb$Transformation[grep("_logdist", long50kb$variable, invert = TRUE)] <- rep("Distance", length(long50kb$variable[grep("_logdist", long50kb$variable, invert = TRUE)]))

table(long50kb$Transformation)

distdensity50kb <- ggplot() + 
  geom_density(data=long50kb[intersect(grep("_dist", long50kb$variable), grep("Average", long50kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long50kb[grep("Average_dist", long50kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,2.5e+07) + 
	ylab("")+ 
  scale_x_continuous(limits = c(0,2.5e+07), labels = c("0.0e+00"="0", "5.0e+06"="5", "1e+07"="10", "1.5e+07"="15", "2.0e+07"="20", "2.0e+07"="25"))+
	scale_fill_manual(values="red")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

logdistdensity50kb <- ggplot() + 
  geom_density(data=long50kb[intersect(grep("logdist", long50kb$variable), grep("Average", long50kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long50kb[grep("Average_logdist", long50kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("")+ 
	scale_fill_manual(values="blue")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

#rescaling nontransformed and combining two plots
long50kb$value[long50kb$Transformation=="Distance"] <- long50kb$value[long50kb$Transformation=="Distance"]/1000000
long50kb <- long50kb[grep("Average", long50kb$variable, invert = TRUE),]
long50kb$variable <- factor(long50kb$variable)
long50kb$Transformation <- factor(long50kb$Transformation)

alldensity50kb <- ggplot(long50kb, aes(x=value, group=variable, fill=Transformation)) + 
  geom_density(alpha=0.01) +
  scale_fill_manual(name = "", labels = c("Untransformed", "Log2"), values = c("red", "blue")) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("50 kb")+ 
  ylim(0,.8)+
  #ggtitle("GM12878") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

###100kb

```{r}
gm12878_100kb <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/training_testing/train.rds")

gm12878_100kb_ld <- gm12878_100kb[,c(1,grep("_dist", names(gm12878_100kb)))]

###############
#gm12878_100kb_ld <- gm12878_100kb_ld[,c(1:5)]
###############

gm12878_100kb_d <- gm12878_100kb_ld
gm12878_100kb_d[,-1] <- exp(gm12878_100kb_ld[,-1]) - 1
gm12878_100kb_d$Average_dist <- rowMeans(gm12878_100kb_d[,-1])

gm12878_100kb_ld$Average_dist <- rowMeans(gm12878_100kb_ld[,-1])
names(gm12878_100kb_ld)[-1] <- paste(substr(names(gm12878_100kb_ld)[-1],
                                           1, 
                                           nchar(names(gm12878_100kb_ld)[-1])-4), 
                                    "logdist", 
                                    sep="")

gm12878_100kb_alldist <- cbind.data.frame(gm12878_100kb_d[,-1], gm12878_100kb_ld[,-1])

long100kb <- melt(gm12878_100kb_alldist)
long100kb$variable <- as.character(long100kb$variable)
long100kb$Transformation <- NA
long100kb$Transformation[grep("_logdist", long100kb$variable)] <- rep("Log Distance", length(long100kb$variable[grep("_logdist", long100kb$variable)]))
long100kb$Transformation[grep("_logdist", long100kb$variable, invert = TRUE)] <- rep("Distance", length(long100kb$variable[grep("_logdist", long100kb$variable, invert = TRUE)]))

table(long100kb$Transformation)

distdensity100kb <- ggplot() + 
  geom_density(data=long100kb[intersect(grep("_dist", long100kb$variable), grep("Average", long100kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long100kb[grep("Average_dist", long100kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("No Transformation") + 
	xlim(0,2.5e+07) + 
	ylab("")+ 
  scale_x_continuous(limits = c(0,2.5e+07), labels = c("0.0e+00"="0", "5.0e+06"="5", "1e+07"="10", "1.5e+07"="15", "2.0e+07"="20", "2.0e+07"="25"))+
	scale_fill_manual(values="red")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

logdistdensity100kb <- ggplot() + 
  geom_density(data=long100kb[intersect(grep("logdist", long100kb$variable), grep("Average", long100kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long100kb[grep("Average_logdist", long100kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("Log Tranformation") + 
	xlim(0,25) + 
	ylab("")+ 
	scale_fill_manual(values="blue")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

#rescaling nontransformed and combining two plots
long100kb$value[long100kb$Transformation=="Distance"] <- long100kb$value[long100kb$Transformation=="Distance"]/1000000
long100kb <- long100kb[grep("Average", long100kb$variable, invert = TRUE),]
long100kb$variable <- factor(long100kb$variable)
long100kb$Transformation <- factor(long100kb$Transformation)

alldensity100kb <- ggplot(long100kb, aes(x=value, group=variable, fill=Transformation)) + 
  geom_density(alpha=0.01) +
  scale_fill_manual(name = "", labels = c("Untransformed", "Log2"), values = c("red", "blue")) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("100 kb")+ 
  ylim(0,.8)+
  #ggtitle("GM12878") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

##K562
###10kb

```{r}
K562_10kb <- readRDS("Z:/TAD_data_analysis/K562/10kb/training_testing/train.rds")

K562_10kb_ld <- K562_10kb[,c(1,grep("_dist", names(K562_10kb)))]

###############
#K562_10kb_ld <- K562_10kb_ld[,c(1:5)]
###############

K562_10kb_d <- K562_10kb_ld
K562_10kb_d[,-1] <- exp(K562_10kb_ld[,-1]) - 1
K562_10kb_d$Average_dist <- rowMeans(K562_10kb_d[,-1])

K562_10kb_ld$Average_dist <- rowMeans(K562_10kb_ld[,-1])
names(K562_10kb_ld)[-1] <- paste(substr(names(K562_10kb_ld)[-1],
                                           1, 
                                           nchar(names(K562_10kb_ld)[-1])-4), 
                                    "logdist", 
                                    sep="")

K562_10kb_alldist <- cbind.data.frame(K562_10kb_d[,-1], K562_10kb_ld[,-1])

long10kb <- melt(K562_10kb_alldist)
long10kb$variable <- as.character(long10kb$variable)
long10kb$Transformation <- NA
long10kb$Transformation[grep("_logdist", long10kb$variable)] <- rep("Log Distance", length(long10kb$variable[grep("_logdist", long10kb$variable)]))
long10kb$Transformation[grep("_logdist", long10kb$variable, invert = TRUE)] <- rep("Distance", length(long10kb$variable[grep("_logdist", long10kb$variable, invert = TRUE)]))

table(long10kb$Transformation)

distdensity10kb <- ggplot() + 
  geom_density(data=long10kb[intersect(grep("_dist", long10kb$variable), grep("Average", long10kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long10kb[grep("Average_dist", long10kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,2.5e+07) + 
	ylab("")+ 
  scale_x_continuous(limits = c(0,2.5e+07), labels = c("0.0e+00"="0", "5.0e+06"="5", "1e+07"="10", "1.5e+07"="15", "2.0e+07"="20", "2.0e+07"="25"))+
	scale_fill_manual(values="red")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

logdistdensity10kb <- ggplot() + 
  geom_density(data=long10kb[intersect(grep("logdist", long10kb$variable), grep("Average", long10kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long10kb[grep("Average_logdist", long10kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("")+ 
	scale_fill_manual(values="blue")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

#rescaling nontransformed and combining two plots
long10kb$value[long10kb$Transformation=="Distance"] <- long10kb$value[long10kb$Transformation=="Distance"]/1000000
long10kb <- long10kb[grep("Average", long10kb$variable, invert = TRUE),]
long10kb$variable <- factor(long10kb$variable)
long10kb$Transformation <- factor(long10kb$Transformation)

alldensity10kb_k562 <- ggplot(long10kb, aes(x=value, group=variable, fill=Transformation)) + 
  geom_density(alpha=0.01) +
  scale_fill_manual(name = "", labels = c("Untransformed", "Log2"), values = c("red", "blue")) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	#ylab("10 kb")+ 
  ylim(0,.8)+
  ggtitle("K562") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

###25kb

```{r}
K562_25kb <- readRDS("Z:/TAD_data_analysis/K562/25kb/training_testing/train.rds")

K562_25kb_ld <- K562_25kb[,c(1,grep("_dist", names(K562_25kb)))]

###############
#K562_25kb_ld <- K562_25kb_ld[,c(1:5)]
###############

K562_25kb_d <- K562_25kb_ld
K562_25kb_d[,-1] <- exp(K562_25kb_ld[,-1]) - 1
K562_25kb_d$Average_dist <- rowMeans(K562_25kb_d[,-1])

K562_25kb_ld$Average_dist <- rowMeans(K562_25kb_ld[,-1])
names(K562_25kb_ld)[-1] <- paste(substr(names(K562_25kb_ld)[-1],
                                           1, 
                                           nchar(names(K562_25kb_ld)[-1])-4), 
                                    "logdist", 
                                    sep="")

K562_25kb_alldist <- cbind.data.frame(K562_25kb_d[,-1], K562_25kb_ld[,-1])

long25kb <- melt(K562_25kb_alldist)
long25kb$variable <- as.character(long25kb$variable)
long25kb$Transformation <- NA
long25kb$Transformation[grep("_logdist", long25kb$variable)] <- rep("Log Distance", length(long25kb$variable[grep("_logdist", long25kb$variable)]))
long25kb$Transformation[grep("_logdist", long25kb$variable, invert = TRUE)] <- rep("Distance", length(long25kb$variable[grep("_logdist", long25kb$variable, invert = TRUE)]))

table(long25kb$Transformation)

distdensity25kb <- ggplot() + 
  geom_density(data=long25kb[intersect(grep("_dist", long25kb$variable), grep("Average", long25kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long25kb[grep("Average_dist", long25kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,2.5e+07) + 
	ylab("")+ 
  scale_x_continuous(limits = c(0,2.5e+07), labels = c("0.0e+00"="0", "5.0e+06"="5", "1e+07"="10", "1.5e+07"="15", "2.0e+07"="20", "2.0e+07"="25"))+
	scale_fill_manual(values="red")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

logdistdensity25kb <- ggplot() + 
  geom_density(data=long25kb[intersect(grep("logdist", long25kb$variable), grep("Average", long25kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long25kb[grep("Average_logdist", long25kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("")+ 
	scale_fill_manual(values="blue")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

#rescaling nontransformed and combining two plots
long25kb$value[long25kb$Transformation=="Distance"] <- long25kb$value[long25kb$Transformation=="Distance"]/1000000
long25kb <- long25kb[grep("Average", long25kb$variable, invert = TRUE),]
long25kb$variable <- factor(long25kb$variable)
long25kb$Transformation <- factor(long25kb$Transformation)

alldensity25kb_k562 <- ggplot(long25kb, aes(x=value, group=variable, fill=Transformation)) + 
  geom_density(alpha=0.01) +
  scale_fill_manual(name = "", labels = c("Untransformed", "Log2"), values = c("red", "blue")) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	#ylab("25 kb")+ 
  ylim(0,.8)+
  #ggtitle("K562") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

###50kb

```{r}
K562_50kb <- readRDS("Z:/TAD_data_analysis/K562/50kb/training_testing/train.rds")

K562_50kb_ld <- K562_50kb[,c(1,grep("_dist", names(K562_50kb)))]

###############
#K562_50kb_ld <- K562_50kb_ld[,c(1:5)]
###############

K562_50kb_d <- K562_50kb_ld
K562_50kb_d[,-1] <- exp(K562_50kb_ld[,-1]) - 1
K562_50kb_d$Average_dist <- rowMeans(K562_50kb_d[,-1])

K562_50kb_ld$Average_dist <- rowMeans(K562_50kb_ld[,-1])
names(K562_50kb_ld)[-1] <- paste(substr(names(K562_50kb_ld)[-1],
                                           1, 
                                           nchar(names(K562_50kb_ld)[-1])-4), 
                                    "logdist", 
                                    sep="")

K562_50kb_alldist <- cbind.data.frame(K562_50kb_d[,-1], K562_50kb_ld[,-1])

long50kb <- melt(K562_50kb_alldist)
long50kb$variable <- as.character(long50kb$variable)
long50kb$Transformation <- NA
long50kb$Transformation[grep("_logdist", long50kb$variable)] <- rep("Log Distance", length(long50kb$variable[grep("_logdist", long50kb$variable)]))
long50kb$Transformation[grep("_logdist", long50kb$variable, invert = TRUE)] <- rep("Distance", length(long50kb$variable[grep("_logdist", long50kb$variable, invert = TRUE)]))

table(long50kb$Transformation)

distdensity50kb <- ggplot() + 
  geom_density(data=long50kb[intersect(grep("_dist", long50kb$variable), grep("Average", long50kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long50kb[grep("Average_dist", long50kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,2.5e+07) + 
	ylab("")+ 
  scale_x_continuous(limits = c(0,2.5e+07), labels = c("0.0e+00"="0", "5.0e+06"="5", "1e+07"="10", "1.5e+07"="15", "2.0e+07"="20", "2.0e+07"="25"))+
	scale_fill_manual(values="red")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

logdistdensity50kb <- ggplot() + 
  geom_density(data=long50kb[intersect(grep("logdist", long50kb$variable), grep("Average", long50kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long50kb[grep("Average_logdist", long50kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("")+ 
	scale_fill_manual(values="blue")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

#rescaling nontransformed and combining two plots
long50kb$value[long50kb$Transformation=="Distance"] <- long50kb$value[long50kb$Transformation=="Distance"]/1000000
long50kb <- long50kb[grep("Average", long50kb$variable, invert = TRUE),]
long50kb$variable <- factor(long50kb$variable)
long50kb$Transformation <- factor(long50kb$Transformation)

alldensity50kb_k562 <- ggplot(long50kb, aes(x=value, group=variable, fill=Transformation)) + 
  geom_density(alpha=0.01) +
  scale_fill_manual(name = "", labels = c("Untransformed", "Log2"), values = c("red", "blue")) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	#ylab("50 kb")+ 
  ylim(0,.8)+
  #ggtitle("K562") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

###100kb

```{r}
K562_100kb <- readRDS("Z:/TAD_data_analysis/K562/100kb/training_testing/train.rds")

K562_100kb_ld <- K562_100kb[,c(1,grep("_dist", names(K562_100kb)))]

###############
#K562_100kb_ld <- K562_100kb_ld[,c(1:5)]
###############

K562_100kb_d <- K562_100kb_ld
K562_100kb_d[,-1] <- exp(K562_100kb_ld[,-1]) - 1
K562_100kb_d$Average_dist <- rowMeans(K562_100kb_d[,-1])

K562_100kb_ld$Average_dist <- rowMeans(K562_100kb_ld[,-1])
names(K562_100kb_ld)[-1] <- paste(substr(names(K562_100kb_ld)[-1],
                                           1, 
                                           nchar(names(K562_100kb_ld)[-1])-4), 
                                    "logdist", 
                                    sep="")

K562_100kb_alldist <- cbind.data.frame(K562_100kb_d[,-1], K562_100kb_ld[,-1])

long100kb <- melt(K562_100kb_alldist)
long100kb$variable <- as.character(long100kb$variable)
long100kb$Transformation <- NA
long100kb$Transformation[grep("_logdist", long100kb$variable)] <- rep("Log Distance", length(long100kb$variable[grep("_logdist", long100kb$variable)]))
long100kb$Transformation[grep("_logdist", long100kb$variable, invert = TRUE)] <- rep("Distance", length(long100kb$variable[grep("_logdist", long100kb$variable, invert = TRUE)]))

table(long100kb$Transformation)

distdensity100kb <- ggplot() + 
  geom_density(data=long100kb[intersect(grep("_dist", long100kb$variable), grep("Average", long100kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long100kb[grep("Average_dist", long100kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,2.5e+07) + 
	ylab("")+ 
  scale_x_continuous(limits = c(0,2.5e+07), labels = c("0.0e+00"="0", "5.0e+06"="5", "1e+07"="10", "1.5e+07"="15", "2.0e+07"="20", "2.0e+07"="25"))+
	scale_fill_manual(values="red")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

logdistdensity100kb <- ggplot() + 
  geom_density(data=long100kb[intersect(grep("logdist", long100kb$variable), grep("Average", long100kb$variable, invert = TRUE)),]
               ,aes(x=value, 
                    group=variable, 
                    fill=Transformation), alpha=0.01, size=0.5) +
  #geom_density(data=long100kb[grep("Average_logdist", long100kb$variable),]
  #             ,aes(x=value), size = 1) +
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	ylab("")+ 
	scale_fill_manual(values="blue")+
		#labels=as.character(
		#		unique(
		#			sort(
		#				as.numeric(
		#					substr(as.character(domains$Chromosome),4,nchar(as.character(domains$Chromosome))))))),
		#values=rainbow(22)) +
	theme_minimal() +
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())
  #theme(axis.line = element_line(color="black", size = .5))

#rescaling nontransformed and combining two plots
long100kb$value[long100kb$Transformation=="Distance"] <- long100kb$value[long100kb$Transformation=="Distance"]/1000000
long100kb <- long100kb[grep("Average", long100kb$variable, invert = TRUE),]
long100kb$variable <- factor(long100kb$variable)
long100kb$Transformation <- factor(long100kb$Transformation)

alldensity100kb_k562 <- ggplot(long100kb, aes(x=value, group=variable, fill=Transformation)) + 
  geom_density(alpha=0.01) +
  scale_fill_manual(name = "", labels = c("Untransformed", "Log2"), values = c("red", "blue")) +
  guides(fill = guide_legend(override.aes = list(alpha=1)))+
  geom_vline(xintercept=0, size=1)+
  geom_hline(yintercept=0, size=1)+
	xlab("") + 
	xlim(0,25) + 
	#ylab("100 kb")+ 
  ylim(0,.8)+
  #ggtitle("K562") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

##Combined

```{r}
#alldistplots <- grid.arrange(distdensity10kb, logdistdensity10kb,
#             distdensity25kb, logdistdensity25kb,
#             distdensity50kb, logdistdensity50kb,
#             distdensity100kb, logdistdensity100kb, ncol=2, nrow=4)

alldistplots <- ggarrange(alldensity10kb,
                          alldensity10kb_k562,
                          alldensity25kb,
                          alldensity25kb_k562,
                          alldensity50kb,
                          alldensity50kb_k562,
                          alldensity100kb,
                          alldensity100kb_k562, 
          ncol=2, 
          nrow=4,
          #labels = c("A","", "B","", "C","", "D",""),
          legend = "top",
          common.legend = TRUE)

annotate_figure(alldistplots,
                bottom = text_grob("Distance to Nearest TAD Boundary \n (in millions of basepairs)", 
                                color = "black", 
                                face = "bold", 
                                size = 15),
                left = text_grob("Density", 
                                 color = "black",
                                 size = 15,
                                 rot = 90))
```

#Evaluating Imbalance Ratios
##GM12878
###10kb

```{r}
train <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/train.rds")
test <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/test.rds")

set.seed(123)
sampids <- sample(which(train$y=="No"),
                        length(which(train$y=="Yes")),
                        replace = FALSE)
data.rus <- rbind.data.frame(train[which(train$y=="Yes"),],
                           train[sampids,])

set.seed(123)		
sampids <- sample(which(train$y=="Yes"),
                        length(which(train$y=="No")),
                        replace = TRUE)
data.ros <- rbind.data.frame(train[which(train$y=="No"),],
                           train[sampids,])

set.seed(123)
data.sm <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

names(train)[-1] <- gsub("-","_",names(train)[-1])
data.rose <- ROSE(y ~ ., data=train,seed=1,p=0.5)$data

#set.seed(123)
#n <- 1000
#nr <- nrow(train)
#splitdat <- split(train, rep(1:ceiling(nr/n), each=n, length.out=nr))
#meds <- max(unlist(lapply(splitdat, function(x){length(which(x$y=="Yes"))})))
#medoids <- list(as.numeric(),length(splitdat))
#for(i in 1:length(splitdat)){
#  print(i)
#  data <- splitdat[[i]]
#  majority <- data[which(data$y=="No"),-1]
#  minnum <- length(which(data$y=="Yes"))
# 
#  pam.med <- pam(majority, k = minnum, diss = FALSE,
#                 metric = "manhattan", 
#                 #medoids = NULL, 
#                 #stand = FALSE, 
#                 cluster.only = FALSE,
#                 do.swap = TRUE,
#                 keep.diss = FALSE,
#                 keep.data = FALSE,
#                 pamonce = FALSE, trace.lev = 2)
#  
#  medoids[[i]] <- as.numeric(rownames(pam.med$medoids))
#  
#}
#medoids <- unlist(medoids)
#data.km <- rbind.data.frame(train[which(as.numeric(rownames(train)) %in% medoids),],
#                               train[which(train$y=="Yes"),])

x10 <- data.frame(IR = c((length(train$y[which(train$y=="Yes")])+length(test$y[which(test$y=="Yes")]))/(length(train$y[which(train$y=="No")])+length(test$y[which(test$y=="No")])), 
					  length(train$y[which(train$y=="Yes")])/length(train$y[which(train$y=="No")]),
					  length(test$y[which(test$y=="Yes")])/length(test$y[which(test$y=="No")]),
					  length(data.rus$y[which(data.rus$y=="Yes")])/length(data.rus$y[which(data.rus$y=="No")]),
					  length(data.ros$y[which(data.ros$y=="Yes")])/length(data.ros$y[which(data.ros$y=="No")]),
					  length(data.sm$y[which(data.sm$y=="Yes")])/length(data.sm$y[which(data.sm$y=="No")]),
					  length(data.rose$y[which(data.rose$y=="Yes")])/length(data.rose$y[which(data.rose$y=="No")])),
					  #length(data.km$y[which(data.km$y=="No")]), 
					  #length(data.km$y[which(data.km$y=="Yes")])),
			    CB = c("Complete",
					   "Train",
					   "Test",
					   "Under-Sampling",
					   "Over-Sampling",
					   "SMOTE",
					   "ROSE"))

p10kb <- ggplot(x10, aes(y=IR, x=factor(CB, levels = unique(x10$CB)))) + 
  geom_bar( stat="identity") +
  ylab("Imbalance Ratio")+
  xlab("")+
  #ylim(0, 30000)+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15, 
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

###25kb

```{r}
train <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/training_testing/train.rds")
test <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/training_testing/test.rds")

set.seed(123)
sampids <- sample(which(train$y=="No"),
                        length(which(train$y=="Yes")),
                        replace = FALSE)
data.rus <- rbind.data.frame(train[which(train$y=="Yes"),],
                           train[sampids,])

set.seed(123)		
sampids <- sample(which(train$y=="Yes"),
                        length(which(train$y=="No")),
                        replace = TRUE)
data.ros <- rbind.data.frame(train[which(train$y=="No"),],
                           train[sampids,])

set.seed(123)
data.sm <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

names(train)[-1] <- gsub("-","_",names(train)[-1])
data.rose <- ROSE(y ~ ., data=train,seed=1,p=0.5)$data

#set.seed(123)
#n <- 1000
#nr <- nrow(train)
#splitdat <- split(train, rep(1:ceiling(nr/n), each=n, length.out=nr))
#meds <- max(unlist(lapply(splitdat, function(x){length(which(x$y=="Yes"))})))
#medoids <- list(as.numeric(),length(splitdat))
#for(i in 1:length(splitdat)){
#  print(i)
#  data <- splitdat[[i]]
#  majority <- data[which(data$y=="No"),-1]
#  minnum <- length(which(data$y=="Yes"))
# 
#  pam.med <- pam(majority, k = minnum, diss = FALSE,
#                 metric = "manhattan", 
#                 #medoids = NULL, 
#                 #stand = FALSE, 
#                 cluster.only = FALSE,
#                 do.swap = TRUE,
#                 keep.diss = FALSE,
#                 keep.data = FALSE,
#                 pamonce = FALSE, trace.lev = 2)
#  
#  medoids[[i]] <- as.numeric(rownames(pam.med$medoids))
#  
#}
#medoids <- unlist(medoids)
#data.km <- rbind.data.frame(train[which(as.numeric(rownames(train)) %in% medoids),],
#                               train[which(train$y=="Yes"),])

x25 <- data.frame(IR = c((length(train$y[which(train$y=="Yes")])+length(test$y[which(test$y=="Yes")]))/(length(train$y[which(train$y=="No")])+length(test$y[which(test$y=="No")])), 
					  length(train$y[which(train$y=="Yes")])/length(train$y[which(train$y=="No")]),
					  length(test$y[which(test$y=="Yes")])/length(test$y[which(test$y=="No")]),
					  length(data.rus$y[which(data.rus$y=="Yes")])/length(data.rus$y[which(data.rus$y=="No")]),
					  length(data.ros$y[which(data.ros$y=="Yes")])/length(data.ros$y[which(data.ros$y=="No")]),
					  length(data.sm$y[which(data.sm$y=="Yes")])/length(data.sm$y[which(data.sm$y=="No")]),
					  length(data.rose$y[which(data.rose$y=="Yes")])/length(data.rose$y[which(data.rose$y=="No")])),
					  #length(data.km$y[which(data.km$y=="No")]), 
					  #length(data.km$y[which(data.km$y=="Yes")])),
			    CB = c("Complete",
					   "Train",
					   "Test",
					   "Under-Sampling",
					   "Over-Sampling",
					   "SMOTE",
					   "ROSE"))

p25kb <- ggplot(x25, aes(y=IR, x=factor(CB, levels = unique(x25$CB)))) + 
  geom_bar( stat="identity") +
  ylab("Imbalance Ratio")+
  xlab("")+
  #ylim(0, 30000)+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15, 
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

###50kb

```{r}
train <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/training_testing/train.rds")
test <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/training_testing/test.rds")

set.seed(123)
sampids <- sample(which(train$y=="No"),
                        length(which(train$y=="Yes")),
                        replace = FALSE)
data.rus <- rbind.data.frame(train[which(train$y=="Yes"),],
                           train[sampids,])

set.seed(123)		
sampids <- sample(which(train$y=="Yes"),
                        length(which(train$y=="No")),
                        replace = TRUE)
data.ros <- rbind.data.frame(train[which(train$y=="No"),],
                           train[sampids,])

set.seed(123)
data.sm <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

names(train)[-1] <- gsub("-","_",names(train)[-1])
data.rose <- ROSE(y ~ ., data=train,seed=1,p=0.5)$data

#set.seed(123)
#n <- 1000
#nr <- nrow(train)
#splitdat <- split(train, rep(1:ceiling(nr/n), each=n, length.out=nr))
#meds <- max(unlist(lapply(splitdat, function(x){length(which(x$y=="Yes"))})))
#medoids <- list(as.numeric(),length(splitdat))
#for(i in 1:length(splitdat)){
#  print(i)
#  data <- splitdat[[i]]
#  majority <- data[which(data$y=="No"),-1]
#  minnum <- length(which(data$y=="Yes"))
# 
#  pam.med <- pam(majority, k = minnum, diss = FALSE,
#                 metric = "manhattan", 
#                 #medoids = NULL, 
#                 #stand = FALSE, 
#                 cluster.only = FALSE,
#                 do.swap = TRUE,
#                 keep.diss = FALSE,
#                 keep.data = FALSE,
#                 pamonce = FALSE, trace.lev = 2)
#  
#  medoids[[i]] <- as.numeric(rownames(pam.med$medoids))
#  
#}
#medoids <- unlist(medoids)
#data.km <- rbind.data.frame(train[which(as.numeric(rownames(train)) %in% medoids),],
#                               train[which(train$y=="Yes"),])

x50 <- data.frame(IR = c((length(train$y[which(train$y=="Yes")])+length(test$y[which(test$y=="Yes")]))/(length(train$y[which(train$y=="No")])+length(test$y[which(test$y=="No")])), 
					  length(train$y[which(train$y=="Yes")])/length(train$y[which(train$y=="No")]),
					  length(test$y[which(test$y=="Yes")])/length(test$y[which(test$y=="No")]),
					  length(data.rus$y[which(data.rus$y=="Yes")])/length(data.rus$y[which(data.rus$y=="No")]),
					  length(data.ros$y[which(data.ros$y=="Yes")])/length(data.ros$y[which(data.ros$y=="No")]),
					  length(data.sm$y[which(data.sm$y=="Yes")])/length(data.sm$y[which(data.sm$y=="No")]),
					  length(data.rose$y[which(data.rose$y=="Yes")])/length(data.rose$y[which(data.rose$y=="No")])),
					  #length(data.km$y[which(data.km$y=="No")]), 
					  #length(data.km$y[which(data.km$y=="Yes")])),
			    CB = c("Complete",
					   "Train",
					   "Test",
					   "Under-Sampling",
					   "Over-Sampling",
					   "SMOTE",
					   "ROSE"))

p50kb <- ggplot(x50, aes(y=IR, x=factor(CB, levels = unique(x50$CB)))) + 
  geom_bar( stat="identity") +
  ylab("Imbalance Ratio")+
  xlab("")+
  #ylim(0, 30000)+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15, 
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

###100kb

```{r}
train <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/training_testing/train.rds")
test <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/training_testing/test.rds")

set.seed(123)
sampids <- sample(which(train$y=="No"),
                        length(which(train$y=="Yes")),
                        replace = FALSE)
data.rus <- rbind.data.frame(train[which(train$y=="Yes"),],
                           train[sampids,])

set.seed(123)		
sampids <- sample(which(train$y=="Yes"),
                        length(which(train$y=="No")),
                        replace = TRUE)
data.ros <- rbind.data.frame(train[which(train$y=="No"),],
                           train[sampids,])

set.seed(123)
data.sm <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

names(train)[-1] <- gsub("-","_",names(train)[-1])
data.rose <- ROSE(y ~ ., data=train,seed=1,p=0.5)$data

#set.seed(123)
#n <- 1000
#nr <- nrow(train)
#splitdat <- split(train, rep(1:ceiling(nr/n), each=n, length.out=nr))
#meds <- max(unlist(lapply(splitdat, function(x){length(which(x$y=="Yes"))})))
#medoids <- list(as.numeric(),length(splitdat))
#for(i in 1:length(splitdat)){
#  print(i)
#  data <- splitdat[[i]]
#  majority <- data[which(data$y=="No"),-1]
#  minnum <- length(which(data$y=="Yes"))
# 
#  pam.med <- pam(majority, k = minnum, diss = FALSE,
#                 metric = "manhattan", 
#                 #medoids = NULL, 
#                 #stand = FALSE, 
#                 cluster.only = FALSE,
#                 do.swap = TRUE,
#                 keep.diss = FALSE,
#                 keep.data = FALSE,
#                 pamonce = FALSE, trace.lev = 2)
#  
#  medoids[[i]] <- as.numeric(rownames(pam.med$medoids))
#  
#}
#medoids <- unlist(medoids)
#data.km <- rbind.data.frame(train[which(as.numeric(rownames(train)) %in% medoids),],
#                               train[which(train$y=="Yes"),])

x100 <- data.frame(IR = c((length(train$y[which(train$y=="Yes")])+length(test$y[which(test$y=="Yes")]))/(length(train$y[which(train$y=="No")])+length(test$y[which(test$y=="No")])), 
					  length(train$y[which(train$y=="Yes")])/length(train$y[which(train$y=="No")]),
					  length(test$y[which(test$y=="Yes")])/length(test$y[which(test$y=="No")]),
					  length(data.rus$y[which(data.rus$y=="Yes")])/length(data.rus$y[which(data.rus$y=="No")]),
					  length(data.ros$y[which(data.ros$y=="Yes")])/length(data.ros$y[which(data.ros$y=="No")]),
					  length(data.sm$y[which(data.sm$y=="Yes")])/length(data.sm$y[which(data.sm$y=="No")]),
					  length(data.rose$y[which(data.rose$y=="Yes")])/length(data.rose$y[which(data.rose$y=="No")])),
					  #length(data.km$y[which(data.km$y=="No")]), 
					  #length(data.km$y[which(data.km$y=="Yes")])),
			    CB = c("Complete",
					   "Train",
					   "Test",
					   "Under-Sampling",
					   "Over-Sampling",
					   "SMOTE",
					   "ROSE"))

p100kb <- ggplot(x100, aes(y=IR, x=factor(CB, levels = unique(x100$CB)))) + 
  geom_bar( stat="identity") +
  ylab("Imbalance Ratio")+
  xlab("")+
  #ylim(0, 30000)+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15, 
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

##K562
###10kb

```{r}
train <- readRDS("Z:/TAD_data_analysis/K562/10kb/training_testing/train.rds")
test <- readRDS("Z:/TAD_data_analysis/K562/10kb/training_testing/test.rds")

set.seed(123)
sampids <- sample(which(train$y=="No"),
                        length(which(train$y=="Yes")),
                        replace = FALSE)
data.rus <- rbind.data.frame(train[which(train$y=="Yes"),],
                           train[sampids,])

set.seed(123)		
sampids <- sample(which(train$y=="Yes"),
                        length(which(train$y=="No")),
                        replace = TRUE)
data.ros <- rbind.data.frame(train[which(train$y=="No"),],
                           train[sampids,])

set.seed(123)
data.sm <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

names(train)[-1] <- gsub("-","_",names(train)[-1])
data.rose <- ROSE(y ~ ., data=train,seed=1,p=0.5)$data

#set.seed(123)
#n <- 1000
#nr <- nrow(train)
#splitdat <- split(train, rep(1:ceiling(nr/n), each=n, length.out=nr))
#meds <- max(unlist(lapply(splitdat, function(x){length(which(x$y=="Yes"))})))
#medoids <- list(as.numeric(),length(splitdat))
#for(i in 1:length(splitdat)){
#  print(i)
#  data <- splitdat[[i]]
#  majority <- data[which(data$y=="No"),-1]
#  minnum <- length(which(data$y=="Yes"))
# 
#  pam.med <- pam(majority, k = minnum, diss = FALSE,
#                 metric = "manhattan", 
#                 #medoids = NULL, 
#                 #stand = FALSE, 
#                 cluster.only = FALSE,
#                 do.swap = TRUE,
#                 keep.diss = FALSE,
#                 keep.data = FALSE,
#                 pamonce = FALSE, trace.lev = 2)
#  
#  medoids[[i]] <- as.numeric(rownames(pam.med$medoids))
#  
#}
#medoids <- unlist(medoids)
#data.km <- rbind.data.frame(train[which(as.numeric(rownames(train)) %in% medoids),],
#                               train[which(train$y=="Yes"),])

x10_k562 <- data.frame(IR = c((length(train$y[which(train$y=="Yes")])+length(test$y[which(test$y=="Yes")]))/(length(train$y[which(train$y=="No")])+length(test$y[which(test$y=="No")])), 
					  length(train$y[which(train$y=="Yes")])/length(train$y[which(train$y=="No")]),
					  length(test$y[which(test$y=="Yes")])/length(test$y[which(test$y=="No")]),
					  length(data.rus$y[which(data.rus$y=="Yes")])/length(data.rus$y[which(data.rus$y=="No")]),
					  length(data.ros$y[which(data.ros$y=="Yes")])/length(data.ros$y[which(data.ros$y=="No")]),
					  length(data.sm$y[which(data.sm$y=="Yes")])/length(data.sm$y[which(data.sm$y=="No")]),
					  length(data.rose$y[which(data.rose$y=="Yes")])/length(data.rose$y[which(data.rose$y=="No")])),
					  #length(data.km$y[which(data.km$y=="No")]), 
					  #length(data.km$y[which(data.km$y=="Yes")])),
			    CB = c("Complete",
					   "Train",
					   "Test",
					   "Under-Sampling",
					   "Over-Sampling",
					   "SMOTE",
					   "ROSE"))

p10kb_k562 <- ggplot(x10_k562, aes(y=IR, x=factor(CB, levels = unique(x10_k562$CB)))) + 
  geom_bar( stat="identity") +
  ylab("Imbalance Ratio")+
  xlab("")+
  #ylim(0, 30000)+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15, 
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

###25kb

```{r}
train <- readRDS("Z:/TAD_data_analysis/K562/25kb/training_testing/train.rds")
test <- readRDS("Z:/TAD_data_analysis/K562/25kb/training_testing/test.rds")

set.seed(123)
sampids <- sample(which(train$y=="No"),
                        length(which(train$y=="Yes")),
                        replace = FALSE)
data.rus <- rbind.data.frame(train[which(train$y=="Yes"),],
                           train[sampids,])

set.seed(123)		
sampids <- sample(which(train$y=="Yes"),
                        length(which(train$y=="No")),
                        replace = TRUE)
data.ros <- rbind.data.frame(train[which(train$y=="No"),],
                           train[sampids,])

set.seed(123)
data.sm <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

names(train)[-1] <- gsub("-","_",names(train)[-1])
data.rose <- ROSE(y ~ ., data=train,seed=1,p=0.5)$data

#set.seed(123)
#n <- 1000
#nr <- nrow(train)
#splitdat <- split(train, rep(1:ceiling(nr/n), each=n, length.out=nr))
#meds <- max(unlist(lapply(splitdat, function(x){length(which(x$y=="Yes"))})))
#medoids <- list(as.numeric(),length(splitdat))
#for(i in 1:length(splitdat)){
#  print(i)
#  data <- splitdat[[i]]
#  majority <- data[which(data$y=="No"),-1]
#  minnum <- length(which(data$y=="Yes"))
# 
#  pam.med <- pam(majority, k = minnum, diss = FALSE,
#                 metric = "manhattan", 
#                 #medoids = NULL, 
#                 #stand = FALSE, 
#                 cluster.only = FALSE,
#                 do.swap = TRUE,
#                 keep.diss = FALSE,
#                 keep.data = FALSE,
#                 pamonce = FALSE, trace.lev = 2)
#  
#  medoids[[i]] <- as.numeric(rownames(pam.med$medoids))
#  
#}
#medoids <- unlist(medoids)
#data.km <- rbind.data.frame(train[which(as.numeric(rownames(train)) %in% medoids),],
#                               train[which(train$y=="Yes"),])

x25_k562 <- data.frame(IR = c((length(train$y[which(train$y=="Yes")])+length(test$y[which(test$y=="Yes")]))/(length(train$y[which(train$y=="No")])+length(test$y[which(test$y=="No")])), 
					  length(train$y[which(train$y=="Yes")])/length(train$y[which(train$y=="No")]),
					  length(test$y[which(test$y=="Yes")])/length(test$y[which(test$y=="No")]),
					  length(data.rus$y[which(data.rus$y=="Yes")])/length(data.rus$y[which(data.rus$y=="No")]),
					  length(data.ros$y[which(data.ros$y=="Yes")])/length(data.ros$y[which(data.ros$y=="No")]),
					  length(data.sm$y[which(data.sm$y=="Yes")])/length(data.sm$y[which(data.sm$y=="No")]),
					  length(data.rose$y[which(data.rose$y=="Yes")])/length(data.rose$y[which(data.rose$y=="No")])),
					  #length(data.km$y[which(data.km$y=="No")]), 
					  #length(data.km$y[which(data.km$y=="Yes")])),
			    CB = c("Complete",
					   "Train",
					   "Test",
					   "Under-Sampling",
					   "Over-Sampling",
					   "SMOTE",
					   "ROSE"))

p25kb_k562 <- ggplot(x25_k562, aes(y=IR, x=factor(CB, levels = unique(x25_k562$CB)))) + 
  geom_bar( stat="identity") +
  ylab("Imbalance Ratio")+
  xlab("")+
  #ylim(0, 30000)+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15, 
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

###50kb

```{r}
train <- readRDS("Z:/TAD_data_analysis/K562/50kb/training_testing/train.rds")
test <- readRDS("Z:/TAD_data_analysis/K562/50kb/training_testing/test.rds")

set.seed(123)
sampids <- sample(which(train$y=="No"),
                        length(which(train$y=="Yes")),
                        replace = FALSE)
data.rus <- rbind.data.frame(train[which(train$y=="Yes"),],
                           train[sampids,])

set.seed(123)		
sampids <- sample(which(train$y=="Yes"),
                        length(which(train$y=="No")),
                        replace = TRUE)
data.ros <- rbind.data.frame(train[which(train$y=="No"),],
                           train[sampids,])

set.seed(123)
data.sm <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

names(train)[-1] <- gsub("-","_",names(train)[-1])
data.rose <- ROSE(y ~ ., data=train,seed=1,p=0.5)$data

#set.seed(123)
#n <- 1000
#nr <- nrow(train)
#splitdat <- split(train, rep(1:ceiling(nr/n), each=n, length.out=nr))
#meds <- max(unlist(lapply(splitdat, function(x){length(which(x$y=="Yes"))})))
#medoids <- list(as.numeric(),length(splitdat))
#for(i in 1:length(splitdat)){
#  print(i)
#  data <- splitdat[[i]]
#  majority <- data[which(data$y=="No"),-1]
#  minnum <- length(which(data$y=="Yes"))
# 
#  pam.med <- pam(majority, k = minnum, diss = FALSE,
#                 metric = "manhattan", 
#                 #medoids = NULL, 
#                 #stand = FALSE, 
#                 cluster.only = FALSE,
#                 do.swap = TRUE,
#                 keep.diss = FALSE,
#                 keep.data = FALSE,
#                 pamonce = FALSE, trace.lev = 2)
#  
#  medoids[[i]] <- as.numeric(rownames(pam.med$medoids))
#  
#}
#medoids <- unlist(medoids)
#data.km <- rbind.data.frame(train[which(as.numeric(rownames(train)) %in% medoids),],
#                               train[which(train$y=="Yes"),])

x50_k562 <- data.frame(IR = c((length(train$y[which(train$y=="Yes")])+length(test$y[which(test$y=="Yes")]))/(length(train$y[which(train$y=="No")])+length(test$y[which(test$y=="No")])), 
					  length(train$y[which(train$y=="Yes")])/length(train$y[which(train$y=="No")]),
					  length(test$y[which(test$y=="Yes")])/length(test$y[which(test$y=="No")]),
					  length(data.rus$y[which(data.rus$y=="Yes")])/length(data.rus$y[which(data.rus$y=="No")]),
					  length(data.ros$y[which(data.ros$y=="Yes")])/length(data.ros$y[which(data.ros$y=="No")]),
					  length(data.sm$y[which(data.sm$y=="Yes")])/length(data.sm$y[which(data.sm$y=="No")]),
					  length(data.rose$y[which(data.rose$y=="Yes")])/length(data.rose$y[which(data.rose$y=="No")])),
					  #length(data.km$y[which(data.km$y=="No")]), 
					  #length(data.km$y[which(data.km$y=="Yes")])),
			    CB = c("Complete",
					   "Train",
					   "Test",
					   "Under-Sampling",
					   "Over-Sampling",
					   "SMOTE",
					   "ROSE"))

p50kb_k562 <- ggplot(x50_k562, aes(y=IR, x=factor(CB, levels = unique(x50_k562$CB)))) + 
  geom_bar( stat="identity") +
  ylab("Imbalance Ratio")+
  xlab("")+
  #ylim(0, 30000)+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15, 
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

###100kb

```{r}
train <- readRDS("Z:/TAD_data_analysis/K562/100kb/training_testing/train.rds")
test <- readRDS("Z:/TAD_data_analysis/K562/100kb/training_testing/test.rds")

set.seed(123)
sampids <- sample(which(train$y=="No"),
                        length(which(train$y=="Yes")),
                        replace = FALSE)
data.rus <- rbind.data.frame(train[which(train$y=="Yes"),],
                           train[sampids,])

set.seed(123)		
sampids <- sample(which(train$y=="Yes"),
                        length(which(train$y=="No")),
                        replace = TRUE)
data.ros <- rbind.data.frame(train[which(train$y=="No"),],
                           train[sampids,])

set.seed(123)
data.sm <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

names(train)[-1] <- gsub("-","_",names(train)[-1])
data.rose <- ROSE(y ~ ., data=train,seed=1,p=0.5)$data

#set.seed(123)
#n <- 1000
#nr <- nrow(train)
#splitdat <- split(train, rep(1:ceiling(nr/n), each=n, length.out=nr))
#meds <- max(unlist(lapply(splitdat, function(x){length(which(x$y=="Yes"))})))
#medoids <- list(as.numeric(),length(splitdat))
#for(i in 1:length(splitdat)){
#  print(i)
#  data <- splitdat[[i]]
#  majority <- data[which(data$y=="No"),-1]
#  minnum <- length(which(data$y=="Yes"))
# 
#  pam.med <- pam(majority, k = minnum, diss = FALSE,
#                 metric = "manhattan", 
#                 #medoids = NULL, 
#                 #stand = FALSE, 
#                 cluster.only = FALSE,
#                 do.swap = TRUE,
#                 keep.diss = FALSE,
#                 keep.data = FALSE,
#                 pamonce = FALSE, trace.lev = 2)
#  
#  medoids[[i]] <- as.numeric(rownames(pam.med$medoids))
#  
#}
#medoids <- unlist(medoids)
#data.km <- rbind.data.frame(train[which(as.numeric(rownames(train)) %in% medoids),],
#                               train[which(train$y=="Yes"),])

x100_k562 <- data.frame(IR = c((length(train$y[which(train$y=="Yes")])+length(test$y[which(test$y=="Yes")]))/(length(train$y[which(train$y=="No")])+length(test$y[which(test$y=="No")])), 
					  length(train$y[which(train$y=="Yes")])/length(train$y[which(train$y=="No")]),
					  length(test$y[which(test$y=="Yes")])/length(test$y[which(test$y=="No")]),
					  length(data.rus$y[which(data.rus$y=="Yes")])/length(data.rus$y[which(data.rus$y=="No")]),
					  length(data.ros$y[which(data.ros$y=="Yes")])/length(data.ros$y[which(data.ros$y=="No")]),
					  length(data.sm$y[which(data.sm$y=="Yes")])/length(data.sm$y[which(data.sm$y=="No")]),
					  length(data.rose$y[which(data.rose$y=="Yes")])/length(data.rose$y[which(data.rose$y=="No")])),
					  #length(data.km$y[which(data.km$y=="No")]), 
					  #length(data.km$y[which(data.km$y=="Yes")])),
			    CB = c("Complete",
					   "Train",
					   "Test",
					   "Under-Sampling",
					   "Over-Sampling",
					   "SMOTE",
					   "ROSE"))

p100kb_k562 <- ggplot(x100_k562, aes(y=IR, x=factor(CB, levels = unique(x100_k562$CB)))) + 
  geom_bar( stat="identity") +
  ylab("Imbalance Ratio")+
  xlab("")+
  #ylim(0, 30000)+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15, 
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15))

```

##Combined
```{r}

IR=c(x10$IR[x10$CB=="Complete"],
                              x25$IR[x25$CB=="Complete"],
                              x50$IR[x50$CB=="Complete"],
                              x100$IR[x100$CB=="Complete"],
                              x10_k562$IR[x10_k562$CB=="Complete"], 
                              x25_k562$IR[x25_k562$CB=="Complete"],
                              x50_k562$IR[x50_k562$CB=="Complete"],
                              x100_k562$IR[x100_k562$CB=="Complete"])
CL <- c(rep("GM12878", 4), 
                  rep("K562", 4))
Resolution <- c(rep(c("10 kb", "25 kb", "50 kb", "100 kb"), 2))

allirdata <- cbind.data.frame(IR, CL, Resolution)
allirdata$CL <- factor(allirdata$CL, levels=c("GM12878", "K562"))
allirdata$Resolution <- factor(allirdata$Resolution, c("10 kb", "25 kb", "50 kb", "100 kb"))

allirplots <- ggplot(allirdata, aes(x = Resolution, y = IR,fill = CL)) + 
  geom_bar(position = "dodge", stat="identity") + 
  ylab("Imbalance Ratio") +
  xlab("Resolution") +
  #scale_fill_manual(name="Cell Line", values=c("red", "blue"))+
  labs(fill="Cell Line")+
  theme_bw() +
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

#Evaluating Feature Engineering

## PR Curves

###GM12878

####10kb

```{r}

prall <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/op/RFE/prcurve.rds")

#plot(prall, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, rand.plot = TRUE)
plot(prall$curve, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")
lines(prdist$curve, auc.main= FALSE, legend=FALSE, col='red', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")#, add=TRUE)
lines(proc$curve, auc.main= FALSE, legend=FALSE, col='blue', panel.first= grid(), cex.main = 1.5, type="l",lwd=2, add=TRUE)
lines(prop$curve, auc.main= FALSE, legend=FALSE, col='green', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, add=TRUE)

test <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/test.rds")
abline(h = sum(test$y=="Yes")/(sum(test$y=="Yes") + sum(test$y=="No")), col="black", lty=2)

#legend("topright", title="Predictor Type", legend=c("Distance (AUPRC: 0.431)","All (AUPRC: 0.358)", "Counts (AUPRC: 0.209)","Percents (AUPRC: 0.202)", "Baseline"),
#       col=c("red", "black", "blue", "green", "black"), lty=c(1,1,1,1,2), lwd=c(3,3,3,3,1), cex=0.8)

```

####25kb

```{r}

prall <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/op/RFE/prcurve.rds")

#plot(prall, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, rand.plot = TRUE)
plot(prall$curve, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")
lines(prdist$curve, auc.main= FALSE, legend=FALSE, col='red', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")#, add=TRUE)
lines(proc$curve, auc.main= FALSE, legend=FALSE, col='blue', panel.first= grid(), cex.main = 1.5, type="l",lwd=2, add=TRUE)
lines(prop$curve, auc.main= FALSE, legend=FALSE, col='green', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, add=TRUE)

test <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/test.rds")
abline(h = sum(test$y=="Yes")/(sum(test$y=="Yes") + sum(test$y=="No")), col="black", lty=2)

#legend("topright", title="Predictor Type", legend=c("Distance (AUPRC: 0.436)","All (AUPRC: 0.380)", "Percents (AUPRC: 0.239)","Counts (AUPRC: 0.231)", "Baseline"),
#       col=c("red", "black", "green", "blue", "black"), lty=c(1,1,1,1,2), lwd=c(3,3,3,3,1), cex=0.8)

```

####50kb

```{r}

prall <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/op/RFE/prcurve.rds")

#plot(prall, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, rand.plot = TRUE)
plot(prall$curve, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")
lines(prdist$curve, auc.main= FALSE, legend=FALSE, col='red', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")#, add=TRUE)
lines(proc$curve, auc.main= FALSE, legend=FALSE, col='blue', panel.first= grid(), cex.main = 1.5, type="l",lwd=2, add=TRUE)
lines(prop$curve, auc.main= FALSE, legend=FALSE, col='green', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, add=TRUE)

test <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/test.rds")
abline(h = sum(test$y=="Yes")/(sum(test$y=="Yes") + sum(test$y=="No")), col="black", lty=2)

#legend("topright", title="Predictor Type", legend=c("Distance (AUPRC: 0.405)","All (AUPRC: 0.367)", "Percents (AUPRC: 0.218)","Counts (AUPRC: 0.212)", "Baseline"),
#       col=c("red", "black", "green", "blue", "black"), lty=c(1,1,1,1,2), lwd=c(3,3,3,3,1), cex=0.8)

```

####100kb

```{r}

prall <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/op/RFE/prcurve.rds")

#plot(prall, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, rand.plot = TRUE)
plot(prall$curve, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")
lines(prdist$curve, auc.main= FALSE, legend=FALSE, col='red', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")#, add=TRUE)
lines(proc$curve, auc.main= FALSE, legend=FALSE, col='blue', panel.first= grid(), cex.main = 1.5, type="l",lwd=2, add=TRUE)
lines(prop$curve, auc.main= FALSE, legend=FALSE, col='green', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, add=TRUE)

test <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/test.rds")
abline(h = sum(test$y=="Yes")/(sum(test$y=="Yes") + sum(test$y=="No")), col="black", lty=2)

#legend("topright", title="Predictor Type", legend=c("Distance (AUPRC: 0.361)","All (AUPRC: 0.301)", "Counts (AUPRC: 0.175)", "Percents (AUPRC: 0.156)", "Baseline"),
#       col=c("red", "black", "blue", "green", "black"), lty=c(1,1,1,1,2), lwd=c(3,3,3,3,1), cex=0.8)

```

###K562

####10kb

```{r}
prall <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/op/RFE/prcurve.rds")

#plot(prall, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, rand.plot = TRUE)
plot(prall$curve, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")
lines(prdist$curve, auc.main= FALSE, legend=FALSE, col='red', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")#, add=TRUE)
lines(proc$curve, auc.main= FALSE, legend=FALSE, col='blue', panel.first= grid(), cex.main = 1.5, type="l",lwd=2, add=TRUE)
lines(prop$curve, auc.main= FALSE, legend=FALSE, col='green', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, add=TRUE)

test <- readRDS("Z:/TAD_data_analysis/K562/10kb/training_testing/test.rds")
abline(h = sum(test$y=="Yes")/(sum(test$y=="Yes") + sum(test$y=="No")), col="black", lty=2)

#legend("topright", title="Predictor Type", legend=c("Distance (AUPRC: )","All (AUPRC: )", "Percents (AUPRC: )", "Counts (AUPRC: )", "Baseline"),
#       col=c("red", "black", "green", "blue", "black"), lty=c(1,1,1,1,2), lwd=c(3,3,3,3,1), cex=0.8)

```

####25kb

```{r}
prall <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/op/RFE/prcurve.rds")

#plot(prall, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, rand.plot = TRUE)
plot(prall$curve, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")
lines(prdist$curve, auc.main= FALSE, legend=FALSE, col='red', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")#, add=TRUE)
lines(proc$curve, auc.main= FALSE, legend=FALSE, col='blue', panel.first= grid(), cex.main = 1.5, type="l",lwd=2, add=TRUE)
lines(prop$curve, auc.main= FALSE, legend=FALSE, col='green', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, add=TRUE)

test <- readRDS("Z:/TAD_data_analysis/K562/25kb/training_testing/test.rds")
abline(h = sum(test$y=="Yes")/(sum(test$y=="Yes") + sum(test$y=="No")), col="black", lty=2)

#legend("topright", title="Predictor Type", legend=c("Distance (AUPRC: )","All (AUPRC: )","Percents (AUPRC: )", "Counts (AUPRC: )", "Baseline"),
#       col=c("red", "black", "green", "blue", "black"), lty=c(1,1,1,1,2), lwd=c(3,3,3,3,1), cex=0.8)

```

####50kb

```{r}
prall <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/op/RFE/prcurve.rds")

#plot(prall, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, rand.plot = TRUE)
plot(prall$curve, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")
lines(prdist$curve, auc.main= FALSE, legend=FALSE, col='red', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")#, add=TRUE)
lines(proc$curve, auc.main= FALSE, legend=FALSE, col='blue', panel.first= grid(), cex.main = 1.5, type="l",lwd=2, add=TRUE)
lines(prop$curve, auc.main= FALSE, legend=FALSE, col='green', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, add=TRUE)

test <- readRDS("Z:/TAD_data_analysis/K562/50kb/training_testing/test.rds")
abline(h = sum(test$y=="Yes")/(sum(test$y=="Yes") + sum(test$y=="No")), col="black", lty=2)

#legend("topright", title="Predictor Type", legend=c("Distance (AUPRC: )","All (AUPRC: )","Percents (AUPRC: )", "Counts (AUPRC: )", "Baseline"),
#       col=c("red", "black", "green", "blue", "black"), lty=c(1,1,1,1,2), lwd=c(3,3,3,3,1), cex=0.8)

```

####100kb

```{r}
prall <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/op/RFE/prcurve.rds")

#plot(prall, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, rand.plot = TRUE)
plot(prall$curve, auc.main= FALSE, legend=FALSE, col='black', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")
lines(prdist$curve, auc.main= FALSE, legend=FALSE, col='red', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, xlab="Recall", ylab="Precision")#, add=TRUE)
lines(proc$curve, auc.main= FALSE, legend=FALSE, col='blue', panel.first= grid(), cex.main = 1.5, type="l",lwd=2, add=TRUE)
lines(prop$curve, auc.main= FALSE, legend=FALSE, col='green', panel.first= grid(), cex.main = 1.5, type="l", ylim=c(0,1), lwd=2, add=TRUE)

test <- readRDS("Z:/TAD_data_analysis/K562/100kb/training_testing/test.rds")
abline(h = sum(test$y=="Yes")/(sum(test$y=="Yes") + sum(test$y=="No")), col="black", lty=2)

#legend("topright", title="Predictor Type", legend=c("Distance (AUPRC: )","All (AUPRC: )", "Percents (AUPRC: )", "Counts (AUPRC: )", "Baseline"),
#       col=c("red", "black", "green", "blue", "black"), lty=c(1,1,1,1,2), lwd=c(3,3,3,3,1), cex=0.8)

```

###Combined

```{r}
#10kb
prall <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/op/RFE/prcurve.rds")

prall_k562 <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/all/RFE/prcurve.rds")
prdist_k562 <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc_k562 <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/oc/RFE/prcurve.rds")
prop_k562 <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/op/RFE/prcurve.rds")

prdata_10kb <- rbind.data.frame(prall$curve,
                                prdist$curve,
                                prop$curve,
                                proc$curve,
                                prall_k562$curve,
                                prdist_k562$curve,
                                prop_k562$curve,
                                proc_k562$curve)
names(prdata_10kb) <- c("Recall", "Precision", "Threshold")
prdata_10kb$Predictor <- c(rep("All", dim(prall$curve)[1]),
                           rep("Distance", dim(prdist$curve)[1]),
                           rep("OP", dim(prop$curve)[1]),
                           rep("OC", dim(proc$curve)[1]),
                           rep("All", dim(prall_k562$curve)[1]),
                           rep("Distance", dim(prdist_k562$curve)[1]),
                           rep("OP", dim(prop_k562$curve)[1]),
                           rep("OC", dim(proc_k562$curve)[1]))
prdata_10kb$CL <- c(rep("GM12878", (dim(prall$curve)[1]+
                                      dim(prdist$curve)[1]+
                                      dim(prop$curve)[1]+
                                      dim(proc$curve)[1])),
                    rep("K562", (dim(prall_k562$curve)[1]+
                                      dim(prdist_k562$curve)[1]+
                                      dim(prop_k562$curve)[1]+
                                      dim(proc_k562$curve)[1])))
prdata_10kb$Predictor <- factor(prdata_10kb$Predictor)
prdata_10kb$CL <- factor(prdata_10kb$CL)

prplot_10kb <- ggplot(prdata_10kb, aes(x=Recall, y=Precision, color=Predictor, linetype=CL)) +
  geom_line(size=.75)+
  scale_linetype_manual(name=("Cell Line"), values=c("solid", "dashed"))+
  scale_color_discrete(name = "Predictor Type") +
  theme_minimal()+
  theme_bw()

#25kb
prall <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/op/RFE/prcurve.rds")

prall_k562 <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/all/RFE/prcurve.rds")
prdist_k562 <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc_k562 <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/oc/RFE/prcurve.rds")
prop_k562 <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/op/RFE/prcurve.rds")

prdata_25kb <- rbind.data.frame(prall$curve,
                                prdist$curve,
                                prop$curve,
                                proc$curve,
                                prall_k562$curve,
                                prdist_k562$curve,
                                prop_k562$curve,
                                proc_k562$curve)
names(prdata_25kb) <- c("Recall", "Precision", "Threshold")
prdata_25kb$Predictor <- c(rep("All", dim(prall$curve)[1]),
                           rep("Distance", dim(prdist$curve)[1]),
                           rep("OP", dim(prop$curve)[1]),
                           rep("OC", dim(proc$curve)[1]),
                           rep("All", dim(prall_k562$curve)[1]),
                           rep("Distance", dim(prdist_k562$curve)[1]),
                           rep("OP", dim(prop_k562$curve)[1]),
                           rep("OC", dim(proc_k562$curve)[1]))
prdata_25kb$CL <- c(rep("GM12878", (dim(prall$curve)[1]+
                                      dim(prdist$curve)[1]+
                                      dim(prop$curve)[1]+
                                      dim(proc$curve)[1])),
                    rep("K562", (dim(prall_k562$curve)[1]+
                                      dim(prdist_k562$curve)[1]+
                                      dim(prop_k562$curve)[1]+
                                      dim(proc_k562$curve)[1])))
prdata_25kb$Predictor <- factor(prdata_25kb$Predictor)
prdata_25kb$CL <- factor(prdata_25kb$CL)

prplot_25kb <- ggplot(prdata_25kb, aes(x=Recall, y=Precision, color=Predictor, linetype=CL)) +
  geom_line(size=.75)+
  scale_linetype_manual(name=("Cell Line"), values=c("solid", "dashed"))+
  scale_color_discrete(name = "Predictor Type") +
  theme_minimal()+
  theme_bw()

#50kb
prall <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/op/RFE/prcurve.rds")

prall_k562 <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/all/RFE/prcurve.rds")
prdist_k562 <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc_k562 <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/oc/RFE/prcurve.rds")
prop_k562 <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/op/RFE/prcurve.rds")

prdata_50kb <- rbind.data.frame(prall$curve,
                                prdist$curve,
                                prop$curve,
                                proc$curve,
                                prall_k562$curve,
                                prdist_k562$curve,
                                prop_k562$curve,
                                proc_k562$curve)
names(prdata_50kb) <- c("Recall", "Precision", "Threshold")
prdata_50kb$Predictor <- c(rep("All", dim(prall$curve)[1]),
                           rep("Distance", dim(prdist$curve)[1]),
                           rep("OP", dim(prop$curve)[1]),
                           rep("OC", dim(proc$curve)[1]),
                           rep("All", dim(prall_k562$curve)[1]),
                           rep("Distance", dim(prdist_k562$curve)[1]),
                           rep("OP", dim(prop_k562$curve)[1]),
                           rep("OC", dim(proc_k562$curve)[1]))
prdata_50kb$CL <- c(rep("GM12878", (dim(prall$curve)[1]+
                                      dim(prdist$curve)[1]+
                                      dim(prop$curve)[1]+
                                      dim(proc$curve)[1])),
                    rep("K562", (dim(prall_k562$curve)[1]+
                                      dim(prdist_k562$curve)[1]+
                                      dim(prop_k562$curve)[1]+
                                      dim(proc_k562$curve)[1])))
prdata_50kb$Predictor <- factor(prdata_50kb$Predictor)
prdata_50kb$CL <- factor(prdata_50kb$CL)

prplot_50kb <- ggplot(prdata_50kb, aes(x=Recall, y=Precision, color=Predictor, linetype=CL)) +
  geom_line(size=.75)+
  scale_linetype_manual(name=("Cell Line"), values=c("solid", "dashed"))+
  scale_color_discrete(name = "Predictor Type") +
  theme_minimal()+
  theme_bw()

#100kb
prall <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/all/RFE/prcurve.rds")
prdist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/oc/RFE/prcurve.rds")
prop <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/op/RFE/prcurve.rds")

prall_k562 <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/all/RFE/prcurve.rds")
prdist_k562 <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/RFE/prcurve.rds")
proc_k562 <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/oc/RFE/prcurve.rds")
prop_k562 <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/op/RFE/prcurve.rds")

prdata_100kb <- rbind.data.frame(prall$curve,
                                prdist$curve,
                                prop$curve,
                                proc$curve,
                                prall_k562$curve,
                                prdist_k562$curve,
                                prop_k562$curve,
                                proc_k562$curve)
names(prdata_100kb) <- c("Recall", "Precision", "Threshold")
prdata_100kb$Predictor <- c(rep("All", dim(prall$curve)[1]),
                           rep("Distance", dim(prdist$curve)[1]),
                           rep("OP", dim(prop$curve)[1]),
                           rep("OC", dim(proc$curve)[1]),
                           rep("All", dim(prall_k562$curve)[1]),
                           rep("Distance", dim(prdist_k562$curve)[1]),
                           rep("OP", dim(prop_k562$curve)[1]),
                           rep("OC", dim(proc_k562$curve)[1]))
prdata_100kb$CL <- c(rep("GM12878", (dim(prall$curve)[1]+
                                      dim(prdist$curve)[1]+
                                      dim(prop$curve)[1]+
                                      dim(proc$curve)[1])),
                    rep("K562", (dim(prall_k562$curve)[1]+
                                      dim(prdist_k562$curve)[1]+
                                      dim(prop_k562$curve)[1]+
                                      dim(proc_k562$curve)[1])))
prdata_100kb$Predictor <- factor(prdata_100kb$Predictor)
prdata_100kb$CL <- factor(prdata_100kb$CL)

prplot_100kb <- ggplot(prdata_100kb, aes(x=Recall, y=Precision, color=Predictor, linetype=CL)) +
  geom_line(size=.75)+
  scale_linetype_manual(name=("Cell Line"), values=c("solid", "dashed"))+
  scale_color_discrete(name = "Predictor Type") +
  theme_minimal()+
  theme_bw()

ggarrange(prplot_25kb, prplot_50kb, prplot_100kb,
          nrow=2,ncol=2,
          common.legend = TRUE, legend = "right")
```

## RFE

###GM12878

####10kb

```{r}
#rfeModel_all <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/all/RFE/rfeModel_all.rds")
rfeModel_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
#rfeModel_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/RFE/rfeModel_oc.rds")
#rfeModel_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/RFE/rfeModel_op.rds")

#resampdata <- rbind.data.frame(rfeModel_all$results,rfeModel_dist$results,rfeModel_oc$results,rfeModel_op$results)

#resampdata$`Predictor Type` <- c(rep("Combined", nrow(rfeModel_all$results)),
#                                 rep("Distance", nrow(rfeModel_dist$results)),
#                                 rep("OC", nrow(rfeModel_oc$results)),
#                                 rep("OP", nrow(rfeModel_op$results)))

#resampdata$`Predictor Type` <- factor(resampdata$`Predictor Type`)

resampdata_10kb <- rfeModel_dist$results[1:6,]

pt10kb <- ggplot(resampdata_10kb, aes(x=Variables, y=AUPRC)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  theme_minimal()+
  ylim(0,.5)

```

####25kb

```{r}
#rfeModel_all <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/all/RFE/rfeModel_all.rds")
rfeModel_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
#rfeModel_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OC/RFE/rfeModel_oc.rds")
#rfeModel_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OP/RFE/rfeModel_op.rds")

#resampdata <- rbind.data.frame(rfeModel_all$results,rfeModel_dist$results,rfeModel_oc$results,rfeModel_op$results)

#resampdata$`Predictor Type` <- c(rep("Combined", nrow(rfeModel_all$results)),
#                                 rep("Distance", nrow(rfeModel_dist$results)),
#                                 rep("OC", nrow(rfeModel_oc$results)),
#                                 rep("OP", nrow(rfeModel_op$results)))

#resampdata$`Predictor Type` <- factor(resampdata$`Predictor Type`)

resampdata_25kb <- rfeModel_dist$results[1:6,]

pt25kb <- ggplot(resampdata_25kb, aes(x=Variables, y=AUPRC)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  theme_minimal()+
  ylim(0,.5)

```

####50kb

```{r}
#rfeModel_all <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/all/RFE/rfeModel_all.rds")
rfeModel_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
#rfeModel_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OC/RFE/rfeModel_oc.rds")
#rfeModel_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OP/RFE/rfeModel_op.rds")

#resampdata <- rbind.data.frame(rfeModel_all$results,rfeModel_dist$results,rfeModel_oc$results,rfeModel_op$results)

#resampdata$`Predictor Type` <- c(rep("Combined", nrow(rfeModel_all$results)),
#                                 rep("Distance", nrow(rfeModel_dist$results)),
#                                 rep("OC", nrow(rfeModel_oc$results)),
#                                 rep("OP", nrow(rfeModel_op$results)))

#resampdata$`Predictor Type` <- factor(resampdata$`Predictor Type`)

resampdata_50kb <- rfeModel_dist$results[1:6,]

pt50kb <- ggplot(resampdata_50kb, aes(x=Variables, y=AUPRC)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  theme_minimal()+
  ylim(0,.5)

```

####100kb

```{r}
#rfeModel_all <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/all/RFE/rfeModel_all.rds")
rfeModel_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
#rfeModel_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OC/RFE/rfeModel_oc.rds")
#rfeModel_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OP/RFE/rfeModel_op.rds")

#resampdata <- rbind.data.frame(rfeModel_all$results,rfeModel_dist$results,rfeModel_oc$results,rfeModel_op$results)

#resampdata$`Predictor Type` <- c(rep("Combined", nrow(rfeModel_all$results)),
#                                 rep("Distance", nrow(rfeModel_dist$results)),
#                                 rep("OC", nrow(rfeModel_oc$results)),
#                                 rep("OP", nrow(rfeModel_op$results)))

#resampdata$`Predictor Type` <- factor(resampdata$`Predictor Type`)

resampdata_100kb <- rfeModel_dist$results[1:6,]

pt100kb <- ggplot(resampdata_100kb, aes(x=Variables, y=AUPRC)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  theme_minimal()+
  ylim(0,.5)

```

###K562

####10kb

```{r}
#rfeModel_all <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/all/RFE/rfeModel_all.rds")
rfeModel_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
#rfeModel_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OC/RFE/rfeModel_oc.rds")
#rfeModel_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OP/RFE/rfeModel_op.rds")

#resampdata <- rbind.data.frame(rfeModel_all$results,rfeModel_dist$results,rfeModel_oc$results,rfeModel_op$results)

#resampdata$`Predictor Type` <- c(rep("Combined", nrow(rfeModel_all$results)),
#                                 rep("Distance", nrow(rfeModel_dist$results)),
#                                 rep("OC", nrow(rfeModel_oc$results)),
#                                 rep("OP", nrow(rfeModel_op$results)))

#resampdata$`Predictor Type` <- factor(resampdata$`Predictor Type`)

resampdata_10kb_k562 <- rfeModel_dist$results[1:6,]

pt10kb_k562 <- ggplot(resampdata_10kb_k562, aes(x=Variables, y=AUPRC)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  theme_minimal()+
  ylim(0,.5)

```

####25kb

```{r}
#rfeModel_all <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/all/RFE/rfeModel_all.rds")
rfeModel_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
#rfeModel_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OC/RFE/rfeModel_oc.rds")
#rfeModel_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OP/RFE/rfeModel_op.rds")

#resampdata <- rbind.data.frame(rfeModel_all$results,rfeModel_dist$results,rfeModel_oc$results,rfeModel_op$results)

#resampdata$`Predictor Type` <- c(rep("Combined", nrow(rfeModel_all$results)),
#                                 rep("Distance", nrow(rfeModel_dist$results)),
#                                 rep("OC", nrow(rfeModel_oc$results)),
#                                 rep("OP", nrow(rfeModel_op$results)))

#resampdata$`Predictor Type` <- factor(resampdata$`Predictor Type`)

resampdata_25kb_k562 <- rfeModel_dist$results[1:6,]

pt25kb <- ggplot(resampdata_25kb_k562, aes(x=Variables, y=AUPRC)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  theme_minimal()+
  ylim(0,.5)

```

####50kb

```{r}
#rfeModel_all <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/all/RFE/rfeModel_all.rds")
rfeModel_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
#rfeModel_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OC/RFE/rfeModel_oc.rds")
#rfeModel_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OP/RFE/rfeModel_op.rds")

#resampdata <- rbind.data.frame(rfeModel_all$results,rfeModel_dist$results,rfeModel_oc$results,rfeModel_op$results)

#resampdata$`Predictor Type` <- c(rep("Combined", nrow(rfeModel_all$results)),
#                                 rep("Distance", nrow(rfeModel_dist$results)),
#                                 rep("OC", nrow(rfeModel_oc$results)),
#                                 rep("OP", nrow(rfeModel_op$results)))

#resampdata$`Predictor Type` <- factor(resampdata$`Predictor Type`)

resampdata_50kb_k562 <- rfeModel_dist$results[1:6,]

pt50kb <- ggplot(resampdata_50kb_k562, aes(x=Variables, y=AUPRC)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  theme_minimal()+
  ylim(0,.5)

```

####100kb

```{r}
#rfeModel_all <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/all/RFE/rfeModel_all.rds")
rfeModel_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
#rfeModel_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OC/RFE/rfeModel_oc.rds")
#rfeModel_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OP/RFE/rfeModel_op.rds")

#resampdata <- rbind.data.frame(rfeModel_all$results,rfeModel_dist$results,rfeModel_oc$results,rfeModel_op$results)

#resampdata$`Predictor Type` <- c(rep("Combined", nrow(rfeModel_all$results)),
#                                 rep("Distance", nrow(rfeModel_dist$results)),
#                                 rep("OC", nrow(rfeModel_oc$results)),
#                                 rep("OP", nrow(rfeModel_op$results)))

#resampdata$`Predictor Type` <- factor(resampdata$`Predictor Type`)

resampdata_100kb_k562 <- rfeModel_dist$results[1:6,]

pt100kb <- ggplot(resampdata_100kb_k562, aes(x=Variables, y=AUPRC)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  theme_minimal()+
  ylim(0,.5)

```

### Combined

```{r}

allrfe10kb <- rbind.data.frame(resampdata_10kb, resampdata_10kb_k562)
allrfe10kb$CL <- c(rep("GM12878", 6),rep("K562", 6))
allrfe10kb$CL <- factor(allrfe10kb$CL)

rfe10kbplot <- ggplot(data=allrfe10kb, aes(x=Variables, y=AUPRC, color=CL)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  scale_color_discrete(name="Cell Line")+
  theme_bw()+
  #ggtitle("10 kb") +
  ylim(0,.5)+
  theme(plot.title = element_text(hjust = 0.5))

allrfe25kb <- rbind.data.frame(resampdata_25kb, resampdata_25kb_k562)
allrfe25kb$CL <- c(rep("GM12878", 6),rep("K562", 6))
allrfe25kb$CL <- factor(allrfe25kb$CL)

rfe25kbplot <- ggplot(data=allrfe25kb, aes(x=Variables, y=AUPRC, color=CL)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  scale_color_discrete(name="Cell Line")+
  theme_bw()+
  #ggtitle("25 kb") +
  ylim(0,.5)+
  theme(plot.title = element_text(hjust = 0.5))

allrfe50kb <- rbind.data.frame(resampdata_50kb, resampdata_50kb_k562)
allrfe50kb$CL <- c(rep("GM12878", 6),rep("K562", 6))
allrfe50kb$CL <- factor(allrfe50kb$CL)

rfe50kbplot <- ggplot(data=allrfe50kb, aes(x=Variables, y=AUPRC, color=CL)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  scale_color_discrete(name="Cell Line")+
  theme_bw()+
  #ggtitle("50 kb") +
  ylim(0,.5)+
  theme(plot.title = element_text(hjust = 0.5))

allrfe100kb <- rbind.data.frame(resampdata_100kb, resampdata_100kb_k562)
allrfe100kb$CL <- c(rep("GM12878", 6),rep("K562", 6))
allrfe100kb$CL <- factor(allrfe100kb$CL)

rfe100kbplot <- ggplot(data=allrfe100kb, aes(x=Variables, y=AUPRC, color=CL)) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=AUPRC-AUPRCSD, ymax=AUPRC+AUPRCSD), width=1)+
  scale_x_discrete(name="Number of Features", 
                   breaks=c("2", "4", "8", "16", "32", "64"), 
                   labels=c("2", "4", "8", "16", "32", "64"), 
                   limits=as.character(c(1:64)))+
  scale_color_discrete(name="Cell Line")+
  theme_minimal()+
  theme_bw()+
  #ggtitle("100 kb") +
  ylim(0,.5)+
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(rfe25kbplot, rfe50kbplot, rfe100kbplot,
          nrow=2,ncol=2,
          common.legend = TRUE, legend = "right")

#comparing top predictors across resolution and cell line
rfeModel_dist_10kb <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
rfe10kbvars <- rfeModel_dist_10kb$optVariables[1:32]

rfeModel_dist_25kb <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
rfe25kbvars <- rfeModel_dist_25kb$optVariables[1:32]

rfeModel_dist_50kb <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
rfe50kbvars <- rfeModel_dist_50kb$optVariables[1:32]

rfeModel_dist_100kb <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
rfe100kbvars <- rfeModel_dist_100kb$optVariables[1:32]

rfeModel_dist_10kb_k562 <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
rfe10kbvars_k562 <- rfeModel_dist_10kb_k562$optVariables[1:32]

rfeModel_dist_25kb_k562 <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
rfe25kbvars_k562 <- rfeModel_dist_25kb_k562$optVariables[1:32]

rfeModel_dist_50kb_k562 <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
rfe50kbvars_k562 <- rfeModel_dist_50kb_k562$optVariables[1:32]

rfeModel_dist_100kb_k562 <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/RFE/rfeModel_dist.rds")
rfe100kbvars_k562 <- rfeModel_dist_100kb_k562$variables$var[rfeModel_dist_100kb_k562$variables$Variables==32 & rfeModel_dist_100kb_k562$variables$Resample=="Fold1.Rep1"]

rfe10kbvars <- gsub(paste(c("Gm12878-", 
                            "_dist", 
                            "-Sydh", 
                            "-Haib", 
                            "-Broad",
                            "-Uchicago",
                            "-Uta",
                            "-Uw"), collapse = "|"),"", rfe10kbvars)
rfe25kbvars <- gsub(paste(c("Gm12878-", 
                            "_dist", 
                            "-Sydh", 
                            "-Haib", 
                            "-Broad",
                            "-Uchicago",
                            "-Uta",
                            "-Uw"), collapse = "|"),"", rfe25kbvars)
rfe50kbvars <- gsub(paste(c("Gm12878-", 
                            "_dist", 
                            "-Sydh", 
                            "-Haib", 
                            "-Broad",
                            "-Uchicago",
                            "-Uta",
                            "-Uw"), collapse = "|"),"", rfe50kbvars)
rfe100kbvars <- gsub(paste(c("Gm12878-", 
                            "_dist", 
                            "-Sydh", 
                            "-Haib", 
                            "-Broad",
                            "-Uchicago",
                            "-Uta",
                            "-Uw"), collapse = "|"),"", rfe100kbvars)

rfe10kbvars_k562 <- gsub(paste(c("K562-", 
                            "_dist", 
                            "-Sydh", 
                            "-Haib", 
                            "-Broad",
                            "-Uchicago",
                            "-Uta",
                            "-Uw"), collapse = "|"),"", rfe10kbvars_k562)
rfe25kbvars_k562 <- gsub(paste(c("K562-", 
                            "_dist", 
                            "-Sydh", 
                            "-Haib", 
                            "-Broad",
                            "-Uchicago",
                            "-Uta",
                            "-Uw"), collapse = "|"),"", rfe25kbvars_k562)
rfe50kbvars_k562 <- gsub(paste(c("K562-", 
                            "_dist", 
                            "-Sydh", 
                            "-Haib", 
                            "-Broad",
                            "-Uchicago",
                            "-Uta",
                            "-Uw"), collapse = "|"),"", rfe50kbvars_k562)
rfe100kbvars_k562 <- gsub(paste(c("K562-", 
                            "_dist", 
                            "-Sydh", 
                            "-Haib", 
                            "-Broad",
                            "-Uchicago",
                            "-Uta",
                            "-Uw"), collapse = "|"),"", rfe100kbvars_k562)

alltopvars <- data.frame(rfe10kbvars,
                         rfe10kbvars_k562,
                         rfe25kbvars,
                         rfe25kbvars_k562,
                         rfe50kbvars,
                         rfe50kbvars_k562,
                         rfe100kbvars,
                         rfe100kbvars_k562)

rfe10kbvars_o <- sort(rfe10kbvars)
rfe25kbvars_o <- sort(rfe25kbvars)
rfe50kbvars_o <- sort(rfe50kbvars)
rfe100kbvars_o <- sort(rfe100kbvars)
rfe10kbvars_k562_o <- sort(rfe10kbvars_k562)
rfe25kbvars_k562_o <- sort(rfe25kbvars_k562)
rfe50kbvars_k562_o <- sort(rfe50kbvars_k562)
rfe100kbvars_k562_o <- sort(rfe100kbvars_k562)
alltopvars_o <- data.frame(rfe10kbvars_o,rfe25kbvars_o,rfe50kbvars_o,rfe100kbvars_o,
                         rfe10kbvars_k562_o,rfe25kbvars_k562_o,rfe50kbvars_k562_o,rfe100kbvars_k562_o)

grid.newpage()
venn.plot <- draw.quad.venn(
	area1 = length(rfe10kbvars), #The size of the first set
	area2 = length(rfe25kbvars), #The size of the second set
	area3 = length(rfe50kbvars), #The size of the third set
	area4 = length(rfe100kbvars), #The size of the fourth set
	n12 = length(intersect(rfe10kbvars,rfe25kbvars)), #The size of the intersection between the first and the second set
	n13 = length(intersect(rfe10kbvars,rfe50kbvars)), #The size of the intersection between the first and the third set
	n14 = length(intersect(rfe10kbvars,rfe100kbvars)), #The size of the intersection between the first and the fourth set
	n23 = length(intersect(rfe25kbvars,rfe50kbvars)), #The size of the intersection between the second and the third set
	n24 = length(intersect(rfe25kbvars,rfe100kbvars)), #The size of the intersection between the second and the fourth set
	n34 = length(intersect(rfe50kbvars,rfe100kbvars)), #The size of the intersection between the third and the fourth set
	n123 = length(Reduce(intersect, list(rfe10kbvars,rfe25kbvars,rfe50kbvars))), #The size of the intersection between the first, second and third sets
	n124 = length(Reduce(intersect, list(rfe10kbvars,rfe25kbvars,rfe100kbvars))), #The size of the intersection between the first, second and fourth sets
	n134 = length(Reduce(intersect, list(rfe10kbvars,rfe50kbvars,rfe100kbvars))), #The size of the intersection between the first, third and fourth sets
	n234 = length(Reduce(intersect, list(rfe25kbvars,rfe50kbvars,rfe100kbvars))), #The size of the intersection between the second, third and fourth sets
	n1234 = length(Reduce(intersect, list(rfe10kbvars,rfe25kbvars,rfe50kbvars,rfe100kbvars))), #The size of the intersection between all four sets
	category = c("10 kb", "25 kb", "50 kb", "100 kb"),
	fill = c("orange", "red", "green", "blue"),
	lty = "dashed",
	cex = 2,
	cat.cex = 2,
	cat.col = c("orange", "red", "green", "blue")
	);

grid.newpage()
venn.plot.k562 <- draw.quad.venn(
	area1 = length(rfe10kbvars_k562), #The size of the first set
	area2 = length(rfe25kbvars_k562), #The size of the second set
	area3 = length(rfe50kbvars_k562), #The size of the third set
	area4 = length(rfe100kbvars_k562), #The size of the fourth set
	n12 = length(intersect(rfe10kbvars_k562,rfe25kbvars_k562)), #The size of the intersection between the first and the second set
	n13 = length(intersect(rfe10kbvars_k562,rfe50kbvars_k562)), #The size of the intersection between the first and the third set
	n14 = length(intersect(rfe10kbvars_k562,rfe100kbvars_k562)), #The size of the intersection between the first and the fourth set
	n23 = length(intersect(rfe25kbvars_k562,rfe50kbvars_k562)), #The size of the intersection between the second and the third set
	n24 = length(intersect(rfe25kbvars_k562,rfe100kbvars_k562)), #The size of the intersection between the second and the fourth set
	n34 = length(intersect(rfe50kbvars_k562,rfe100kbvars_k562)), #The size of the intersection between the third and the fourth set
	n123 = length(Reduce(intersect, list(rfe10kbvars_k562,rfe25kbvars_k562,rfe50kbvars_k562))), #The size of the intersection between the first, second and third sets
	n124 = length(Reduce(intersect, list(rfe10kbvars_k562,rfe25kbvars_k562,rfe100kbvars_k562))), #The size of the intersection between the first, second and fourth sets
	n134 = length(Reduce(intersect, list(rfe10kbvars_k562,rfe50kbvars_k562,rfe100kbvars_k562))), #The size of the intersection between the first, third and fourth sets
	n234 = length(Reduce(intersect, list(rfe25kbvars_k562,rfe50kbvars_k562,rfe100kbvars_k562))), #The size of the intersection between the second, third and fourth sets
	n1234 = length(Reduce(intersect, list(rfe10kbvars_k562,rfe25kbvars_k562,rfe50kbvars_k562,rfe100kbvars_k562))), #The size of the intersection between all four sets
	category = c("10 kb", "25 kb", "50 kb", "100 kb"),
	fill = c("orange", "red", "green", "blue"),
	lty = "dashed",
	cex = 2,
	cat.cex = 2,
	cat.col = c("orange", "red", "green", "blue")
	);

grid.newpage()
venn.plot.10kb <- draw.pairwise.venn(
	area1 = length(rfe10kbvars),
	area2 = length(rfe10kbvars_k562),
	cross.area = length(intersect(rfe10kbvars,rfe10kbvars_k562)),
	category = c("GM12878", "K562"),
	fill = c("blue", "red"),
	lty = "blank",
	cex = 2,
	cat.cex = 2,
	cat.just = list(c(-.5, -3), c(2, -3)))

grid.newpage()
venn.plot.25kb <- draw.pairwise.venn(
	area1 = length(rfe25kbvars),
	area2 = length(rfe25kbvars_k562),
	cross.area = length(intersect(rfe25kbvars,rfe25kbvars_k562)),
	category = c("GM12878", "K562"),
	fill = c("blue", "red"),
	lty = "blank",
	cex = 2,
	cat.cex = 2,
	cat.just = list(c(-.5, -3), c(2, -3)))

grid.newpage()
venn.plot.50kb <- draw.pairwise.venn(
	area1 = length(rfe50kbvars),
	area2 = length(rfe50kbvars_k562),
	cross.area = length(intersect(rfe50kbvars,rfe50kbvars_k562)),
	category = c("GM12878", "K562"),
	fill = c("blue", "red"),
	lty = "blank",
	cex = 2,
	cat.cex = 2,
	cat.just = list(c(-.5, -3), c(2, -3)))

grid.newpage()
venn.plot.100kb <- draw.pairwise.venn(
	area1 = length(rfe100kbvars),
	area2 = length(rfe100kbvars_k562),
	cross.area = length(intersect(rfe100kbvars,rfe100kbvars_k562)),
	category = c("GM12878", "K562"),
	fill = c("blue", "red"),
	lty = "blank",
	cex = 2,
	cat.cex = 2,
	cat.just = list(c(-.5, -3), c(2, -3)))

```

##Heatmaps of top predictors

###10 kb

```{r}
topvars10kbgm12878 <- cbind.data.frame(names(rfeModel_dist_10kb$fit$importance[,1]), rfeModel_dist_10kb$fit$importance)
topvars10kbgm12878 <- topvars10kbgm12878[1:32,]
rownames(topvars10kbgm12878) <- NULL
names(topvars10kbgm12878) <- c("Feature", "Importance")
topvars10kbgm12878$Feature <- gsub(paste(c("Gm12878-",
                                           "_dist",
                                           "-Haib", 
                                           "-Sydh",
                                           "-Uta",
                                           "-Uw",
                                           "-Broad",
                                           "-Uchicago"), collapse = "|"),
                                   "",
                                   topvars10kbgm12878$Feature)

topvars10kbk562 <- cbind.data.frame(names(rfeModel_dist_10kb_k562$fit$importance[,1]), rfeModel_dist_10kb_k562$fit$importance[,4])
topvars10kbk562 <- topvars10kbk562[1:32,]
rownames(topvars10kbk562) <- NULL
names(topvars10kbk562) <- c("Feature", "Importance")
topvars10kbk562$Feature <- gsub(paste(c("K562-",
                                           "_dist",
                                           "-Haib", 
                                           "-Sydh",
                                           "-Uta",
                                           "-Uw",
                                           "-Broad",
                                           "-Uchicago"), collapse = "|"),
                                   "",
                                   topvars10kbk562$Feature)

impdata_10kb <- data.frame(Feature = Reduce(union, 
                                            list(topvars10kbgm12878$Feature,
                                                 topvars10kbk562$Feature)))
impdata_10kb$gm12878_10kb <- NA
impdata_10kb$k562_10kb <- NA

impdata_10kb$gm12878_10kb[which(impdata_10kb$Feature %in% topvars10kbgm12878$Feature)] <- topvars10kbgm12878$Importance
impdata_10kb$k562_10kb[which(impdata_10kb$Feature %in% topvars10kbk562$Feature)] <- topvars10kbk562$Importance

impdata_10kb <- impdata_10kb[order(impdata_10kb$Feature),]

#preObj <- preProcess(impdata_10kb, method=c("center", "scale"))
#impdata_10kb <- predict(preObj, impdata_10kb)

impdata_10kb_long <- melt(impdata_10kb)
impdata_10kb_long <- ddply(impdata_10kb_long, .(variable), transform,   
                           rescale = rescale(value))

topvars10kbplot <- ggplot(impdata_10kb_long, aes(variable, Feature)) + 
    geom_tile(aes(fill = rescale),
              colour = "white") + scale_fill_gradient2(low = "white", 
                                                      #mid = "white",
                                                      high = muted("red"), 
                                                      #midpoint = 0, 
                                                      space = "Lab",
                                                      na.value = "grey50", 
                                                      guide = "legend", 
                                                      aesthetics = "fill",
                                                      name="Importance",
                                                      labels=c("Very Low",
                                                               "Low",
                                                               "Medium",
                                                               "High",
                                                               "Very High"))+
  scale_x_discrete(name="Cell Line", labels=c("GM12878", "K562"))+
  theme_bw()+
  theme_minimal()
```

###25 kb

```{r}
topvars25kbgm12878 <- cbind.data.frame(names(rfeModel_dist_25kb$fit$importance[,1]), rfeModel_dist_25kb$fit$importance)
topvars25kbgm12878 <- topvars25kbgm12878[1:32,]
rownames(topvars25kbgm12878) <- NULL
names(topvars25kbgm12878) <- c("Feature", "Importance")
topvars25kbgm12878$Feature <- gsub(paste(c("Gm12878-",
                                           "_dist",
                                           "-Haib", 
                                           "-Sydh",
                                           "-Uta",
                                           "-Uw",
                                           "-Broad",
                                           "-Uchicago"), collapse = "|"),
                                   "",
                                   topvars25kbgm12878$Feature)

topvars25kbk562 <- cbind.data.frame(names(rfeModel_dist_25kb_k562$fit$importance[,1]), rfeModel_dist_25kb_k562$fit$importance[,1])
topvars25kbk562 <- topvars25kbk562[1:32,]
rownames(topvars25kbk562) <- NULL
names(topvars25kbk562) <- c("Feature", "Importance")
topvars25kbk562$Feature <- gsub(paste(c("K562-",
                                           "_dist",
                                           "-Haib", 
                                           "-Sydh",
                                           "-Uta",
                                           "-Uw",
                                           "-Broad",
                                           "-Uchicago"), collapse = "|"),
                                   "",
                                   topvars25kbk562$Feature)

impdata_25kb <- data.frame(Feature = Reduce(union, 
                                            list(topvars25kbgm12878$Feature,
                                                 topvars25kbk562$Feature)))
impdata_25kb$gm12878_25kb <- NA
impdata_25kb$k562_25kb <- NA

impdata_25kb$gm12878_25kb[which(impdata_25kb$Feature %in% topvars25kbgm12878$Feature)] <- topvars25kbgm12878$Importance
impdata_25kb$k562_25kb[which(impdata_25kb$Feature %in% topvars25kbk562$Feature)] <- topvars25kbk562$Importance

impdata_25kb <- impdata_25kb[order(impdata_25kb$Feature),]

#preObj <- preProcess(impdata_25kb, method=c("center", "scale"))
#impdata_25kb <- predict(preObj, impdata_25kb)

impdata_25kb_long <- melt(impdata_25kb)
impdata_25kb_long <- ddply(impdata_25kb_long, .(variable), transform,   
                           rescale = rescale(value))

topvars25kbplot <- ggplot(impdata_25kb_long, aes(variable, Feature)) + 
    geom_tile(aes(fill = rescale),
              colour = "white") + scale_fill_gradient2(low = "white", 
                                                      #mid = "white",
                                                      high = muted("red"), 
                                                      #midpoint = 0, 
                                                      space = "Lab",
                                                      na.value = "grey50", 
                                                      guide = "legend", 
                                                      aesthetics = "fill",
                                                      name="Importance",
                                                      labels=c("Very Low",
                                                               "Low",
                                                               "Medium",
                                                               "High",
                                                               "Very High"))+
  scale_x_discrete(name="Cell Line", labels=c("GM12878", "K562"))+
  theme_bw()+
  theme_minimal()
```

###50kb

```{r}
topvars50kbgm12878 <- cbind.data.frame(names(rfeModel_dist_50kb$fit$importance[,1]), rfeModel_dist_50kb$fit$importance)
topvars50kbgm12878 <- topvars50kbgm12878[1:32,]
rownames(topvars50kbgm12878) <- NULL
names(topvars50kbgm12878) <- c("Feature", "Importance")
topvars50kbgm12878$Feature <- gsub(paste(c("Gm12878-",
                                           "_dist",
                                           "-Haib", 
                                           "-Sydh",
                                           "-Uta",
                                           "-Uw",
                                           "-Broad",
                                           "-Uchicago"), collapse = "|"),
                                   "",
                                   topvars50kbgm12878$Feature)

topvars50kbk562 <- cbind.data.frame(names(rfeModel_dist_50kb_k562$fit$importance[,1]), rfeModel_dist_50kb_k562$fit$importance[,1])
topvars50kbk562 <- topvars50kbk562[1:32,]
rownames(topvars50kbk562) <- NULL
names(topvars50kbk562) <- c("Feature", "Importance")
topvars50kbk562$Feature <- gsub(paste(c("K562-",
                                           "_dist",
                                           "-Haib", 
                                           "-Sydh",
                                           "-Uta",
                                           "-Uw",
                                           "-Broad",
                                           "-Uchicago"), collapse = "|"),
                                   "",
                                   topvars50kbk562$Feature)

impdata_50kb <- data.frame(Feature = Reduce(union, 
                                            list(topvars50kbgm12878$Feature,
                                                 topvars50kbk562$Feature)))
impdata_50kb$gm12878_50kb <- NA
impdata_50kb$k562_50kb <- NA

impdata_50kb$gm12878_50kb[which(impdata_50kb$Feature %in% topvars50kbgm12878$Feature)] <- topvars50kbgm12878$Importance
impdata_50kb$k562_50kb[which(impdata_50kb$Feature %in% topvars50kbk562$Feature)] <- topvars50kbk562$Importance

impdata_50kb <- impdata_50kb[order(impdata_50kb$Feature),]

#preObj <- preProcess(impdata_50kb, method=c("center", "scale"))
#impdata_50kb <- predict(preObj, impdata_50kb)

impdata_50kb_long <- melt(impdata_50kb)
impdata_50kb_long <- ddply(impdata_50kb_long, .(variable), transform,   
                           rescale = rescale(value))

topvars50kbplot <- ggplot(impdata_50kb_long, aes(variable, Feature)) + 
    geom_tile(aes(fill = rescale),
              colour = "white") + scale_fill_gradient2(low = "white", 
                                                      #mid = "white",
                                                      high = muted("red"), 
                                                      #midpoint = 0, 
                                                      space = "Lab",
                                                      na.value = "grey50", 
                                                      guide = "legend", 
                                                      aesthetics = "fill",
                                                      name="Importance",
                                                      labels=c("Very Low",
                                                               "Low",
                                                               "Medium",
                                                               "High",
                                                               "Very High"))+
  scale_x_discrete(name="Cell Line", labels=c("GM12878", "K562"))+
  theme_bw()+
  theme_minimal()
```

###100kb

```{r}
topvars100kbgm12878 <- cbind.data.frame(names(rfeModel_dist_100kb$fit$importance[,1]), rfeModel_dist_100kb$fit$importance)
topvars100kbgm12878 <- topvars100kbgm12878[1:32,]
rownames(topvars100kbgm12878) <- NULL
names(topvars100kbgm12878) <- c("Feature", "Importance")
topvars100kbgm12878$Feature <- gsub(paste(c("Gm12878-",
                                           "_dist",
                                           "-Haib", 
                                           "-Sydh",
                                           "-Uta",
                                           "-Uw",
                                           "-Broad",
                                           "-Uchicago"), collapse = "|"),
                                   "",
                                   topvars100kbgm12878$Feature)

topvars100kbk562 <- cbind.data.frame(rfeModel_dist_100kb_k562$variables$var[1:32], rfeModel_dist_100kb_k562$variables$Overall[1:32])
topvars100kbk562 <- topvars100kbk562[1:32,]
rownames(topvars100kbk562) <- NULL
names(topvars100kbk562) <- c("Feature", "Importance")
topvars100kbk562$Feature <- gsub(paste(c("K562-",
                                           "_dist",
                                           "-Haib", 
                                           "-Sydh",
                                           "-Uta",
                                           "-Uw",
                                           "-Broad",
                                           "-Uchicago"), collapse = "|"),
                                   "",
                                   topvars100kbk562$Feature)

impdata_100kb <- data.frame(Feature = Reduce(union, 
                                            list(topvars100kbgm12878$Feature,
                                                 topvars100kbk562$Feature)))
impdata_100kb$gm12878_100kb <- NA
impdata_100kb$k562_100kb <- NA

impdata_100kb$gm12878_100kb[which(impdata_100kb$Feature %in% topvars100kbgm12878$Feature)] <- topvars100kbgm12878$Importance
impdata_100kb$k562_100kb[which(impdata_100kb$Feature %in% topvars100kbk562$Feature)] <- topvars100kbk562$Importance

impdata_100kb <- impdata_100kb[order(impdata_100kb$Feature),]

#preObj <- preProcess(impdata_100kb, method=c("center", "scale"))
#impdata_100kb <- predict(preObj, impdata_100kb)

impdata_100kb_long <- melt(impdata_100kb)
impdata_100kb_long <- ddply(impdata_100kb_long, .(variable), transform,   
                           rescale = rescale(value))

topvars100kbplot <- ggplot(impdata_100kb_long, aes(variable, Feature)) + 
    geom_tile(aes(fill = rescale),
              colour = "white") + scale_fill_gradient2(low = "white", 
                                                      #mid = "white",
                                                      high = muted("red"), 
                                                      #midpoint = 0, 
                                                      space = "Lab",
                                                      na.value = "grey50", 
                                                      guide = "legend", 
                                                      aesthetics = "fill",
                                                      name="Importance",
                                                      labels=c("Very Low",
                                                               "Low",
                                                               "Medium",
                                                               "High",
                                                               "Very High"))+
  scale_x_discrete(name="Cell Line", labels=c("GM12878", "K562"))+
  theme_bw()+
  theme_minimal()
```

###Combined

```{r}
ggarrange(topvars25kbplot,
          topvars50kbplot,
          topvars100kbplot,
          ncol=3,
          nrow=1,
          common.legend = TRUE,
          legend = "right")
```

#Barplots of performance metrics

##Comparing results for no class balancing

###10kb

```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/LASSO/RF/rfperf.rds")
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/LASSO/RF/rfperf.rds")
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/LASSO/RF/rfperf.rds")

none.10 <- data.frame(Performance = c(rfperf1[9,2],
								rfperf2[9,2],
								rfperf3[9,2],
								rfperf1[17,2],
								rfperf2[17,2],
								rfperf3[17,2],
								rfperf1[16,2],
								rfperf2[16,2],
								rfperf3[16,2],
								rfperf1[15,2],
								rfperf2[15,2],
								rfperf3[15,2]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
none.10$Predictor <- factor(none.10$Predictor, levels=c("OC", "OP", "Distance"))
none.10$Metric <- factor(none.10$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###25kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OC/LASSO/RF/rfperf.rds")
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OP/LASSO/RF/rfperf.rds")
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/LASSO/RF/rfperf.rds")

none.25 <- data.frame(Performance = c(rfperf1[9,2],
								rfperf2[9,2],
								rfperf3[9,2],
								rfperf1[17,2],
								rfperf2[17,2],
								rfperf3[17,2],
								rfperf1[16,2],
								rfperf2[16,2],
								rfperf3[16,2],
								rfperf1[15,2],
								rfperf2[15,2],
								rfperf3[15,2]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
none.25$Predictor <- factor(none.25$Predictor, levels=c("OC", "OP", "Distance"))
none.25$Metric <- factor(none.25$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###50kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OC/LASSO/RF/rfperf.rds")
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OP/LASSO/RF/rfperf.rds")
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/LASSO/RF/rfperf.rds")

none.50 <- data.frame(Performance = c(rfperf1[9,2],
								rfperf2[9,2],
								rfperf3[9,2],
								rfperf1[17,2],
								rfperf2[17,2],
								rfperf3[17,2],
								rfperf1[16,2],
								rfperf2[16,2],
								rfperf3[16,2],
								rfperf1[15,2],
								rfperf2[15,2],
								rfperf3[15,2]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
none.50$Predictor <- factor(none.50$Predictor, levels=c("OC", "OP", "Distance"))
none.50$Metric <- factor(none.50$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###100kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OC/LASSO/RF/rfperf.rds")
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OP/LASSO/RF/rfperf.rds")
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/LASSO/RF/rfperf.rds")

none.100 <- data.frame(Performance = c(rfperf1[9,2],
								rfperf2[9,2],
								rfperf3[9,2],
								rfperf1[17,2],
								rfperf2[17,2],
								rfperf3[17,2],
								rfperf1[16,2],
								rfperf2[16,2],
								rfperf3[16,2],
								rfperf1[15,2],
								rfperf2[15,2],
								rfperf3[15,2]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
none.100$Predictor <- factor(none.100$Predictor, levels=c("OC", "OP", "Distance"))
none.100$Metric <- factor(none.100$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###Plots
```{r}
none.data <- rbind.data.frame(none.10,none.25,none.50,none.100) 
none.data$Resolution <- c(rep("10 kb", 12), rep("25 kb", 12), rep("50 kb", 12), rep("100 kb", 12))
none.data$Resolution <- factor(none.data$Resolution, levels=c("10 kb","25 kb","50 kb","100 kb"))

allnoneplots <- ggplot(none.data, aes(fill=Resolution, y=Performance, x=Resolution)) +
	geom_bar( stat="identity", position="dodge") +
	facet_grid(Predictor ~ Metric) +
	ylab("Performance")+
	xlab("Resolution")+
	theme_bw() +
  guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15), #, angle = 45, margin = ggplot2::margin(t = 35)),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
		legend.position="bottom")

```

###Tables

```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/LASSO/RF/rfperf.rds")
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/LASSO/RF/rfperf.rds")
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/LASSO/RF/rfperf.rds")

rfperf4 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OC/LASSO/RF/rfperf.rds")
rfperf5 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OP/LASSO/RF/rfperf.rds")
rfperf6 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/LASSO/RF/rfperf.rds")

rfperf7 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OC/LASSO/RF/rfperf.rds")
rfperf8 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OP/LASSO/RF/rfperf.rds")
rfperf9 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/LASSO/RF/rfperf.rds")

rfperf10 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OC/LASSO/RF/rfperf.rds")
rfperf11 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OP/LASSO/RF/rfperf.rds")
rfperf12 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/LASSO/RF/rfperf.rds")

nonetab <- cbind.data.frame(rfperf1$Metric,
                            rfperf1$Performance,
                            rfperf2$Performance,
                            rfperf3$Performance,
                            rfperf4$Performance,
                            rfperf5$Performance,
                            rfperf6$Performance,
                            rfperf7$Performance,
                            rfperf8$Performance,
                            rfperf9$Performance,
                            rfperf10$Performance,
                            rfperf11$Performance,
                            rfperf12$Performance)
nonetab[,-1] <- round(nonetab[,-1], 3)
datatable(nonetab)
```

##Comparing results for RUS

###10kb

```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

RUS.10 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
RUS.10$Predictor <- factor(RUS.10$Predictor, levels=c("OC", "OP", "Distance"))
RUS.10$Metric <- factor(RUS.10$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###25kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

RUS.25 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
RUS.25$Predictor <- factor(RUS.25$Predictor, levels=c("OC", "OP", "Distance"))
RUS.25$Metric <- factor(RUS.25$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###50kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

RUS.50 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
RUS.50$Predictor <- factor(RUS.50$Predictor, levels=c("OC", "OP", "Distance"))
RUS.50$Metric <- factor(RUS.50$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###100kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

RUS.100 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
RUS.100$Predictor <- factor(RUS.100$Predictor, levels=c("OC", "OP", "Distance"))
RUS.100$Metric <- factor(RUS.100$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###Plots
```{r}
RUS.data <- rbind.data.frame(RUS.10,RUS.25,RUS.50,RUS.100) 
RUS.data$Resolution <- c(rep("10 kb", 12), rep("25 kb", 12), rep("50 kb", 12), rep("100 kb", 12))
RUS.data$Resolution <- factor(RUS.data$Resolution, levels=c("10 kb","25 kb","50 kb","100 kb"))

allRUSplots <- ggplot(RUS.data, aes(fill=Resolution, y=Performance, x=Resolution)) +
	geom_bar( stat="identity", position="dodge") +
	facet_grid(Predictor ~ Metric) +
	ylab("Performance")+
	xlab("Resolution")+
	theme_bw() +
  guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15), #, angle = 45, margin = ggplot2::margin(t = 35)),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
		legend.position="bottom")

```

###Tables
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

rfperf4 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf4 <- rowMeans(rfperf4)
rfperf5 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf5 <- rowMeans(rfperf5)
rfperf6 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf6 <- rowMeans(rfperf6)

rfperf7 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf7 <- rowMeans(rfperf7)
rfperf8 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf8 <- rowMeans(rfperf8)
rfperf9 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf9 <- rowMeans(rfperf9)

rfperf10 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf10 <- rowMeans(rfperf10)
rfperf11 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf11 <- rowMeans(rfperf11)
rfperf12 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf12 <- rowMeans(rfperf12)

RUStab <- cbind.data.frame(names(rfperf1),
                            rfperf1,
                            rfperf2,
                            rfperf3,
                            rfperf4,
                            rfperf5,
                            rfperf6,
                            rfperf7,
                            rfperf8,
                            rfperf9,
                            rfperf10,
                            rfperf11,
                            rfperf12)
RUStab[,-1] <- round(RUStab[,-1], 3)
datatable(RUStab)
```

##Comparing results for SMOTE

###10kb

```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

SMOTE.10 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
SMOTE.10$Predictor <- factor(SMOTE.10$Predictor, levels=c("OC", "OP", "Distance"))
SMOTE.10$Metric <- factor(SMOTE.10$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###25kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

SMOTE.25 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
SMOTE.25$Predictor <- factor(SMOTE.25$Predictor, levels=c("OC", "OP", "Distance"))
SMOTE.25$Metric <- factor(SMOTE.25$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###50kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

SMOTE.50 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
SMOTE.50$Predictor <- factor(SMOTE.50$Predictor, levels=c("OC", "OP", "Distance"))
SMOTE.50$Metric <- factor(SMOTE.50$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###100kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

SMOTE.100 <- data.frame(Performance = c(rfperf1[9],
								NA,#rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								NA,#rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								NA,#rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								NA,#rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
SMOTE.100$Predictor <- factor(SMOTE.100$Predictor, levels=c("OC", "OP", "Distance"))
SMOTE.100$Metric <- factor(SMOTE.100$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###Plots
```{r}
SMOTE.data <- rbind.data.frame(SMOTE.10,SMOTE.25,SMOTE.50,SMOTE.100) 
SMOTE.data$Resolution <- c(rep("10 kb", 12), rep("25 kb", 12), rep("50 kb", 12), rep("100 kb", 12))
SMOTE.data$Resolution <- factor(SMOTE.data$Resolution, levels=c("10 kb","25 kb","50 kb","100 kb"))

allSMOTEplots <- ggplot(SMOTE.data, aes(fill=Resolution, y=Performance, x=Resolution)) +
	geom_bar( stat="identity", position="dodge") +
	facet_grid(Predictor ~ Metric) +
	ylab("Performance")+
	xlab("Resolution")+
	theme_bw() +
  guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15), #, angle = 45, margin = ggplot2::margin(t = 35)),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
		legend.position="bottom")

```

###Tables
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

rfperf4 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OC/LASSO/RF/rfperf.rds")
rfperf4 <- rowMeans(rfperf4)
rfperf5 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OP/LASSO/RF/rfperf.rds")
rfperf5 <- rowMeans(rfperf5)
rfperf6 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/Distance/LASSO/RF/rfperf.rds")
rfperf6 <- rowMeans(rfperf6)

rfperf7 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OC/LASSO/RF/rfperf.rds")
rfperf7 <- rowMeans(rfperf7)
rfperf8 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OP/LASSO/RF/rfperf.rds")
rfperf8 <- rowMeans(rfperf8)
rfperf9 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/Distance/LASSO/RF/rfperf.rds")
rfperf9 <- rowMeans(rfperf9)

rfperf10 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OC/LASSO/RF/rfperf.rds")
rfperf10 <- rowMeans(rfperf10)
rfperf11 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OP/LASSO/RF/rfperf.rds")
rfperf11 <- rowMeans(rfperf11)
rfperf12 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/Distance/LASSO/RF/rfperf.rds")
rfperf12 <- rowMeans(rfperf12)

SMOTEtab <- cbind.data.frame(names(rfperf1),
                            rfperf1,
                            rfperf2,
                            rfperf3,
                            rfperf4,
                            rfperf5,
                            rfperf6,
                            rfperf7,
                            rfperf8,
                            rfperf9,
                            rfperf10,
                            rep(NA,18),#rfperf11,
                            rfperf12)
SMOTEtab[,-c(1)] <- round(SMOTEtab[,-c(1)], 3)
SMOTEtab[,-c(1,12)] <- round(SMOTEtab[,-c(1,12)], 3)
datatable(SMOTEtab)
```

##Comparing results for SMOTE

###10kb

```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

ROS.10 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
ROS.10$Predictor <- factor(ROS.10$Predictor, levels=c("OC", "OP", "Distance"))
ROS.10$Metric <- factor(ROS.10$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###25kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

ROS.25 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
ROS.25$Predictor <- factor(ROS.25$Predictor, levels=c("OC", "OP", "Distance"))
ROS.25$Metric <- factor(ROS.25$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###50kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

ROS.50 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
ROS.50$Predictor <- factor(ROS.50$Predictor, levels=c("OC", "OP", "Distance"))
ROS.50$Metric <- factor(ROS.50$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###100kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

ROS.100 <- data.frame(Performance = c(rfperf1[9],
								rfperf2[9],
								rfperf3[9],
								rfperf1[17],
								rfperf2[17],
								rfperf3[17],
								rfperf1[16],
								rfperf2[16],
								rfperf3[16],
								rfperf1[15],
								rfperf2[15],
								rfperf3[15]),
				Predictor = rep(c("OC", "OP", "Distance"), 4),
				Metric = c(rep("Accuracy", 3), rep("AUC", 3), rep("F1-Score", 3), rep("MCC", 3)))
ROS.100$Predictor <- factor(ROS.100$Predictor, levels=c("OC", "OP", "Distance"))
ROS.100$Metric <- factor(ROS.100$Metric, levels=c("Accuracy", "AUC", "F1-Score", "MCC"))
```

###Plots
```{r}
ROS.data <- rbind.data.frame(ROS.10,ROS.25,ROS.50,ROS.100) 
ROS.data$Resolution <- c(rep("10 kb", 12), rep("25 kb", 12), rep("50 kb", 12), rep("100 kb", 12))
ROS.data$Resolution <- factor(ROS.data$Resolution, levels=c("10 kb","25 kb","50 kb","100 kb"))

allROSplots <- ggplot(ROS.data, aes(fill=Resolution, y=Performance, x=Resolution)) +
	geom_bar( stat="identity", position="dodge") +
	facet_grid(Predictor ~ Metric) +
	ylab("Performance")+
	xlab("Resolution")+
	theme_bw() +
  guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15), #, angle = 45, margin = ggplot2::margin(t = 35)),
        axis.text.y = element_text(size = 15),
        #axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
		legend.position="bottom")

```

###Tables
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- rowMeans(rfperf1)
rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- rowMeans(rfperf2)
rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- rowMeans(rfperf3)

rfperf4 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OC/LASSO/RF/rfperf.rds")
rfperf4 <- rowMeans(rfperf4)
rfperf5 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OP/LASSO/RF/rfperf.rds")
rfperf5 <- rowMeans(rfperf5)
rfperf6 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/Distance/LASSO/RF/rfperf.rds")
rfperf6 <- rowMeans(rfperf6)

rfperf7 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OC/LASSO/RF/rfperf.rds")
rfperf7 <- rowMeans(rfperf7)
rfperf8 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OP/LASSO/RF/rfperf.rds")
rfperf8 <- rowMeans(rfperf8)
rfperf9 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/Distance/LASSO/RF/rfperf.rds")
rfperf9 <- rowMeans(rfperf9)

rfperf10 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OC/LASSO/RF/rfperf.rds")
rfperf10 <- rowMeans(rfperf10)
rfperf11 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OP/LASSO/RF/rfperf.rds")
rfperf11 <- rowMeans(rfperf11)
rfperf12 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/Distance/LASSO/RF/rfperf.rds")
rfperf12 <- rowMeans(rfperf12)

ROStab <- cbind.data.frame(names(rfperf1),
                            rfperf1,
                            rfperf2,
                            rfperf3,
                            rfperf4,
                            rfperf5,
                            rfperf6,
                            rfperf7,
                            rfperf8,
                            rfperf9,
                            rfperf10,
                            rfperf11,
                            rfperf12)
ROStab[,-1] <- round(ROStab[,-1], 3)
datatable(ROStab)
```

#Comparing performance for distance predictors using RUS 
##10kb
```{r}
rfperf1 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf1 <- cbind.data.frame(rownames(rfperf1),round(rowMeans(rfperf1),3))

rfperf2 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf2 <- cbind.data.frame(rownames(rfperf2),round(rowMeans(rfperf2),3))

rfperf3 <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf3 <- cbind.data.frame(rownames(rfperf3),round(rowMeans(rfperf3),3))

```
##25kb
```{r}
rfperf4 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf4 <- cbind.data.frame(rownames(rfperf4),round(rowMeans(rfperf4),3))

rfperf5 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf5 <- cbind.data.frame(rownames(rfperf5),round(rowMeans(rfperf5),3))

rfperf6 <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf6 <- cbind.data.frame(rownames(rfperf6),round(rowMeans(rfperf6),3))
```
##50kb
```{r}
rfperf7 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf7 <- cbind.data.frame(rownames(rfperf7),round(rowMeans(rfperf7),3))

rfperf8 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf8 <- cbind.data.frame(rownames(rfperf8),round(rowMeans(rfperf8),3))

rfperf9 <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf9 <- cbind.data.frame(rownames(rfperf9),round(rowMeans(rfperf9),3))

```
##100kb
```{r}
rfperf10 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OC/LASSO/RF/rfperf.rds")
rfperf10 <- cbind.data.frame(rownames(rfperf10),round(rowMeans(rfperf10),3))

rfperf11 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OP/LASSO/RF/rfperf.rds")
rfperf11 <- cbind.data.frame(rownames(rfperf11),round(rowMeans(rfperf11),3))

rfperf12 <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/Distance/LASSO/RF/rfperf.rds")
rfperf12 <- cbind.data.frame(rownames(rfperf12),round(rowMeans(rfperf12),3))

```

##Plots
```{r}
distperftab <- data.frame(Performance =
                    c(rfperf1[9,2],
											rfperf1[17,2],
											rfperf1[16,2],
											rfperf1[15,2],
											rfperf2[9,2],
											rfperf2[17,2],
											rfperf2[16,2],
											rfperf2[15,2],
											rfperf3[9,2],
											rfperf3[17,2],
											rfperf3[16,2],
											rfperf3[15,2],
											rfperf4[9,2],
											rfperf4[17,2],
											rfperf4[16,2],
											rfperf4[15,2],
											rfperf5[9,2],
											rfperf5[17,2],
											rfperf5[16,2],
											rfperf5[15,2],
											rfperf6[9,2],
											rfperf6[17,2],
											rfperf6[16,2],
											rfperf6[15,2],
											rfperf7[9,2],
											rfperf7[17,2],
											rfperf7[16,2],
											rfperf7[15,2],
											rfperf8[9,2],
											rfperf8[17,2],
											rfperf8[16,2],
											rfperf8[15,2],
											rfperf9[9,2],
											rfperf9[17,2],
											rfperf9[16,2],
											rfperf9[15,2],
											rfperf10[9,2],
											rfperf10[17,2],
											rfperf10[16,2],
											rfperf10[15,2],
											rfperf11[9,2],
											rfperf11[17,2],
											rfperf11[16,2],
											rfperf11[15,2],
											rfperf12[9,2],
											rfperf12[17,2],
											rfperf12[16,2],
											rfperf12[15,2]),
							Metric = rep(c("Accuracy", "AUC", "F1-Score", "MCC"), 12),
							Resolution = c(rep("10kb", 12), rep("25kb", 12), rep("50kb", 12), rep("100kb", 12)),
							Predictor = rep(c(rep("OC", 4), rep("OP", 4), rep("Distance", 4)), 4))
distperftab$Metric <- factor(distperftab$Metric, levels = c("Accuracy", "AUC", "F1-Score", "MCC"))
distperftab$Resolution <- factor(distperftab$Resolution, levels = c("10kb","25kb","50kb","100kb"))
distperftab$Predictor <- factor(distperftab$Predictor, levels = c("OC", "OP", "Distance"))

p.distperfsbar <- ggplot(distperftab, aes(fill=Predictor, y=Performance, x=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
	facet_grid(Resolution ~ Metric) +
  ylab("Performance")+
	xlab("Predictor Type")+
  theme_bw() +
  guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15), #, angle = 45, margin = ggplot2::margin(t = 35)),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
		legend.position="bottom")

#ggsave("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/results/GM12878/p.distperfsbarline.png")
```

#Comparing performances across resolutions for RUS and distance

#Variable Importance Plots for distance predictors with ENET reduction

##5kb
###Random Under-Sampling
```{r}
rfimpvars <- readRDS("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878/5kb/CHR1/rus/distance/enet/rfimpvars.rds")

rf5kb <- data.frame(Feature = as.character(rownames(rfimpvars)),
					Importance = rowMeans(rfimpvars))
rf5kb <- rf5kb[order(rf5kb$Importance, decreasing=TRUE),]
rownames(rf) <- NULL

varimp.rf <- rf5kb$Importance
Labels <- as.character(rf5kb$Feature)
Labels[grep("Gm12878-", Labels)] <- gsub("Gm12878-","",Labels[grep("Gm12878-", Labels)])
Labels[grep("_dist", Labels)] <- gsub("_dist","",Labels[grep("_dist", Labels)])
varimp.rf.df.5kb <- cbind.data.frame(Feature=as.character(Labels),
                                 Importance=varimp.rf)
varimp.rf.df.5kb$Feature <- as.character(varimp.rf.df.5kb$Feature)							 
varimp.rf.df.5kb <-varimp.rf.df.5kb[order(varimp.rf.df.5kb$Importance),]
varimp.rf.df.5kb <- varimp.rf.df.5kb[(dim(varimp.rf.df.5kb)[1]-14):dim(varimp.rf.df.5kb)[1],]
varimp.rf.df.5kb$Feature <- factor(varimp.rf.df.5kb$Feature,
                                     levels=as.character(varimp.rf.df.5kb$Feature))

p.rf.5kb <- ggplot(varimp.rf.df.5kb, aes(x=Feature, y=Importance)) +
  xlab("Predictors") +
  ylab("Importance") +
  ggtitle("(A)") +
  geom_bar(stat="identity", 
           width=.5, 
           position="dodge",
           fill="red") +
  theme_minimal() +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=20,face="bold")) +
  coord_flip()

  ggsave("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/results/GM12878/5kb/p.rf.5kb.png")
```
##25kb
###Random Under-Sampling
```{r}
  rfimpvars <- readRDS("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878/25kb/CHR1/rus/distance/enet/rfimpvars.rds")

rf <- data.frame(Feature = as.character(rownames(rfimpvars)),
					Importance = rowMeans(rfimpvars))
rf <- rf[order(rf$Importance, decreasing=TRUE),]
rownames(rf) <- NULL

varimp.rf <- rf$Importance
Labels <- as.character(rf$Feature)
Labels[grep("Gm12878-", Labels)] <- gsub("Gm12878-","",Labels[grep("Gm12878-", Labels)])
Labels[grep("_dist", Labels)] <- gsub("_dist","",Labels[grep("_dist", Labels)])
varimp.rf.df.25kb <- cbind.data.frame(Feature=as.character(Labels),
                                 Importance=varimp.rf)
varimp.rf.df.25kb$Feature <- as.character(varimp.rf.df.25kb$Feature)							 
varimp.rf.df.25kb <-varimp.rf.df.25kb[order(varimp.rf.df.25kb$Importance),]
varimp.rf.df.25kb <- varimp.rf.df.25kb[(dim(varimp.rf.df.25kb)[1]-14):dim(varimp.rf.df.25kb)[1],]
varimp.rf.df.25kb$Feature <- factor(varimp.rf.df.25kb$Feature,
                                     levels=as.character(varimp.rf.df.25kb$Feature))

p.rf.25kb <- ggplot(varimp.rf.df.25kb, aes(x=Feature, y=Importance)) +
  xlab("Predictors") +
  ylab("Importance") +
  ggtitle("(B)") +
  geom_bar(stat="identity", 
           width=.5, 
           position="dodge",
           fill="blue") +
  theme_minimal() +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=20,face="bold")) +
  coord_flip()

   ggsave("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/results/GM12878/25kb/p.rf.25kb.png")
```
##50kb
###Random Under-Sampling
```{r}
rfimpvars <- readRDS("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878/50kb/CHR1/rus/distance/enet/rfimpvars.rds")

rf <- data.frame(Feature = as.character(rownames(rfimpvars)),
					Importance = rowMeans(rfimpvars))
rf <- rf[order(rf$Importance, decreasing=TRUE),]
rownames(rf) <- NULL

varimp.rf <- rf$Importance
Labels <- as.character(rf$Feature)
Labels[grep("Gm12878-", Labels)] <- gsub("Gm12878-","",Labels[grep("Gm12878-", Labels)])
Labels[grep("_dist", Labels)] <- gsub("_dist","",Labels[grep("_dist", Labels)])
varimp.rf.df.50kb <- cbind.data.frame(Feature=as.character(Labels),
                                 Importance=varimp.rf)
varimp.rf.df.50kb$Feature <- as.character(varimp.rf.df.50kb$Feature)							 
varimp.rf.df.50kb <-varimp.rf.df.50kb[order(varimp.rf.df.50kb$Importance),]
varimp.rf.df.50kb <- varimp.rf.df.50kb[(dim(varimp.rf.df.50kb)[1]-14):dim(varimp.rf.df.50kb)[1],]
varimp.rf.df.50kb$Feature <- factor(varimp.rf.df.50kb$Feature,
                                     levels=as.character(varimp.rf.df.50kb$Feature))

p.rf.50kb <- ggplot(varimp.rf.df.50kb, aes(x=Feature, y=Importance)) +
  xlab("Predictors") +
  ylab("Importance") +
  ggtitle("(C)") +
  geom_bar(stat="identity", 
           width=.5, 
           position="dodge",
           fill="green") +
  theme_minimal() +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=20,face="bold")) +
  coord_flip()

   ggsave("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/results/GM12878/50kb/p.rf.50kb.png")
```
##100kb
###Random Under-Sampling
```{r}
rfimpvars <- readRDS("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878/100kb/CHR1/rus/distance/enet/rfimpvars.rds")

rf <- data.frame(Feature = as.character(rownames(rfimpvars)),
					Importance = rowMeans(rfimpvars))
rf <- rf[order(rf$Importance, decreasing=TRUE),]
rownames(rf) <- NULL

varimp.rf <- rf$Importance
Labels <- as.character(rf$Feature)
Labels[grep("Gm12878-", Labels)] <- gsub("Gm12878-","",Labels[grep("Gm12878-", Labels)])
Labels[grep("_dist", Labels)] <- gsub("_dist","",Labels[grep("_dist", Labels)])
varimp.rf.df.100kb <- cbind.data.frame(Feature=as.character(Labels),
                                 Importance=varimp.rf)
varimp.rf.df.100kb$Feature <- as.character(varimp.rf.df.100kb$Feature)							 
varimp.rf.df.100kb <-varimp.rf.df.100kb[order(varimp.rf.df.100kb$Importance),]
#varimp.rf.df.100kb <- varimp.rf.df.100kb[(dim(varimp.rf.df.100kb)[1]-14):dim(varimp.rf.df.100kb)[1],]
varimp.rf.df.100kb$Feature <- factor(varimp.rf.df.100kb$Feature,
                                     levels=as.character(varimp.rf.df.100kb$Feature))

p.rf.100kb <- ggplot(varimp.rf.df.100kb, aes(x=Feature, y=Importance)) +
  xlab("Predictors") +
  ylab("Importance") +
  ggtitle("(D)") +
  geom_bar(stat="identity", 
           width=.5, 
           position="dodge",
           fill="orange") +
  theme_minimal() +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=20,face="bold")) +
  coord_flip()

   ggsave("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/results/GM12878/100kb/p.rf.100kb.png")

allvarimps <- ggarrange(p.rf.5kb,p.rf.25kb,p.rf.50kb,p.rf.100kb, ncol=2, nrow=2)

ggsave("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/results/GM12878/allvarimps.png")
```

#Network Analysis

```{r}
##venn diagram of top predictors 
#grid.newpage()
##vennp <- draw.triple.venn(area1 = 10, area2 = 10, area3 = 10, n12 = length(intersect(us,sm)), n23 = length(intersect(sm,km)), n13 = length(intersect(km,us)), 
##                 n123 = length(intersect(us,intersect(sm,km))), category = c("Under- \n Sampling", "SMOTE", "K-Medoids"), lty = "blank", 
##                 fill = c("red", "green", "blue"),cex=3,cat.cex = 3)
#
#vennp <- draw.pairwise.venn(length(varimp.rf.df.gm12878$Feature),
#						  length(varimp.rf.df.k562$Feature),
#						  length(intersect(varimp.rf.df.gm12878$Feature, varimp.rf.df.k562$Feature)),
#						  category = c("GM12878", "K562"),
#						  lty = "blank", 
#                          fill = c("red", "blue"),
#						  cex=3,
#						  cat.cex = 3,
#						  cat.pos = c(0,0),
#						  scaled = FALSE)

##Network Visualization between resolutions
rfimpvars <- readRDS("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878/5kb/CHR1/rus/distance/enet/rfimpvars.rds")

rf5kb <- data.frame(Feature = as.character(rownames(rfimpvars)),
					Importance = rowMeans(rfimpvars))
rownames(rf5kb) <- NULL
rf5kb$Feature <- as.character(rf5kb$Feature)
rf5kb$Feature[grep("Gm12878-", rf5kb$Feature)] <- gsub("Gm12878-","",rf5kb$Feature[grep("Gm12878-", rf5kb$Feature)])
rf5kb$Feature[grep("_dist", rf5kb$Feature)] <- gsub("_dist","",rf5kb$Feature[grep("_dist", rf5kb$Feature)])
rf5kb$Importance <- rf5kb$Importance/sum(rf5kb$Importance)

rfimpvars <- readRDS("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878/25kb/CHR1/rus/distance/enet/rfimpvars.rds")

rf25kb <- data.frame(Feature = as.character(rownames(rfimpvars)),
					Importance = rowMeans(rfimpvars))
rownames(rf25kb) <- NULL
rf25kb$Feature <- as.character(rf25kb$Feature)
rf25kb$Feature[grep("Gm12878-", rf25kb$Feature)] <- gsub("Gm12878-","",rf25kb$Feature[grep("Gm12878-", rf25kb$Feature)])
rf25kb$Feature[grep("_dist", rf25kb$Feature)] <- gsub("_dist","",rf25kb$Feature[grep("_dist", rf25kb$Feature)])
rf25kb$Importance <- rf25kb$Importance/sum(rf25kb$Importance)

rfimpvars <- readRDS("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878/50kb/CHR1/rus/distance/enet/rfimpvars.rds")

rf50kb <- data.frame(Feature = as.character(rownames(rfimpvars)),
					Importance = rowMeans(rfimpvars))
rownames(rf50kb) <- NULL
rf50kb$Feature <- as.character(rf50kb$Feature)
rf50kb$Feature[grep("Gm12878-", rf50kb$Feature)] <- gsub("Gm12878-","",rf50kb$Feature[grep("Gm12878-", rf50kb$Feature)])
rf50kb$Feature[grep("_dist", rf50kb$Feature)] <- gsub("_dist","",rf50kb$Feature[grep("_dist", rf50kb$Feature)])
rf50kb$Importance <- rf50kb$Importance/sum(rf50kb$Importance)

rfimpvars <- readRDS("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878/100kb/CHR1/rus/distance/enet/rfimpvars.rds")

rf100kb <- data.frame(Feature = as.character(rownames(rfimpvars)),
					Importance = rowMeans(rfimpvars))
rownames(rf100kb) <- NULL
rf100kb$Feature <- as.character(rf100kb$Feature)
rf100kb$Feature[grep("Gm12878-", rf100kb$Feature)] <- gsub("Gm12878-","",rf100kb$Feature[grep("Gm12878-", rf100kb$Feature)])
rf100kb$Feature[grep("_dist", rf100kb$Feature)] <- gsub("_dist","",rf100kb$Feature[grep("_dist", rf100kb$Feature)])
rf100kb$Importance <- rf100kb$Importance/sum(rf100kb$Importance)

commonannots <- union(rf5kb$Feature,union(rf25kb$Feature,union(rf50kb$Feature,rf100kb$Feature)))

networkmat <- matrix(ncol=4, nrow=length(commonannots), dimnames=list(commonannots, c("5kb", "25kb", "50kb", "100kb")))

networkmat[which(rownames(networkmat) %in% rf5kb$Feature),1] <- rf5kb$Importance
networkmat[which(rownames(networkmat) %in% rf25kb$Feature),2] <- rf25kb$Importance
networkmat[which(rownames(networkmat) %in% rf50kb$Feature),3] <- rf50kb$Importance
networkmat[which(rownames(networkmat) %in% rf100kb$Feature),4] <- rf100kb$Importance

networkmat[is.na(networkmat)] <- 0
networkmat <- networkmat*10

networkdf <- as.data.frame(networkmat)

net = network(networkdf,
              matrix.type = "bipartite",
              ignore.eval = FALSE,
              names.eval = "weights")

col = c("actor" = "green", "event" = "blue")

networkplot <- ggnet2(net, color = "mode", palette = col, edge.size = "weights", label=TRUE)

ggsave("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/results/GM12878/networkplot.png")

networkdf <- as.data.frame(t(networkdf))

net = network(networkdf,
              matrix.type = "bipartite",
              ignore.eval = FALSE,
              names.eval = "weights")

col = c("actor" = "blue", "event" = "green")

networkplot2 <- ggnet2(net, color = "mode", palette = col, edge.size = "weights", label=TRUE)

ggsave("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/results/GM12878/networkplot2.png")

```