### 2019-03-08 Scripts/evaluating_smote_performances.Rmd
script for comparing models with distance type predictors

---
title: "Comparing top performing models"
author: "Spiro Stilianoudakis"
date: "March 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# Comparing models with distance type predictors

## SMOTE

### GM12878

```{r}

###10kb
rfperf_gm12878_10kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

###25kb
rfperf_gm12878_25kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

###50kb
rfperf_gm12878_50kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

###100kb
rfperf_gm12878_100kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

```

### K562

```{r}

###10kb
rfperf_k562_10kb_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

###25kb
rfperf_k562_25kb_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

###50kb
rfperf_k562_50kb_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

###100kb
rfperf_k562_100kb_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

```

### Combined F1 and MCC and balanced acc

```{r}
all_smote_perfs <- data.frame(Performance = c(rfperf_gm12878_10kb_smote_dist[16,2],
                                              rfperf_gm12878_10kb_smote_dist[15,2],
                                              (rfperf_gm12878_10kb_smote_dist[6,2] + rfperf_gm12878_10kb_smote_dist[7,2])/2,
                                              rfperf_gm12878_25kb_smote_dist[16,2],
                                              rfperf_gm12878_25kb_smote_dist[15,2],
                                              (rfperf_gm12878_25kb_smote_dist[6,2] + rfperf_gm12878_25kb_smote_dist[7,2])/2,
                                              rfperf_gm12878_50kb_smote_dist[16,2],
                                              rfperf_gm12878_50kb_smote_dist[15,2],
                                              (rfperf_gm12878_50kb_smote_dist[6,2] + rfperf_gm12878_50kb_smote_dist[7,2])/2,
                                              rfperf_gm12878_100kb_smote_dist[16,2],
                                              rfperf_gm12878_100kb_smote_dist[15,2],
                                              (rfperf_gm12878_100kb_smote_dist[6,2] + rfperf_gm12878_100kb_smote_dist[7,2])/2,
                                              rfperf_k562_10kb_smote_dist[16,2],
                                              rfperf_k562_10kb_smote_dist[15,2],
                                              (rfperf_k562_10kb_smote_dist[6,2] + rfperf_k562_10kb_smote_dist[7,2])/2,
                                              rfperf_k562_25kb_smote_dist[16,2],
                                              rfperf_k562_25kb_smote_dist[15,2],
                                              (rfperf_k562_25kb_smote_dist[6,2] + rfperf_k562_25kb_smote_dist[7,2])/2,
                                              rfperf_k562_50kb_smote_dist[16,2],
                                              rfperf_k562_50kb_smote_dist[15,2],
                                              (rfperf_k562_50kb_smote_dist[6,2] + rfperf_k562_50kb_smote_dist[7,2])/2,
                                              rfperf_k562_100kb_smote_dist[16,2],
                                              rfperf_k562_100kb_smote_dist[15,2],
                                              (rfperf_k562_100kb_smote_dist[6,2] + rfperf_k562_100kb_smote_dist[7,2])/2),
                              CL = c(rep("GM12878", 12),
                                     rep("K562", 12)),
                              Resolution = rep(c(rep("10 kb", 3),
                                             rep("25 kb", 3),
                                             rep("50 kb", 3),
                                             rep("100 kb", 3)), 2),
                              Metric = rep(c("F1-Score", "MCC", "Balanced Acc"), 8))

all_smote_perfs$CL <- factor(all_smote_perfs$CL, levels=c("GM12878", "K562"))
all_smote_perfs$Resolution <- factor(all_smote_perfs$Resolution, levels=c("10 kb", "25 kb", "50 kb", "100 kb"))
all_smote_perfs$Metric <- factor(all_smote_perfs$Metric, levels=c("F1-Score", "MCC", "Balanced Acc"))

ggplot(all_smote_perfs, aes(x=Resolution, y=Performance, group=CL, color=CL)) +
  geom_line(aes(linetype=CL))+
  geom_point(aes(shape=CL))+
  facet_grid(~ Metric)+
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Cell Line")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15
	                                 #angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   #hjust = 1
	                                 ),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))
```

## None

### GM12878

```{r}

###10kb
rfperf_gm12878_10kb_None_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

###25kb
rfperf_gm12878_25kb_None_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

###50kb
rfperf_gm12878_50kb_None_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

###100kb
rfperf_gm12878_100kb_None_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

```

### K562

```{r}

###10kb
rfperf_k562_10kb_None_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

###25kb
rfperf_k562_25kb_None_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

###50kb
rfperf_k562_50kb_None_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

###100kb
rfperf_k562_100kb_None_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

```

### Combined F1 and MCC and balanced acc

```{r}
all_None_perfs <- data.frame(Performance = c(rfperf_gm12878_10kb_None_dist[16,2],
                                              rfperf_gm12878_10kb_None_dist[15,2],
                                              (rfperf_gm12878_10kb_None_dist[6,2] + rfperf_gm12878_10kb_None_dist[7,2])/2,
                                              rfperf_gm12878_25kb_None_dist[16,2],
                                              rfperf_gm12878_25kb_None_dist[15,2],
                                              (rfperf_gm12878_25kb_None_dist[6,2] + rfperf_gm12878_25kb_None_dist[7,2])/2,
                                              rfperf_gm12878_50kb_None_dist[16,2],
                                              rfperf_gm12878_50kb_None_dist[15,2],
                                              (rfperf_gm12878_50kb_None_dist[6,2] + rfperf_gm12878_50kb_None_dist[7,2])/2,
                                              rfperf_gm12878_100kb_None_dist[16,2],
                                              rfperf_gm12878_100kb_None_dist[15,2],
                                              (rfperf_gm12878_100kb_None_dist[6,2] + rfperf_gm12878_100kb_None_dist[7,2])/2,
                                              rfperf_k562_10kb_None_dist[16,2],
                                              rfperf_k562_10kb_None_dist[15,2],
                                              (rfperf_k562_10kb_None_dist[6,2] + rfperf_k562_10kb_None_dist[7,2])/2,
                                              rfperf_k562_25kb_None_dist[16,2],
                                              rfperf_k562_25kb_None_dist[15,2],
                                              (rfperf_k562_25kb_None_dist[6,2] + rfperf_k562_25kb_None_dist[7,2])/2,
                                              rfperf_k562_50kb_None_dist[16,2],
                                              rfperf_k562_50kb_None_dist[15,2],
                                              (rfperf_k562_50kb_None_dist[6,2] + rfperf_k562_50kb_None_dist[7,2])/2,
                                              rfperf_k562_100kb_None_dist[16,2],
                                              rfperf_k562_100kb_None_dist[15,2],
                                              (rfperf_k562_100kb_None_dist[6,2] + rfperf_k562_100kb_None_dist[7,2])/2),
                              CL = c(rep("GM12878", 12),
                                     rep("K562", 12)),
                              Resolution = rep(c(rep("10 kb", 3),
                                             rep("25 kb", 3),
                                             rep("50 kb", 3),
                                             rep("100 kb", 3)), 2),
                              Metric = rep(c("F1-Score", "MCC", "Balanced Acc"), 8))

all_None_perfs$CL <- factor(all_None_perfs$CL, levels=c("GM12878", "K562"))
all_None_perfs$Resolution <- factor(all_None_perfs$Resolution, levels=c("10 kb", "25 kb", "50 kb", "100 kb"))
all_None_perfs$Metric <- factor(all_None_perfs$Metric, levels=c("F1-Score", "MCC", "Balanced Acc"))

ggplot(all_None_perfs, aes(x=Resolution, y=Performance, group=CL, color=CL)) +
  geom_line(aes(linetype=CL))+
  geom_point(aes(shape=CL))+
  facet_grid(~ Metric)+
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Cell Line")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15
	                                 #angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   #hjust = 1
	                                 ),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))
```

##ROS

### GM12878

```{r}

###10kb
rfperf_gm12878_10kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

###25kb
rfperf_gm12878_25kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

###50kb
rfperf_gm12878_50kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

###100kb
rfperf_gm12878_100kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

```

### K562

```{r}

###10kb
rfperf_k562_10kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

###25kb
rfperf_k562_25kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

###50kb
rfperf_k562_50kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

###100kb
rfperf_k562_100kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

```

### Combined F1 and MCC and balanced acc

```{r}
all_ROS_perfs <- data.frame(Performance = c(rfperf_gm12878_10kb_ROS_dist[16,2],
                                              rfperf_gm12878_10kb_ROS_dist[15,2],
                                              (rfperf_gm12878_10kb_ROS_dist[6,2] + rfperf_gm12878_10kb_ROS_dist[7,2])/2,
                                              rfperf_gm12878_25kb_ROS_dist[16,2],
                                              rfperf_gm12878_25kb_ROS_dist[15,2],
                                              (rfperf_gm12878_25kb_ROS_dist[6,2] + rfperf_gm12878_25kb_ROS_dist[7,2])/2,
                                              rfperf_gm12878_50kb_ROS_dist[16,2],
                                              rfperf_gm12878_50kb_ROS_dist[15,2],
                                              (rfperf_gm12878_50kb_ROS_dist[6,2] + rfperf_gm12878_50kb_ROS_dist[7,2])/2,
                                              rfperf_gm12878_100kb_ROS_dist[16,2],
                                              rfperf_gm12878_100kb_ROS_dist[15,2],
                                              (rfperf_gm12878_100kb_ROS_dist[6,2] + rfperf_gm12878_100kb_ROS_dist[7,2])/2,
                                              rfperf_k562_10kb_ROS_dist[16,2],
                                              rfperf_k562_10kb_ROS_dist[15,2],
                                              (rfperf_k562_10kb_ROS_dist[6,2] + rfperf_k562_10kb_ROS_dist[7,2])/2,
                                              rfperf_k562_25kb_ROS_dist[16,2],
                                              rfperf_k562_25kb_ROS_dist[15,2],
                                              (rfperf_k562_25kb_ROS_dist[6,2] + rfperf_k562_25kb_ROS_dist[7,2])/2,
                                              rfperf_k562_50kb_ROS_dist[16,2],
                                              rfperf_k562_50kb_ROS_dist[15,2],
                                              (rfperf_k562_50kb_ROS_dist[6,2] + rfperf_k562_50kb_ROS_dist[7,2])/2,
                                              rfperf_k562_100kb_ROS_dist[16,2],
                                              rfperf_k562_100kb_ROS_dist[15,2],
                                              (rfperf_k562_100kb_ROS_dist[6,2] + rfperf_k562_100kb_ROS_dist[7,2])/2),
                              CL = c(rep("GM12878", 12),
                                     rep("K562", 12)),
                              Resolution = rep(c(rep("10 kb", 3),
                                             rep("25 kb", 3),
                                             rep("50 kb", 3),
                                             rep("100 kb", 3)), 2),
                              Metric = rep(c("F1-Score", "MCC", "Balanced Acc"), 8))

all_ROS_perfs$CL <- factor(all_ROS_perfs$CL, levels=c("GM12878", "K562"))
all_ROS_perfs$Resolution <- factor(all_ROS_perfs$Resolution, levels=c("10 kb", "25 kb", "50 kb", "100 kb"))
all_ROS_perfs$Metric <- factor(all_ROS_perfs$Metric, levels=c("F1-Score", "MCC", "Balanced Acc"))

ggplot(all_ROS_perfs, aes(x=Resolution, y=Performance, group=CL, color=CL)) +
  geom_line(aes(linetype=CL))+
  geom_point(aes(shape=CL))+
  facet_grid(~ Metric)+
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Cell Line")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15
	                                 #angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   #hjust = 1
	                                 ),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))
```

##RUS

### GM12878

```{r}

###10kb
rfperf_gm12878_10kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

###25kb
rfperf_gm12878_25kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

###50kb
rfperf_gm12878_50kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

###100kb
rfperf_gm12878_100kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

```

### K562

```{r}

###10kb
rfperf_k562_10kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

###25kb
rfperf_k562_25kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

###50kb
rfperf_k562_50kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

###100kb
rfperf_k562_100kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

```

### Combined F1 and MCC and balanced acc

```{r}
all_RUS_perfs <- data.frame(Performance = c(rfperf_gm12878_10kb_RUS_dist[16,2],
                                              rfperf_gm12878_10kb_RUS_dist[15,2],
                                              (rfperf_gm12878_10kb_RUS_dist[6,2] + rfperf_gm12878_10kb_RUS_dist[7,2])/2,
                                              rfperf_gm12878_25kb_RUS_dist[16,2],
                                              rfperf_gm12878_25kb_RUS_dist[15,2],
                                              (rfperf_gm12878_25kb_RUS_dist[6,2] + rfperf_gm12878_25kb_RUS_dist[7,2])/2,
                                              rfperf_gm12878_50kb_RUS_dist[16,2],
                                              rfperf_gm12878_50kb_RUS_dist[15,2],
                                              (rfperf_gm12878_50kb_RUS_dist[6,2] + rfperf_gm12878_50kb_RUS_dist[7,2])/2,
                                              rfperf_gm12878_100kb_RUS_dist[16,2],
                                              rfperf_gm12878_100kb_RUS_dist[15,2],
                                              (rfperf_gm12878_100kb_RUS_dist[6,2] + rfperf_gm12878_100kb_RUS_dist[7,2])/2,
                                              rfperf_k562_10kb_RUS_dist[16,2],
                                              rfperf_k562_10kb_RUS_dist[15,2],
                                              (rfperf_k562_10kb_RUS_dist[6,2] + rfperf_k562_10kb_RUS_dist[7,2])/2,
                                              rfperf_k562_25kb_RUS_dist[16,2],
                                              rfperf_k562_25kb_RUS_dist[15,2],
                                              (rfperf_k562_25kb_RUS_dist[6,2] + rfperf_k562_25kb_RUS_dist[7,2])/2,
                                              rfperf_k562_50kb_RUS_dist[16,2],
                                              rfperf_k562_50kb_RUS_dist[15,2],
                                              (rfperf_k562_50kb_RUS_dist[6,2] + rfperf_k562_50kb_RUS_dist[7,2])/2,
                                              rfperf_k562_100kb_RUS_dist[16,2],
                                              rfperf_k562_100kb_RUS_dist[15,2],
                                              (rfperf_k562_100kb_RUS_dist[6,2] + rfperf_k562_100kb_RUS_dist[7,2])/2),
                              CL = c(rep("GM12878", 12),
                                     rep("K562", 12)),
                              Resolution = rep(c(rep("10 kb", 3),
                                             rep("25 kb", 3),
                                             rep("50 kb", 3),
                                             rep("100 kb", 3)), 2),
                              Metric = rep(c("F1-Score", "MCC", "Balanced Acc"), 8))

all_RUS_perfs$CL <- factor(all_RUS_perfs$CL, levels=c("GM12878", "K562"))
all_RUS_perfs$Resolution <- factor(all_RUS_perfs$Resolution, levels=c("10 kb", "25 kb", "50 kb", "100 kb"))
all_RUS_perfs$Metric <- factor(all_RUS_perfs$Metric, levels=c("F1-Score", "MCC", "Balanced Acc"))

ggplot(all_RUS_perfs, aes(x=Resolution, y=Performance, group=CL, color=CL)) +
  geom_line(aes(linetype=CL))+
  geom_point(aes(shape=CL))+
  facet_grid(~ Metric)+
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Cell Line")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15
	                                 #angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   #hjust = 1
	                                 ),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))
```

```{r}
ggarrange(p.none, p.ros, p.rus, 
          ncol=1, nrow=3, 
          common.legend = TRUE, 
          legend="bottom",
          labels="AUTO",
          font.label = list(size=20))
```