### 2019-03-14 Scripts/evaluating_pca.Rmd
script for evaluating pca results

---
title: "Performing Cluster Analysis on TAD boundaries & Annotations"
author: "Spiro Stilianoudakis"
date: "March 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(factoextra)
library(FactoMineR)

```

# Boundaries as rows; Annotations as columns; clustering annotations

## GM12878

### 10kb

#### Overlap counts data

```{r}
pcadat_gm12878_10kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_count.rds")

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_10kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_10kb_count_f <- pcadat_gm12878_10kb_count[, -which(colnames(pcadat_gm12878_10kb_count) %in% nzvar)]

dim(pcadat_gm12878_10kb_count) 
dim(pcadat_gm12878_10kb_count_f) 
```

#### Distance data

```{r}
pcadat_gm12878_10kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_dist.rds")
```