### 2019-03-15 Scripts/evaluating_pca.Rmd
color coded pca plots by annotation group

---
title: "Performing Cluster Analysis on TAD boundaries & Annotations"
author: "Spiro Stilianoudakis"
date: "March 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(factoextra)
library(FactoMineR)

```

# Clustering Annotations

## GM12878

### 10kb

#### Overlap counts data

```{r}
pcadat_gm12878_10kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_count.rds")
dim(pcadat_gm12878_10kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_10kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_10kb_count_f <- pcadat_gm12878_10kb_count[, -which(colnames(pcadat_gm12878_10kb_count) %in% nzvar)]

dim(pcadat_gm12878_10kb_count) 
dim(pcadat_gm12878_10kb_count_f)

colnames(pcadat_gm12878_10kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_10kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_10kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_10kb_count_f),
          #rowsep=1:nrow(pcadat_gm12878_10kb_count_f),
          keysize = 1)

##create annotation groups
cs <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/BroadHMM")))
hm <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")))

#PCA
pcadat_gm12878_10kb_count_f <- t(pcadat_gm12878_10kb_count_f)

Group <- ifelse(rownames(pcadat_gm12878_10kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_10kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_10kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_10kb_count_f <- cbind.data.frame(pcadat_gm12878_10kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_f[,-(dim(pcadat_gm12878_10kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_f[,-(dim(pcadat_gm12878_10kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count_f[,-(dim(pcadat_gm12878_10kb_count_f2)[2])], 4),
         data=pcadat_gm12878_10kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_gm12878_10kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_perc.rds")
dim(pcadat_gm12878_10kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_10kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_10kb_perc_f <- pcadat_gm12878_10kb_perc[, -which(colnames(pcadat_gm12878_10kb_perc) %in% nzvar)]

dim(pcadat_gm12878_10kb_perc) 
dim(pcadat_gm12878_10kb_perc_f)

colnames(pcadat_gm12878_10kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_10kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_10kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_10kb_perc_f),
          #rowsep=1:nrow(pcadat_gm12878_10kb_perc_f),
          keysize = 1)

#PCA
pcadat_gm12878_10kb_perc_f <- t(pcadat_gm12878_10kb_perc_f)

Group <- ifelse(rownames(pcadat_gm12878_10kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_10kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_10kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_10kb_perc_f <- cbind.data.frame(pcadat_gm12878_10kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_f[,-(dim(pcadat_gm12878_10kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_f[,-(dim(pcadat_gm12878_10kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_perc_f[,-(dim(pcadat_gm12878_10kb_perc_f2)[2])], 4),
         data=pcadat_gm12878_10kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

#### Distance data

```{r}
pcadat_gm12878_10kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_dist.rds")
dim(pcadat_gm12878_10kb_dist)

#performing log2 transformation
pcadat_gm12878_10kb_dist <- apply(pcadat_gm12878_10kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_gm12878_10kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_gm12878_10kb_dist_f <- pcadat_gm12878_10kb_dist[, -which(colnames(pcadat_gm12878_10kb_dist) %in% nzvar)]
pcadat_gm12878_10kb_dist_f <- pcadat_gm12878_10kb_dist

dim(pcadat_gm12878_10kb_dist) 
dim(pcadat_gm12878_10kb_dist_f)

colnames(pcadat_gm12878_10kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_10kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_10kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_10kb_dist_f),
          #rowsep=1:nrow(pcadat_gm12878_10kb_dist_f),
          keysize = 1)

#PCA
pcadat_gm12878_10kb_dist_f <- t(pcadat_gm12878_10kb_dist_f)

Group <- ifelse(rownames(pcadat_gm12878_10kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_10kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_10kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_10kb_dist_f <- cbind.data.frame(pcadat_gm12878_10kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_f[,-(dim(pcadat_gm12878_10kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_f[,-(dim(pcadat_gm12878_10kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_10kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_dist_f[,-(dim(pcadat_gm12878_10kb_dist_f2)[2])], 4),
         data=pcadat_gm12878_10kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

### 25kb

#### Overlap counts data

```{r}
pcadat_gm12878_25kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_count.rds")
dim(pcadat_gm12878_25kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_25kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_25kb_count_f <- pcadat_gm12878_25kb_count[, -which(colnames(pcadat_gm12878_25kb_count) %in% nzvar)]

dim(pcadat_gm12878_25kb_count) 
dim(pcadat_gm12878_25kb_count_f)

colnames(pcadat_gm12878_25kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_25kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_25kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_25kb_count_f),
          #rowsep=1:nrow(pcadat_gm12878_25kb_count_f),
          keysize = 1)

#PCA
pcadat_gm12878_25kb_count_f <- t(pcadat_gm12878_25kb_count_f)

Group <- ifelse(rownames(pcadat_gm12878_25kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_25kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_25kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_25kb_count_f <- cbind.data.frame(pcadat_gm12878_25kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_f[,-(dim(pcadat_gm12878_25kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_f[,-(dim(pcadat_gm12878_25kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_count_f[,-(dim(pcadat_gm12878_25kb_count_f2)[2])], 4),
         data=pcadat_gm12878_25kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_gm12878_25kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_perc.rds")
dim(pcadat_gm12878_25kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_25kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_25kb_perc_f <- pcadat_gm12878_25kb_perc[, -which(colnames(pcadat_gm12878_25kb_perc) %in% nzvar)]

dim(pcadat_gm12878_25kb_perc) 
dim(pcadat_gm12878_25kb_perc_f)

colnames(pcadat_gm12878_25kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_25kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_25kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_25kb_perc_f),
          #rowsep=1:nrow(pcadat_gm12878_25kb_perc_f),
          keysize = 1)

#PCA
pcadat_gm12878_25kb_perc_f <- t(pcadat_gm12878_25kb_perc_f)

Group <- ifelse(rownames(pcadat_gm12878_25kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_25kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_25kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_25kb_perc_f <- cbind.data.frame(pcadat_gm12878_25kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_f[,-(dim(pcadat_gm12878_25kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_f[,-(dim(pcadat_gm12878_25kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_perc_f[,-(dim(pcadat_gm12878_25kb_perc_f2)[2])], 4),
         data=pcadat_gm12878_25kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Distance data

```{r}
pcadat_gm12878_25kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_dist.rds")
dim(pcadat_gm12878_25kb_dist)

#performing log2 transformation
pcadat_gm12878_25kb_dist <- apply(pcadat_gm12878_25kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_gm12878_25kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_gm12878_25kb_dist_f <- pcadat_gm12878_25kb_dist[, -which(colnames(pcadat_gm12878_25kb_dist) %in% nzvar)]
pcadat_gm12878_25kb_dist_f <- pcadat_gm12878_25kb_dist

dim(pcadat_gm12878_25kb_dist) 
dim(pcadat_gm12878_25kb_dist_f)

colnames(pcadat_gm12878_25kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_25kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_25kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_25kb_dist_f),
          #rowsep=1:nrow(pcadat_gm12878_25kb_dist_f),
          keysize = 1)

#PCA
pcadat_gm12878_25kb_dist_f <- t(pcadat_gm12878_25kb_dist_f)

Group <- ifelse(rownames(pcadat_gm12878_25kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_25kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_25kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_25kb_dist_f <- cbind.data.frame(pcadat_gm12878_25kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_f[,-(dim(pcadat_gm12878_25kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_f[,-(dim(pcadat_gm12878_25kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_25kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_dist_f[,-(dim(pcadat_gm12878_25kb_dist_f2)[2])], 4),
         data=pcadat_gm12878_25kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

### 50kb

#### Overlap counts data

```{r}
pcadat_gm12878_50kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/PCA/pcadat_gm12878_50kb_count.rds")
dim(pcadat_gm12878_50kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_50kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_50kb_count_f <- pcadat_gm12878_50kb_count[, -which(colnames(pcadat_gm12878_50kb_count) %in% nzvar)]

dim(pcadat_gm12878_50kb_count) 
dim(pcadat_gm12878_50kb_count_f)

colnames(pcadat_gm12878_50kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_50kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_50kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_50kb_count_f),
          #rowsep=1:nrow(pcadat_gm12878_50kb_count_f),
          keysize = 1)

#PCA
pcadat_gm12878_50kb_count_f <- t(pcadat_gm12878_50kb_count_f)

Group <- ifelse(rownames(pcadat_gm12878_50kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_50kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_50kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_50kb_count_f <- cbind.data.frame(pcadat_gm12878_50kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_f[,-(dim(pcadat_gm12878_50kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_50kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_f[,-(dim(pcadat_gm12878_50kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_50kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_count_f[,-(dim(pcadat_gm12878_50kb_count_f2)[2])], 4),
         data=pcadat_gm12878_50kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_gm12878_50kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/PCA/pcadat_gm12878_50kb_perc.rds")
dim(pcadat_gm12878_50kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_50kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_50kb_perc_f <- pcadat_gm12878_50kb_perc[, -which(colnames(pcadat_gm12878_50kb_perc) %in% nzvar)]

dim(pcadat_gm12878_50kb_perc) 
dim(pcadat_gm12878_50kb_perc_f)

colnames(pcadat_gm12878_50kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_50kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_50kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_50kb_perc_f),
          #rowsep=1:nrow(pcadat_gm12878_50kb_perc_f),
          keysize = 1)

#PCA
pcadat_gm12878_50kb_perc_f <- t(pcadat_gm12878_50kb_perc_f)

Group <- ifelse(rownames(pcadat_gm12878_50kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_50kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_50kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_50kb_perc_f <- cbind.data.frame(pcadat_gm12878_50kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_perc_f[,-(dim(pcadat_gm12878_50kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_50kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_perc_f[,-(dim(pcadat_gm12878_50kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_50kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_perc_f[,-(dim(pcadat_gm12878_50kb_perc_f2)[2])], 4),
         data=pcadat_gm12878_50kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

#### Distance data

```{r}
pcadat_gm12878_50kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/PCA/pcadat_gm12878_50kb_dist.rds")
dim(pcadat_gm12878_50kb_dist)

#performing log2 transformation
pcadat_gm12878_50kb_dist <- apply(pcadat_gm12878_50kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_gm12878_50kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_gm12878_50kb_dist_f <- pcadat_gm12878_50kb_dist[, -which(colnames(pcadat_gm12878_50kb_dist) %in% nzvar)]
pcadat_gm12878_50kb_dist_f <- pcadat_gm12878_50kb_dist

dim(pcadat_gm12878_50kb_dist) 
dim(pcadat_gm12878_50kb_dist_f)

colnames(pcadat_gm12878_50kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_50kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_50kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_50kb_dist_f),
          #rowsep=1:nrow(pcadat_gm12878_50kb_dist_f),
          keysize = 1)

#PCA
pcadat_gm12878_50kb_dist_f <- t(pcadat_gm12878_50kb_dist_f)

Group <- ifelse(rownames(pcadat_gm12878_50kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_50kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_50kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_50kb_dist_f <- cbind.data.frame(pcadat_gm12878_50kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_dist_f[,-(dim(pcadat_gm12878_50kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_50kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_dist_f[,-(dim(pcadat_gm12878_50kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_50kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_dist_f[,-(dim(pcadat_gm12878_50kb_dist_f2)[2])], 4),
         data=pcadat_gm12878_50kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

### 100kb

#### Overlap counts data

```{r}
pcadat_gm12878_100kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/PCA/pcadat_gm12878_100kb_count.rds")
dim(pcadat_gm12878_100kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_100kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_100kb_count_f <- pcadat_gm12878_100kb_count[, -which(colnames(pcadat_gm12878_100kb_count) %in% nzvar)]

dim(pcadat_gm12878_100kb_count) 
dim(pcadat_gm12878_100kb_count_f)

colnames(pcadat_gm12878_100kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_100kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_100kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_100kb_count_f),
          #rowsep=1:nrow(pcadat_gm12878_100kb_count_f),
          keysize = 1)

#PCA
pcadat_gm12878_100kb_count_f <- t(pcadat_gm12878_100kb_count_f)

Group <- ifelse(rownames(pcadat_gm12878_100kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_100kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_100kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_100kb_count_f <- cbind.data.frame(pcadat_gm12878_100kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_100kb_count_f[,-(dim(pcadat_gm12878_100kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_100kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_100kb_count_f[,-(dim(pcadat_gm12878_100kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_100kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_100kb_count_f[,-(dim(pcadat_gm12878_100kb_count_f2)[2])], 4),
         data=pcadat_gm12878_100kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_gm12878_100kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/PCA/pcadat_gm12878_100kb_perc.rds")
dim(pcadat_gm12878_100kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_gm12878_100kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_gm12878_100kb_perc_f <- pcadat_gm12878_100kb_perc[, -which(colnames(pcadat_gm12878_100kb_perc) %in% nzvar)]

dim(pcadat_gm12878_100kb_perc) 
dim(pcadat_gm12878_100kb_perc_f)

colnames(pcadat_gm12878_100kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_100kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_100kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_100kb_perc_f),
          #rowsep=1:nrow(pcadat_gm12878_100kb_perc_f),
          keysize = 1)

#PCA
pcadat_gm12878_100kb_perc_f <- t(pcadat_gm12878_100kb_perc_f)

Group <- ifelse(rownames(pcadat_gm12878_100kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_100kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_100kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_100kb_perc_f <- cbind.data.frame(pcadat_gm12878_100kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_100kb_perc_f[,-(dim(pcadat_gm12878_100kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_100kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_100kb_perc_f[,-(dim(pcadat_gm12878_100kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_100kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_100kb_perc_f[,-(dim(pcadat_gm12878_100kb_perc_f2)[2])], 4),
         data=pcadat_gm12878_100kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Distance data

```{r}
pcadat_gm12878_100kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/PCA/pcadat_gm12878_100kb_dist.rds")
dim(pcadat_gm12878_100kb_dist)

#performing log2 transformation
pcadat_gm12878_100kb_dist <- apply(pcadat_gm12878_100kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_gm12878_100kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_gm12878_100kb_dist_f <- pcadat_gm12878_100kb_dist[, -which(colnames(pcadat_gm12878_100kb_dist) %in% nzvar)]
pcadat_gm12878_100kb_dist_f <- pcadat_gm12878_100kb_dist

dim(pcadat_gm12878_100kb_dist) 
dim(pcadat_gm12878_100kb_dist_f)

colnames(pcadat_gm12878_100kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_gm12878_100kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_gm12878_100kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_gm12878_100kb_dist_f),
          #rowsep=1:nrow(pcadat_gm12878_100kb_dist_f),
          keysize = 1)

#PCA
pcadat_gm12878_100kb_dist_f <- t(pcadat_gm12878_100kb_dist_f)

Group <- ifelse(rownames(pcadat_gm12878_100kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_gm12878_100kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_gm12878_100kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_gm12878_100kb_dist_f <- cbind.data.frame(pcadat_gm12878_100kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_gm12878_100kb_dist_f[,-(dim(pcadat_gm12878_100kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_100kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_100kb_dist_f[,-(dim(pcadat_gm12878_100kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_gm12878_100kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_gm12878_100kb_dist_f[,-(dim(pcadat_gm12878_100kb_dist_f2)[2])], 4),
         data=pcadat_gm12878_100kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

## K562

### 10kb

#### Overlap counts data

```{r}
pcadat_K562_10kb_count <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/PCA/pcadat_K562_10kb_count.rds")
dim(pcadat_K562_10kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_10kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_10kb_count_f <- pcadat_K562_10kb_count[, -which(colnames(pcadat_K562_10kb_count) %in% nzvar)]

dim(pcadat_K562_10kb_count) 
dim(pcadat_K562_10kb_count_f)

colnames(pcadat_K562_10kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_10kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_10kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_10kb_count_f),
          #rowsep=1:nrow(pcadat_K562_10kb_count_f),
          keysize = 1)

##create annotation groups
cs <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/BroadHMM")))
hm <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")))

#PCA
pcadat_K562_10kb_count_f <- t(pcadat_K562_10kb_count_f)

Group <- ifelse(rownames(pcadat_K562_10kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_10kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_10kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_10kb_count_f <- cbind.data.frame(pcadat_K562_10kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_10kb_count_f[,-(dim(pcadat_K562_10kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_10kb_count_f[,-(dim(pcadat_K562_10kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_10kb_count_f[,-(dim(pcadat_K562_10kb_count_f2)[2])], 4),
         data=pcadat_K562_10kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_K562_10kb_perc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/PCA/pcadat_K562_10kb_perc.rds")
dim(pcadat_K562_10kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_10kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_10kb_perc_f <- pcadat_K562_10kb_perc[, -which(colnames(pcadat_K562_10kb_perc) %in% nzvar)]

dim(pcadat_K562_10kb_perc) 
dim(pcadat_K562_10kb_perc_f)

colnames(pcadat_K562_10kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_10kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_10kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_10kb_perc_f),
          #rowsep=1:nrow(pcadat_K562_10kb_perc_f),
          keysize = 1)

#PCA
pcadat_K562_10kb_perc_f <- t(pcadat_K562_10kb_perc_f)

Group <- ifelse(rownames(pcadat_K562_10kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_10kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_10kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_10kb_perc_f <- cbind.data.frame(pcadat_K562_10kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_10kb_perc_f[,-(dim(pcadat_K562_10kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_10kb_perc_f[,-(dim(pcadat_K562_10kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_10kb_perc_f[,-(dim(pcadat_K562_10kb_perc_f2)[2])], 4),
         data=pcadat_K562_10kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

#### Distance data

```{r}
pcadat_K562_10kb_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/PCA/pcadat_K562_10kb_dist.rds")
dim(pcadat_K562_10kb_dist)

#performing log2 transformation
pcadat_K562_10kb_dist <- apply(pcadat_K562_10kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_K562_10kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_K562_10kb_dist_f <- pcadat_K562_10kb_dist[, -which(colnames(pcadat_K562_10kb_dist) %in% nzvar)]
pcadat_K562_10kb_dist_f <- pcadat_K562_10kb_dist

dim(pcadat_K562_10kb_dist) 
dim(pcadat_K562_10kb_dist_f)

colnames(pcadat_K562_10kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_10kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_10kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_10kb_dist_f),
          #rowsep=1:nrow(pcadat_K562_10kb_dist_f),
          keysize = 1)

#PCA
pcadat_K562_10kb_dist_f <- t(pcadat_K562_10kb_dist_f)

Group <- ifelse(rownames(pcadat_K562_10kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_10kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_10kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_10kb_dist_f <- cbind.data.frame(pcadat_K562_10kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_10kb_dist_f[,-(dim(pcadat_K562_10kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_10kb_dist_f[,-(dim(pcadat_K562_10kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_10kb_dist_f[,-(dim(pcadat_K562_10kb_dist_f2)[2])], 4),
         data=pcadat_K562_10kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

### 25kb

#### Overlap counts data

```{r}
pcadat_K562_25kb_count <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/PCA/pcadat_K562_25kb_count.rds")
dim(pcadat_K562_25kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_25kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_25kb_count_f <- pcadat_K562_25kb_count[, -which(colnames(pcadat_K562_25kb_count) %in% nzvar)]

dim(pcadat_K562_25kb_count) 
dim(pcadat_K562_25kb_count_f)

colnames(pcadat_K562_25kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_25kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_25kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_25kb_count_f),
          #rowsep=1:nrow(pcadat_K562_25kb_count_f),
          keysize = 1)

#PCA
pcadat_K562_25kb_count_f <- t(pcadat_K562_25kb_count_f)

Group <- ifelse(rownames(pcadat_K562_25kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_25kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_25kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_25kb_count_f <- cbind.data.frame(pcadat_K562_25kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_25kb_count_f[,-(dim(pcadat_K562_25kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_25kb_count_f[,-(dim(pcadat_K562_25kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_25kb_count_f[,-(dim(pcadat_K562_25kb_count_f2)[2])], 4),
         data=pcadat_K562_25kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_K562_25kb_perc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/PCA/pcadat_K562_25kb_perc.rds")
dim(pcadat_K562_25kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_25kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_25kb_perc_f <- pcadat_K562_25kb_perc[, -which(colnames(pcadat_K562_25kb_perc) %in% nzvar)]

dim(pcadat_K562_25kb_perc) 
dim(pcadat_K562_25kb_perc_f)

colnames(pcadat_K562_25kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_25kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_25kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_25kb_perc_f),
          #rowsep=1:nrow(pcadat_K562_25kb_perc_f),
          keysize = 1)

#PCA
pcadat_K562_25kb_perc_f <- t(pcadat_K562_25kb_perc_f)

Group <- ifelse(rownames(pcadat_K562_25kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_25kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_25kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_25kb_perc_f <- cbind.data.frame(pcadat_K562_25kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_25kb_perc_f[,-(dim(pcadat_K562_25kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_25kb_perc_f[,-(dim(pcadat_K562_25kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_25kb_perc_f[,-(dim(pcadat_K562_25kb_perc_f2)[2])], 4),
         data=pcadat_K562_25kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Distance data

```{r}
pcadat_K562_25kb_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/PCA/pcadat_K562_25kb_dist.rds")
dim(pcadat_K562_25kb_dist)

#performing log2 transformation
pcadat_K562_25kb_dist <- apply(pcadat_K562_25kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_K562_25kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_K562_25kb_dist_f <- pcadat_K562_25kb_dist[, -which(colnames(pcadat_K562_25kb_dist) %in% nzvar)]
pcadat_K562_25kb_dist_f <- pcadat_K562_25kb_dist

dim(pcadat_K562_25kb_dist) 
dim(pcadat_K562_25kb_dist_f)

colnames(pcadat_K562_25kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_25kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_25kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_25kb_dist_f),
          #rowsep=1:nrow(pcadat_K562_25kb_dist_f),
          keysize = 1)

#PCA
pcadat_K562_25kb_dist_f <- t(pcadat_K562_25kb_dist_f)

Group <- ifelse(rownames(pcadat_K562_25kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_25kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_25kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_25kb_dist_f <- cbind.data.frame(pcadat_K562_25kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_25kb_dist_f[,-(dim(pcadat_K562_25kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_25kb_dist_f[,-(dim(pcadat_K562_25kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_25kb_dist_f[,-(dim(pcadat_K562_25kb_dist_f2)[2])], 4),
         data=pcadat_K562_25kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

### 50kb

#### Overlap counts data

```{r}
pcadat_K562_50kb_count <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/PCA/pcadat_K562_50kb_count.rds")
dim(pcadat_K562_50kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_50kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_50kb_count_f <- pcadat_K562_50kb_count[, -which(colnames(pcadat_K562_50kb_count) %in% nzvar)]

dim(pcadat_K562_50kb_count) 
dim(pcadat_K562_50kb_count_f)

colnames(pcadat_K562_50kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_50kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_50kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_50kb_count_f),
          #rowsep=1:nrow(pcadat_K562_50kb_count_f),
          keysize = 1)

#PCA
pcadat_K562_50kb_count_f <- t(pcadat_K562_50kb_count_f)

Group <- ifelse(rownames(pcadat_K562_50kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_50kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_50kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_50kb_count_f <- cbind.data.frame(pcadat_K562_50kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_50kb_count_f[,-(dim(pcadat_K562_50kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_50kb_count_f[,-(dim(pcadat_K562_50kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_50kb_count_f[,-(dim(pcadat_K562_50kb_count_f2)[2])], 4),
         data=pcadat_K562_50kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_K562_50kb_perc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/PCA/pcadat_K562_50kb_perc.rds")
dim(pcadat_K562_50kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_50kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_50kb_perc_f <- pcadat_K562_50kb_perc[, -which(colnames(pcadat_K562_50kb_perc) %in% nzvar)]

dim(pcadat_K562_50kb_perc) 
dim(pcadat_K562_50kb_perc_f)

colnames(pcadat_K562_50kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_50kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_50kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_50kb_perc_f),
          #rowsep=1:nrow(pcadat_K562_50kb_perc_f),
          keysize = 1)

#PCA
pcadat_K562_50kb_perc_f <- t(pcadat_K562_50kb_perc_f)

Group <- ifelse(rownames(pcadat_K562_50kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_50kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_50kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_50kb_perc_f <- cbind.data.frame(pcadat_K562_50kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_50kb_perc_f[,-(dim(pcadat_K562_50kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_50kb_perc_f[,-(dim(pcadat_K562_50kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_50kb_perc_f[,-(dim(pcadat_K562_50kb_perc_f2)[2])], 4),
         data=pcadat_K562_50kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

#### Distance data

```{r}
pcadat_K562_50kb_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/PCA/pcadat_K562_50kb_dist.rds")
dim(pcadat_K562_50kb_dist)

#performing log2 transformation
pcadat_K562_50kb_dist <- apply(pcadat_K562_50kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_K562_50kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_K562_50kb_dist_f <- pcadat_K562_50kb_dist[, -which(colnames(pcadat_K562_50kb_dist) %in% nzvar)]
pcadat_K562_50kb_dist_f <- pcadat_K562_50kb_dist

dim(pcadat_K562_50kb_dist) 
dim(pcadat_K562_50kb_dist_f)

colnames(pcadat_K562_50kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_50kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_50kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_50kb_dist_f),
          #rowsep=1:nrow(pcadat_K562_50kb_dist_f),
          keysize = 1)

#PCA
pcadat_K562_50kb_dist_f <- t(pcadat_K562_50kb_dist_f)

Group <- ifelse(rownames(pcadat_K562_50kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_50kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_50kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_50kb_dist_f <- cbind.data.frame(pcadat_K562_50kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_50kb_dist_f[,-(dim(pcadat_K562_50kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_50kb_dist_f[,-(dim(pcadat_K562_50kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_50kb_dist_f[,-(dim(pcadat_K562_50kb_dist_f2)[2])], 4),
         data=pcadat_K562_50kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

### 100kb

#### Overlap counts data

```{r}
pcadat_K562_100kb_count <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/PCA/pcadat_K562_100kb_count.rds")
dim(pcadat_K562_100kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_100kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_100kb_count_f <- pcadat_K562_100kb_count[, -which(colnames(pcadat_K562_100kb_count) %in% nzvar)]

dim(pcadat_K562_100kb_count) 
dim(pcadat_K562_100kb_count_f)

colnames(pcadat_K562_100kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_100kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_100kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_100kb_count_f),
          #rowsep=1:nrow(pcadat_K562_100kb_count_f),
          keysize = 1)

#PCA
pcadat_K562_100kb_count_f <- t(pcadat_K562_100kb_count_f)

Group <- ifelse(rownames(pcadat_K562_100kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_100kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_100kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_100kb_count_f <- cbind.data.frame(pcadat_K562_100kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_100kb_count_f[,-(dim(pcadat_K562_100kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_100kb_count_f[,-(dim(pcadat_K562_100kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_100kb_count_f[,-(dim(pcadat_K562_100kb_count_f2)[2])], 4),
         data=pcadat_K562_100kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_K562_100kb_perc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/PCA/pcadat_K562_100kb_perc.rds")
dim(pcadat_K562_100kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_100kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_100kb_perc_f <- pcadat_K562_100kb_perc[, -which(colnames(pcadat_K562_100kb_perc) %in% nzvar)]

dim(pcadat_K562_100kb_perc) 
dim(pcadat_K562_100kb_perc_f)

colnames(pcadat_K562_100kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_100kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_100kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_100kb_perc_f),
          #rowsep=1:nrow(pcadat_K562_100kb_perc_f),
          keysize = 1)

#PCA
pcadat_K562_100kb_perc_f <- t(pcadat_K562_100kb_perc_f)

Group <- ifelse(rownames(pcadat_K562_100kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_100kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_100kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_100kb_perc_f <- cbind.data.frame(pcadat_K562_100kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_100kb_perc_f[,-(dim(pcadat_K562_100kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_100kb_perc_f[,-(dim(pcadat_K562_100kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_100kb_perc_f[,-(dim(pcadat_K562_100kb_perc_f2)[2])], 4),
         data=pcadat_K562_100kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Distance data

```{r}
pcadat_K562_100kb_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/PCA/pcadat_K562_100kb_dist.rds")
dim(pcadat_K562_100kb_dist)

#performing log2 transformation
pcadat_K562_100kb_dist <- apply(pcadat_K562_100kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_K562_100kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_K562_100kb_dist_f <- pcadat_K562_100kb_dist[, -which(colnames(pcadat_K562_100kb_dist) %in% nzvar)]
pcadat_K562_100kb_dist_f <- pcadat_K562_100kb_dist

dim(pcadat_K562_100kb_dist) 
dim(pcadat_K562_100kb_dist_f)

colnames(pcadat_K562_100kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_100kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_100kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_100kb_dist_f),
          #rowsep=1:nrow(pcadat_K562_100kb_dist_f),
          keysize = 1)

#PCA
pcadat_K562_100kb_dist_f <- t(pcadat_K562_100kb_dist_f)

Group <- ifelse(rownames(pcadat_K562_100kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_100kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_100kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_100kb_dist_f <- cbind.data.frame(pcadat_K562_100kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_100kb_dist_f[,-(dim(pcadat_K562_100kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_100kb_dist_f[,-(dim(pcadat_K562_100kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_100kb_dist_f[,-(dim(pcadat_K562_100kb_dist_f2)[2])], 4),
         data=pcadat_K562_100kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

## Saving plots

```{r}

```