### 2019-03-22 Scripts/evaluating_pca.Rmd
editted pca script

---
title: "Performing Cluster Analysis on TAD boundaries & Annotations"
author: "Spiro Stilianoudakis"
date: "March 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(factoextra)
library(FactoMineR)
library(ggfortify)
```

# Clustering Annotations

## GM12878

### 10kb

```{r}
pcadat_gm12878_10kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_count.rds")
pcadat_gm12878_10kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_perc.rds")
pcadat_gm12878_10kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/PCA/pcadat_gm12878_10kb_dist.rds")

#performing log2 transformation on distances
pcadat_gm12878_10kb_dist <- apply(pcadat_gm12878_10kb_dist,2,function(x){log(x + 1, base=2)})

cleancolnames <- function(x){gsub(paste(c("`", 
                                          "_",
                                          #"-",
                                          #"Gm12878", 
                                          #"Haib", 
                                          #"Sydh",
                                          #"Uta",
                                          #"Uw",
                                          #"Broad",
                                          #"Uchicago",
                                          "dist",
                                          "count",
                                          "perc"), collapse = "|"),"", 
                                  colnames(x))}

colnames(pcadat_gm12878_10kb_count) <- cleancolnames(pcadat_gm12878_10kb_count)
colnames(pcadat_gm12878_10kb_perc) <- cleancolnames(pcadat_gm12878_10kb_perc)
colnames(pcadat_gm12878_10kb_dist) <- cleancolnames(pcadat_gm12878_10kb_dist)

pcadat_gm12878_10kb_count <- t(pcadat_gm12878_10kb_count)
pcadat_gm12878_10kb_perc <- t(pcadat_gm12878_10kb_perc)
pcadat_gm12878_10kb_dist <- t(pcadat_gm12878_10kb_dist)

#pcadat_gm12878_10kb_count <- data.frame(pcadat_gm12878_10kb_count)
#pcadat_gm12878_10kb_perc <- data.frame(pcadat_gm12878_10kb_perc)
#pcadat_gm12878_10kb_dist <- data.frame(pcadat_gm12878_10kb_dist)

#pcadat_gm12878_10kb_count$Type <- "Counts"
#pcadat_gm12878_10kb_perc$Type <- "Percent"
#pcadat_gm12878_10kb_dist$Type <- "Distance"

tfbsvars <-  gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878"))

```

#### Chromatin States

```{r}
chromstatevars <- gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/BroadHMM"))
```

##### Overlap Counts

```{r}
pcadat_gm12878_10kb_count_chromstates <- pcadat_gm12878_10kb_count[which(rownames(pcadat_gm12878_10kb_count) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count_chromstates)-1)*sum(apply(pcadat_gm12878_10kb_count_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count_chromstates, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_count_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

##### Overlap Percents

```{r}
pcadat_gm12878_10kb_perc_chromstates <- pcadat_gm12878_10kb_perc[which(rownames(pcadat_gm12878_10kb_perc) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_perc_chromstates)-1)*sum(apply(pcadat_gm12878_10kb_perc_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_perc_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_perc_chromstates, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_perc_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

#####Distance

```{r}
pcadat_gm12878_10kb_dist_chromstates <- pcadat_gm12878_10kb_dist[which(rownames(pcadat_gm12878_10kb_dist) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_dist_chromstates)-1)*sum(apply(pcadat_gm12878_10kb_dist_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_dist_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_dist_chromstates, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_dist_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

#### Histone Modifications

```{r}
histonevars <- gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/HistoneModifications"))
```

##### Overlap Counts

```{r}
pcadat_gm12878_10kb_count_histones <- pcadat_gm12878_10kb_count[which(rownames(pcadat_gm12878_10kb_count) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count_histones)-1)*sum(apply(pcadat_gm12878_10kb_count_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count_histones, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_count_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

##### Overlap Percents

```{r}
pcadat_gm12878_10kb_perc_histones <- pcadat_gm12878_10kb_perc[which(rownames(pcadat_gm12878_10kb_perc) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_perc_histones)-1)*sum(apply(pcadat_gm12878_10kb_perc_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_perc_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_perc_histones, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_perc_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

#####Distance

```{r}
pcadat_gm12878_10kb_dist_histones <- pcadat_gm12878_10kb_dist[which(rownames(pcadat_gm12878_10kb_dist) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_dist_histones)-1)*sum(apply(pcadat_gm12878_10kb_dist_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_dist_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_dist_histones, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_dist_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

#### TFBS

```{r}
tfbsvars <-  gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878"))
```

##### Overlap Counts

```{r}
pcadat_gm12878_10kb_count_tfbs <- pcadat_gm12878_10kb_count[which(rownames(pcadat_gm12878_10kb_count) %in% tfbsvars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_count_tfbs,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_count_tfbs, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_count_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

##### Overlap Percents

```{r}
pcadat_gm12878_10kb_perc_tfbs <- pcadat_gm12878_10kb_perc[which(rownames(pcadat_gm12878_10kb_perc) %in% tfbsvars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_perc_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_perc_tfbs,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_perc_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_perc_tfbs, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_perc_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

#####Distance

```{r}
pcadat_gm12878_10kb_dist_tfbs <- pcadat_gm12878_10kb_dist[which(rownames(pcadat_gm12878_10kb_dist) %in% tfbsvars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_dist_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_dist_tfbs,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_dist_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_dist_tfbs, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_dist_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```

## K562

### 10kb

#### Overlap counts data

```{r}
pcadat_K562_10kb_count <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/PCA/pcadat_K562_10kb_count.rds")
dim(pcadat_K562_10kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_10kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_10kb_count_f <- pcadat_K562_10kb_count[, -which(colnames(pcadat_K562_10kb_count) %in% nzvar)]

dim(pcadat_K562_10kb_count) 
dim(pcadat_K562_10kb_count_f)

colnames(pcadat_K562_10kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_10kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_10kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_10kb_count_f),
          #rowsep=1:nrow(pcadat_K562_10kb_count_f),
          keysize = 1)

##create annotation groups
cs <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/BroadHMM")))
hm <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")))

#PCA
pcadat_K562_10kb_count_f <- t(pcadat_K562_10kb_count_f)

Group <- ifelse(rownames(pcadat_K562_10kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_10kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_10kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_10kb_count_f <- cbind.data.frame(pcadat_K562_10kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_10kb_count_f[,-(dim(pcadat_K562_10kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_10kb_count_f[,-(dim(pcadat_K562_10kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_10kb_count_f[,-(dim(pcadat_K562_10kb_count_f2)[2])], 4),
         data=pcadat_K562_10kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_K562_10kb_perc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/PCA/pcadat_K562_10kb_perc.rds")
dim(pcadat_K562_10kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_10kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_10kb_perc_f <- pcadat_K562_10kb_perc[, -which(colnames(pcadat_K562_10kb_perc) %in% nzvar)]

dim(pcadat_K562_10kb_perc) 
dim(pcadat_K562_10kb_perc_f)

colnames(pcadat_K562_10kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_10kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_10kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_10kb_perc_f),
          #rowsep=1:nrow(pcadat_K562_10kb_perc_f),
          keysize = 1)

#PCA
pcadat_K562_10kb_perc_f <- t(pcadat_K562_10kb_perc_f)

Group <- ifelse(rownames(pcadat_K562_10kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_10kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_10kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_10kb_perc_f <- cbind.data.frame(pcadat_K562_10kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_10kb_perc_f[,-(dim(pcadat_K562_10kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_10kb_perc_f[,-(dim(pcadat_K562_10kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_10kb_perc_f[,-(dim(pcadat_K562_10kb_perc_f2)[2])], 4),
         data=pcadat_K562_10kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

#### Distance data

```{r}
pcadat_K562_10kb_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/PCA/pcadat_K562_10kb_dist.rds")
dim(pcadat_K562_10kb_dist)

#performing log2 transformation
pcadat_K562_10kb_dist <- apply(pcadat_K562_10kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_K562_10kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_K562_10kb_dist_f <- pcadat_K562_10kb_dist[, -which(colnames(pcadat_K562_10kb_dist) %in% nzvar)]
pcadat_K562_10kb_dist_f <- pcadat_K562_10kb_dist

dim(pcadat_K562_10kb_dist) 
dim(pcadat_K562_10kb_dist_f)

colnames(pcadat_K562_10kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_10kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_10kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_10kb_dist_f),
          #rowsep=1:nrow(pcadat_K562_10kb_dist_f),
          keysize = 1)

#PCA
pcadat_K562_10kb_dist_f <- t(pcadat_K562_10kb_dist_f)

Group <- ifelse(rownames(pcadat_K562_10kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_10kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_10kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_10kb_dist_f <- cbind.data.frame(pcadat_K562_10kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_10kb_dist_f[,-(dim(pcadat_K562_10kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_10kb_dist_f[,-(dim(pcadat_K562_10kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_10kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_10kb_dist_f[,-(dim(pcadat_K562_10kb_dist_f2)[2])], 4),
         data=pcadat_K562_10kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

### 25kb

#### Overlap counts data

```{r}
pcadat_K562_25kb_count <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/PCA/pcadat_K562_25kb_count.rds")
dim(pcadat_K562_25kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_25kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_25kb_count_f <- pcadat_K562_25kb_count[, -which(colnames(pcadat_K562_25kb_count) %in% nzvar)]

dim(pcadat_K562_25kb_count) 
dim(pcadat_K562_25kb_count_f)

colnames(pcadat_K562_25kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_25kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_25kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_25kb_count_f),
          #rowsep=1:nrow(pcadat_K562_25kb_count_f),
          keysize = 1)

#PCA
pcadat_K562_25kb_count_f <- t(pcadat_K562_25kb_count_f)

Group <- ifelse(rownames(pcadat_K562_25kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_25kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_25kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_25kb_count_f <- cbind.data.frame(pcadat_K562_25kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_25kb_count_f[,-(dim(pcadat_K562_25kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_25kb_count_f[,-(dim(pcadat_K562_25kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_25kb_count_f[,-(dim(pcadat_K562_25kb_count_f2)[2])], 4),
         data=pcadat_K562_25kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_K562_25kb_perc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/PCA/pcadat_K562_25kb_perc.rds")
dim(pcadat_K562_25kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_25kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_25kb_perc_f <- pcadat_K562_25kb_perc[, -which(colnames(pcadat_K562_25kb_perc) %in% nzvar)]

dim(pcadat_K562_25kb_perc) 
dim(pcadat_K562_25kb_perc_f)

colnames(pcadat_K562_25kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_25kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_25kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_25kb_perc_f),
          #rowsep=1:nrow(pcadat_K562_25kb_perc_f),
          keysize = 1)

#PCA
pcadat_K562_25kb_perc_f <- t(pcadat_K562_25kb_perc_f)

Group <- ifelse(rownames(pcadat_K562_25kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_25kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_25kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_25kb_perc_f <- cbind.data.frame(pcadat_K562_25kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_25kb_perc_f[,-(dim(pcadat_K562_25kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_25kb_perc_f[,-(dim(pcadat_K562_25kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_25kb_perc_f[,-(dim(pcadat_K562_25kb_perc_f2)[2])], 4),
         data=pcadat_K562_25kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Distance data

```{r}
pcadat_K562_25kb_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/PCA/pcadat_K562_25kb_dist.rds")
dim(pcadat_K562_25kb_dist)

#performing log2 transformation
pcadat_K562_25kb_dist <- apply(pcadat_K562_25kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_K562_25kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_K562_25kb_dist_f <- pcadat_K562_25kb_dist[, -which(colnames(pcadat_K562_25kb_dist) %in% nzvar)]
pcadat_K562_25kb_dist_f <- pcadat_K562_25kb_dist

dim(pcadat_K562_25kb_dist) 
dim(pcadat_K562_25kb_dist_f)

colnames(pcadat_K562_25kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_25kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_25kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_25kb_dist_f),
          #rowsep=1:nrow(pcadat_K562_25kb_dist_f),
          keysize = 1)

#PCA
pcadat_K562_25kb_dist_f <- t(pcadat_K562_25kb_dist_f)

Group <- ifelse(rownames(pcadat_K562_25kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_25kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_25kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_25kb_dist_f <- cbind.data.frame(pcadat_K562_25kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_25kb_dist_f[,-(dim(pcadat_K562_25kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_25kb_dist_f[,-(dim(pcadat_K562_25kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_25kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_25kb_dist_f[,-(dim(pcadat_K562_25kb_dist_f2)[2])], 4),
         data=pcadat_K562_25kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

### 50kb

#### Overlap counts data

```{r}
pcadat_K562_50kb_count <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/PCA/pcadat_K562_50kb_count.rds")
dim(pcadat_K562_50kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_50kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_50kb_count_f <- pcadat_K562_50kb_count[, -which(colnames(pcadat_K562_50kb_count) %in% nzvar)]

dim(pcadat_K562_50kb_count) 
dim(pcadat_K562_50kb_count_f)

colnames(pcadat_K562_50kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_50kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_50kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_50kb_count_f),
          #rowsep=1:nrow(pcadat_K562_50kb_count_f),
          keysize = 1)

#PCA
pcadat_K562_50kb_count_f <- t(pcadat_K562_50kb_count_f)

Group <- ifelse(rownames(pcadat_K562_50kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_50kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_50kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_50kb_count_f <- cbind.data.frame(pcadat_K562_50kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_50kb_count_f[,-(dim(pcadat_K562_50kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_50kb_count_f[,-(dim(pcadat_K562_50kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_50kb_count_f[,-(dim(pcadat_K562_50kb_count_f2)[2])], 4),
         data=pcadat_K562_50kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_K562_50kb_perc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/PCA/pcadat_K562_50kb_perc.rds")
dim(pcadat_K562_50kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_50kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_50kb_perc_f <- pcadat_K562_50kb_perc[, -which(colnames(pcadat_K562_50kb_perc) %in% nzvar)]

dim(pcadat_K562_50kb_perc) 
dim(pcadat_K562_50kb_perc_f)

colnames(pcadat_K562_50kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_50kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_50kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_50kb_perc_f),
          #rowsep=1:nrow(pcadat_K562_50kb_perc_f),
          keysize = 1)

#PCA
pcadat_K562_50kb_perc_f <- t(pcadat_K562_50kb_perc_f)

Group <- ifelse(rownames(pcadat_K562_50kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_50kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_50kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_50kb_perc_f <- cbind.data.frame(pcadat_K562_50kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_50kb_perc_f[,-(dim(pcadat_K562_50kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_50kb_perc_f[,-(dim(pcadat_K562_50kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_50kb_perc_f[,-(dim(pcadat_K562_50kb_perc_f2)[2])], 4),
         data=pcadat_K562_50kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

#### Distance data

```{r}
pcadat_K562_50kb_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/PCA/pcadat_K562_50kb_dist.rds")
dim(pcadat_K562_50kb_dist)

#performing log2 transformation
pcadat_K562_50kb_dist <- apply(pcadat_K562_50kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_K562_50kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_K562_50kb_dist_f <- pcadat_K562_50kb_dist[, -which(colnames(pcadat_K562_50kb_dist) %in% nzvar)]
pcadat_K562_50kb_dist_f <- pcadat_K562_50kb_dist

dim(pcadat_K562_50kb_dist) 
dim(pcadat_K562_50kb_dist_f)

colnames(pcadat_K562_50kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_50kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_50kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_50kb_dist_f),
          #rowsep=1:nrow(pcadat_K562_50kb_dist_f),
          keysize = 1)

#PCA
pcadat_K562_50kb_dist_f <- t(pcadat_K562_50kb_dist_f)

Group <- ifelse(rownames(pcadat_K562_50kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_50kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_50kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_50kb_dist_f <- cbind.data.frame(pcadat_K562_50kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_50kb_dist_f[,-(dim(pcadat_K562_50kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_50kb_dist_f[,-(dim(pcadat_K562_50kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_50kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_50kb_dist_f[,-(dim(pcadat_K562_50kb_dist_f2)[2])], 4),
         data=pcadat_K562_50kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

### 100kb

#### Overlap counts data

```{r}
pcadat_K562_100kb_count <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/PCA/pcadat_K562_100kb_count.rds")
dim(pcadat_K562_100kb_count)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_100kb_count , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_100kb_count_f <- pcadat_K562_100kb_count[, -which(colnames(pcadat_K562_100kb_count) %in% nzvar)]

dim(pcadat_K562_100kb_count) 
dim(pcadat_K562_100kb_count_f)

colnames(pcadat_K562_100kb_count_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_100kb_count_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_100kb_count_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_100kb_count_f),
          #rowsep=1:nrow(pcadat_K562_100kb_count_f),
          keysize = 1)

#PCA
pcadat_K562_100kb_count_f <- t(pcadat_K562_100kb_count_f)

Group <- ifelse(rownames(pcadat_K562_100kb_count_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_100kb_count_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_100kb_count_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_100kb_count_f <- cbind.data.frame(pcadat_K562_100kb_count_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_100kb_count_f[,-(dim(pcadat_K562_100kb_count_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_count_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_100kb_count_f[,-(dim(pcadat_K562_100kb_count_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_count_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_100kb_count_f[,-(dim(pcadat_K562_100kb_count_f2)[2])], 4),
         data=pcadat_K562_100kb_count_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Overlap percent data

```{r}
pcadat_K562_100kb_perc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/PCA/pcadat_K562_100kb_perc.rds")
dim(pcadat_K562_100kb_perc)

#filtering low variance predictors
nzv <- nearZeroVar(pcadat_K562_100kb_perc , saveMetrics= TRUE)
nzvar <- rownames(nzv[nzv$nzv,])
nzvar

pcadat_K562_100kb_perc_f <- pcadat_K562_100kb_perc[, -which(colnames(pcadat_K562_100kb_perc) %in% nzvar)]

dim(pcadat_K562_100kb_perc) 
dim(pcadat_K562_100kb_perc_f)

colnames(pcadat_K562_100kb_perc_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "perc",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_100kb_perc_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_100kb_perc_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_100kb_perc_f),
          #rowsep=1:nrow(pcadat_K562_100kb_perc_f),
          keysize = 1)

#PCA
pcadat_K562_100kb_perc_f <- t(pcadat_K562_100kb_perc_f)

Group <- ifelse(rownames(pcadat_K562_100kb_perc_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_100kb_perc_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_100kb_perc_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_100kb_perc_f <- cbind.data.frame(pcadat_K562_100kb_perc_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_100kb_perc_f[,-(dim(pcadat_K562_100kb_perc_f2)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_perc_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_100kb_perc_f[,-(dim(pcadat_K562_100kb_perc_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_perc_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_100kb_perc_f[,-(dim(pcadat_K562_100kb_perc_f2)[2])], 4),
         data=pcadat_K562_100kb_perc_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()

```

#### Distance data

```{r}
pcadat_K562_100kb_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/PCA/pcadat_K562_100kb_dist.rds")
dim(pcadat_K562_100kb_dist)

#performing log2 transformation
pcadat_K562_100kb_dist <- apply(pcadat_K562_100kb_dist,2,function(x){log(x + 1, base=2)})

#filtering low variance predictors
#nzv <- nearZeroVar(pcadat_K562_100kb_dist , saveMetrics= TRUE)
#nzvar <- rownames(nzv[nzv$nzv,])
#nzvar
#no near zero variance predictors

pcadat_K562_100kb_dist_f <- pcadat_K562_100kb_dist[, -which(colnames(pcadat_K562_100kb_dist) %in% nzvar)]
pcadat_K562_100kb_dist_f <- pcadat_K562_100kb_dist

dim(pcadat_K562_100kb_dist) 
dim(pcadat_K562_100kb_dist_f)

colnames(pcadat_K562_100kb_dist_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadat_K562_100kb_dist_f))

#heatmap 
my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
heatmap.2(pcadat_K562_100kb_dist_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadat_K562_100kb_dist_f),
          #rowsep=1:nrow(pcadat_K562_100kb_dist_f),
          keysize = 1)

#PCA
pcadat_K562_100kb_dist_f <- t(pcadat_K562_100kb_dist_f)

Group <- ifelse(rownames(pcadat_K562_100kb_dist_f) %in% cs, "Chromatin State", ###chromatin states
                                            ifelse(rownames(pcadat_K562_100kb_dist_f)=="DNase", "DNase I", ###dnase
                                                   ifelse(rownames(pcadat_K562_100kb_dist_f) %in% hm, "Histone Modification", ###histone modifications
                                                          "TFBS"))) ###transcription factor binding sites

pcadat_K562_100kb_dist_f <- cbind.data.frame(pcadat_K562_100kb_dist_f, Group)
##with loadings
autoplot(prcomp(pcadat_K562_100kb_dist_f[,-(dim(pcadat_K562_100kb_dist_f)[2])]),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_dist_f,
         colour='Group') +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_K562_100kb_dist_f[,-(dim(pcadat_K562_100kb_dist_f)[2])]),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T,
         data=pcadat_K562_100kb_dist_f,
         colour='Group') +
  theme_bw()
##kmeans clustering algorithm to group annotations
set.seed(1)
autoplot(kmeans(pcadat_K562_100kb_dist_f[,-(dim(pcadat_K562_100kb_dist_f2)[2])], 4),
         data=pcadat_K562_100kb_dist_f,
         frame = TRUE,
         frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         #colour='Group',
         label.repel=T) + 
  theme_bw()
```

## Saving plots

```{r}
fileslst <- c(paste("Z:/TAD_data_analysis/GM12878/",c("10kb", "25kb", "50kb", "100kb"), "/Ensemble/PCA/pcadat_GM12878_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_count.rds",sep=""),
              paste("Z:/TAD_data_analysis/GM12878/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_GM12878_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_perc.rds",sep=""),
              paste("Z:/TAD_data_analysis/GM12878/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_GM12878_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_dist.rds",sep=""),
              paste("Z:/TAD_data_analysis/K562/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_K562_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_count.rds",sep=""),
              paste("Z:/TAD_data_analysis/K562/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_K562_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_perc.rds",sep=""),
              paste("Z:/TAD_data_analysis/K562/",c("10kb", "25kb", "50kb", "100kb"),"/Ensemble/PCA/pcadat_K562_", 
                    c("10kb", "25kb", "50kb", "100kb"), 
                    "_dist.rds",sep=""))

#file names for heatmaps
heatnames <- c(paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_count", "_heatmap.png", sep=""),
               paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_perc", "_heatmap.png", sep=""),
               paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_dist", "_heatmap.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_count", "_heatmap.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_perc", "_heatmap.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_dist", "_heatmap.png", sep=""))

#file names for pca plots
scorepnames <- c(paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_count", "_scoreplots.png", sep=""),
               paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_perc", "_scoreplots.png", sep=""),
               paste("gm12878", c("_10kb", "_25kb", "_50kb", "_100kb"), "_dist", "_scoreplots.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_count", "_scoreplots.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_perc", "_scoreplots.png", sep=""),
               paste("k562", c("_10kb", "_25kb", "_50kb", "_100kb"), "_dist", "_scoreplots.png", sep=""))

##create annotation groups
cs <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/BroadHMM")))
hm <- unique(gsub(paste(c("Gm12878-", "K562-", ".bed"), 
                 sep = "",
                 collapse = "|"),
           "", 
           list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")))

for(i in 1:length(fileslst)){
  print(i)
  pcadata <- readRDS(fileslst[i])

  #performing log2 transformation only for distance predictor types
  if(grepl("_dist", colnames(pcadata)[1])){
    pcadata <- apply(pcadata,2,function(x){log(x + 1, base=2)})
  }

  #filtering low variance predictors
  nzv <- nearZeroVar(pcadata, saveMetrics= TRUE)
  nzvar <- rownames(nzv[nzv$nzv,])
  if(length(nzvar) > 0){ 
    pcadata_f <- pcadata[, -which(colnames(pcadata) %in% nzvar)]
    }else{pcadata_f <- pcadata}

  #cleaning column names
  colnames(pcadata_f) <- gsub(paste(c("`", 
                                                    "_",
                                                    "-",
                                                    "Gm12878",
                                                    "K562", 
                                                    "dist",
                                                    "count",
                                                    "perc",
                                                    "Haib", 
                                                    "Sydh",
                                                    "Uta",
                                                    "Uw",
                                                    "Broad",
                                                    "Uchicago"), collapse = "|"),"", colnames(pcadata_f))

  #heatmap 
  my_palette <- rev(colorRampPalette(brewer.pal(11, "RdBu"))(100))
  png(filename = paste("C:/Users/stili/Documents/TAD_miscellaneous/figures/pca_plots/", heatnames[i], sep=""),
      width=11, height=8, units="in", res = 300)
  heatmap.2(pcadata_f,
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          #margins =c(5,5),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          #na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=TRUE,
          dendrogram="column",     # only draw a row dendrogram
          cexRow=0.5,
          cexCol=0.5,
          labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          #sepwidth=c(0.001,0.001),
          #sepcolor="black",
          #colsep=1:ncol(pcadata_f),
          #rowsep=1:nrow(pcadata_f),
          keysize = 1) 
  dev.off()

  #PCA
  #pcadata_f <- t(pcadata_f)
  #Group <- ifelse(rownames(pcadata_f) %in% cs, "Chromatin State", ###chromatin states
  #                                          ifelse(rownames(pcadata_f)=="DNase", "DNase I", ###dnase
  #                                                 ifelse(rownames(pcadata_f) %in% hm, "Histone Modification", ###histone modifications
  #                                                        "TFBS"))) ###transcription factor binding sites
  #pcadata_f <- cbind.data.frame(pcadata_f, Group)
  #
  ###Plots without loadings
  #p <- autoplot(prcomp(pcadata_f[,-(dim(pcadata_f)[2])]),
  #       #loadings = TRUE,
  #       label = TRUE, 
  #       label.size = 3,
  #       label.repel=T,
  #       data=pcadata_f,
  #       colour='Group',
  #       label.show.legend = FALSE) +
  #theme_bw()

  #ggsave(p, filename = paste("C:/Users/stili/Documents/TAD_miscellaneous/figures/pca_plots/", 
  #                           scorepnames[i], sep=""),
  #       width=8, height=11, units="in", dpi=300)

}

```