### 2019-04-07 Scripts/evaluating_model_performances.Rmd
editted results scripts

---
title: "Evaluating Model Performances"
author: "Spiro Stilianoudakis"
date: "March 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

#Barplots of performance metrics for models

## GM12878

###10kb

```{r}
#OC
##None
rfperf_gm12878_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_10kb_rus_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_10kb_ros_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_10kb_smote_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_10kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_10kb_rus_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_10kb_ros_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_10kb_smote_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_10kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_10kb_rus_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_10kb_ros_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_10kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_gm12878_10kb <- data.frame(Resampling=c(rep("None", 18),
                                       rep("RUS", 18),
                                       rep("ROS", 18),
                                       rep("SMOTE", 18)),
                          Predictor=rep(c(rep("OC", 6),
                                      rep("OP", 6),
                                      rep("Distance", 6)),4),
                          Metric=rep(c("Accuracy", 
                                       "AUPRC", 
                                       "F1-Score", 
                                       "MCC",
                                       "Informedness",
                                       "Balanced Acc"), 12))

all_rfperfs_gm12878_10kb$Performance <- c(round(rfperf_gm12878_10kb_none_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_none_oc[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_none_oc[16,2],2), #f1
                                round(rfperf_gm12878_10kb_none_oc[15,2],2), #mcc
								round(rfperf_gm12878_10kb_none_oc[18,2],2), #youden
								(round(rfperf_gm12878_10kb_none_oc[6,2],2) + round(rfperf_gm12878_10kb_none_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_10kb_none_op[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_none_op[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_none_op[16,2],2), #f1
                                round(rfperf_gm12878_10kb_none_op[15,2],2), #mcc
								round(rfperf_gm12878_10kb_none_op[18,2],2), #youden
								(round(rfperf_gm12878_10kb_none_op[6,2],2) + round(rfperf_gm12878_10kb_none_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_10kb_none_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_none_dist[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_none_dist[16,2],2), #f1
                                round(rfperf_gm12878_10kb_none_dist[15,2],2), #mcc
								round(rfperf_gm12878_10kb_none_dist[18,2],2), #youden
								(round(rfperf_gm12878_10kb_none_dist[6,2],2) + round(rfperf_gm12878_10kb_none_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_10kb_rus_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_rus_oc[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_rus_oc[16,2],2), #f1
                                round(rfperf_gm12878_10kb_rus_oc[15,2],2), #mcc
								round(rfperf_gm12878_10kb_rus_oc[18,2],2), #youden
								(round(rfperf_gm12878_10kb_rus_oc[6,2],2) + round(rfperf_gm12878_10kb_rus_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_10kb_rus_op[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_rus_op[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_rus_op[16,2],2), #f1
                                round(rfperf_gm12878_10kb_rus_op[15,2],2), #mcc
								round(rfperf_gm12878_10kb_rus_op[18,2],2), #youden
								(round(rfperf_gm12878_10kb_rus_op[6,2],2) + round(rfperf_gm12878_10kb_rus_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_10kb_rus_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_rus_dist[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_rus_dist[16,2],2), #f1
                                round(rfperf_gm12878_10kb_rus_dist[15,2],2), #mcc
								round(rfperf_gm12878_10kb_rus_dist[18,2],2), #youden
								(round(rfperf_gm12878_10kb_rus_dist[6,2],2) + round(rfperf_gm12878_10kb_rus_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_10kb_ros_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_ros_oc[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_ros_oc[16,2],2), #f1
                                round(rfperf_gm12878_10kb_ros_oc[15,2],2), #mcc
								round(rfperf_gm12878_10kb_ros_oc[18,2],2), #youden
								(round(rfperf_gm12878_10kb_ros_oc[6,2],2) + round(rfperf_gm12878_10kb_ros_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_10kb_ros_op[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_ros_op[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_ros_op[16,2],2), #f1
                                round(rfperf_gm12878_10kb_ros_op[15,2],2), #mcc
								round(rfperf_gm12878_10kb_ros_op[18,2],2), #youden
								(round(rfperf_gm12878_10kb_ros_op[6,2],2) + round(rfperf_gm12878_10kb_ros_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_10kb_ros_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_ros_dist[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_ros_dist[16,2],2), #f1
                                round(rfperf_gm12878_10kb_ros_dist[15,2],2), #mcc
								round(rfperf_gm12878_10kb_ros_dist[18,2],2), #youden
								(round(rfperf_gm12878_10kb_ros_dist[6,2],2) + round(rfperf_gm12878_10kb_ros_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_10kb_smote_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_smote_oc[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_smote_oc[16,2],2), #f1
                                round(rfperf_gm12878_10kb_smote_oc[15,2],2), #mcc
								round(rfperf_gm12878_10kb_smote_oc[18,2],2), #youden
								(round(rfperf_gm12878_10kb_smote_oc[6,2],2) + round(rfperf_gm12878_10kb_smote_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_10kb_smote_op[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_smote_op[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_smote_op[16,2],2), #f1
                                round(rfperf_gm12878_10kb_smote_op[15,2],2), #mcc
								round(rfperf_gm12878_10kb_smote_op[18,2],2), #youden
								(round(rfperf_gm12878_10kb_smote_op[6,2],2) + round(rfperf_gm12878_10kb_smote_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_10kb_smote_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_10kb_smote_dist[19,2],2), #auprc
                                round(rfperf_gm12878_10kb_smote_dist[16,2],2), #f1
                                round(rfperf_gm12878_10kb_smote_dist[15,2],2), #mcc
								round(rfperf_gm12878_10kb_smote_dist[18,2],2), #youden
								(round(rfperf_gm12878_10kb_smote_dist[6,2],2) + round(rfperf_gm12878_10kb_smote_dist[7,2],2))/2 #balanced accuracy
								)
all_rfperfs_gm12878_10kb$Resampling <- factor(all_rfperfs_gm12878_10kb$Resampling, levels=c("None","ROS","RUS","SMOTE"))
all_rfperfs_gm12878_10kb$Predictor <- factor(all_rfperfs_gm12878_10kb$Predictor, levels = c("OC", "OP", "Distance"))
all_rfperfs_gm12878_10kb$Metric <- factor(all_rfperfs_gm12878_10kb$Metric, levels = c("Accuracy",
                                                                                      "AUPRC",
                                                                                      "F1-Score",
                                                                                      "MCC",
                                                                                      "Informedness",
                                                                                      "Balanced Acc"))

#comparing only f1 and mcc and balanced acc

ggplot(all_rfperfs_gm12878_10kb[which(all_rfperfs_gm12878_10kb$Metric=="F1-Score" |
                                               all_rfperfs_gm12878_10kb$Metric=="MCC" |
                                        all_rfperfs_gm12878_10kb$Metric=="Balanced Acc"),],
                       aes(x=Resampling, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(~ Metric) +
  xlab("Resampling Technique") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

###25kb

```{r}
#OC
##None
rfperf_gm12878_25kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_25kb_rus_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_25kb_ros_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_25kb_smote_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_25kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_25kb_rus_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_25kb_ros_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_25kb_smote_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_25kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_25kb_rus_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_25kb_ros_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_25kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_gm12878_25kb <- data.frame(Resampling=c(rep("None", 18),
                                       rep("RUS", 18),
                                       rep("ROS", 18),
                                       rep("SMOTE", 18)),
                          Predictor=rep(c(rep("OC", 6),
                                      rep("OP", 6),
                                      rep("Distance", 6)),4),
                          Metric=rep(c("Accuracy", 
                                       "AUPRC", 
                                       "F1-Score", 
                                       "MCC",
                                       "Informedness",
                                       "Balanced Acc"), 12))

all_rfperfs_gm12878_25kb$Performance <- c(round(rfperf_gm12878_25kb_none_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_none_oc[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_none_oc[16,2],2), #f1
                                round(rfperf_gm12878_25kb_none_oc[15,2],2), #mcc
								round(rfperf_gm12878_25kb_none_oc[18,2],2), #youden
								(round(rfperf_gm12878_25kb_none_oc[6,2],2) + round(rfperf_gm12878_25kb_none_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_25kb_none_op[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_none_op[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_none_op[16,2],2), #f1
                                round(rfperf_gm12878_25kb_none_op[15,2],2), #mcc
								round(rfperf_gm12878_25kb_none_op[18,2],2), #youden
								(round(rfperf_gm12878_25kb_none_op[6,2],2) + round(rfperf_gm12878_25kb_none_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_25kb_none_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_none_dist[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_none_dist[16,2],2), #f1
                                round(rfperf_gm12878_25kb_none_dist[15,2],2), #mcc
								round(rfperf_gm12878_25kb_none_dist[18,2],2), #youden
								(round(rfperf_gm12878_25kb_none_dist[6,2],2) + round(rfperf_gm12878_25kb_none_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_25kb_rus_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_rus_oc[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_rus_oc[16,2],2), #f1
                                round(rfperf_gm12878_25kb_rus_oc[15,2],2), #mcc
								round(rfperf_gm12878_25kb_rus_oc[18,2],2), #youden
								(round(rfperf_gm12878_25kb_rus_oc[6,2],2) + round(rfperf_gm12878_25kb_rus_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_25kb_rus_op[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_rus_op[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_rus_op[16,2],2), #f1
                                round(rfperf_gm12878_25kb_rus_op[15,2],2), #mcc
								round(rfperf_gm12878_25kb_rus_op[18,2],2), #youden
								(round(rfperf_gm12878_25kb_rus_op[6,2],2) + round(rfperf_gm12878_25kb_rus_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_25kb_rus_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_rus_dist[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_rus_dist[16,2],2), #f1
                                round(rfperf_gm12878_25kb_rus_dist[15,2],2), #mcc
								round(rfperf_gm12878_25kb_rus_dist[18,2],2), #youden
								(round(rfperf_gm12878_25kb_rus_dist[6,2],2) + round(rfperf_gm12878_25kb_rus_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_25kb_ros_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_ros_oc[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_ros_oc[16,2],2), #f1
                                round(rfperf_gm12878_25kb_ros_oc[15,2],2), #mcc
								round(rfperf_gm12878_25kb_ros_oc[18,2],2), #youden
								(round(rfperf_gm12878_25kb_ros_oc[6,2],2) + round(rfperf_gm12878_25kb_ros_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_25kb_ros_op[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_ros_op[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_ros_op[16,2],2), #f1
                                round(rfperf_gm12878_25kb_ros_op[15,2],2), #mcc
								round(rfperf_gm12878_25kb_ros_op[18,2],2), #youden
								(round(rfperf_gm12878_25kb_ros_op[6,2],2) + round(rfperf_gm12878_25kb_ros_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_25kb_ros_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_ros_dist[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_ros_dist[16,2],2), #f1
                                round(rfperf_gm12878_25kb_ros_dist[15,2],2), #mcc
								round(rfperf_gm12878_25kb_ros_dist[18,2],2), #youden
								(round(rfperf_gm12878_25kb_ros_dist[6,2],2) + round(rfperf_gm12878_25kb_ros_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_25kb_smote_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_smote_oc[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_smote_oc[16,2],2), #f1
                                round(rfperf_gm12878_25kb_smote_oc[15,2],2), #mcc
								round(rfperf_gm12878_25kb_smote_oc[18,2],2), #youden
								(round(rfperf_gm12878_25kb_smote_oc[6,2],2) + round(rfperf_gm12878_25kb_smote_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_25kb_smote_op[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_smote_op[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_smote_op[16,2],2), #f1
                                round(rfperf_gm12878_25kb_smote_op[15,2],2), #mcc
								round(rfperf_gm12878_25kb_smote_op[18,2],2), #youden
								(round(rfperf_gm12878_25kb_smote_op[6,2],2) + round(rfperf_gm12878_25kb_smote_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_25kb_smote_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_25kb_smote_dist[19,2],2), #auprc
                                round(rfperf_gm12878_25kb_smote_dist[16,2],2), #f1
                                round(rfperf_gm12878_25kb_smote_dist[15,2],2), #mcc
								round(rfperf_gm12878_25kb_smote_dist[18,2],2), #youden
								(round(rfperf_gm12878_25kb_smote_dist[6,2],2) + round(rfperf_gm12878_25kb_smote_dist[7,2],2))/2 #balanced accuracy
								)
all_rfperfs_gm12878_25kb$Resampling <- factor(all_rfperfs_gm12878_25kb$Resampling, levels=c("None","ROS","RUS","SMOTE"))
all_rfperfs_gm12878_25kb$Predictor <- factor(all_rfperfs_gm12878_25kb$Predictor, levels = c("OC", "OP", "Distance"))
all_rfperfs_gm12878_25kb$Metric <- factor(all_rfperfs_gm12878_25kb$Metric, levels = c("Accuracy",
                                                                                      "AUPRC",
                                                                                      "F1-Score",
                                                                                      "MCC",
                                                                                      "Informedness",
                                                                                      "Balanced Acc"))

#comparing only f1 and mcc and balanced acc

ggplot(all_rfperfs_gm12878_25kb[which(all_rfperfs_gm12878_25kb$Metric=="F1-Score" |
                                               all_rfperfs_gm12878_25kb$Metric=="MCC" |
                                        all_rfperfs_gm12878_25kb$Metric=="Balanced Acc"),],
                       aes(x=Resampling, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(~ Metric) +
  xlab("Resampling Technique") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

###50kb

```{r}
#OC
##None
rfperf_gm12878_50kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_50kb_rus_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_50kb_ros_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_50kb_smote_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_50kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_50kb_rus_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_50kb_ros_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_50kb_smote_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_50kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_50kb_rus_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_50kb_ros_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_50kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_gm12878_50kb <- data.frame(Resampling=c(rep("None", 18),
                                       rep("RUS", 18),
                                       rep("ROS", 18),
                                       rep("SMOTE", 18)),
                          Predictor=rep(c(rep("OC", 6),
                                      rep("OP", 6),
                                      rep("Distance", 6)),4),
                          Metric=rep(c("Accuracy", 
                                       "AUPRC", 
                                       "F1-Score", 
                                       "MCC",
                                       "Informedness",
                                       "Balanced Acc"), 12))

all_rfperfs_gm12878_50kb$Performance <- c(round(rfperf_gm12878_50kb_none_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_none_oc[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_none_oc[16,2],2), #f1
                                round(rfperf_gm12878_50kb_none_oc[15,2],2), #mcc
								round(rfperf_gm12878_50kb_none_oc[18,2],2), #youden
								(round(rfperf_gm12878_50kb_none_oc[6,2],2) + round(rfperf_gm12878_50kb_none_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_50kb_none_op[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_none_op[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_none_op[16,2],2), #f1
                                round(rfperf_gm12878_50kb_none_op[15,2],2), #mcc
								round(rfperf_gm12878_50kb_none_op[18,2],2), #youden
								(round(rfperf_gm12878_50kb_none_op[6,2],2) + round(rfperf_gm12878_50kb_none_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_50kb_none_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_none_dist[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_none_dist[16,2],2), #f1
                                round(rfperf_gm12878_50kb_none_dist[15,2],2), #mcc
								round(rfperf_gm12878_50kb_none_dist[18,2],2), #youden
								(round(rfperf_gm12878_50kb_none_dist[6,2],2) + round(rfperf_gm12878_50kb_none_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_50kb_rus_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_rus_oc[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_rus_oc[16,2],2), #f1
                                round(rfperf_gm12878_50kb_rus_oc[15,2],2), #mcc
								round(rfperf_gm12878_50kb_rus_oc[18,2],2), #youden
								(round(rfperf_gm12878_50kb_rus_oc[6,2],2) + round(rfperf_gm12878_50kb_rus_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_50kb_rus_op[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_rus_op[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_rus_op[16,2],2), #f1
                                round(rfperf_gm12878_50kb_rus_op[15,2],2), #mcc
								round(rfperf_gm12878_50kb_rus_op[18,2],2), #youden
								(round(rfperf_gm12878_50kb_rus_op[6,2],2) + round(rfperf_gm12878_50kb_rus_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_50kb_rus_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_rus_dist[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_rus_dist[16,2],2), #f1
                                round(rfperf_gm12878_50kb_rus_dist[15,2],2), #mcc
								round(rfperf_gm12878_50kb_rus_dist[18,2],2), #youden
								(round(rfperf_gm12878_50kb_rus_dist[6,2],2) + round(rfperf_gm12878_50kb_rus_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_50kb_ros_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_ros_oc[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_ros_oc[16,2],2), #f1
                                round(rfperf_gm12878_50kb_ros_oc[15,2],2), #mcc
								round(rfperf_gm12878_50kb_ros_oc[18,2],2), #youden
								(round(rfperf_gm12878_50kb_ros_oc[6,2],2) + round(rfperf_gm12878_50kb_ros_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_50kb_ros_op[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_ros_op[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_ros_op[16,2],2), #f1
                                round(rfperf_gm12878_50kb_ros_op[15,2],2), #mcc
								round(rfperf_gm12878_50kb_ros_op[18,2],2), #youden
								(round(rfperf_gm12878_50kb_ros_op[6,2],2) + round(rfperf_gm12878_50kb_ros_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_50kb_ros_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_ros_dist[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_ros_dist[16,2],2), #f1
                                round(rfperf_gm12878_50kb_ros_dist[15,2],2), #mcc
								round(rfperf_gm12878_50kb_ros_dist[18,2],2), #youden
								(round(rfperf_gm12878_50kb_ros_dist[6,2],2) + round(rfperf_gm12878_50kb_ros_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_50kb_smote_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_smote_oc[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_smote_oc[16,2],2), #f1
                                round(rfperf_gm12878_50kb_smote_oc[15,2],2), #mcc
								round(rfperf_gm12878_50kb_smote_oc[18,2],2), #youden
								(round(rfperf_gm12878_50kb_smote_oc[6,2],2) + round(rfperf_gm12878_50kb_smote_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_50kb_smote_op[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_smote_op[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_smote_op[16,2],2), #f1
                                round(rfperf_gm12878_50kb_smote_op[15,2],2), #mcc
								round(rfperf_gm12878_50kb_smote_op[18,2],2), #youden
								(round(rfperf_gm12878_50kb_smote_op[6,2],2) + round(rfperf_gm12878_50kb_smote_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_50kb_smote_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_50kb_smote_dist[19,2],2), #auprc
                                round(rfperf_gm12878_50kb_smote_dist[16,2],2), #f1
                                round(rfperf_gm12878_50kb_smote_dist[15,2],2), #mcc
								round(rfperf_gm12878_50kb_smote_dist[18,2],2), #youden
								(round(rfperf_gm12878_50kb_smote_dist[6,2],2) + round(rfperf_gm12878_50kb_smote_dist[7,2],2))/2 #balanced accuracy
								)
all_rfperfs_gm12878_50kb$Resampling <- factor(all_rfperfs_gm12878_50kb$Resampling, levels=c("None","ROS","RUS","SMOTE"))
all_rfperfs_gm12878_50kb$Predictor <- factor(all_rfperfs_gm12878_50kb$Predictor, levels = c("OC", "OP", "Distance"))
all_rfperfs_gm12878_50kb$Metric <- factor(all_rfperfs_gm12878_50kb$Metric, levels = c("Accuracy",
                                                                                      "AUPRC",
                                                                                      "F1-Score",
                                                                                      "MCC",
                                                                                      "Informedness",
                                                                                      "Balanced Acc"))

#comparing only f1 and mcc and balanced acc

ggplot(all_rfperfs_gm12878_50kb[which(all_rfperfs_gm12878_50kb$Metric=="F1-Score" |
                                               all_rfperfs_gm12878_50kb$Metric=="MCC" |
                                        all_rfperfs_gm12878_50kb$Metric=="Balanced Acc"),],
                       aes(x=Resampling, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(~ Metric) +
  xlab("Resampling Technique") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

###100kb

```{r}
#OC
##None
rfperf_gm12878_100kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_100kb_rus_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_100kb_ros_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_100kb_smote_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_100kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_100kb_rus_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_100kb_ros_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_100kb_smote_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_100kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

##RUS
rfperf_gm12878_100kb_rus_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

##ROS
rfperf_gm12878_100kb_ros_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

##SMOTE
rfperf_gm12878_100kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_gm12878_100kb <- data.frame(Resampling=c(rep("None", 18),
                                       rep("RUS", 18),
                                       rep("ROS", 18),
                                       rep("SMOTE", 18)),
                          Predictor=rep(c(rep("OC", 6),
                                      rep("OP", 6),
                                      rep("Distance", 6)),4),
                          Metric=rep(c("Accuracy", 
                                       "AUPRC", 
                                       "F1-Score", 
                                       "MCC",
                                       "Informedness",
                                       "Balanced Acc"), 12))

all_rfperfs_gm12878_100kb$Performance <- c(round(rfperf_gm12878_100kb_none_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_none_oc[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_none_oc[16,2],2), #f1
                                round(rfperf_gm12878_100kb_none_oc[15,2],2), #mcc
								round(rfperf_gm12878_100kb_none_oc[18,2],2), #youden
								(round(rfperf_gm12878_100kb_none_oc[6,2],2) + round(rfperf_gm12878_100kb_none_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_100kb_none_op[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_none_op[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_none_op[16,2],2), #f1
                                round(rfperf_gm12878_100kb_none_op[15,2],2), #mcc
								round(rfperf_gm12878_100kb_none_op[18,2],2), #youden
								(round(rfperf_gm12878_100kb_none_op[6,2],2) + round(rfperf_gm12878_100kb_none_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_100kb_none_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_none_dist[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_none_dist[16,2],2), #f1
                                round(rfperf_gm12878_100kb_none_dist[15,2],2), #mcc
								round(rfperf_gm12878_100kb_none_dist[18,2],2), #youden
								(round(rfperf_gm12878_100kb_none_dist[6,2],2) + round(rfperf_gm12878_100kb_none_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_100kb_rus_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_rus_oc[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_rus_oc[16,2],2), #f1
                                round(rfperf_gm12878_100kb_rus_oc[15,2],2), #mcc
								round(rfperf_gm12878_100kb_rus_oc[18,2],2), #youden
								(round(rfperf_gm12878_100kb_rus_oc[6,2],2) + round(rfperf_gm12878_100kb_rus_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_100kb_rus_op[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_rus_op[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_rus_op[16,2],2), #f1
                                round(rfperf_gm12878_100kb_rus_op[15,2],2), #mcc
								round(rfperf_gm12878_100kb_rus_op[18,2],2), #youden
								(round(rfperf_gm12878_100kb_rus_op[6,2],2) + round(rfperf_gm12878_100kb_rus_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_100kb_rus_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_rus_dist[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_rus_dist[16,2],2), #f1
                                round(rfperf_gm12878_100kb_rus_dist[15,2],2), #mcc
								round(rfperf_gm12878_100kb_rus_dist[18,2],2), #youden
								(round(rfperf_gm12878_100kb_rus_dist[6,2],2) + round(rfperf_gm12878_100kb_rus_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_100kb_ros_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_ros_oc[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_ros_oc[16,2],2), #f1
                                round(rfperf_gm12878_100kb_ros_oc[15,2],2), #mcc
								round(rfperf_gm12878_100kb_ros_oc[18,2],2), #youden
								(round(rfperf_gm12878_100kb_ros_oc[6,2],2) + round(rfperf_gm12878_100kb_ros_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_100kb_ros_op[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_ros_op[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_ros_op[16,2],2), #f1
                                round(rfperf_gm12878_100kb_ros_op[15,2],2), #mcc
								round(rfperf_gm12878_100kb_ros_op[18,2],2), #youden
								(round(rfperf_gm12878_100kb_ros_op[6,2],2) + round(rfperf_gm12878_100kb_ros_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_100kb_ros_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_ros_dist[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_ros_dist[16,2],2), #f1
                                round(rfperf_gm12878_100kb_ros_dist[15,2],2), #mcc
								round(rfperf_gm12878_100kb_ros_dist[18,2],2), #youden
								(round(rfperf_gm12878_100kb_ros_dist[6,2],2) + round(rfperf_gm12878_100kb_ros_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_100kb_smote_oc[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_smote_oc[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_smote_oc[16,2],2), #f1
                                round(rfperf_gm12878_100kb_smote_oc[15,2],2), #mcc
								round(rfperf_gm12878_100kb_smote_oc[18,2],2), #youden
								(round(rfperf_gm12878_100kb_smote_oc[6,2],2) + round(rfperf_gm12878_100kb_smote_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_gm12878_100kb_smote_op[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_smote_op[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_smote_op[16,2],2), #f1
                                round(rfperf_gm12878_100kb_smote_op[15,2],2), #mcc
								round(rfperf_gm12878_100kb_smote_op[18,2],2), #youden
								(round(rfperf_gm12878_100kb_smote_op[6,2],2) + round(rfperf_gm12878_100kb_smote_op[7,2],2))/2, #balanced accuracy

								round(rfperf_gm12878_100kb_smote_dist[9,2],2), #accuracy
                                round(rfperf_gm12878_100kb_smote_dist[19,2],2), #auprc
                                round(rfperf_gm12878_100kb_smote_dist[16,2],2), #f1
                                round(rfperf_gm12878_100kb_smote_dist[15,2],2), #mcc
								round(rfperf_gm12878_100kb_smote_dist[18,2],2), #youden
								(round(rfperf_gm12878_100kb_smote_dist[6,2],2) + round(rfperf_gm12878_100kb_smote_dist[7,2],2))/2 #balanced accuracy
								)
all_rfperfs_gm12878_100kb$Resampling <- factor(all_rfperfs_gm12878_100kb$Resampling, levels=c("None","ROS","RUS","SMOTE"))
all_rfperfs_gm12878_100kb$Predictor <- factor(all_rfperfs_gm12878_100kb$Predictor, levels = c("OC", "OP", "Distance"))
all_rfperfs_gm12878_100kb$Metric <- factor(all_rfperfs_gm12878_100kb$Metric, levels = c("Accuracy",
                                                                                      "AUPRC",
                                                                                      "F1-Score",
                                                                                      "MCC",
                                                                                      "Informedness",
                                                                                      "Balanced Acc"))

#comparing only f1 and mcc and balanced acc

ggplot(all_rfperfs_gm12878_100kb[which(all_rfperfs_gm12878_100kb$Metric=="F1-Score" |
                                               all_rfperfs_gm12878_100kb$Metric=="MCC" |
                                         all_rfperfs_gm12878_100kb$Metric=="Balanced Acc"),],
                       aes(x=Resampling, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(~ Metric) +
  xlab("Resampling Technique") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

## K562

###10kb

```{r}
#OC
##None
rfperf_K562_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_10kb_rus_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_10kb_ros_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_10kb_smote_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_10kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_10kb_rus_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_10kb_ros_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_10kb_smote_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_10kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_10kb_rus_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_10kb_ros_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_10kb_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_K562_10kb <- data.frame(Resampling=c(rep("None", 18),
                                       rep("RUS", 18),
                                       rep("ROS", 18),
                                       rep("SMOTE", 18)),
                          Predictor=rep(c(rep("OC", 6),
                                      rep("OP", 6),
                                      rep("Distance", 6)),4),
                          Metric=rep(c("Accuracy", 
                                       "AUPRC", 
                                       "F1-Score", 
                                       "MCC",
                                       "Informedness",
                                       "Balanced Acc"), 12))

all_rfperfs_K562_10kb$Performance <- c(round(rfperf_K562_10kb_none_oc[9,2],2), #accuracy
                                round(rfperf_K562_10kb_none_oc[19,2],2), #auprc
                                round(rfperf_K562_10kb_none_oc[16,2],2), #f1
                                round(rfperf_K562_10kb_none_oc[15,2],2), #mcc
								round(rfperf_K562_10kb_none_oc[18,2],2), #youden
								(round(rfperf_K562_10kb_none_oc[6,2],2) + round(rfperf_K562_10kb_none_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_10kb_none_op[9,2],2), #accuracy
                                round(rfperf_K562_10kb_none_op[19,2],2), #auprc
                                round(rfperf_K562_10kb_none_op[16,2],2), #f1
                                round(rfperf_K562_10kb_none_op[15,2],2), #mcc
								round(rfperf_K562_10kb_none_op[18,2],2), #youden
								(round(rfperf_K562_10kb_none_op[6,2],2) + round(rfperf_K562_10kb_none_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_10kb_none_dist[9,2],2), #accuracy
                                round(rfperf_K562_10kb_none_dist[19,2],2), #auprc
                                round(rfperf_K562_10kb_none_dist[16,2],2), #f1
                                round(rfperf_K562_10kb_none_dist[15,2],2), #mcc
								round(rfperf_K562_10kb_none_dist[18,2],2), #youden
								(round(rfperf_K562_10kb_none_dist[6,2],2) + round(rfperf_K562_10kb_none_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_10kb_rus_oc[9,2],2), #accuracy
                                round(rfperf_K562_10kb_rus_oc[19,2],2), #auprc
                                round(rfperf_K562_10kb_rus_oc[16,2],2), #f1
                                round(rfperf_K562_10kb_rus_oc[15,2],2), #mcc
								round(rfperf_K562_10kb_rus_oc[18,2],2), #youden
								(round(rfperf_K562_10kb_rus_oc[6,2],2) + round(rfperf_K562_10kb_rus_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_10kb_rus_op[9,2],2), #accuracy
                                round(rfperf_K562_10kb_rus_op[19,2],2), #auprc
                                round(rfperf_K562_10kb_rus_op[16,2],2), #f1
                                round(rfperf_K562_10kb_rus_op[15,2],2), #mcc
								round(rfperf_K562_10kb_rus_op[18,2],2), #youden
								(round(rfperf_K562_10kb_rus_op[6,2],2) + round(rfperf_K562_10kb_rus_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_10kb_rus_dist[9,2],2), #accuracy
                                round(rfperf_K562_10kb_rus_dist[19,2],2), #auprc
                                round(rfperf_K562_10kb_rus_dist[16,2],2), #f1
                                round(rfperf_K562_10kb_rus_dist[15,2],2), #mcc
								round(rfperf_K562_10kb_rus_dist[18,2],2), #youden
								(round(rfperf_K562_10kb_rus_dist[6,2],2) + round(rfperf_K562_10kb_rus_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_10kb_ros_oc[9,2],2), #accuracy
                                round(rfperf_K562_10kb_ros_oc[19,2],2), #auprc
                                round(rfperf_K562_10kb_ros_oc[16,2],2), #f1
                                round(rfperf_K562_10kb_ros_oc[15,2],2), #mcc
								round(rfperf_K562_10kb_ros_oc[18,2],2), #youden
								(round(rfperf_K562_10kb_ros_oc[6,2],2) + round(rfperf_K562_10kb_ros_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_10kb_ros_op[9,2],2), #accuracy
                                round(rfperf_K562_10kb_ros_op[19,2],2), #auprc
                                round(rfperf_K562_10kb_ros_op[16,2],2), #f1
                                round(rfperf_K562_10kb_ros_op[15,2],2), #mcc
								round(rfperf_K562_10kb_ros_op[18,2],2), #youden
								(round(rfperf_K562_10kb_ros_op[6,2],2) + round(rfperf_K562_10kb_ros_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_10kb_ros_dist[9,2],2), #accuracy
                                round(rfperf_K562_10kb_ros_dist[19,2],2), #auprc
                                round(rfperf_K562_10kb_ros_dist[16,2],2), #f1
                                round(rfperf_K562_10kb_ros_dist[15,2],2), #mcc
								round(rfperf_K562_10kb_ros_dist[18,2],2), #youden
								(round(rfperf_K562_10kb_ros_dist[6,2],2) + round(rfperf_K562_10kb_ros_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_10kb_smote_oc[9,2],2), #accuracy
                                round(rfperf_K562_10kb_smote_oc[19,2],2), #auprc
                                round(rfperf_K562_10kb_smote_oc[16,2],2), #f1
                                round(rfperf_K562_10kb_smote_oc[15,2],2), #mcc
								round(rfperf_K562_10kb_smote_oc[18,2],2), #youden
								(round(rfperf_K562_10kb_smote_oc[6,2],2) + round(rfperf_K562_10kb_smote_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_10kb_smote_op[9,2],2), #accuracy
                                round(rfperf_K562_10kb_smote_op[19,2],2), #auprc
                                round(rfperf_K562_10kb_smote_op[16,2],2), #f1
                                round(rfperf_K562_10kb_smote_op[15,2],2), #mcc
								round(rfperf_K562_10kb_smote_op[18,2],2), #youden
								(round(rfperf_K562_10kb_smote_op[6,2],2) + round(rfperf_K562_10kb_smote_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_10kb_smote_dist[9,2],2), #accuracy
                                round(rfperf_K562_10kb_smote_dist[19,2],2), #auprc
                                round(rfperf_K562_10kb_smote_dist[16,2],2), #f1
                                round(rfperf_K562_10kb_smote_dist[15,2],2), #mcc
								round(rfperf_K562_10kb_smote_dist[18,2],2), #youden
								(round(rfperf_K562_10kb_smote_dist[6,2],2) + round(rfperf_K562_10kb_smote_dist[7,2],2))/2 #balanced accuracy
								)
all_rfperfs_K562_10kb$Resampling <- factor(all_rfperfs_K562_10kb$Resampling, levels=c("None","ROS","RUS","SMOTE"))
all_rfperfs_K562_10kb$Predictor <- factor(all_rfperfs_K562_10kb$Predictor, levels = c("OC", "OP", "Distance"))
all_rfperfs_K562_10kb$Metric <- factor(all_rfperfs_K562_10kb$Metric, levels = c("Accuracy",
                                                                                      "AUPRC",
                                                                                      "F1-Score",
                                                                                      "MCC",
                                                                                      "Informedness",
                                                                                      "Balanced Acc"))

#comparing only f1 and mcc and balanced acc

ggplot(all_rfperfs_K562_10kb[which(all_rfperfs_K562_10kb$Metric=="F1-Score" |
                                               all_rfperfs_K562_10kb$Metric=="MCC" |
                                     all_rfperfs_K562_10kb$Metric=="Balanced Acc"),],
                       aes(x=Resampling, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(~ Metric) +
  xlab("Resampling Technique") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

###25kb

```{r}
#OC
##None
rfperf_K562_25kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_25kb_rus_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_25kb_ros_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_25kb_smote_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_25kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_25kb_rus_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_25kb_ros_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_25kb_smote_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_25kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_25kb_rus_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_25kb_ros_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_25kb_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_K562_25kb <- data.frame(Resampling=c(rep("None", 18),
                                       rep("RUS", 18),
                                       rep("ROS", 18),
                                       rep("SMOTE", 18)),
                          Predictor=rep(c(rep("OC", 6),
                                      rep("OP", 6),
                                      rep("Distance", 6)),4),
                          Metric=rep(c("Accuracy", 
                                       "AUPRC", 
                                       "F1-Score", 
                                       "MCC",
                                       "Informedness",
                                       "Balanced Acc"), 12))

all_rfperfs_K562_25kb$Performance <- c(round(rfperf_K562_25kb_none_oc[9,2],2), #accuracy
                                round(rfperf_K562_25kb_none_oc[19,2],2), #auprc
                                round(rfperf_K562_25kb_none_oc[16,2],2), #f1
                                round(rfperf_K562_25kb_none_oc[15,2],2), #mcc
								round(rfperf_K562_25kb_none_oc[18,2],2), #youden
								(round(rfperf_K562_25kb_none_oc[6,2],2) + round(rfperf_K562_25kb_none_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_25kb_none_op[9,2],2), #accuracy
                                round(rfperf_K562_25kb_none_op[19,2],2), #auprc
                                round(rfperf_K562_25kb_none_op[16,2],2), #f1
                                round(rfperf_K562_25kb_none_op[15,2],2), #mcc
								round(rfperf_K562_25kb_none_op[18,2],2), #youden
								(round(rfperf_K562_25kb_none_op[6,2],2) + round(rfperf_K562_25kb_none_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_25kb_none_dist[9,2],2), #accuracy
                                round(rfperf_K562_25kb_none_dist[19,2],2), #auprc
                                round(rfperf_K562_25kb_none_dist[16,2],2), #f1
                                round(rfperf_K562_25kb_none_dist[15,2],2), #mcc
								round(rfperf_K562_25kb_none_dist[18,2],2), #youden
								(round(rfperf_K562_25kb_none_dist[6,2],2) + round(rfperf_K562_25kb_none_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_25kb_rus_oc[9,2],2), #accuracy
                                round(rfperf_K562_25kb_rus_oc[19,2],2), #auprc
                                round(rfperf_K562_25kb_rus_oc[16,2],2), #f1
                                round(rfperf_K562_25kb_rus_oc[15,2],2), #mcc
								round(rfperf_K562_25kb_rus_oc[18,2],2), #youden
								(round(rfperf_K562_25kb_rus_oc[6,2],2) + round(rfperf_K562_25kb_rus_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_25kb_rus_op[9,2],2), #accuracy
                                round(rfperf_K562_25kb_rus_op[19,2],2), #auprc
                                round(rfperf_K562_25kb_rus_op[16,2],2), #f1
                                round(rfperf_K562_25kb_rus_op[15,2],2), #mcc
								round(rfperf_K562_25kb_rus_op[18,2],2), #youden
								(round(rfperf_K562_25kb_rus_op[6,2],2) + round(rfperf_K562_25kb_rus_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_25kb_rus_dist[9,2],2), #accuracy
                                round(rfperf_K562_25kb_rus_dist[19,2],2), #auprc
                                round(rfperf_K562_25kb_rus_dist[16,2],2), #f1
                                round(rfperf_K562_25kb_rus_dist[15,2],2), #mcc
								round(rfperf_K562_25kb_rus_dist[18,2],2), #youden
								(round(rfperf_K562_25kb_rus_dist[6,2],2) + round(rfperf_K562_25kb_rus_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_25kb_ros_oc[9,2],2), #accuracy
                                round(rfperf_K562_25kb_ros_oc[19,2],2), #auprc
                                round(rfperf_K562_25kb_ros_oc[16,2],2), #f1
                                round(rfperf_K562_25kb_ros_oc[15,2],2), #mcc
								round(rfperf_K562_25kb_ros_oc[18,2],2), #youden
								(round(rfperf_K562_25kb_ros_oc[6,2],2) + round(rfperf_K562_25kb_ros_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_25kb_ros_op[9,2],2), #accuracy
                                round(rfperf_K562_25kb_ros_op[19,2],2), #auprc
                                round(rfperf_K562_25kb_ros_op[16,2],2), #f1
                                round(rfperf_K562_25kb_ros_op[15,2],2), #mcc
								round(rfperf_K562_25kb_ros_op[18,2],2), #youden
								(round(rfperf_K562_25kb_ros_op[6,2],2) + round(rfperf_K562_25kb_ros_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_25kb_ros_dist[9,2],2), #accuracy
                                round(rfperf_K562_25kb_ros_dist[19,2],2), #auprc
                                round(rfperf_K562_25kb_ros_dist[16,2],2), #f1
                                round(rfperf_K562_25kb_ros_dist[15,2],2), #mcc
								round(rfperf_K562_25kb_ros_dist[18,2],2), #youden
								(round(rfperf_K562_25kb_ros_dist[6,2],2) + round(rfperf_K562_25kb_ros_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_25kb_smote_oc[9,2],2), #accuracy
                                round(rfperf_K562_25kb_smote_oc[19,2],2), #auprc
                                round(rfperf_K562_25kb_smote_oc[16,2],2), #f1
                                round(rfperf_K562_25kb_smote_oc[15,2],2), #mcc
								round(rfperf_K562_25kb_smote_oc[18,2],2), #youden
								(round(rfperf_K562_25kb_smote_oc[6,2],2) + round(rfperf_K562_25kb_smote_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_25kb_smote_op[9,2],2), #accuracy
                                round(rfperf_K562_25kb_smote_op[19,2],2), #auprc
                                round(rfperf_K562_25kb_smote_op[16,2],2), #f1
                                round(rfperf_K562_25kb_smote_op[15,2],2), #mcc
								round(rfperf_K562_25kb_smote_op[18,2],2), #youden
								(round(rfperf_K562_25kb_smote_op[6,2],2) + round(rfperf_K562_25kb_smote_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_25kb_smote_dist[9,2],2), #accuracy
                                round(rfperf_K562_25kb_smote_dist[19,2],2), #auprc
                                round(rfperf_K562_25kb_smote_dist[16,2],2), #f1
                                round(rfperf_K562_25kb_smote_dist[15,2],2), #mcc
								round(rfperf_K562_25kb_smote_dist[18,2],2), #youden
								(round(rfperf_K562_25kb_smote_dist[6,2],2) + round(rfperf_K562_25kb_smote_dist[7,2],2))/2 #balanced accuracy
								)
all_rfperfs_K562_25kb$Resampling <- factor(all_rfperfs_K562_25kb$Resampling, levels=c("None","ROS","RUS","SMOTE"))
all_rfperfs_K562_25kb$Predictor <- factor(all_rfperfs_K562_25kb$Predictor, levels = c("OC", "OP", "Distance"))
all_rfperfs_K562_25kb$Metric <- factor(all_rfperfs_K562_25kb$Metric, levels = c("Accuracy",
                                                                                      "AUPRC",
                                                                                      "F1-Score",
                                                                                      "MCC",
                                                                                      "Informedness",
                                                                                      "Balanced Acc"))

#comparing only f1 and mcc and balanced acc

ggplot(all_rfperfs_K562_25kb[which(all_rfperfs_K562_25kb$Metric=="F1-Score" |
                                               all_rfperfs_K562_25kb$Metric=="MCC" |
                                     all_rfperfs_K562_25kb$Metric=="Balanced Acc"),],
                       aes(x=Resampling, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(~ Metric) +
  xlab("Resampling Technique") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

###50kb

```{r}
#OC
##None
rfperf_K562_50kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_50kb_rus_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_50kb_ros_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_50kb_smote_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_50kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_50kb_rus_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_50kb_ros_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_50kb_smote_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_50kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_50kb_rus_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_50kb_ros_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_50kb_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_K562_50kb <- data.frame(Resampling=c(rep("None", 18),
                                       rep("RUS", 18),
                                       rep("ROS", 18),
                                       rep("SMOTE", 18)),
                          Predictor=rep(c(rep("OC", 6),
                                      rep("OP", 6),
                                      rep("Distance", 6)),4),
                          Metric=rep(c("Accuracy", 
                                       "AUPRC", 
                                       "F1-Score", 
                                       "MCC",
                                       "Informedness",
                                       "Balanced Acc"), 12))

all_rfperfs_K562_50kb$Performance <- c(round(rfperf_K562_50kb_none_oc[9,2],2), #accuracy
                                round(rfperf_K562_50kb_none_oc[19,2],2), #auprc
                                round(rfperf_K562_50kb_none_oc[16,2],2), #f1
                                round(rfperf_K562_50kb_none_oc[15,2],2), #mcc
								round(rfperf_K562_50kb_none_oc[18,2],2), #youden
								(round(rfperf_K562_50kb_none_oc[6,2],2) + round(rfperf_K562_50kb_none_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_50kb_none_op[9,2],2), #accuracy
                                round(rfperf_K562_50kb_none_op[19,2],2), #auprc
                                round(rfperf_K562_50kb_none_op[16,2],2), #f1
                                round(rfperf_K562_50kb_none_op[15,2],2), #mcc
								round(rfperf_K562_50kb_none_op[18,2],2), #youden
								(round(rfperf_K562_50kb_none_op[6,2],2) + round(rfperf_K562_50kb_none_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_50kb_none_dist[9,2],2), #accuracy
                                round(rfperf_K562_50kb_none_dist[19,2],2), #auprc
                                round(rfperf_K562_50kb_none_dist[16,2],2), #f1
                                round(rfperf_K562_50kb_none_dist[15,2],2), #mcc
								round(rfperf_K562_50kb_none_dist[18,2],2), #youden
								(round(rfperf_K562_50kb_none_dist[6,2],2) + round(rfperf_K562_50kb_none_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_50kb_rus_oc[9,2],2), #accuracy
                                round(rfperf_K562_50kb_rus_oc[19,2],2), #auprc
                                round(rfperf_K562_50kb_rus_oc[16,2],2), #f1
                                round(rfperf_K562_50kb_rus_oc[15,2],2), #mcc
								round(rfperf_K562_50kb_rus_oc[18,2],2), #youden
								(round(rfperf_K562_50kb_rus_oc[6,2],2) + round(rfperf_K562_50kb_rus_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_50kb_rus_op[9,2],2), #accuracy
                                round(rfperf_K562_50kb_rus_op[19,2],2), #auprc
                                round(rfperf_K562_50kb_rus_op[16,2],2), #f1
                                round(rfperf_K562_50kb_rus_op[15,2],2), #mcc
								round(rfperf_K562_50kb_rus_op[18,2],2), #youden
								(round(rfperf_K562_50kb_rus_op[6,2],2) + round(rfperf_K562_50kb_rus_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_50kb_rus_dist[9,2],2), #accuracy
                                round(rfperf_K562_50kb_rus_dist[19,2],2), #auprc
                                round(rfperf_K562_50kb_rus_dist[16,2],2), #f1
                                round(rfperf_K562_50kb_rus_dist[15,2],2), #mcc
								round(rfperf_K562_50kb_rus_dist[18,2],2), #youden
								(round(rfperf_K562_50kb_rus_dist[6,2],2) + round(rfperf_K562_50kb_rus_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_50kb_ros_oc[9,2],2), #accuracy
                                round(rfperf_K562_50kb_ros_oc[19,2],2), #auprc
                                round(rfperf_K562_50kb_ros_oc[16,2],2), #f1
                                round(rfperf_K562_50kb_ros_oc[15,2],2), #mcc
								round(rfperf_K562_50kb_ros_oc[18,2],2), #youden
								(round(rfperf_K562_50kb_ros_oc[6,2],2) + round(rfperf_K562_50kb_ros_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_50kb_ros_op[9,2],2), #accuracy
                                round(rfperf_K562_50kb_ros_op[19,2],2), #auprc
                                round(rfperf_K562_50kb_ros_op[16,2],2), #f1
                                round(rfperf_K562_50kb_ros_op[15,2],2), #mcc
								round(rfperf_K562_50kb_ros_op[18,2],2), #youden
								(round(rfperf_K562_50kb_ros_op[6,2],2) + round(rfperf_K562_50kb_ros_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_50kb_ros_dist[9,2],2), #accuracy
                                round(rfperf_K562_50kb_ros_dist[19,2],2), #auprc
                                round(rfperf_K562_50kb_ros_dist[16,2],2), #f1
                                round(rfperf_K562_50kb_ros_dist[15,2],2), #mcc
								round(rfperf_K562_50kb_ros_dist[18,2],2), #youden
								(round(rfperf_K562_50kb_ros_dist[6,2],2) + round(rfperf_K562_50kb_ros_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_50kb_smote_oc[9,2],2), #accuracy
                                round(rfperf_K562_50kb_smote_oc[19,2],2), #auprc
                                round(rfperf_K562_50kb_smote_oc[16,2],2), #f1
                                round(rfperf_K562_50kb_smote_oc[15,2],2), #mcc
								round(rfperf_K562_50kb_smote_oc[18,2],2), #youden
								(round(rfperf_K562_50kb_smote_oc[6,2],2) + round(rfperf_K562_50kb_smote_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_50kb_smote_op[9,2],2), #accuracy
                                round(rfperf_K562_50kb_smote_op[19,2],2), #auprc
                                round(rfperf_K562_50kb_smote_op[16,2],2), #f1
                                round(rfperf_K562_50kb_smote_op[15,2],2), #mcc
								round(rfperf_K562_50kb_smote_op[18,2],2), #youden
								(round(rfperf_K562_50kb_smote_op[6,2],2) + round(rfperf_K562_50kb_smote_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_50kb_smote_dist[9,2],2), #accuracy
                                round(rfperf_K562_50kb_smote_dist[19,2],2), #auprc
                                round(rfperf_K562_50kb_smote_dist[16,2],2), #f1
                                round(rfperf_K562_50kb_smote_dist[15,2],2), #mcc
								round(rfperf_K562_50kb_smote_dist[18,2],2), #youden
								(round(rfperf_K562_50kb_smote_dist[6,2],2) + round(rfperf_K562_50kb_smote_dist[7,2],2))/2 #balanced accuracy
								)
all_rfperfs_K562_50kb$Resampling <- factor(all_rfperfs_K562_50kb$Resampling, levels=c("None","ROS","RUS","SMOTE"))
all_rfperfs_K562_50kb$Predictor <- factor(all_rfperfs_K562_50kb$Predictor, levels = c("OC", "OP", "Distance"))
all_rfperfs_K562_50kb$Metric <- factor(all_rfperfs_K562_50kb$Metric, levels = c("Accuracy",
                                                                                      "AUPRC",
                                                                                      "F1-Score",
                                                                                      "MCC",
                                                                                      "Informedness",
                                                                                      "Balanced Acc"))

#comparing only f1 and mcc and balanced acc

ggplot(all_rfperfs_K562_50kb[which(all_rfperfs_K562_50kb$Metric=="F1-Score" |
                                               all_rfperfs_K562_50kb$Metric=="MCC" |
                                     all_rfperfs_K562_50kb$Metric=="Balanced Acc"),],
                       aes(x=Resampling, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(~ Metric) +
  xlab("Resampling Technique") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

###100kb

```{r}
#OC
##None
rfperf_K562_100kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_100kb_rus_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_100kb_ros_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_100kb_smote_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_100kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_100kb_rus_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_100kb_ros_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_100kb_smote_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_100kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

##RUS
rfperf_K562_100kb_rus_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

##ROS
rfperf_K562_100kb_ros_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

##SMOTE
rfperf_K562_100kb_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_K562_100kb <- data.frame(Resampling=c(rep("None", 18),
                                       rep("RUS", 18),
                                       rep("ROS", 18),
                                       rep("SMOTE", 18)),
                          Predictor=rep(c(rep("OC", 6),
                                      rep("OP", 6),
                                      rep("Distance", 6)),4),
                          Metric=rep(c("Accuracy", 
                                       "AUPRC", 
                                       "F1-Score", 
                                       "MCC",
                                       "Informedness",
                                       "Balanced Acc"), 12))

all_rfperfs_K562_100kb$Performance <- c(round(rfperf_K562_100kb_none_oc[9,2],2), #accuracy
                                round(rfperf_K562_100kb_none_oc[19,2],2), #auprc
                                round(rfperf_K562_100kb_none_oc[16,2],2), #f1
                                round(rfperf_K562_100kb_none_oc[15,2],2), #mcc
								round(rfperf_K562_100kb_none_oc[18,2],2), #youden
								(round(rfperf_K562_100kb_none_oc[6,2],2) + round(rfperf_K562_100kb_none_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_100kb_none_op[9,2],2), #accuracy
                                round(rfperf_K562_100kb_none_op[19,2],2), #auprc
                                round(rfperf_K562_100kb_none_op[16,2],2), #f1
                                round(rfperf_K562_100kb_none_op[15,2],2), #mcc
								round(rfperf_K562_100kb_none_op[18,2],2), #youden
								(round(rfperf_K562_100kb_none_op[6,2],2) + round(rfperf_K562_100kb_none_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_100kb_none_dist[9,2],2), #accuracy
                                round(rfperf_K562_100kb_none_dist[19,2],2), #auprc
                                round(rfperf_K562_100kb_none_dist[16,2],2), #f1
                                round(rfperf_K562_100kb_none_dist[15,2],2), #mcc
								round(rfperf_K562_100kb_none_dist[18,2],2), #youden
								(round(rfperf_K562_100kb_none_dist[6,2],2) + round(rfperf_K562_100kb_none_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_100kb_rus_oc[9,2],2), #accuracy
                                round(rfperf_K562_100kb_rus_oc[19,2],2), #auprc
                                round(rfperf_K562_100kb_rus_oc[16,2],2), #f1
                                round(rfperf_K562_100kb_rus_oc[15,2],2), #mcc
								round(rfperf_K562_100kb_rus_oc[18,2],2), #youden
								(round(rfperf_K562_100kb_rus_oc[6,2],2) + round(rfperf_K562_100kb_rus_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_100kb_rus_op[9,2],2), #accuracy
                                round(rfperf_K562_100kb_rus_op[19,2],2), #auprc
                                round(rfperf_K562_100kb_rus_op[16,2],2), #f1
                                round(rfperf_K562_100kb_rus_op[15,2],2), #mcc
								round(rfperf_K562_100kb_rus_op[18,2],2), #youden
								(round(rfperf_K562_100kb_rus_op[6,2],2) + round(rfperf_K562_100kb_rus_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_100kb_rus_dist[9,2],2), #accuracy
                                round(rfperf_K562_100kb_rus_dist[19,2],2), #auprc
                                round(rfperf_K562_100kb_rus_dist[16,2],2), #f1
                                round(rfperf_K562_100kb_rus_dist[15,2],2), #mcc
								round(rfperf_K562_100kb_rus_dist[18,2],2), #youden
								(round(rfperf_K562_100kb_rus_dist[6,2],2) + round(rfperf_K562_100kb_rus_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_100kb_ros_oc[9,2],2), #accuracy
                                round(rfperf_K562_100kb_ros_oc[19,2],2), #auprc
                                round(rfperf_K562_100kb_ros_oc[16,2],2), #f1
                                round(rfperf_K562_100kb_ros_oc[15,2],2), #mcc
								round(rfperf_K562_100kb_ros_oc[18,2],2), #youden
								(round(rfperf_K562_100kb_ros_oc[6,2],2) + round(rfperf_K562_100kb_ros_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_100kb_ros_op[9,2],2), #accuracy
                                round(rfperf_K562_100kb_ros_op[19,2],2), #auprc
                                round(rfperf_K562_100kb_ros_op[16,2],2), #f1
                                round(rfperf_K562_100kb_ros_op[15,2],2), #mcc
								round(rfperf_K562_100kb_ros_op[18,2],2), #youden
								(round(rfperf_K562_100kb_ros_op[6,2],2) + round(rfperf_K562_100kb_ros_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_100kb_ros_dist[9,2],2), #accuracy
                                round(rfperf_K562_100kb_ros_dist[19,2],2), #auprc
                                round(rfperf_K562_100kb_ros_dist[16,2],2), #f1
                                round(rfperf_K562_100kb_ros_dist[15,2],2), #mcc
								round(rfperf_K562_100kb_ros_dist[18,2],2), #youden
								(round(rfperf_K562_100kb_ros_dist[6,2],2) + round(rfperf_K562_100kb_ros_dist[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_100kb_smote_oc[9,2],2), #accuracy
                                round(rfperf_K562_100kb_smote_oc[19,2],2), #auprc
                                round(rfperf_K562_100kb_smote_oc[16,2],2), #f1
                                round(rfperf_K562_100kb_smote_oc[15,2],2), #mcc
								round(rfperf_K562_100kb_smote_oc[18,2],2), #youden
								(round(rfperf_K562_100kb_smote_oc[6,2],2) + round(rfperf_K562_100kb_smote_oc[7,2],2))/2, #balanced accuracy

                                round(rfperf_K562_100kb_smote_op[9,2],2), #accuracy
                                round(rfperf_K562_100kb_smote_op[19,2],2), #auprc
                                round(rfperf_K562_100kb_smote_op[16,2],2), #f1
                                round(rfperf_K562_100kb_smote_op[15,2],2), #mcc
								round(rfperf_K562_100kb_smote_op[18,2],2), #youden
								(round(rfperf_K562_100kb_smote_op[6,2],2) + round(rfperf_K562_100kb_smote_op[7,2],2))/2, #balanced accuracy

								round(rfperf_K562_100kb_smote_dist[9,2],2), #accuracy
                                round(rfperf_K562_100kb_smote_dist[19,2],2), #auprc
                                round(rfperf_K562_100kb_smote_dist[16,2],2), #f1
                                round(rfperf_K562_100kb_smote_dist[15,2],2), #mcc
								round(rfperf_K562_100kb_smote_dist[18,2],2), #youden
								(round(rfperf_K562_100kb_smote_dist[6,2],2) + round(rfperf_K562_100kb_smote_dist[7,2],2))/2 #balanced accuracy
								)
all_rfperfs_K562_100kb$Resampling <- factor(all_rfperfs_K562_100kb$Resampling, levels=c("None","ROS","RUS","SMOTE"))
all_rfperfs_K562_100kb$Predictor <- factor(all_rfperfs_K562_100kb$Predictor, levels = c("OC", "OP", "Distance"))
all_rfperfs_K562_100kb$Metric <- factor(all_rfperfs_K562_100kb$Metric, levels = c("Accuracy",
                                                                                      "AUPRC",
                                                                                      "F1-Score",
                                                                                      "MCC",
                                                                                      "Informedness",
                                                                                      "Balanced Acc"))

#comparing only f1 and mcc and balanced acc

ggplot(all_rfperfs_K562_100kb[which(all_rfperfs_K562_100kb$Metric=="F1-Score" |
                                               all_rfperfs_K562_100kb$Metric=="MCC" |
                                      all_rfperfs_K562_100kb$Metric=="Balanced Acc"),],
                       aes(x=Resampling, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(~ Metric) +
  xlab("Resampling Technique") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

##Plots

```{r}

figure <- ggarrange(p1,p2,ncol = 2, nrow = 1, common.legend = TRUE, legend = "top")
figure
annotate_figure(figure,
    bottom = text_grob("Resampling Technique", color = "black", size = 20),
    left = text_grob("Performance", color = "black", rot = 90, size=20)
)

figure <- ggarrange(p25, p25.2, p50, p50.2, p100, p100.2, 
          ncol=2, nrow=3, 
          common.legend = TRUE, 
          legend = "top", 
          labels = c("A", "D", "B", "E", "C", "F"), 
          font.label = list(size=20))
figure
annotate_figure(figure,
    bottom = text_grob("Resampling Technique", color = "black", size = 20),
    left = text_grob("Performance", color = "black", rot = 90, size=20)
)

```

##gm12878

```{r}
#OC
##None
rfperf_gm12878_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_10kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_10kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_gm12878_25kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_25kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_25kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_gm12878_50kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_50kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_50kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_gm12878_100kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_100kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_100kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_10kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_10kb_ROS_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_10kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_25kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_25kb_ROS_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_25kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_50kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_50kb_ROS_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_50kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_100kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_100kb_ROS_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_100kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_10kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_10kb_RUS_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_10kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_25kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_25kb_RUS_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_25kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_50kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_50kb_RUS_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_50kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_100kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_100kb_RUS_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_100kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_10kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_10kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_10kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_25kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_25kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_25kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_50kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_50kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_50kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_100kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_100kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_100kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_gm12878 <- data.frame(Resolution=rep(c(rep("10 kb", 12),
                                                        rep("25 kb", 12),
                                                        rep("50 kb", 12),
                                                        rep("100 kb", 12)),4),
                                       Predictor=rep(c("OC","OP","Distance"),64),
                                       Metric=c(rep("AUC",48),
                                                rep("AUPRC",48),
                                                rep("F1-Score",48), 
                                                rep("MCC",48)),
                                       Resampling=rep(c(rep("None",3),
                                                    rep("ROS",3),
                                                    rep("RUS",3),
                                                    rep("SMOTE",3)),16))
all_rfperfs_gm12878$Resolution <- factor(all_rfperfs_gm12878$Resolution, levels=c("10 kb","25 kb","50 kb","100 kb"))
all_rfperfs_gm12878$Predictor <- factor(all_rfperfs_gm12878$Predictor, levels=c("OC", "OP", "Distance"))
all_rfperfs_gm12878$Metric <- factor(all_rfperfs_gm12878$Metric, levels=c("AUC", "AUPRC", "F1-Score", "MCC"))
all_rfperfs_gm12878$Resampling <- factor(all_rfperfs_gm12878$Resampling, levels=c("None", "ROS", "RUS", "SMOTE"))

all_rfperfs_gm12878$Performance <- c(round(rfperf_gm12878_10kb_none_oc[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_none_op[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_none_dist[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_ROS_op[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_RUS_op[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_gm12878_25kb_none_oc[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_none_op[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_none_dist[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_ROS_op[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_RUS_op[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_gm12878_50kb_none_oc[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_none_op[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_none_dist[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_ROS_op[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_RUS_op[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_gm12878_100kb_none_oc[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_none_op[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_none_dist[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_ROS_op[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_RUS_op[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_gm12878_10kb_none_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_none_op[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_none_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_gm12878_25kb_none_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_none_op[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_none_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_gm12878_50kb_none_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_none_op[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_none_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_gm12878_100kb_none_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_none_op[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_none_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_gm12878_10kb_none_oc[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_none_op[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_none_dist[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_ROS_op[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_RUS_op[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_gm12878_25kb_none_oc[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_none_op[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_none_dist[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_ROS_op[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_RUS_op[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_gm12878_50kb_none_oc[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_none_op[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_none_dist[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_ROS_op[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_RUS_op[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_gm12878_100kb_none_oc[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_none_op[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_none_dist[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_ROS_op[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_RUS_op[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_gm12878_10kb_none_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_none_op[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_none_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_gm12878_25kb_none_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_none_op[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_none_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_gm12878_50kb_none_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_none_op[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_none_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_gm12878_100kb_none_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_none_op[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_none_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_SMOTE_dist[15,2],2) #mcc
)

p1 <- ggplot(all_rfperfs_gm12878,
       aes(x=Resolution, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(Metric ~ Resampling) +
  xlab("") + ylab("") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))
```

##k562

```{r}
#OC
##None
rfperf_K562_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_10kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_10kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_K562_25kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_25kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_25kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_K562_50kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_50kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_50kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_K562_100kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_100kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_100kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_K562_10kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_K562_10kb_ROS_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_K562_10kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_K562_25kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_K562_25kb_ROS_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_K562_25kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_K562_50kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_K562_50kb_ROS_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_K562_50kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_K562_100kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_K562_100kb_ROS_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_K562_100kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_K562_10kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_K562_10kb_RUS_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_K562_10kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_K562_25kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_K562_25kb_RUS_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_K562_25kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_K562_50kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_K562_50kb_RUS_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_K562_50kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_K562_100kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_K562_100kb_RUS_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_K562_100kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_K562_10kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_K562_10kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_K562_10kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_K562_25kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_K562_25kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_K562_25kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_K562_50kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_K562_50kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_K562_50kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_K562_100kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_K562_100kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_K562_100kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_K562 <- data.frame(Resolution=rep(c(rep("10 kb", 12),
                                                        rep("25 kb", 12),
                                                        rep("50 kb", 12),
                                                        rep("100 kb", 12)),4),
                                       Predictor=rep(c("OC","OP","Distance"),64),
                                       Metric=c(rep("AUC",48),
                                                rep("AUPRC",48),
                                                rep("F1-Score",48), 
                                                rep("MCC",48)),
                                       Resampling=rep(c(rep("None",3),
                                                    rep("ROS",3),
                                                    rep("RUS",3),
                                                    rep("SMOTE",3)),16))
all_rfperfs_K562$Resolution <- factor(all_rfperfs_K562$Resolution, levels=c("10 kb","25 kb","50 kb","100 kb"))
all_rfperfs_K562$Predictor <- factor(all_rfperfs_K562$Predictor, levels=c("OC", "OP", "Distance"))
all_rfperfs_K562$Metric <- factor(all_rfperfs_K562$Metric, levels=c("AUC", "AUPRC", "F1-Score", "MCC"))
all_rfperfs_K562$Resampling <- factor(all_rfperfs_K562$Resampling, levels=c("None", "ROS", "RUS", "SMOTE"))

all_rfperfs_K562$Performance <- c(round(rfperf_K562_10kb_none_oc[17,2],2), #auc
                                          round(rfperf_K562_10kb_none_op[17,2],2), #auc
                                          round(rfperf_K562_10kb_none_dist[17,2],2), #auc
                                          round(rfperf_K562_10kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_K562_10kb_ROS_op[17,2],2), #auc
                                          round(rfperf_K562_10kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_K562_10kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_K562_10kb_RUS_op[17,2],2), #auc
                                          round(rfperf_K562_10kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_K562_10kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_K562_10kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_K562_10kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_K562_25kb_none_oc[17,2],2), #auc
                                          round(rfperf_K562_25kb_none_op[17,2],2), #auc
                                          round(rfperf_K562_25kb_none_dist[17,2],2), #auc
                                          round(rfperf_K562_25kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_K562_25kb_ROS_op[17,2],2), #auc
                                          round(rfperf_K562_25kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_K562_25kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_K562_25kb_RUS_op[17,2],2), #auc
                                          round(rfperf_K562_25kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_K562_25kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_K562_25kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_K562_25kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_K562_50kb_none_oc[17,2],2), #auc
                                          round(rfperf_K562_50kb_none_op[17,2],2), #auc
                                          round(rfperf_K562_50kb_none_dist[17,2],2), #auc
                                          round(rfperf_K562_50kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_K562_50kb_ROS_op[17,2],2), #auc
                                          round(rfperf_K562_50kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_K562_50kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_K562_50kb_RUS_op[17,2],2), #auc
                                          round(rfperf_K562_50kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_K562_50kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_K562_50kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_K562_50kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_K562_100kb_none_oc[17,2],2), #auc
                                          round(rfperf_K562_100kb_none_op[17,2],2), #auc
                                          round(rfperf_K562_100kb_none_dist[17,2],2), #auc
                                          round(rfperf_K562_100kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_K562_100kb_ROS_op[17,2],2), #auc
                                          round(rfperf_K562_100kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_K562_100kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_K562_100kb_RUS_op[17,2],2), #auc
                                          round(rfperf_K562_100kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_K562_100kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_K562_100kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_K562_100kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_K562_10kb_none_oc[19,2],2), #auprc
                                          round(rfperf_K562_10kb_none_op[19,2],2), #auprc
                                          round(rfperf_K562_10kb_none_dist[19,2],2), #auprc
                                          round(rfperf_K562_10kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_K562_10kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_K562_10kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_K562_10kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_K562_10kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_K562_10kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_K562_10kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_K562_10kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_K562_10kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_K562_25kb_none_oc[19,2],2), #auprc
                                          round(rfperf_K562_25kb_none_op[19,2],2), #auprc
                                          round(rfperf_K562_25kb_none_dist[19,2],2), #auprc
                                          round(rfperf_K562_25kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_K562_25kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_K562_25kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_K562_25kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_K562_25kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_K562_25kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_K562_25kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_K562_25kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_K562_25kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_K562_50kb_none_oc[19,2],2), #auprc
                                          round(rfperf_K562_50kb_none_op[19,2],2), #auprc
                                          round(rfperf_K562_50kb_none_dist[19,2],2), #auprc
                                          round(rfperf_K562_50kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_K562_50kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_K562_50kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_K562_50kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_K562_50kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_K562_50kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_K562_50kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_K562_50kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_K562_50kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_K562_100kb_none_oc[19,2],2), #auprc
                                          round(rfperf_K562_100kb_none_op[19,2],2), #auprc
                                          round(rfperf_K562_100kb_none_dist[19,2],2), #auprc
                                          round(rfperf_K562_100kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_K562_100kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_K562_100kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_K562_100kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_K562_100kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_K562_100kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_K562_100kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_K562_100kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_K562_100kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_K562_10kb_none_oc[16,2],2), #f1
                                          round(rfperf_K562_10kb_none_op[16,2],2), #f1
                                          round(rfperf_K562_10kb_none_dist[16,2],2), #f1
                                          round(rfperf_K562_10kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_K562_10kb_ROS_op[16,2],2), #f1
                                          round(rfperf_K562_10kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_K562_10kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_K562_10kb_RUS_op[16,2],2), #f1
                                          round(rfperf_K562_10kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_K562_10kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_K562_10kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_K562_10kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_K562_25kb_none_oc[16,2],2), #f1
                                          round(rfperf_K562_25kb_none_op[16,2],2), #f1
                                          round(rfperf_K562_25kb_none_dist[16,2],2), #f1
                                          round(rfperf_K562_25kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_K562_25kb_ROS_op[16,2],2), #f1
                                          round(rfperf_K562_25kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_K562_25kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_K562_25kb_RUS_op[16,2],2), #f1
                                          round(rfperf_K562_25kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_K562_25kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_K562_25kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_K562_25kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_K562_50kb_none_oc[16,2],2), #f1
                                          round(rfperf_K562_50kb_none_op[16,2],2), #f1
                                          round(rfperf_K562_50kb_none_dist[16,2],2), #f1
                                          round(rfperf_K562_50kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_K562_50kb_ROS_op[16,2],2), #f1
                                          round(rfperf_K562_50kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_K562_50kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_K562_50kb_RUS_op[16,2],2), #f1
                                          round(rfperf_K562_50kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_K562_50kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_K562_50kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_K562_50kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_K562_100kb_none_oc[16,2],2), #f1
                                          round(rfperf_K562_100kb_none_op[16,2],2), #f1
                                          round(rfperf_K562_100kb_none_dist[16,2],2), #f1
                                          round(rfperf_K562_100kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_K562_100kb_ROS_op[16,2],2), #f1
                                          round(rfperf_K562_100kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_K562_100kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_K562_100kb_RUS_op[16,2],2), #f1
                                          round(rfperf_K562_100kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_K562_100kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_K562_100kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_K562_100kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_K562_10kb_none_oc[15,2],2), #mcc
                                          round(rfperf_K562_10kb_none_op[15,2],2), #mcc
                                          round(rfperf_K562_10kb_none_dist[15,2],2), #mcc
                                          round(rfperf_K562_10kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_K562_10kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_K562_10kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_K562_10kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_K562_10kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_K562_10kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_K562_10kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_K562_10kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_K562_10kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_K562_25kb_none_oc[15,2],2), #mcc
                                          round(rfperf_K562_25kb_none_op[15,2],2), #mcc
                                          round(rfperf_K562_25kb_none_dist[15,2],2), #mcc
                                          round(rfperf_K562_25kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_K562_25kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_K562_25kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_K562_25kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_K562_25kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_K562_25kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_K562_25kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_K562_25kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_K562_25kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_K562_50kb_none_oc[15,2],2), #mcc
                                          round(rfperf_K562_50kb_none_op[15,2],2), #mcc
                                          round(rfperf_K562_50kb_none_dist[15,2],2), #mcc
                                          round(rfperf_K562_50kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_K562_50kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_K562_50kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_K562_50kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_K562_50kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_K562_50kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_K562_50kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_K562_50kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_K562_50kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_K562_100kb_none_oc[15,2],2), #mcc
                                          round(rfperf_K562_100kb_none_op[15,2],2), #mcc
                                          round(rfperf_K562_100kb_none_dist[15,2],2), #mcc
                                          round(rfperf_K562_100kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_K562_100kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_K562_100kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_K562_100kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_K562_100kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_K562_100kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_K562_100kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_K562_100kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_K562_100kb_SMOTE_dist[15,2],2) #mcc
)

p2 <- ggplot(all_rfperfs_K562,
       aes(x=Resolution, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(Metric ~ Resampling) +
  xlab("") + ylab("") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))
```

##Plots

```{r}
p1 <- p1 + labs(title = "GM12878") + theme(plot.title = element_text(size=20))
p2 <- p2 + labs(title = "K562") + theme(plot.title = element_text(size=20))

figure <- ggarrange(p1,p2,
                ncol = 2, nrow = 1, 
                common.legend = TRUE, 
                legend = "top",
                labels = "AUTO",
                font.label = list(size=20))
figure
annotate_figure(figure,
    bottom = text_grob("Resolution", color = "black", size = 20),
    left = text_grob("Performance", color = "black", rot = 90, size=20)
)

```

##Plots

```{r}
c <- ggplot(all_rfperfs_gm12878[which(all_rfperfs_gm12878$Predictor=="OC" & (all_rfperfs_gm12878$Metric=="F1-Score" | all_rfperfs_gm12878$Metric=="MCC")),], aes(x=Resolution, y=Performance, color=Resampling, group=Resampling)) +
    geom_line(size=2)+
    geom_point(size=5)+
    facet_grid(~ Metric)+
    theme_minimal() +
    theme_bw()+
    xlab("") + ylab("") + ylim(-.02,.5) +
    ggtitle("Overlap Counts") +
    scale_color_discrete(name="Re-Sampling Technique")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    plot.title = element_text(size=20))

p <- ggplot(all_rfperfs_gm12878[which(all_rfperfs_gm12878$Predictor=="OP" & (all_rfperfs_gm12878$Metric=="F1-Score" | all_rfperfs_gm12878$Metric=="MCC")),], aes(x=Resolution, y=Performance, color=Resampling, group=Resampling)) +
    geom_line(size=2)+
    geom_point(size=5)+
    facet_grid(~ Metric)+
    theme_minimal() +
    theme_bw()+
    xlab("") + ylab("") + ylim(-.02,.5) +
    ggtitle("Overlap Percents") +
    scale_color_discrete(name="Re-Sampling Technique")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    plot.title = element_text(size=20))

d <- ggplot(all_rfperfs_gm12878[which(all_rfperfs_gm12878$Predictor=="Distance" & (all_rfperfs_gm12878$Metric=="F1-Score" | all_rfperfs_gm12878$Metric=="MCC")),], aes(x=Resolution, y=Performance, color=Resampling, group=Resampling)) +
    geom_line(size=2)+
    geom_point(size=5)+
    facet_grid(~ Metric)+
    theme_minimal() +
    theme_bw()+
    xlab("") + ylab("") + ylim(-.02,.5) +
    ggtitle("Distance") +
    scale_color_discrete(name="Re-Sampling Technique")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    plot.title = element_text(size=20))

figure <- ggarrange(c,p,d,
                ncol = 3, nrow = 1, 
                common.legend = TRUE, 
                legend = "top",
                labels = "AUTO",
                font.label = list(size=20))
figure
annotate_figure(figure,
    bottom = text_grob("Resolution", color = "black", size = 20),
    left = text_grob("Performance", color = "black", rot = 90, size=20)
)

```