### 2019-04-07 Scripts/evaluating_models_with_dist_preds.Rmd

---
title: "Comparing top performing models"
author: "Spiro Stilianoudakis"
date: "March 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# Comparing models with distance type predictors

##gm12878

```{r}
#OC
##None
rfperf_gm12878_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_10kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_10kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_gm12878_25kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_25kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_25kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_gm12878_50kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_50kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_50kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_gm12878_100kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_100kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_100kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_10kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_10kb_ROS_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_10kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_25kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_25kb_ROS_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_25kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_50kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_50kb_ROS_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_50kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_100kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_100kb_ROS_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_100kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_10kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_10kb_RUS_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_10kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_25kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_25kb_RUS_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_25kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_50kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_50kb_RUS_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_50kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_100kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_100kb_RUS_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_100kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_10kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_10kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_10kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_25kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_25kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_25kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_50kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_50kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_50kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_100kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_100kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_100kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_gm12878 <- data.frame(Resolution=rep(c(rep("10 kb", 12),
                                                        rep("25 kb", 12),
                                                        rep("50 kb", 12),
                                                        rep("100 kb", 12)),4),
                                       Predictor=rep(c("OC","OP","Distance"),64),
                                       Metric=c(rep("AUC",48),
                                                rep("AUPRC",48),
                                                rep("F1-Score",48), 
                                                rep("MCC",48)),
                                       Resampling=rep(c(rep("None",3),
                                                    rep("ROS",3),
                                                    rep("RUS",3),
                                                    rep("SMOTE",3)),16))
all_rfperfs_gm12878$Resolution <- factor(all_rfperfs_gm12878$Resolution, levels=c("10 kb","25 kb","50 kb","100 kb"))
all_rfperfs_gm12878$Predictor <- factor(all_rfperfs_gm12878$Predictor, levels=c("OC", "OP", "Distance"))
all_rfperfs_gm12878$Metric <- factor(all_rfperfs_gm12878$Metric, levels=c("AUC", "AUPRC", "F1-Score", "MCC"))
all_rfperfs_gm12878$Resampling <- factor(all_rfperfs_gm12878$Resampling, levels=c("None", "ROS", "RUS", "SMOTE"))

all_rfperfs_gm12878$Performance <- c(round(rfperf_gm12878_10kb_none_oc[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_none_op[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_none_dist[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_ROS_op[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_RUS_op[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_gm12878_10kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_gm12878_25kb_none_oc[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_none_op[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_none_dist[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_ROS_op[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_RUS_op[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_gm12878_25kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_gm12878_50kb_none_oc[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_none_op[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_none_dist[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_ROS_op[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_RUS_op[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_gm12878_50kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_gm12878_100kb_none_oc[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_none_op[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_none_dist[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_ROS_op[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_RUS_op[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_gm12878_100kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_gm12878_10kb_none_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_none_op[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_none_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_gm12878_10kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_gm12878_25kb_none_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_none_op[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_none_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_gm12878_25kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_gm12878_50kb_none_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_none_op[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_none_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_gm12878_50kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_gm12878_100kb_none_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_none_op[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_none_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_gm12878_100kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_gm12878_10kb_none_oc[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_none_op[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_none_dist[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_ROS_op[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_RUS_op[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_gm12878_10kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_gm12878_25kb_none_oc[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_none_op[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_none_dist[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_ROS_op[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_RUS_op[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_gm12878_25kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_gm12878_50kb_none_oc[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_none_op[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_none_dist[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_ROS_op[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_RUS_op[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_gm12878_50kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_gm12878_100kb_none_oc[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_none_op[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_none_dist[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_ROS_op[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_RUS_op[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_gm12878_100kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_gm12878_10kb_none_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_none_op[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_none_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_gm12878_10kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_gm12878_25kb_none_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_none_op[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_none_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_gm12878_25kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_gm12878_50kb_none_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_none_op[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_none_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_gm12878_50kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_gm12878_100kb_none_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_none_op[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_none_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_gm12878_100kb_SMOTE_dist[15,2],2) #mcc
)

p1 <- ggplot(all_rfperfs_gm12878[which(all_rfperfs_gm12878$Predictor=="Distance"),], aes(x=Resolution, y=Performance, color=Resampling, group=Resampling)) +
    geom_line(size=2)+
    geom_point(size=5)+
    facet_grid(~ Metric)+
    theme_minimal() +
    theme_bw()+
    xlab("") + ylab("") +
  scale_color_discrete(name="Re-Sampling Technique")+
	theme(axis.text.x = element_text(size=15,
	                                 angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
	                                 ),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
	      plot.title = element_text(size=20))

```

##k562

```{r}
#OC
##None
rfperf_K562_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_10kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_10kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_K562_25kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_25kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_25kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_K562_50kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_50kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_50kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##None
rfperf_K562_100kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_K562_100kb_none_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_K562_100kb_none_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_K562_10kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_K562_10kb_ROS_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_K562_10kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_K562_25kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_K562_25kb_ROS_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_K562_25kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_K562_50kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_K562_50kb_ROS_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_K562_50kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_K562_100kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_K562_100kb_ROS_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_K562_100kb_ROS_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_K562_10kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_K562_10kb_RUS_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_K562_10kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_K562_25kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_K562_25kb_RUS_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_K562_25kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_K562_50kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_K562_50kb_RUS_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_K562_50kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_K562_100kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_K562_100kb_RUS_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_K562_100kb_RUS_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_K562_10kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_K562_10kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_K562_10kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_K562_25kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_K562_25kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_K562_25kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_K562_50kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_K562_50kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_K562_50kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_K562_100kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_K562_100kb_SMOTE_op <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_K562_100kb_SMOTE_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_K562 <- data.frame(Resolution=rep(c(rep("10 kb", 12),
                                                        rep("25 kb", 12),
                                                        rep("50 kb", 12),
                                                        rep("100 kb", 12)),4),
                                       Predictor=rep(c("OC","OP","Distance"),64),
                                       Metric=c(rep("AUC",48),
                                                rep("AUPRC",48),
                                                rep("F1-Score",48), 
                                                rep("MCC",48)),
                                       Resampling=rep(c(rep("None",3),
                                                    rep("ROS",3),
                                                    rep("RUS",3),
                                                    rep("SMOTE",3)),16))
all_rfperfs_K562$Resolution <- factor(all_rfperfs_K562$Resolution, levels=c("10 kb","25 kb","50 kb","100 kb"))
all_rfperfs_K562$Predictor <- factor(all_rfperfs_K562$Predictor, levels=c("OC", "OP", "Distance"))
all_rfperfs_K562$Metric <- factor(all_rfperfs_K562$Metric, levels=c("AUC", "AUPRC", "F1-Score", "MCC"))
all_rfperfs_K562$Resampling <- factor(all_rfperfs_K562$Resampling, levels=c("None", "ROS", "RUS", "SMOTE"))

all_rfperfs_K562$Performance <- c(round(rfperf_K562_10kb_none_oc[17,2],2), #auc
                                          round(rfperf_K562_10kb_none_op[17,2],2), #auc
                                          round(rfperf_K562_10kb_none_dist[17,2],2), #auc
                                          round(rfperf_K562_10kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_K562_10kb_ROS_op[17,2],2), #auc
                                          round(rfperf_K562_10kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_K562_10kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_K562_10kb_RUS_op[17,2],2), #auc
                                          round(rfperf_K562_10kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_K562_10kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_K562_10kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_K562_10kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_K562_25kb_none_oc[17,2],2), #auc
                                          round(rfperf_K562_25kb_none_op[17,2],2), #auc
                                          round(rfperf_K562_25kb_none_dist[17,2],2), #auc
                                          round(rfperf_K562_25kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_K562_25kb_ROS_op[17,2],2), #auc
                                          round(rfperf_K562_25kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_K562_25kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_K562_25kb_RUS_op[17,2],2), #auc
                                          round(rfperf_K562_25kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_K562_25kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_K562_25kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_K562_25kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_K562_50kb_none_oc[17,2],2), #auc
                                          round(rfperf_K562_50kb_none_op[17,2],2), #auc
                                          round(rfperf_K562_50kb_none_dist[17,2],2), #auc
                                          round(rfperf_K562_50kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_K562_50kb_ROS_op[17,2],2), #auc
                                          round(rfperf_K562_50kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_K562_50kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_K562_50kb_RUS_op[17,2],2), #auc
                                          round(rfperf_K562_50kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_K562_50kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_K562_50kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_K562_50kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_K562_100kb_none_oc[17,2],2), #auc
                                          round(rfperf_K562_100kb_none_op[17,2],2), #auc
                                          round(rfperf_K562_100kb_none_dist[17,2],2), #auc
                                          round(rfperf_K562_100kb_ROS_oc[17,2],2), #auc
                                          round(rfperf_K562_100kb_ROS_op[17,2],2), #auc
                                          round(rfperf_K562_100kb_ROS_dist[17,2],2), #auc
                                          round(rfperf_K562_100kb_RUS_oc[17,2],2), #auc
                                          round(rfperf_K562_100kb_RUS_op[17,2],2), #auc
                                          round(rfperf_K562_100kb_RUS_dist[17,2],2), #auc
                                          round(rfperf_K562_100kb_SMOTE_oc[17,2],2), #auc
                                          round(rfperf_K562_100kb_SMOTE_op[17,2],2), #auc
                                          round(rfperf_K562_100kb_SMOTE_dist[17,2],2), #auc

                                     round(rfperf_K562_10kb_none_oc[19,2],2), #auprc
                                          round(rfperf_K562_10kb_none_op[19,2],2), #auprc
                                          round(rfperf_K562_10kb_none_dist[19,2],2), #auprc
                                          round(rfperf_K562_10kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_K562_10kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_K562_10kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_K562_10kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_K562_10kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_K562_10kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_K562_10kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_K562_10kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_K562_10kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_K562_25kb_none_oc[19,2],2), #auprc
                                          round(rfperf_K562_25kb_none_op[19,2],2), #auprc
                                          round(rfperf_K562_25kb_none_dist[19,2],2), #auprc
                                          round(rfperf_K562_25kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_K562_25kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_K562_25kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_K562_25kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_K562_25kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_K562_25kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_K562_25kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_K562_25kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_K562_25kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_K562_50kb_none_oc[19,2],2), #auprc
                                          round(rfperf_K562_50kb_none_op[19,2],2), #auprc
                                          round(rfperf_K562_50kb_none_dist[19,2],2), #auprc
                                          round(rfperf_K562_50kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_K562_50kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_K562_50kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_K562_50kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_K562_50kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_K562_50kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_K562_50kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_K562_50kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_K562_50kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_K562_100kb_none_oc[19,2],2), #auprc
                                          round(rfperf_K562_100kb_none_op[19,2],2), #auprc
                                          round(rfperf_K562_100kb_none_dist[19,2],2), #auprc
                                          round(rfperf_K562_100kb_ROS_oc[19,2],2), #auprc
                                          round(rfperf_K562_100kb_ROS_op[19,2],2), #auprc
                                          round(rfperf_K562_100kb_ROS_dist[19,2],2), #auprc
                                          round(rfperf_K562_100kb_RUS_oc[19,2],2), #auprc
                                          round(rfperf_K562_100kb_RUS_op[19,2],2), #auprc
                                          round(rfperf_K562_100kb_RUS_dist[19,2],2), #auprc
                                          round(rfperf_K562_100kb_SMOTE_oc[19,2],2), #auprc
                                          round(rfperf_K562_100kb_SMOTE_op[19,2],2), #auprc
                                          round(rfperf_K562_100kb_SMOTE_dist[19,2],2), #auprc

                                     round(rfperf_K562_10kb_none_oc[16,2],2), #f1
                                          round(rfperf_K562_10kb_none_op[16,2],2), #f1
                                          round(rfperf_K562_10kb_none_dist[16,2],2), #f1
                                          round(rfperf_K562_10kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_K562_10kb_ROS_op[16,2],2), #f1
                                          round(rfperf_K562_10kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_K562_10kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_K562_10kb_RUS_op[16,2],2), #f1
                                          round(rfperf_K562_10kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_K562_10kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_K562_10kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_K562_10kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_K562_25kb_none_oc[16,2],2), #f1
                                          round(rfperf_K562_25kb_none_op[16,2],2), #f1
                                          round(rfperf_K562_25kb_none_dist[16,2],2), #f1
                                          round(rfperf_K562_25kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_K562_25kb_ROS_op[16,2],2), #f1
                                          round(rfperf_K562_25kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_K562_25kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_K562_25kb_RUS_op[16,2],2), #f1
                                          round(rfperf_K562_25kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_K562_25kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_K562_25kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_K562_25kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_K562_50kb_none_oc[16,2],2), #f1
                                          round(rfperf_K562_50kb_none_op[16,2],2), #f1
                                          round(rfperf_K562_50kb_none_dist[16,2],2), #f1
                                          round(rfperf_K562_50kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_K562_50kb_ROS_op[16,2],2), #f1
                                          round(rfperf_K562_50kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_K562_50kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_K562_50kb_RUS_op[16,2],2), #f1
                                          round(rfperf_K562_50kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_K562_50kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_K562_50kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_K562_50kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_K562_100kb_none_oc[16,2],2), #f1
                                          round(rfperf_K562_100kb_none_op[16,2],2), #f1
                                          round(rfperf_K562_100kb_none_dist[16,2],2), #f1
                                          round(rfperf_K562_100kb_ROS_oc[16,2],2), #f1
                                          round(rfperf_K562_100kb_ROS_op[16,2],2), #f1
                                          round(rfperf_K562_100kb_ROS_dist[16,2],2), #f1
                                          round(rfperf_K562_100kb_RUS_oc[16,2],2), #f1
                                          round(rfperf_K562_100kb_RUS_op[16,2],2), #f1
                                          round(rfperf_K562_100kb_RUS_dist[16,2],2), #f1
                                          round(rfperf_K562_100kb_SMOTE_oc[16,2],2), #f1
                                          round(rfperf_K562_100kb_SMOTE_op[16,2],2), #f1
                                          round(rfperf_K562_100kb_SMOTE_dist[16,2],2), #f1

                                     round(rfperf_K562_10kb_none_oc[15,2],2), #mcc
                                          round(rfperf_K562_10kb_none_op[15,2],2), #mcc
                                          round(rfperf_K562_10kb_none_dist[15,2],2), #mcc
                                          round(rfperf_K562_10kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_K562_10kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_K562_10kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_K562_10kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_K562_10kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_K562_10kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_K562_10kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_K562_10kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_K562_10kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_K562_25kb_none_oc[15,2],2), #mcc
                                          round(rfperf_K562_25kb_none_op[15,2],2), #mcc
                                          round(rfperf_K562_25kb_none_dist[15,2],2), #mcc
                                          round(rfperf_K562_25kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_K562_25kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_K562_25kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_K562_25kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_K562_25kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_K562_25kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_K562_25kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_K562_25kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_K562_25kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_K562_50kb_none_oc[15,2],2), #mcc
                                          round(rfperf_K562_50kb_none_op[15,2],2), #mcc
                                          round(rfperf_K562_50kb_none_dist[15,2],2), #mcc
                                          round(rfperf_K562_50kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_K562_50kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_K562_50kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_K562_50kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_K562_50kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_K562_50kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_K562_50kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_K562_50kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_K562_50kb_SMOTE_dist[15,2],2), #mcc

                                     round(rfperf_K562_100kb_none_oc[15,2],2), #mcc
                                          round(rfperf_K562_100kb_none_op[15,2],2), #mcc
                                          round(rfperf_K562_100kb_none_dist[15,2],2), #mcc
                                          round(rfperf_K562_100kb_ROS_oc[15,2],2), #mcc
                                          round(rfperf_K562_100kb_ROS_op[15,2],2), #mcc
                                          round(rfperf_K562_100kb_ROS_dist[15,2],2), #mcc
                                          round(rfperf_K562_100kb_RUS_oc[15,2],2), #mcc
                                          round(rfperf_K562_100kb_RUS_op[15,2],2), #mcc
                                          round(rfperf_K562_100kb_RUS_dist[15,2],2), #mcc
                                          round(rfperf_K562_100kb_SMOTE_oc[15,2],2), #mcc
                                          round(rfperf_K562_100kb_SMOTE_op[15,2],2), #mcc
                                          round(rfperf_K562_100kb_SMOTE_dist[15,2],2) #mcc
)

p2 <- ggplot(all_rfperfs_K562,
       aes(x=Resolution, y=Performance, fill=Predictor)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid(Metric ~ Resampling) +
  xlab("") + ylab("") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Predictor Type")+
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))
```

##Plots

```{r}
p1 <- p1 + labs(title = "GM12878") + theme(plot.title = element_text(size=20))
p2 <- p2 + labs(title = "K562") + theme(plot.title = element_text(size=20))

figure <- ggarrange(p1,p2,
                ncol = 2, nrow = 1, 
                common.legend = TRUE, 
                legend = "top",
                labels = "AUTO",
                font.label = list(size=20))
figure
annotate_figure(figure,
    bottom = text_grob("Resolution", color = "black", size = 20),
    left = text_grob("Performance", color = "black", rot = 90, size=20)
)

```