### 2019-04-11 Scripts/distribution_of_tads.Rmd
added script to plot distribution of tads and create table

---
title: "Distribution of TADs"
author: "Spiro Stilianoudakis"
date: "April 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# GM12878

## 10kb

```{r}
tad_gm12878_10kb <- read.table("Z:/TAD_data_analysis/GM12878/10kb/GM12878_domain_data_10000.b.txt")
tad_gm12878_10kb <- tad_gm12878_10kb[,1:3]

tad_gm12878_10kb <- tad_gm12878_10kb %>%
  group_by(V1) %>%
  mutate(Length = V3-V2)

tad_gm12878_10kb$V1 <- factor(tad_gm12878_10kb$V1, levels = c(1:22))

tad_data <- tad_gm12878_10kb %>% 
  group_by(V1) %>% 
  summarise(Count = length(V1))

gm12878_10kb_tabnum <- ggplot(tad_data, aes(x=V1, y=Count, fill=V1))+
  geom_bar(stat="identity") +
  scale_fill_discrete(name="Chromosome") +
  guides(fill=FALSE) +
  ylab("Number of TADs") +
  xlab("Chromosome") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

gm12878_10kb_tabdist <- ggplot(tad_gm12878_10kb, aes(x=Length, color=V1)) +
  geom_density() +
  scale_color_discrete(name="Chromosome") +
  ylab("Count") +
  xlab("Length of TAD") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

## 25kb

```{r}
tad_gm12878_25kb <- read.table("Z:/TAD_data_analysis/GM12878/25kb/GM12878_domain_data_25000.b.txt")
tad_gm12878_25kb <- tad_gm12878_25kb[,1:3]

tad_gm12878_25kb <- tad_gm12878_25kb %>%
  group_by(V1) %>%
  mutate(Length = V3-V2)

tad_gm12878_25kb$V1 <- factor(tad_gm12878_25kb$V1, levels = c(1:22))

tad_data <- tad_gm12878_25kb %>% 
  group_by(V1) %>% 
  summarise(Count = length(V1))

gm12878_25kb_tabnum <- ggplot(tad_data, aes(x=V1, y=Count, fill=V1))+
  geom_bar(stat="identity") +
  scale_fill_discrete(name="Chromosome") +
  guides(fill=FALSE) +
  ylab("Number of TADs") +
  xlab("Chromosome") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

gm12878_25kb_tabdist <- ggplot(tad_gm12878_25kb, aes(x=Length, color=V1)) +
  geom_density() +
  scale_color_discrete(name="Chromosome") +
  ylab("Count") +
  xlab("Length of TAD") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

## 50kb

```{r}
tad_gm12878_50kb <- read.table("Z:/TAD_data_analysis/GM12878/50kb/GM12878_domain_data_50000.b.txt")
tad_gm12878_50kb <- tad_gm12878_50kb[,1:3]

tad_gm12878_50kb <- tad_gm12878_50kb %>%
  group_by(V1) %>%
  mutate(Length = V3-V2)

tad_gm12878_50kb$V1 <- factor(tad_gm12878_50kb$V1, levels = c(1:22))

tad_data <- tad_gm12878_50kb %>% 
  group_by(V1) %>% 
  summarise(Count = length(V1))

gm12878_50kb_tabnum <- ggplot(tad_data, aes(x=V1, y=Count, fill=V1))+
  geom_bar(stat="identity") +
  scale_fill_discrete(name="Chromosome") +
  guides(fill=FALSE) +
  ylab("Number of TADs") +
  xlab("Chromosome") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

gm12878_50kb_tabdist <- ggplot(tad_gm12878_50kb, aes(x=Length, color=V1)) +
  geom_density() +
  scale_color_discrete(name="Chromosome") +
  ylab("Count") +
  xlab("Length of TAD") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

## 100kb

```{r}
tad_gm12878_100kb <- read.table("Z:/TAD_data_analysis/GM12878/100kb/GM12878_domain_data_100000.b.txt")
tad_gm12878_100kb <- tad_gm12878_100kb[,1:3]

tad_gm12878_100kb <- tad_gm12878_100kb %>%
  group_by(V1) %>%
  mutate(Length = V3-V2)

tad_gm12878_100kb$V1 <- factor(tad_gm12878_100kb$V1, levels = c(1:22))

tad_data <- tad_gm12878_100kb %>% 
  group_by(V1) %>% 
  summarise(Count = length(V1))

gm12878_100kb_tabnum <- ggplot(tad_data, aes(x=V1, y=Count, fill=V1))+
  geom_bar(stat="identity") +
  scale_fill_discrete(name="Chromosome") +
  guides(fill=FALSE) +
  ylab("Number of TADs") +
  xlab("Chromosome") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

gm12878_100kb_tabdist <- ggplot(tad_gm12878_100kb, aes(x=Length, color=V1)) +
  geom_density() +
  scale_color_discrete(name="Chromosome") +
  ylab("Count") +
  xlab("Length of TAD") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

# K562

## 10kb

```{r}
tad_K562_10kb <- read.table("Z:/TAD_data_analysis/K562/10kb/K562_domain_data_10000.b.txt")
tad_K562_10kb <- tad_K562_10kb[,1:3]

tad_K562_10kb <- tad_K562_10kb %>%
  group_by(V1) %>%
  mutate(Length = V3-V2)

tad_K562_10kb$V1 <- factor(tad_K562_10kb$V1, levels = c(1:22))

tad_data <- tad_K562_10kb %>% 
  group_by(V1) %>% 
  summarise(Count = length(V1))

K562_10kb_tabnum <- ggplot(tad_data, aes(x=V1, y=Count, fill=V1))+
  geom_bar(stat="identity") +
  scale_fill_discrete(name="Chromosome") +
  guides(fill=FALSE) +
  ylab("Number of TADs") +
  xlab("Chromosome") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

K562_10kb_tabdist <- ggplot(tad_K562_10kb, aes(x=Length, color=V1)) +
  geom_density() +
  scale_color_discrete(name="Chromosome") +
  ylab("Count") +
  xlab("Length of TAD") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

## 25kb

```{r}
tad_K562_25kb <- read.table("Z:/TAD_data_analysis/K562/25kb/K562_domain_data_25000.b.txt")
tad_K562_25kb <- tad_K562_25kb[,1:3]

tad_K562_25kb <- tad_K562_25kb %>%
  group_by(V1) %>%
  mutate(Length = V3-V2)

tad_K562_25kb$V1 <- factor(tad_K562_25kb$V1, levels = c(1:22))

tad_data <- tad_K562_25kb %>% 
  group_by(V1) %>% 
  summarise(Count = length(V1))

K562_25kb_tabnum <- ggplot(tad_data, aes(x=V1, y=Count, fill=V1))+
  geom_bar(stat="identity") +
  scale_fill_discrete(name="Chromosome") +
  guides(fill=FALSE) +
  ylab("Number of TADs") +
  xlab("Chromosome") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

K562_25kb_tabdist <- ggplot(tad_K562_25kb, aes(x=Length, color=V1)) +
  geom_density() +
  scale_color_discrete(name="Chromosome") +
  ylab("Count") +
  xlab("Length of TAD") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

## 50kb

```{r}
tad_K562_50kb <- read.table("Z:/TAD_data_analysis/K562/50kb/K562_domain_data_50000.b.txt")
tad_K562_50kb <- tad_K562_50kb[,1:3]

tad_K562_50kb <- tad_K562_50kb %>%
  group_by(V1) %>%
  mutate(Length = V3-V2)

tad_K562_50kb$V1 <- factor(tad_K562_50kb$V1, levels = c(1:22))

tad_data <- tad_K562_50kb %>% 
  group_by(V1) %>% 
  summarise(Count = length(V1))

K562_50kb_tabnum <- ggplot(tad_data, aes(x=V1, y=Count, fill=V1))+
  geom_bar(stat="identity") +
  scale_fill_discrete(name="Chromosome") +
  guides(fill=FALSE) +
  ylab("Number of TADs") +
  xlab("Chromosome") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

K562_50kb_tabdist <- ggplot(tad_K562_50kb, aes(x=Length, color=V1)) +
  geom_density() +
  scale_color_discrete(name="Chromosome") +
  ylab("Count") +
  xlab("Length of TAD") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

## 100kb

```{r}
tad_K562_100kb <- read.table("Z:/TAD_data_analysis/K562/100kb/K562_domain_data_100000.b.txt")
tad_K562_100kb <- tad_K562_100kb[,1:3]

tad_K562_100kb <- tad_K562_100kb %>%
  group_by(V1) %>%
  mutate(Length = V3-V2)

tad_K562_100kb$V1 <- factor(tad_K562_100kb$V1, levels = c(1:22))

tad_data <- tad_K562_100kb %>% 
  group_by(V1) %>% 
  summarise(Count = length(V1))

K562_100kb_tabnum <- ggplot(tad_data, aes(x=V1, y=Count, fill=V1))+
  geom_bar(stat="identity") +
  scale_fill_discrete(name="Chromosome") +
  guides(fill=FALSE) +
  ylab("Number of TADs") +
  xlab("Chromosome") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

K562_100kb_tabdist <- ggplot(tad_K562_100kb, aes(x=Length, color=V1)) +
  geom_density() +
  scale_color_discrete(name="Chromosome") +
  ylab("Count") +
  xlab("Length of TAD") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

# Plots

```{r}
figure <- ggarrange(gm12878_10kb_tabnum,gm12878_10kb_tabdist,
                    gm12878_25kb_tabnum,gm12878_25kb_tabdist,
                    gm12878_50kb_tabnum,gm12878_50kb_tabdist,
                    gm12878_100kb_tabnum,gm12878_100kb_tabdist,
                    ncol=2,nrow=4,
                    common.legend = TRUE,
                    legend="right",
                    labels = c("A", "","B", "","C", "","D", ""),
                    font.label = list(size=20))

figure <- ggarrange(K562_10kb_tabnum,K562_10kb_tabdist,
                    K562_25kb_tabnum,K562_25kb_tabdist,
                    K562_50kb_tabnum,K562_50kb_tabdist,
                    K562_100kb_tabnum,K562_100kb_tabdist,
                    ncol=2,nrow=4,
                    common.legend = TRUE,
                    legend="right",
                    labels = "AUTO",
                    font.label = list(size=20))
```

# Creating Table of Summaraies

```{r}
tad_gm12878_10kb$CL <- "GM12878"
tad_gm12878_25kb$CL <- "GM12878"
tad_gm12878_50kb$CL <- "GM12878"
tad_gm12878_100kb$CL <- "GM12878"
tad_K562_10kb$CL <- "K562"
tad_K562_25kb$CL <- "K562"
tad_K562_50kb$CL <- "K562"
tad_K562_100kb$CL <- "K562"

tad_data_summaries <- rbind.data.frame(aggregate(tad_gm12878_10kb$Length, by=list(tad_gm12878_10kb$V1), FUN=summary),
                 aggregate(tad_gm12878_25kb$Length, by=list(tad_gm12878_25kb$V1), FUN=summary),
                 aggregate(tad_gm12878_50kb$Length, by=list(tad_gm12878_50kb$V1), FUN=summary),
                 aggregate(tad_gm12878_100kb$Length, by=list(tad_gm12878_100kb$V1), FUN=summary),
                 aggregate(tad_K562_10kb$Length, by=list(tad_K562_10kb$V1), FUN=summary),
                 aggregate(tad_K562_25kb$Length, by=list(tad_K562_25kb$V1), FUN=summary),
                 aggregate(tad_K562_50kb$Length, by=list(tad_K562_50kb$V1), FUN=summary),
                 aggregate(tad_K562_100kb$Length, by=list(tad_K562_100kb$V1), FUN=summary))

#chr 9 for k562 at 10 kb has no tad boundaries

tad_data_summaries$CL <- c(rep("GM12878", 22*4), rep("K562", 22*4-1))
tad_data_summaries$Resolution <- c(rep("10 kb", 22),rep("25 kb", 22),rep("50 kb", 22),rep("100 kb", 22),
                                   rep("10 kb", 21),rep("25 kb", 22),rep("50 kb", 22),rep("100 kb", 22))

write.csv(tad_data_summaries, "C:/Users/stili/Downloads/tad_data_summaries.csv")
```