### 2019-04-12 Scripts/comparing_tfbs_specific_models_with_all_annotations.Rmd
script for comparing tfbs specific models with models containing all annotations (at 10 kb)

---
title: "Investigating TFBS Specific Model Performances"
author: "Spiro Stilianoudakis"
date: "April 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# TFBS Specific Model Performances

## GM12878

### 10 kb

```{r}
#OC
##None
rfperf_gm12878_10kb_none_oc_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/tfbs/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_10kb_none_op_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/tfbs/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_10kb_none_dist_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/tfbs/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_10kb_ros_oc_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/tfbs/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_10kb_ros_op_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/tfbs/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_10kb_ros_dist_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/tfbs/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_10kb_rus_oc_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/tfbs/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_10kb_rus_op_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/tfbs/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_10kb_rus_dist_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/tfbs/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_10kb_smote_oc_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/tfbs/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_10kb_smote_op_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/tfbs/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_10kb_smote_dist_tfbs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/tfbs/Distance/ENET/RF/rfperf.rds")

all_rfperfs_gm12878_tfbs <- data.frame(Resampling=rep(c(rep("None",3),
                                                    rep("ROS",3),
                                                    rep("RUS",3),
                                                    rep("SMOTE",3)),2),
                                       Predictor=rep(c("OC","OP","Distance"),8),
                                       Metric=c(rep("F1-Score",12), 
                                                rep("MCC",12)),
                                       Annotations=rep("TFBS Only",24)
                                       )

all_rfperfs_gm12878_tfbs$Predictor <- factor(all_rfperfs_gm12878_tfbs$Predictor, levels=c("OC", "OP", "Distance"))
all_rfperfs_gm12878_tfbs$Metric <- factor(all_rfperfs_gm12878_tfbs$Metric, levels=c("F1-Score", "MCC"))
all_rfperfs_gm12878_tfbs$Resampling <- factor(all_rfperfs_gm12878_tfbs$Resampling, levels=c("None", "ROS", "RUS", "SMOTE"))

all_rfperfs_gm12878_tfbs$Performance <- c(rfperf_gm12878_10kb_none_oc_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_none_op_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_none_dist_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_ros_oc_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_ros_op_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_ros_dist_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_rus_oc_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_rus_op_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_rus_dist_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_smote_oc_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_smote_op_tfbs[16,2], #f1
                                          rfperf_gm12878_10kb_smote_dist_tfbs[16,2], #f1

                                     rfperf_gm12878_10kb_none_oc_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_none_op_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_none_dist_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_ros_oc_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_ros_op_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_ros_dist_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_rus_oc_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_rus_op_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_rus_dist_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_smote_oc_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_smote_op_tfbs[15,2], #mcc
                                          rfperf_gm12878_10kb_smote_dist_tfbs[15,2] #mcc

                                     )

```

# All Annotations

## GM12878

### 10 kb

```{r}
#OC
##None
rfperf_gm12878_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/ENET/RF/rfperf.rds")

#OP
##None
rfperf_gm12878_10kb_none_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/ENET/RF/rfperf.rds")

#Distance
##None
rfperf_gm12878_10kb_none_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/ENET/RF/rfperf.rds")

#OC
##ROS
rfperf_gm12878_10kb_ros_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OC/ENET/RF/rfperf.rds")

#OP
##ROS
rfperf_gm12878_10kb_ros_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OP/ENET/RF/rfperf.rds")

#Distance
##ROS
rfperf_gm12878_10kb_ros_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/ENET/RF/rfperf.rds")

#OC
##RUS
rfperf_gm12878_10kb_rus_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/ENET/RF/rfperf.rds")

#OP
##RUS
rfperf_gm12878_10kb_rus_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/ENET/RF/rfperf.rds")

#Distance
##RUS
rfperf_gm12878_10kb_rus_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/ENET/RF/rfperf.rds")

#OC
##SMOTE
rfperf_gm12878_10kb_smote_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OC/ENET/RF/rfperf.rds")

#OP
##SMOTE
rfperf_gm12878_10kb_smote_op <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OP/ENET/RF/rfperf.rds")

#Distance
##SMOTE
rfperf_gm12878_10kb_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

all_rfperfs_gm12878 <- data.frame(Resampling=rep(c(rep("None",3),
                                                    rep("ROS",3),
                                                    rep("RUS",3),
                                                    rep("SMOTE",3)),2),
                                       Predictor=rep(c("OC","OP","Distance"),8),
                                       Metric=c(rep("F1-Score",12), 
                                                rep("MCC",12)),
                                       Annotations=rep("ALL Types",24)
                                       )

all_rfperfs_gm12878$Predictor <- factor(all_rfperfs_gm12878$Predictor, levels=c("OC", "OP", "Distance"))
all_rfperfs_gm12878$Metric <- factor(all_rfperfs_gm12878$Metric, levels=c("F1-Score", "MCC"))
all_rfperfs_gm12878$Resampling <- factor(all_rfperfs_gm12878$Resampling, levels=c("None", "ROS", "RUS", "SMOTE"))

all_rfperfs_gm12878$Performance <- c(rfperf_gm12878_10kb_none_oc[16,2], #f1
                                          rfperf_gm12878_10kb_none_op[16,2], #f1
                                          rfperf_gm12878_10kb_none_dist[16,2], #f1
                                          rfperf_gm12878_10kb_ros_oc[16,2], #f1
                                          rfperf_gm12878_10kb_ros_op[16,2], #f1
                                          rfperf_gm12878_10kb_ros_dist[16,2], #f1
                                          rfperf_gm12878_10kb_rus_oc[16,2], #f1
                                          rfperf_gm12878_10kb_rus_op[16,2], #f1
                                          rfperf_gm12878_10kb_rus_dist[16,2], #f1
                                          rfperf_gm12878_10kb_smote_oc[16,2], #f1
                                          rfperf_gm12878_10kb_smote_op[16,2], #f1
                                          rfperf_gm12878_10kb_smote_dist[16,2], #f1

                                     rfperf_gm12878_10kb_none_oc[15,2], #mcc
                                          rfperf_gm12878_10kb_none_op[15,2], #mcc
                                          rfperf_gm12878_10kb_none_dist[15,2], #mcc
                                          rfperf_gm12878_10kb_ros_oc[15,2], #mcc
                                          rfperf_gm12878_10kb_ros_op[15,2], #mcc
                                          rfperf_gm12878_10kb_ros_dist[15,2], #mcc
                                          rfperf_gm12878_10kb_rus_oc[15,2], #mcc
                                          rfperf_gm12878_10kb_rus_op[15,2], #mcc
                                          rfperf_gm12878_10kb_rus_dist[15,2], #mcc
                                          rfperf_gm12878_10kb_smote_oc[15,2], #mcc
                                          rfperf_gm12878_10kb_smote_op[15,2], #mcc
                                          rfperf_gm12878_10kb_smote_dist[15,2] #mcc

                                     )

```

# Plots

## Comparing 10 kb

```{r}
all_rfperfs_gm12878_allvtfbs <- rbind.data.frame(all_rfperfs_gm12878,all_rfperfs_gm12878_tfbs)
all_rfperfs_gm12878_allvtfbs$Annotations <- factor(all_rfperfs_gm12878_allvtfbs$Annotations, levels=c("ALL Types", "TFBS Only"))

ggplot(all_rfperfs_gm12878_allvtfbs,
       aes(x=Resampling, y=Performance, fill=Annotations)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid( ~ Predictor + Metric) +
  #xlab("") + ylab("") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Feature Space")+
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        strip.text.x = element_text(size = 15))
```