### 2019-04-16 Scripts/comparing_predictortypes_one_annot_at_a_time.Rmd
editted results scripts

---
title: "Comparing Predictor Types One Annotation At A Time"
author: "Spiro Stilianoudakis"
date: "April 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# GM12878

##10kb

###One annotation at a time RF

```{r}
train <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/train.rds")
train <- train[,-grep("_binary", names(train))]

set.seed(123)
train_smote <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

#Randomly shuffle the data
set.seed(123)
train_smote <- train_smote[sample(nrow(train_smote)),]

#Perform RF with 10 fold cross validation
tunegrid <- expand.grid(mtry=ceiling(sqrt(3)))
#modellist <- list()
#for (ntree in c(50,200,500,1000)) {
#  print(ntree)
#  set.seed(333)
#  fit <- train(y~., data=train_smote, 
#               method="rf", 
#               metric="Accuracy",
#               tuneGrid=tunegrid,  
#               ntree=ntree)
#  key <- toString(ntree)
#  modellist[[key]] <- fit
#}
# compare results
#results <- resamples(modellist)
#summary(results)
#dotplot(results)
#results <- data.frame(summary(results)[3]$statistics$Accuracy)
#results <- results[order(results$Mean, decreasing = TRUE),]

##defining function to extract AUPR curve
auprcSummary <- function(data, lev = NULL, model = NULL){
  lvls <- levels(data$obs) #take the probability of good class
  prob_good <- data[,lvls[2]] 
  the_curve <- pr.curve(scores.class0 = prob_good,
                        weights.class0 = as.numeric(data$obs)-1, #provide the class labels as 0/1
                        curve = FALSE)
  out <- the_curve$auc.integral
  names(out) <- "AUPRC"
  out
}

####set tuning parameters
####set seeds for each cross fols
set.seed(123)
#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)
#(1 is the number of tuning parameter lambda
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 1)
#for the last model
seeds[[11]]<-sample.int(1000, 1)

fitControl <- trainControl(seeds = seeds,
                           method = "cv",
                           number = 10,
                           ## Estimate class probabilities
                           classProbs = TRUE,
                           ## Evaluate performance using 
                           ## the following function
                           #summaryFunction = auprcSummary,
						   summaryFunction = twoClassSummary)

rfimpvarsmat_gm12878_10kb <- matrix(nrow=3, ncol=107,
				dimnames=list("Predictor Type"=c("OC",
												 "Distance",
												 "OP"),
							  "Annotation"=unique(gsub(paste0(c("Gm12878", 
																"_count", 
																"_perc", 
																"_dist",
																"-",
																"Haib", 
																"Sydh",
																"Uta",
																"Uw",
																"Broad",
																"Uchicago"), collapse = "|"), "", names(train_smote)[-1]))))

for(i in 0:106){
  print(names(train_smote)[(i*3+2):(i*3+4)])
	rfModel <- train(y~., data=train_smote[,c(1, (i*3+2):(i*3+4))], 
                 method="rf", 
                 metric="ROC", 
                 tuneGrid=tunegrid, 
                 trControl=fitControl, 
                 ntree=500) #as.numeric(rownames(results)[1]))

	#rfimpvars <- data.frame(Feature=rownames(varImp(rfModel)$importance), Importance=varImp(rfModel)$importance[,1])
	#rfimpvars <- rfimpvars[order(rfimpvars$Importance, decreasing=TRUE),]

	rfimpvarsmat_gm12878_10kb[,i+1] <- varImp(rfModel)$importance[,1][order(rownames(varImp(rfModel)$importance))]
}

```