### 2019-04-16 Scripts/investigating_misclassified_TADs.Rmd

---
title: "Determining what makes some TADs unpredictable"
author: "Spiro Stilianoudakis"
date: "April 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# Evaluating Missclassified TAD boundaries from best performing model

## SMOTE; distance-type predictors

### GM12878

####10kb

```{r}
#reading in test data and reducing
test <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/test.rds")
names(test) <- gsub(paste(c("`", 
                            "Gm12878", 
                            "-", 
                            #"_", 
                            #"dist", "count", "perc",
                            "Haib", 
                            "Sydh",
                            "Uta",
                            "Uw",
                            "Broad",
                            "Uchicago"), collapse = "|"), "", names(test))
enetcoefs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/enetcoefs.rds")
test <- test[,c(1,which(names(test) %in% enetcoefs))]

#reading in model and making predictions
rfmodel_gm12878_10kb <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
predictions <- predict(rfmodel_gm12878_10kb, newdata=test)
preddata <- cbind.data.frame(Prediction = predictions, True=test$y)

#subset the false negatives (bins that were predicted as not overlapping with flanked TAD boundary but really were)
FNdata <- test[which(preddata$Prediction=="No" & preddata$True=="Yes"),]
FNdatamelt <- melt(FNdata[,-1])
FNdatamelt$Prediction <- "False Negatives"

#subset the true positives (bins that were predicted as overlapping with flanked TAD boundary and truely were)
TPdata <- test[which(preddata$Prediction=="Yes" & preddata$True=="Yes"),]
TPdatamelt <- melt(TPdata[,-1])
TPdatamelt$Prediction <- "True Positives"

#Wilcoxon test on differences in distances between FN and TP
FNdata <- FNdata[,-1]
TPdata <- TPdata[,-1]
wilcoxon.results <- numeric(length(colnames(FNdata)))
for(i in 1:length(colnames(FNdata))){
 wilcoxon.results[i] <- wilcox.test(FNdata[,i],TPdata[,i])$p.value
}
wilcox.table <- cbind.data.frame(Annotation=gsub("_dist","",colnames(FNdata)), Wilcoxon.PValue=wilcoxon.results)
wilcox.table <- wilcox.table[order(wilcox.table$Wilcoxon.PValue, decreasing = FALSE),]
wilcox.table$Wilcoxon.PValue <- round(wilcox.table$Wilcoxon.PValue,7)

predictiondata_gm12878_10kb <- rbind.data.frame(FNdatamelt,TPdatamelt)
predictiondata_gm12878_10kb$Prediction <- factor(predictiondata_gm12878_10kb$Prediction, levels=c("False Negatives","True Positives"))

#Ploting densities
ggplot(predictiondata_gm12878_10kb, aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

#looking at specific annotations
##Ctcf
ggplot(predictiondata_gm12878_10kb[grep("Ctcf_dist",predictiondata_gm12878_10kb$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

##ZNF143
ggplot(predictiondata_gm12878_10kb[grep("Znf143166181ap_dist",predictiondata_gm12878_10kb$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

##SMC3a
ggplot(predictiondata_gm12878_10kb[grep("Smc3ab9263_dist",predictiondata_gm12878_10kb$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

##Rad21
ggplot(predictiondata_gm12878_10kb[grep("Rad21_dist",predictiondata_gm12878_10kb$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```