### 2019-04-17 Scripts/comparing_distance_densities_by_genomic_bin_class.Rmd
script for comparing distance densities between genomic bin class for top predictive annotations (gm12878)

---
title: "Comparing distance distributions for top predictive features between bins with and without TADs"
author: "Spiro Stilianoudakis"
date: "April 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(coin)
```

# GM12878

## Function for ploting distance densities between bins with and without TAD boundaries

```{r}
distplotsfunc <- function(cl, resolution, annotation){
  train <- readRDS(paste0("Z:/TAD_data_analysis/", cl, "/", resolution, "/training_testing/train.rds"))
  data <- train[, c(1, which(names(train)==annotation))]

  set.seed(123)
  majorityids <- sample(x=which(data$y=="No"), size=length(which(data$y=="Yes")))
  annotdata <- rbind.data.frame(data[which(data$y=="Yes"),],
                                data[majorityids,])

  #ks.test(annotdata[which(annotdata$y=="No"),2], annotdata[which(annotdata$y=="Yes"),2], alternative = "two.sided")

  perm.pval <- pvalue(oneway_test(annotdata[,2] ~ annotdata[,1],
                     data=annotdata,
                     alternative="greater", 
                     distribution=approximate(B=10000)))[1]

  if(perm.pval==0){perm.pval="< 0.0001"}

  plottitle <- paste(gsub(paste(c("Gm12878",
                                  "-","_",
                                  "dist","perc","count",
                                  "Haib",
                                  "Sydh",
                                  "Uta",
                                  "Uw",
                                  "Broad",
                                  "Uchicago"), 
                                collapse = "|"), 
                          "",
                          annotation))

  p <- ggplot(annotdata, aes(x=annotdata[,2], color=y)) +
    geom_density() +
    scale_color_discrete(name="Class of\nGenomic Bin",
                         labels=c("Does not overlap\nw/ flanked TAD boundary", "Overlapped w/\nflanked TAD boundary")) +
    ylab("Frequency") +
    xlab("Log2 Distance") +
    ggtitle(paste(plottitle)) +
    annotate("text", x=7, y=.2, label= paste("P-Value:", perm.pval, sep = " "), size=5) +
    theme_bw() +
    theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20,
                               hjust=.5))

  ggsave(p, 
         filename = paste0("C:/Users/stili/Documents/TAD_miscellaneous/figures/comparing_distance_by_bin_class/", 
                           cl, 
                           "_", 
                           resolution, 
                           "_", 
                           plottitle, 
                           ".png"),
         width=10,
         height=10,
         units="in")
}

```

### 10kb

#### Top Annotations

```{r}
#### CTCF

distplotsfunc(cl="GM12878", resolution="10kb", annotation="Gm12878-Ctcf-Broad_dist")

#### Smc3ab9263

distplotsfunc(cl="GM12878", resolution="10kb", annotation="Gm12878-Smc3ab9263-Sydh_dist")

#### Znf143166181ap

distplotsfunc(cl="GM12878", resolution="10kb", annotation="Gm12878-Znf143166181ap-Sydh_dist")

#### Rad21

distplotsfunc(cl="GM12878", resolution="10kb", annotation="Gm12878-Rad21-Haib_dist")

#### 08Insulator

distplotsfunc(cl="GM12878", resolution="10kb", annotation="Gm12878-08Insulator_dist")

#### Yy1sc281

distplotsfunc(cl="GM12878", resolution="10kb", annotation="Gm12878-Yy1sc281-Haib_dist")

#### DNase

distplotsfunc(cl="GM12878", resolution="10kb", annotation="Gm12878-DNase_dist")
```

### 25kb

#### Top Annotations

```{r}
#### CTCF

distplotsfunc(cl="GM12878", resolution="25kb", annotation="Gm12878-Ctcf-Broad_dist")

#### Smc3ab9263

distplotsfunc(cl="GM12878", resolution="25kb", annotation="Gm12878-Smc3ab9263-Sydh_dist")

#### Znf143166181ap

distplotsfunc(cl="GM12878", resolution="25kb", annotation="Gm12878-Znf143166181ap-Sydh_dist")

#### Rad21

distplotsfunc(cl="GM12878", resolution="25kb", annotation="Gm12878-Rad21-Haib_dist")

#### 08Insulator

distplotsfunc(cl="GM12878", resolution="25kb", annotation="Gm12878-08Insulator_dist")

#### Yy1sc281

distplotsfunc(cl="GM12878", resolution="25kb", annotation="Gm12878-Yy1sc281-Haib_dist")

#### DNase

distplotsfunc(cl="GM12878", resolution="25kb", annotation="Gm12878-DNase_dist")
```

### 50kb

#### Top Annotations

```{r}
#### CTCF

distplotsfunc(cl="GM12878", resolution="50kb", annotation="Gm12878-Ctcf-Broad_dist")

#### Smc3ab9263

distplotsfunc(cl="GM12878", resolution="50kb", annotation="Gm12878-Smc3ab9263-Sydh_dist")

#### Znf143166181ap

distplotsfunc(cl="GM12878", resolution="50kb", annotation="Gm12878-Znf143166181ap-Sydh_dist")

#### Rad21

distplotsfunc(cl="GM12878", resolution="50kb", annotation="Gm12878-Rad21-Haib_dist")

#### 08Insulator

distplotsfunc(cl="GM12878", resolution="50kb", annotation="Gm12878-08Insulator_dist")

#### Yy1sc281

distplotsfunc(cl="GM12878", resolution="50kb", annotation="Gm12878-Yy1sc281-Haib_dist")

#### DNase

distplotsfunc(cl="GM12878", resolution="50kb", annotation="Gm12878-DNase_dist")
```

### 100kb

#### Top Annotations

```{r}
#### CTCF

distplotsfunc(cl="GM12878", resolution="100kb", annotation="Gm12878-Ctcf-Broad_dist")

#### Smc3ab9263

distplotsfunc(cl="GM12878", resolution="100kb", annotation="Gm12878-Smc3ab9263-Sydh_dist")

#### Znf143166181ap

distplotsfunc(cl="GM12878", resolution="100kb", annotation="Gm12878-Znf143166181ap-Sydh_dist")

#### Rad21

distplotsfunc(cl="GM12878", resolution="100kb", annotation="Gm12878-Rad21-Haib_dist")

#### 08Insulator

distplotsfunc(cl="GM12878", resolution="100kb", annotation="Gm12878-08Insulator_dist")

#### Yy1sc281

distplotsfunc(cl="GM12878", resolution="100kb", annotation="Gm12878-Yy1sc281-Haib_dist")

#### DNase

distplotsfunc(cl="GM12878", resolution="100kb", annotation="Gm12878-DNase_dist")
```