### 2019-04-18 Scripts/validating_models_across_cell_lines_clustered_tfbs.Rmd
script to compare performances when validating models across cell lines for models built on clustered tfbs (smote; distance types)

---
title: "Training and validating models across cell lines using clustered TFBS"
author: "Spiro Stilianoudakis"
date: "April 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(PRROC)
```

# Reading in GM12878; smote; distance types

## Validating on GM12878

```{r}
#10kb
rfperf_gm12878_10kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#25kb
rfperf_gm12878_25kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#50kb
rfperf_gm12878_50kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#100kb
rfperf_gm12878_100kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

```

## Validating on K562

```{r}
#10kb
rfmodel_gm12878_10kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
####reading in test data for K562
test <- readRDS("Z:/TAD_data_analysis/K562/10kb/clustered/training_testing/test.rds")
####reading in enetcoefs for K562
vars <- rfmodel_gm12878_10kb_clustered_smote_dist$finalModel$xNames
####reduce test data to match train data
names(test) <- gsub(paste(c("`", 
                          "Clustered", 
                          "-", 
                          #"_", 
                          #"dist", "count", "perc",
                          "TfbsV3"), collapse = "|"), "", names(test))
test <- test[,c(1,which(names(test) %in% vars))]
rfperf_K562_10kb_clustered_smote_dist <- data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)
pred.rfModel <- as.vector(predict(rfmodel_gm12878_10kb_clustered_smote_dist, 
                                  newdata=test, 
                                  type="prob")[,"Yes"])

fg <- pred.rfModel[test$y == "Yes"]
bg <- pred.rfModel[test$y == "No"]
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)

pred.rfModel2 <- predict(rfmodel_gm12878_10kb_clustered_smote_dist,
                         newdata=test,
                         type="raw")

confMat <- confusionMatrix(data=pred.rfModel2, test$y, positive="Yes")
TN = as.numeric(confMat$table[1,1])
FN = as.numeric(confMat$table[1,2])
FP = as.numeric(confMat$table[2,1])
TP = as.numeric(confMat$table[2,2])
rfperf_K562_10kb_clustered_smote_dist[1,2] <- TN
rfperf_K562_10kb_clustered_smote_dist[2,2] <- FN
rfperf_K562_10kb_clustered_smote_dist[3,2] <- FP
rfperf_K562_10kb_clustered_smote_dist[4,2] <- TP
rfperf_K562_10kb_clustered_smote_dist[5,2] <- sum(confMat$table)
rfperf_K562_10kb_clustered_smote_dist[6,2] <- as.vector(confMat$byClass["Sensitivity"])
rfperf_K562_10kb_clustered_smote_dist[7,2] <- as.vector(confMat$byClass["Specificity"])
rfperf_K562_10kb_clustered_smote_dist[8,2] <- as.vector(confMat$overall["Kappa"])
rfperf_K562_10kb_clustered_smote_dist[9,2] <- as.vector(confMat$overall["Accuracy"])
rfperf_K562_10kb_clustered_smote_dist[10,2] <- TP/(TP+FP)
rfperf_K562_10kb_clustered_smote_dist[11,2] <- FP/(FP+TN)
rfperf_K562_10kb_clustered_smote_dist[12,2] <- FN/(FN+TN)
rfperf_K562_10kb_clustered_smote_dist[13,2] <- FN/(FN+TN)
rfperf_K562_10kb_clustered_smote_dist[14,2] <- TN/(TN+FN)
rfperf_K562_10kb_clustered_smote_dist[15,2] <- (TP*TN - FP*FN)/( sqrt( (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN) ) )
rfperf_K562_10kb_clustered_smote_dist[16,2] <- (2*(TP/(TP+FN))*(TP/(TP+FP)))/((TP/(TP+FN)) + ((TP/(TP+FP))))
rfperf_K562_10kb_clustered_smote_dist[17,2] <- pROC::auc(pROC::roc(test$y, pred.rfModel))			
rfperf_K562_10kb_clustered_smote_dist[18,2] <- (TP/(TP + FN)) + (TN/(TN + FP)) - 1
rfperf_K562_10kb_clustered_smote_dist[19,2] <- pr$auc.integral

#25kb
rfmodel_gm12878_25kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
####reading in test data for K562
test <- readRDS("Z:/TAD_data_analysis/K562/25kb/clustered/training_testing/test.rds")
####reading in enetcoefs for K562
vars <- rfmodel_gm12878_25kb_clustered_smote_dist$finalModel$xNames
####reduce test data to match train data
names(test) <- gsub(paste(c("`", 
                          "Clustered", 
                          "-", 
                          #"_", 
                          #"dist", "count", "perc",
                          "TfbsV3"), collapse = "|"), "", names(test))
test <- test[,c(1,which(names(test) %in% vars))]
rfperf_K562_25kb_clustered_smote_dist <- data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)
pred.rfModel <- as.vector(predict(rfmodel_gm12878_25kb_clustered_smote_dist, 
                                  newdata=test, 
                                  type="prob")[,"Yes"])

fg <- pred.rfModel[test$y == "Yes"]
bg <- pred.rfModel[test$y == "No"]
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)

pred.rfModel2 <- predict(rfmodel_gm12878_25kb_clustered_smote_dist,
                         newdata=test,
                         type="raw")

confMat <- confusionMatrix(data=pred.rfModel2, test$y, positive="Yes")
TN = as.numeric(confMat$table[1,1])
FN = as.numeric(confMat$table[1,2])
FP = as.numeric(confMat$table[2,1])
TP = as.numeric(confMat$table[2,2])
rfperf_K562_25kb_clustered_smote_dist[1,2] <- TN
rfperf_K562_25kb_clustered_smote_dist[2,2] <- FN
rfperf_K562_25kb_clustered_smote_dist[3,2] <- FP
rfperf_K562_25kb_clustered_smote_dist[4,2] <- TP
rfperf_K562_25kb_clustered_smote_dist[5,2] <- sum(confMat$table)
rfperf_K562_25kb_clustered_smote_dist[6,2] <- as.vector(confMat$byClass["Sensitivity"])
rfperf_K562_25kb_clustered_smote_dist[7,2] <- as.vector(confMat$byClass["Specificity"])
rfperf_K562_25kb_clustered_smote_dist[8,2] <- as.vector(confMat$overall["Kappa"])
rfperf_K562_25kb_clustered_smote_dist[9,2] <- as.vector(confMat$overall["Accuracy"])
rfperf_K562_25kb_clustered_smote_dist[10,2] <- TP/(TP+FP)
rfperf_K562_25kb_clustered_smote_dist[11,2] <- FP/(FP+TN)
rfperf_K562_25kb_clustered_smote_dist[12,2] <- FN/(FN+TN)
rfperf_K562_25kb_clustered_smote_dist[13,2] <- FN/(FN+TN)
rfperf_K562_25kb_clustered_smote_dist[14,2] <- TN/(TN+FN)
rfperf_K562_25kb_clustered_smote_dist[15,2] <- (TP*TN - FP*FN)/( sqrt( (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN) ) )
rfperf_K562_25kb_clustered_smote_dist[16,2] <- (2*(TP/(TP+FN))*(TP/(TP+FP)))/((TP/(TP+FN)) + ((TP/(TP+FP))))
rfperf_K562_25kb_clustered_smote_dist[17,2] <- pROC::auc(pROC::roc(test$y, pred.rfModel))			
rfperf_K562_25kb_clustered_smote_dist[18,2] <- (TP/(TP + FN)) + (TN/(TN + FP)) - 1
rfperf_K562_25kb_clustered_smote_dist[19,2] <- pr$auc.integral

#50kb
rfmodel_gm12878_50kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
####reading in test data for K562
test <- readRDS("Z:/TAD_data_analysis/K562/50kb/clustered/training_testing/test.rds")
####reading in enetcoefs for K562
vars <- rfmodel_gm12878_50kb_clustered_smote_dist$finalModel$xNames
####reduce test data to match train data
names(test) <- gsub(paste(c("`", 
                          "Clustered", 
                          "-", 
                          #"_", 
                          #"dist", "count", "perc",
                          "TfbsV3"), collapse = "|"), "", names(test))
test <- test[,c(1,which(names(test) %in% vars))]
rfperf_K562_50kb_clustered_smote_dist <- data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)
pred.rfModel <- as.vector(predict(rfmodel_gm12878_50kb_clustered_smote_dist, 
                                  newdata=test, 
                                  type="prob")[,"Yes"])

fg <- pred.rfModel[test$y == "Yes"]
bg <- pred.rfModel[test$y == "No"]
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)

pred.rfModel2 <- predict(rfmodel_gm12878_50kb_clustered_smote_dist,
                         newdata=test,
                         type="raw")

confMat <- confusionMatrix(data=pred.rfModel2, test$y, positive="Yes")
TN = as.numeric(confMat$table[1,1])
FN = as.numeric(confMat$table[1,2])
FP = as.numeric(confMat$table[2,1])
TP = as.numeric(confMat$table[2,2])
rfperf_K562_50kb_clustered_smote_dist[1,2] <- TN
rfperf_K562_50kb_clustered_smote_dist[2,2] <- FN
rfperf_K562_50kb_clustered_smote_dist[3,2] <- FP
rfperf_K562_50kb_clustered_smote_dist[4,2] <- TP
rfperf_K562_50kb_clustered_smote_dist[5,2] <- sum(confMat$table)
rfperf_K562_50kb_clustered_smote_dist[6,2] <- as.vector(confMat$byClass["Sensitivity"])
rfperf_K562_50kb_clustered_smote_dist[7,2] <- as.vector(confMat$byClass["Specificity"])
rfperf_K562_50kb_clustered_smote_dist[8,2] <- as.vector(confMat$overall["Kappa"])
rfperf_K562_50kb_clustered_smote_dist[9,2] <- as.vector(confMat$overall["Accuracy"])
rfperf_K562_50kb_clustered_smote_dist[10,2] <- TP/(TP+FP)
rfperf_K562_50kb_clustered_smote_dist[11,2] <- FP/(FP+TN)
rfperf_K562_50kb_clustered_smote_dist[12,2] <- FN/(FN+TN)
rfperf_K562_50kb_clustered_smote_dist[13,2] <- FN/(FN+TN)
rfperf_K562_50kb_clustered_smote_dist[14,2] <- TN/(TN+FN)
rfperf_K562_50kb_clustered_smote_dist[15,2] <- (TP*TN - FP*FN)/( sqrt( (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN) ) )
rfperf_K562_50kb_clustered_smote_dist[16,2] <- (2*(TP/(TP+FN))*(TP/(TP+FP)))/((TP/(TP+FN)) + ((TP/(TP+FP))))
rfperf_K562_50kb_clustered_smote_dist[17,2] <- pROC::auc(pROC::roc(test$y, pred.rfModel))			
rfperf_K562_50kb_clustered_smote_dist[18,2] <- (TP/(TP + FN)) + (TN/(TN + FP)) - 1
rfperf_K562_50kb_clustered_smote_dist[19,2] <- pr$auc.integral

#100kb
rfmodel_gm12878_100kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
####reading in test data for K562
test <- readRDS("Z:/TAD_data_analysis/K562/100kb/clustered/training_testing/test.rds")
####reading in enetcoefs for K562
vars <- rfmodel_gm12878_100kb_clustered_smote_dist$finalModel$xNames
####reduce test data to match train data
names(test) <- gsub(paste(c("`", 
                          "Clustered", 
                          "-", 
                          #"_", 
                          #"dist", "count", "perc",
                          "TfbsV3"), collapse = "|"), "", names(test))
test <- test[,c(1,which(names(test) %in% vars))]
rfperf_K562_100kb_clustered_smote_dist <- data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)
pred.rfModel <- as.vector(predict(rfmodel_gm12878_100kb_clustered_smote_dist, 
                                  newdata=test, 
                                  type="prob")[,"Yes"])

fg <- pred.rfModel[test$y == "Yes"]
bg <- pred.rfModel[test$y == "No"]
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)

pred.rfModel2 <- predict(rfmodel_gm12878_100kb_clustered_smote_dist,
                         newdata=test,
                         type="raw")

confMat <- confusionMatrix(data=pred.rfModel2, test$y, positive="Yes")
TN = as.numeric(confMat$table[1,1])
FN = as.numeric(confMat$table[1,2])
FP = as.numeric(confMat$table[2,1])
TP = as.numeric(confMat$table[2,2])
rfperf_K562_100kb_clustered_smote_dist[1,2] <- TN
rfperf_K562_100kb_clustered_smote_dist[2,2] <- FN
rfperf_K562_100kb_clustered_smote_dist[3,2] <- FP
rfperf_K562_100kb_clustered_smote_dist[4,2] <- TP
rfperf_K562_100kb_clustered_smote_dist[5,2] <- sum(confMat$table)
rfperf_K562_100kb_clustered_smote_dist[6,2] <- as.vector(confMat$byClass["Sensitivity"])
rfperf_K562_100kb_clustered_smote_dist[7,2] <- as.vector(confMat$byClass["Specificity"])
rfperf_K562_100kb_clustered_smote_dist[8,2] <- as.vector(confMat$overall["Kappa"])
rfperf_K562_100kb_clustered_smote_dist[9,2] <- as.vector(confMat$overall["Accuracy"])
rfperf_K562_100kb_clustered_smote_dist[10,2] <- TP/(TP+FP)
rfperf_K562_100kb_clustered_smote_dist[11,2] <- FP/(FP+TN)
rfperf_K562_100kb_clustered_smote_dist[12,2] <- FN/(FN+TN)
rfperf_K562_100kb_clustered_smote_dist[13,2] <- FN/(FN+TN)
rfperf_K562_100kb_clustered_smote_dist[14,2] <- TN/(TN+FN)
rfperf_K562_100kb_clustered_smote_dist[15,2] <- (TP*TN - FP*FN)/( sqrt( (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN) ) )
rfperf_K562_100kb_clustered_smote_dist[16,2] <- (2*(TP/(TP+FN))*(TP/(TP+FP)))/((TP/(TP+FN)) + ((TP/(TP+FP))))
rfperf_K562_100kb_clustered_smote_dist[17,2] <- pROC::auc(pROC::roc(test$y, pred.rfModel))			
rfperf_K562_100kb_clustered_smote_dist[18,2] <- (TP/(TP + FN)) + (TN/(TN + FP)) - 1
rfperf_K562_100kb_clustered_smote_dist[19,2] <- pr$auc.integral

```

##Plots

```{r}
rfperf_gm12878_validatedon_k562 <- data.frame(Model.Validation=rep(c("GM12878 validated on GM12878", "GM12878 validated on K562"), 8),
                                              Resolution=rep(c(rep("10 kb", 2),
                                                           rep("25 kb", 2),
                                                           rep("50 kb", 2),
                                                           rep("100 kb", 2)), 2),
                                              Metric=c(rep("F1-Score", 8),
                                                       rep("MCC", 8)))

rfperf_gm12878_validatedon_k562$Performance <- c(rfperf_gm12878_10kb_clustered_smote_dist[16,2], #f1
                                                 rfperf_K562_10kb_clustered_smote_dist[16,2], #f1

                                                 rfperf_gm12878_25kb_clustered_smote_dist[16,2], #f1
                                                 rfperf_K562_25kb_clustered_smote_dist[16,2], #f1

                                                 rfperf_gm12878_50kb_clustered_smote_dist[16,2], #f1
                                                 rfperf_K562_50kb_clustered_smote_dist[16,2], #f1

                                                 rfperf_gm12878_100kb_clustered_smote_dist[16,2], #f1
                                                 rfperf_K562_100kb_clustered_smote_dist[16,2], #f1

                                                 rfperf_gm12878_10kb_clustered_smote_dist[15,2], #mcc
                                                 rfperf_K562_10kb_clustered_smote_dist[15,2], #mcc

                                                 rfperf_gm12878_25kb_clustered_smote_dist[15,2], #mcc
                                                 rfperf_K562_25kb_clustered_smote_dist[15,2], #mcc

                                                 rfperf_gm12878_50kb_clustered_smote_dist[15,2], #mcc
                                                 rfperf_K562_50kb_clustered_smote_dist[15,2], #mcc

                                                 rfperf_gm12878_100kb_clustered_smote_dist[15,2], #mcc
                                                 rfperf_K562_100kb_clustered_smote_dist[15,2] #mcc
                                                 )

rfperf_gm12878_validatedon_k562$Resolution <- factor(rfperf_gm12878_validatedon_k562$Resolution, levels=c("10 kb", "25 kb", "50 kb", "100 kb"))
rfperf_gm12878_validatedon_k562$Metric <- factor(rfperf_gm12878_validatedon_k562$Metric, levels=c("F1-Score", "MCC"))
rfperf_gm12878_validatedon_k562$Model.Validation <- factor(rfperf_gm12878_validatedon_k562$Model.Validation, levels=c("GM12878 validated on GM12878", "GM12878 validated on K562"))

ggplot(rfperf_gm12878_validatedon_k562,
       aes(x=Resolution, y=Performance, fill=Model.Validation)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid( ~ Metric) +
  #xlab("") + ylab("") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Model Validation")+
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        strip.text.x = element_text(size = 15))                                                 
```

# Reading in K562; smote; distance types

## Validating on K562

```{r}
#10kb
rfperf_K562_10kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#25kb
rfperf_K562_25kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#50kb
rfperf_K562_50kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

#100kb
rfperf_K562_100kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfperf.rds")

```

## Validating on GM12878

```{r}
#10kb
rfmodel_K562_10kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/10kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
####reading in test data for GM12878
test <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/clustered/training_testing/test.rds")
####reading in enetcoefs for GM12878
vars <- rfmodel_K562_10kb_clustered_smote_dist$finalModel$xNames
####reduce test data to match train data
names(test) <- gsub(paste(c("`", 
                          "Clustered", 
                          "-", 
                          #"_", 
                          #"dist", "count", "perc",
                          "TfbsV3"), collapse = "|"), "", names(test))
test <- test[,c(1,which(names(test) %in% vars))]
rfperf_GM12878_10kb_clustered_smote_dist <- data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)
pred.rfModel <- as.vector(predict(rfmodel_K562_10kb_clustered_smote_dist, 
                                  newdata=test, 
                                  type="prob")[,"Yes"])

fg <- pred.rfModel[test$y == "Yes"]
bg <- pred.rfModel[test$y == "No"]
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)

pred.rfModel2 <- predict(rfmodel_K562_10kb_clustered_smote_dist,
                         newdata=test,
                         type="raw")

confMat <- confusionMatrix(data=pred.rfModel2, test$y, positive="Yes")
TN = as.numeric(confMat$table[1,1])
FN = as.numeric(confMat$table[1,2])
FP = as.numeric(confMat$table[2,1])
TP = as.numeric(confMat$table[2,2])
rfperf_GM12878_10kb_clustered_smote_dist[1,2] <- TN
rfperf_GM12878_10kb_clustered_smote_dist[2,2] <- FN
rfperf_GM12878_10kb_clustered_smote_dist[3,2] <- FP
rfperf_GM12878_10kb_clustered_smote_dist[4,2] <- TP
rfperf_GM12878_10kb_clustered_smote_dist[5,2] <- sum(confMat$table)
rfperf_GM12878_10kb_clustered_smote_dist[6,2] <- as.vector(confMat$byClass["Sensitivity"])
rfperf_GM12878_10kb_clustered_smote_dist[7,2] <- as.vector(confMat$byClass["Specificity"])
rfperf_GM12878_10kb_clustered_smote_dist[8,2] <- as.vector(confMat$overall["Kappa"])
rfperf_GM12878_10kb_clustered_smote_dist[9,2] <- as.vector(confMat$overall["Accuracy"])
rfperf_GM12878_10kb_clustered_smote_dist[10,2] <- TP/(TP+FP)
rfperf_GM12878_10kb_clustered_smote_dist[11,2] <- FP/(FP+TN)
rfperf_GM12878_10kb_clustered_smote_dist[12,2] <- FN/(FN+TN)
rfperf_GM12878_10kb_clustered_smote_dist[13,2] <- FN/(FN+TN)
rfperf_GM12878_10kb_clustered_smote_dist[14,2] <- TN/(TN+FN)
rfperf_GM12878_10kb_clustered_smote_dist[15,2] <- (TP*TN - FP*FN)/( sqrt( (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN) ) )
rfperf_GM12878_10kb_clustered_smote_dist[16,2] <- (2*(TP/(TP+FN))*(TP/(TP+FP)))/((TP/(TP+FN)) + ((TP/(TP+FP))))
rfperf_GM12878_10kb_clustered_smote_dist[17,2] <- pROC::auc(pROC::roc(test$y, pred.rfModel))			
rfperf_GM12878_10kb_clustered_smote_dist[18,2] <- (TP/(TP + FN)) + (TN/(TN + FP)) - 1
rfperf_GM12878_10kb_clustered_smote_dist[19,2] <- pr$auc.integral

#25kb
rfmodel_K562_25kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/25kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
####reading in test data for GM12878
test <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/clustered/training_testing/test.rds")
####reading in enetcoefs for GM12878
vars <- rfmodel_K562_25kb_clustered_smote_dist$finalModel$xNames
####reduce test data to match train data
names(test) <- gsub(paste(c("`", 
                          "Clustered", 
                          "-", 
                          #"_", 
                          #"dist", "count", "perc",
                          "TfbsV3"), collapse = "|"), "", names(test))
test <- test[,c(1,which(names(test) %in% vars))]
rfperf_GM12878_25kb_clustered_smote_dist <- data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)
pred.rfModel <- as.vector(predict(rfmodel_K562_25kb_clustered_smote_dist, 
                                  newdata=test, 
                                  type="prob")[,"Yes"])

fg <- pred.rfModel[test$y == "Yes"]
bg <- pred.rfModel[test$y == "No"]
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)

pred.rfModel2 <- predict(rfmodel_K562_25kb_clustered_smote_dist,
                         newdata=test,
                         type="raw")

confMat <- confusionMatrix(data=pred.rfModel2, test$y, positive="Yes")
TN = as.numeric(confMat$table[1,1])
FN = as.numeric(confMat$table[1,2])
FP = as.numeric(confMat$table[2,1])
TP = as.numeric(confMat$table[2,2])
rfperf_GM12878_25kb_clustered_smote_dist[1,2] <- TN
rfperf_GM12878_25kb_clustered_smote_dist[2,2] <- FN
rfperf_GM12878_25kb_clustered_smote_dist[3,2] <- FP
rfperf_GM12878_25kb_clustered_smote_dist[4,2] <- TP
rfperf_GM12878_25kb_clustered_smote_dist[5,2] <- sum(confMat$table)
rfperf_GM12878_25kb_clustered_smote_dist[6,2] <- as.vector(confMat$byClass["Sensitivity"])
rfperf_GM12878_25kb_clustered_smote_dist[7,2] <- as.vector(confMat$byClass["Specificity"])
rfperf_GM12878_25kb_clustered_smote_dist[8,2] <- as.vector(confMat$overall["Kappa"])
rfperf_GM12878_25kb_clustered_smote_dist[9,2] <- as.vector(confMat$overall["Accuracy"])
rfperf_GM12878_25kb_clustered_smote_dist[10,2] <- TP/(TP+FP)
rfperf_GM12878_25kb_clustered_smote_dist[11,2] <- FP/(FP+TN)
rfperf_GM12878_25kb_clustered_smote_dist[12,2] <- FN/(FN+TN)
rfperf_GM12878_25kb_clustered_smote_dist[13,2] <- FN/(FN+TN)
rfperf_GM12878_25kb_clustered_smote_dist[14,2] <- TN/(TN+FN)
rfperf_GM12878_25kb_clustered_smote_dist[15,2] <- (TP*TN - FP*FN)/( sqrt( (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN) ) )
rfperf_GM12878_25kb_clustered_smote_dist[16,2] <- (2*(TP/(TP+FN))*(TP/(TP+FP)))/((TP/(TP+FN)) + ((TP/(TP+FP))))
rfperf_GM12878_25kb_clustered_smote_dist[17,2] <- pROC::auc(pROC::roc(test$y, pred.rfModel))			
rfperf_GM12878_25kb_clustered_smote_dist[18,2] <- (TP/(TP + FN)) + (TN/(TN + FP)) - 1
rfperf_GM12878_25kb_clustered_smote_dist[19,2] <- pr$auc.integral

#50kb
rfmodel_K562_50kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/50kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
####reading in test data for GM12878
test <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/clustered/training_testing/test.rds")
####reading in enetcoefs for GM12878
vars <- rfmodel_K562_50kb_clustered_smote_dist$finalModel$xNames
####reduce test data to match train data
names(test) <- gsub(paste(c("`", 
                          "Clustered", 
                          "-", 
                          #"_", 
                          #"dist", "count", "perc",
                          "TfbsV3"), collapse = "|"), "", names(test))
test <- test[,c(1,which(names(test) %in% vars))]
rfperf_GM12878_50kb_clustered_smote_dist <- data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)
pred.rfModel <- as.vector(predict(rfmodel_K562_50kb_clustered_smote_dist, 
                                  newdata=test, 
                                  type="prob")[,"Yes"])

fg <- pred.rfModel[test$y == "Yes"]
bg <- pred.rfModel[test$y == "No"]
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)

pred.rfModel2 <- predict(rfmodel_K562_50kb_clustered_smote_dist,
                         newdata=test,
                         type="raw")

confMat <- confusionMatrix(data=pred.rfModel2, test$y, positive="Yes")
TN = as.numeric(confMat$table[1,1])
FN = as.numeric(confMat$table[1,2])
FP = as.numeric(confMat$table[2,1])
TP = as.numeric(confMat$table[2,2])
rfperf_GM12878_50kb_clustered_smote_dist[1,2] <- TN
rfperf_GM12878_50kb_clustered_smote_dist[2,2] <- FN
rfperf_GM12878_50kb_clustered_smote_dist[3,2] <- FP
rfperf_GM12878_50kb_clustered_smote_dist[4,2] <- TP
rfperf_GM12878_50kb_clustered_smote_dist[5,2] <- sum(confMat$table)
rfperf_GM12878_50kb_clustered_smote_dist[6,2] <- as.vector(confMat$byClass["Sensitivity"])
rfperf_GM12878_50kb_clustered_smote_dist[7,2] <- as.vector(confMat$byClass["Specificity"])
rfperf_GM12878_50kb_clustered_smote_dist[8,2] <- as.vector(confMat$overall["Kappa"])
rfperf_GM12878_50kb_clustered_smote_dist[9,2] <- as.vector(confMat$overall["Accuracy"])
rfperf_GM12878_50kb_clustered_smote_dist[10,2] <- TP/(TP+FP)
rfperf_GM12878_50kb_clustered_smote_dist[11,2] <- FP/(FP+TN)
rfperf_GM12878_50kb_clustered_smote_dist[12,2] <- FN/(FN+TN)
rfperf_GM12878_50kb_clustered_smote_dist[13,2] <- FN/(FN+TN)
rfperf_GM12878_50kb_clustered_smote_dist[14,2] <- TN/(TN+FN)
rfperf_GM12878_50kb_clustered_smote_dist[15,2] <- (TP*TN - FP*FN)/( sqrt( (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN) ) )
rfperf_GM12878_50kb_clustered_smote_dist[16,2] <- (2*(TP/(TP+FN))*(TP/(TP+FP)))/((TP/(TP+FN)) + ((TP/(TP+FP))))
rfperf_GM12878_50kb_clustered_smote_dist[17,2] <- pROC::auc(pROC::roc(test$y, pred.rfModel))			
rfperf_GM12878_50kb_clustered_smote_dist[18,2] <- (TP/(TP + FN)) + (TN/(TN + FP)) - 1
rfperf_GM12878_50kb_clustered_smote_dist[19,2] <- pr$auc.integral

#100kb
rfmodel_K562_100kb_clustered_smote_dist <- readRDS("Z:/TAD_data_analysis/K562/100kb/clustered/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
####reading in test data for GM12878
test <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/clustered/training_testing/test.rds")
####reading in enetcoefs for GM12878
vars <- rfmodel_K562_100kb_clustered_smote_dist$finalModel$xNames
####reduce test data to match train data
names(test) <- gsub(paste(c("`", 
                          "Clustered", 
                          "-", 
                          #"_", 
                          #"dist", "count", "perc",
                          "TfbsV3"), collapse = "|"), "", names(test))
test <- test[,c(1,which(names(test) %in% vars))]
rfperf_GM12878_100kb_clustered_smote_dist <- data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)
pred.rfModel <- as.vector(predict(rfmodel_K562_100kb_clustered_smote_dist, 
                                  newdata=test, 
                                  type="prob")[,"Yes"])

fg <- pred.rfModel[test$y == "Yes"]
bg <- pred.rfModel[test$y == "No"]
pr <- pr.curve(scores.class0 = fg, scores.class1 = bg, curve = T)

pred.rfModel2 <- predict(rfmodel_K562_100kb_clustered_smote_dist,
                         newdata=test,
                         type="raw")

confMat <- confusionMatrix(data=pred.rfModel2, test$y, positive="Yes")
TN = as.numeric(confMat$table[1,1])
FN = as.numeric(confMat$table[1,2])
FP = as.numeric(confMat$table[2,1])
TP = as.numeric(confMat$table[2,2])
rfperf_GM12878_100kb_clustered_smote_dist[1,2] <- TN
rfperf_GM12878_100kb_clustered_smote_dist[2,2] <- FN
rfperf_GM12878_100kb_clustered_smote_dist[3,2] <- FP
rfperf_GM12878_100kb_clustered_smote_dist[4,2] <- TP
rfperf_GM12878_100kb_clustered_smote_dist[5,2] <- sum(confMat$table)
rfperf_GM12878_100kb_clustered_smote_dist[6,2] <- as.vector(confMat$byClass["Sensitivity"])
rfperf_GM12878_100kb_clustered_smote_dist[7,2] <- as.vector(confMat$byClass["Specificity"])
rfperf_GM12878_100kb_clustered_smote_dist[8,2] <- as.vector(confMat$overall["Kappa"])
rfperf_GM12878_100kb_clustered_smote_dist[9,2] <- as.vector(confMat$overall["Accuracy"])
rfperf_GM12878_100kb_clustered_smote_dist[10,2] <- TP/(TP+FP)
rfperf_GM12878_100kb_clustered_smote_dist[11,2] <- FP/(FP+TN)
rfperf_GM12878_100kb_clustered_smote_dist[12,2] <- FN/(FN+TN)
rfperf_GM12878_100kb_clustered_smote_dist[13,2] <- FN/(FN+TN)
rfperf_GM12878_100kb_clustered_smote_dist[14,2] <- TN/(TN+FN)
rfperf_GM12878_100kb_clustered_smote_dist[15,2] <- (TP*TN - FP*FN)/( sqrt( (TP+FP)*(TP+FN)*(TN+FP)*(TN+FN) ) )
rfperf_GM12878_100kb_clustered_smote_dist[16,2] <- (2*(TP/(TP+FN))*(TP/(TP+FP)))/((TP/(TP+FN)) + ((TP/(TP+FP))))
rfperf_GM12878_100kb_clustered_smote_dist[17,2] <- pROC::auc(pROC::roc(test$y, pred.rfModel))			
rfperf_GM12878_100kb_clustered_smote_dist[18,2] <- (TP/(TP + FN)) + (TN/(TN + FP)) - 1
rfperf_GM12878_100kb_clustered_smote_dist[19,2] <- pr$auc.integral

```

## Plots

```{r}
rfperf_k562_validatedon_gm12878 <- data.frame(Model.Validation=rep(c("K562 validated on K562", "K562 validated on GM12878"), 8),
                                              Resolution=rep(c(rep("10 kb", 2),
                                                           rep("25 kb", 2),
                                                           rep("50 kb", 2),
                                                           rep("100 kb", 2)), 2),
                                              Metric=c(rep("F1-Score", 8),
                                                       rep("MCC", 8)))

rfperf_k562_validatedon_gm12878$Performance <- c(rfperf_K562_10kb_clustered_smote_dist[16,2], #f1
                                                 rfperf_GM12878_10kb_clustered_smote_dist[16,2], #f1

                                                 rfperf_K562_25kb_clustered_smote_dist[16,2], #f1
                                                 rfperf_GM12878_25kb_clustered_smote_dist[16,2], #f1

                                                 rfperf_K562_50kb_clustered_smote_dist[16,2], #f1
                                                 rfperf_GM12878_50kb_clustered_smote_dist[16,2], #f1

                                                 rfperf_K562_100kb_clustered_smote_dist[16,2], #f1
                                                 rfperf_GM12878_100kb_clustered_smote_dist[16,2], #f1

                                                 rfperf_K562_10kb_clustered_smote_dist[15,2], #mcc
                                                 rfperf_GM12878_10kb_clustered_smote_dist[15,2], #mcc

                                                 rfperf_K562_25kb_clustered_smote_dist[15,2], #mcc
                                                 rfperf_GM12878_25kb_clustered_smote_dist[15,2], #mcc

                                                 rfperf_K562_50kb_clustered_smote_dist[15,2], #mcc
                                                 rfperf_GM12878_50kb_clustered_smote_dist[15,2], #mcc

                                                 rfperf_K562_100kb_clustered_smote_dist[15,2], #mcc
                                                 rfperf_GM12878_100kb_clustered_smote_dist[15,2] #mcc
                                                 )

rfperf_k562_validatedon_gm12878$Resolution <- factor(rfperf_k562_validatedon_gm12878$Resolution, levels=c("10 kb", "25 kb", "50 kb", "100 kb"))
rfperf_k562_validatedon_gm12878$Metric <- factor(rfperf_k562_validatedon_gm12878$Metric, levels=c("F1-Score", "MCC"))
rfperf_k562_validatedon_gm12878$Model.Validation <- factor(rfperf_k562_validatedon_gm12878$Model.Validation, levels=c("K562 validated on K562", "K562 validated on GM12878"))

ggplot(rfperf_k562_validatedon_gm12878,
       aes(x=Resolution, y=Performance, fill=Model.Validation)) +
  geom_bar( stat="identity", position="dodge") +
  facet_grid( ~ Metric) +
  #xlab("") + ylab("") +
  theme_minimal() +
  theme_bw() +
  scale_fill_discrete(name="Model Validation")+
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        strip.text.x = element_text(size = 15))   
```