### 2019-04-21 Scripts/creating_supplementary_table_of_penalization_parameters.Rmd
script to extract alpha and lambda from all models

---
title: "Creating Tables of Elastic Net Parameters"
author: "Spiro Stilianoudakis"
date: "April 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Extracting alpha and lambda penalty parameters from elastic net

## GM12878

### 10kb

#### none

##### OC

```{r}
eNetModel_gm12878_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OC/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_none_oc_alpha <- as.numeric(eNetModel_gm12878_10kb_none_oc$bestTune[1])
eNetModel_gm12878_10kb_none_oc_lambda <- as.numeric(eNetModel_gm12878_10kb_none_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_10kb_none_OP <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/OP/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_none_OP_alpha <- as.numeric(eNetModel_gm12878_10kb_none_OP$bestTune[1])
eNetModel_gm12878_10kb_none_OP_lambda <- as.numeric(eNetModel_gm12878_10kb_none_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_10kb_none_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/None/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_none_Distance_alpha <- as.numeric(eNetModel_gm12878_10kb_none_Distance$bestTune[1])
eNetModel_gm12878_10kb_none_Distance_lambda <- as.numeric(eNetModel_gm12878_10kb_none_Distance$bestTune[2])
```

#### ROS

##### OC

```{r}
eNetModel_gm12878_10kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OC/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_ROS_oc_alpha <- as.numeric(eNetModel_gm12878_10kb_ROS_oc$bestTune[1])
eNetModel_gm12878_10kb_ROS_oc_lambda <- as.numeric(eNetModel_gm12878_10kb_ROS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_10kb_ROS_OP <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/OP/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_ROS_OP_alpha <- as.numeric(eNetModel_gm12878_10kb_ROS_OP$bestTune[1])
eNetModel_gm12878_10kb_ROS_OP_lambda <- as.numeric(eNetModel_gm12878_10kb_ROS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_10kb_ROS_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/ROS/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_ROS_Distance_alpha <- as.numeric(eNetModel_gm12878_10kb_ROS_Distance$bestTune[1])
eNetModel_gm12878_10kb_ROS_Distance_lambda <- as.numeric(eNetModel_gm12878_10kb_ROS_Distance$bestTune[2])
```

#### RUS

##### OC

```{r}
eNetModel_gm12878_10kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OC/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_RUS_oc_alpha <- as.numeric(eNetModel_gm12878_10kb_RUS_oc$bestTune[1])
eNetModel_gm12878_10kb_RUS_oc_lambda <- as.numeric(eNetModel_gm12878_10kb_RUS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_10kb_RUS_OP <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/OP/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_RUS_OP_alpha <- as.numeric(eNetModel_gm12878_10kb_RUS_OP$bestTune[1])
eNetModel_gm12878_10kb_RUS_OP_lambda <- as.numeric(eNetModel_gm12878_10kb_RUS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_10kb_RUS_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/RUS/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_RUS_Distance_alpha <- as.numeric(eNetModel_gm12878_10kb_RUS_Distance$bestTune[1])
eNetModel_gm12878_10kb_RUS_Distance_lambda <- as.numeric(eNetModel_gm12878_10kb_RUS_Distance$bestTune[2])
```

#### SMOTE

##### OC

```{r}
eNetModel_gm12878_10kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OC/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_SMOTE_oc_alpha <- as.numeric(eNetModel_gm12878_10kb_SMOTE_oc$bestTune[1])
eNetModel_gm12878_10kb_SMOTE_oc_lambda <- as.numeric(eNetModel_gm12878_10kb_SMOTE_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_10kb_SMOTE_OP <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/OP/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_SMOTE_OP_alpha <- as.numeric(eNetModel_gm12878_10kb_SMOTE_OP$bestTune[1])
eNetModel_gm12878_10kb_SMOTE_OP_lambda <- as.numeric(eNetModel_gm12878_10kb_SMOTE_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_10kb_SMOTE_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_10kb_SMOTE_Distance_alpha <- as.numeric(eNetModel_gm12878_10kb_SMOTE_Distance$bestTune[1])
eNetModel_gm12878_10kb_SMOTE_Distance_lambda <- as.numeric(eNetModel_gm12878_10kb_SMOTE_Distance$bestTune[2])
```

### 25kb

#### none

##### OC

```{r}
eNetModel_gm12878_25kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OC/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_none_oc_alpha <- as.numeric(eNetModel_gm12878_25kb_none_oc$bestTune[1])
eNetModel_gm12878_25kb_none_oc_lambda <- as.numeric(eNetModel_gm12878_25kb_none_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_25kb_none_OP <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/OP/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_none_OP_alpha <- as.numeric(eNetModel_gm12878_25kb_none_OP$bestTune[1])
eNetModel_gm12878_25kb_none_OP_lambda <- as.numeric(eNetModel_gm12878_25kb_none_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_25kb_none_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/None/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_none_Distance_alpha <- as.numeric(eNetModel_gm12878_25kb_none_Distance$bestTune[1])
eNetModel_gm12878_25kb_none_Distance_lambda <- as.numeric(eNetModel_gm12878_25kb_none_Distance$bestTune[2])
```

#### ROS

##### OC

```{r}
eNetModel_gm12878_25kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OC/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_ROS_oc_alpha <- as.numeric(eNetModel_gm12878_25kb_ROS_oc$bestTune[1])
eNetModel_gm12878_25kb_ROS_oc_lambda <- as.numeric(eNetModel_gm12878_25kb_ROS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_25kb_ROS_OP <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/OP/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_ROS_OP_alpha <- as.numeric(eNetModel_gm12878_25kb_ROS_OP$bestTune[1])
eNetModel_gm12878_25kb_ROS_OP_lambda <- as.numeric(eNetModel_gm12878_25kb_ROS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_25kb_ROS_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/ROS/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_ROS_Distance_alpha <- as.numeric(eNetModel_gm12878_25kb_ROS_Distance$bestTune[1])
eNetModel_gm12878_25kb_ROS_Distance_lambda <- as.numeric(eNetModel_gm12878_25kb_ROS_Distance$bestTune[2])
```

#### RUS

##### OC

```{r}
eNetModel_gm12878_25kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OC/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_RUS_oc_alpha <- as.numeric(eNetModel_gm12878_25kb_RUS_oc$bestTune[1])
eNetModel_gm12878_25kb_RUS_oc_lambda <- as.numeric(eNetModel_gm12878_25kb_RUS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_25kb_RUS_OP <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/OP/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_RUS_OP_alpha <- as.numeric(eNetModel_gm12878_25kb_RUS_OP$bestTune[1])
eNetModel_gm12878_25kb_RUS_OP_lambda <- as.numeric(eNetModel_gm12878_25kb_RUS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_25kb_RUS_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/RUS/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_RUS_Distance_alpha <- as.numeric(eNetModel_gm12878_25kb_RUS_Distance$bestTune[1])
eNetModel_gm12878_25kb_RUS_Distance_lambda <- as.numeric(eNetModel_gm12878_25kb_RUS_Distance$bestTune[2])
```

#### SMOTE

##### OC

```{r}
eNetModel_gm12878_25kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OC/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_SMOTE_oc_alpha <- as.numeric(eNetModel_gm12878_25kb_SMOTE_oc$bestTune[1])
eNetModel_gm12878_25kb_SMOTE_oc_lambda <- as.numeric(eNetModel_gm12878_25kb_SMOTE_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_25kb_SMOTE_OP <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/OP/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_SMOTE_OP_alpha <- as.numeric(eNetModel_gm12878_25kb_SMOTE_OP$bestTune[1])
eNetModel_gm12878_25kb_SMOTE_OP_lambda <- as.numeric(eNetModel_gm12878_25kb_SMOTE_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_25kb_SMOTE_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/SMOTE/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_25kb_SMOTE_Distance_alpha <- as.numeric(eNetModel_gm12878_25kb_SMOTE_Distance$bestTune[1])
eNetModel_gm12878_25kb_SMOTE_Distance_lambda <- as.numeric(eNetModel_gm12878_25kb_SMOTE_Distance$bestTune[2])
```

### 50kb

#### none

##### OC

```{r}
eNetModel_gm12878_50kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OC/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_none_oc_alpha <- as.numeric(eNetModel_gm12878_50kb_none_oc$bestTune[1])
eNetModel_gm12878_50kb_none_oc_lambda <- as.numeric(eNetModel_gm12878_50kb_none_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_50kb_none_OP <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/OP/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_none_OP_alpha <- as.numeric(eNetModel_gm12878_50kb_none_OP$bestTune[1])
eNetModel_gm12878_50kb_none_OP_lambda <- as.numeric(eNetModel_gm12878_50kb_none_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_50kb_none_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/None/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_none_Distance_alpha <- as.numeric(eNetModel_gm12878_50kb_none_Distance$bestTune[1])
eNetModel_gm12878_50kb_none_Distance_lambda <- as.numeric(eNetModel_gm12878_50kb_none_Distance$bestTune[2])
```

#### ROS

##### OC

```{r}
eNetModel_gm12878_50kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OC/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_ROS_oc_alpha <- as.numeric(eNetModel_gm12878_50kb_ROS_oc$bestTune[1])
eNetModel_gm12878_50kb_ROS_oc_lambda <- as.numeric(eNetModel_gm12878_50kb_ROS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_50kb_ROS_OP <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/OP/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_ROS_OP_alpha <- as.numeric(eNetModel_gm12878_50kb_ROS_OP$bestTune[1])
eNetModel_gm12878_50kb_ROS_OP_lambda <- as.numeric(eNetModel_gm12878_50kb_ROS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_50kb_ROS_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/ROS/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_ROS_Distance_alpha <- as.numeric(eNetModel_gm12878_50kb_ROS_Distance$bestTune[1])
eNetModel_gm12878_50kb_ROS_Distance_lambda <- as.numeric(eNetModel_gm12878_50kb_ROS_Distance$bestTune[2])
```

#### RUS

##### OC

```{r}
eNetModel_gm12878_50kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OC/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_RUS_oc_alpha <- as.numeric(eNetModel_gm12878_50kb_RUS_oc$bestTune[1])
eNetModel_gm12878_50kb_RUS_oc_lambda <- as.numeric(eNetModel_gm12878_50kb_RUS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_50kb_RUS_OP <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/OP/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_RUS_OP_alpha <- as.numeric(eNetModel_gm12878_50kb_RUS_OP$bestTune[1])
eNetModel_gm12878_50kb_RUS_OP_lambda <- as.numeric(eNetModel_gm12878_50kb_RUS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_50kb_RUS_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/RUS/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_RUS_Distance_alpha <- as.numeric(eNetModel_gm12878_50kb_RUS_Distance$bestTune[1])
eNetModel_gm12878_50kb_RUS_Distance_lambda <- as.numeric(eNetModel_gm12878_50kb_RUS_Distance$bestTune[2])
```

#### SMOTE

##### OC

```{r}
eNetModel_gm12878_50kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OC/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_SMOTE_oc_alpha <- as.numeric(eNetModel_gm12878_50kb_SMOTE_oc$bestTune[1])
eNetModel_gm12878_50kb_SMOTE_oc_lambda <- as.numeric(eNetModel_gm12878_50kb_SMOTE_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_50kb_SMOTE_OP <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/OP/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_SMOTE_OP_alpha <- as.numeric(eNetModel_gm12878_50kb_SMOTE_OP$bestTune[1])
eNetModel_gm12878_50kb_SMOTE_OP_lambda <- as.numeric(eNetModel_gm12878_50kb_SMOTE_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_50kb_SMOTE_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/SMOTE/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_50kb_SMOTE_Distance_alpha <- as.numeric(eNetModel_gm12878_50kb_SMOTE_Distance$bestTune[1])
eNetModel_gm12878_50kb_SMOTE_Distance_lambda <- as.numeric(eNetModel_gm12878_50kb_SMOTE_Distance$bestTune[2])
```

### 100kb

#### none

##### OC

```{r}
eNetModel_gm12878_100kb_none_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OC/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_none_oc_alpha <- as.numeric(eNetModel_gm12878_100kb_none_oc$bestTune[1])
eNetModel_gm12878_100kb_none_oc_lambda <- as.numeric(eNetModel_gm12878_100kb_none_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_100kb_none_OP <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/OP/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_none_OP_alpha <- as.numeric(eNetModel_gm12878_100kb_none_OP$bestTune[1])
eNetModel_gm12878_100kb_none_OP_lambda <- as.numeric(eNetModel_gm12878_100kb_none_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_100kb_none_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/None/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_none_Distance_alpha <- as.numeric(eNetModel_gm12878_100kb_none_Distance$bestTune[1])
eNetModel_gm12878_100kb_none_Distance_lambda <- as.numeric(eNetModel_gm12878_100kb_none_Distance$bestTune[2])
```

#### ROS

##### OC

```{r}
eNetModel_gm12878_100kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OC/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_ROS_oc_alpha <- as.numeric(eNetModel_gm12878_100kb_ROS_oc$bestTune[1])
eNetModel_gm12878_100kb_ROS_oc_lambda <- as.numeric(eNetModel_gm12878_100kb_ROS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_100kb_ROS_OP <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/OP/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_ROS_OP_alpha <- as.numeric(eNetModel_gm12878_100kb_ROS_OP$bestTune[1])
eNetModel_gm12878_100kb_ROS_OP_lambda <- as.numeric(eNetModel_gm12878_100kb_ROS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_100kb_ROS_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/ROS/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_ROS_Distance_alpha <- as.numeric(eNetModel_gm12878_100kb_ROS_Distance$bestTune[1])
eNetModel_gm12878_100kb_ROS_Distance_lambda <- as.numeric(eNetModel_gm12878_100kb_ROS_Distance$bestTune[2])
```

#### RUS

##### OC

```{r}
eNetModel_gm12878_100kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OC/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_RUS_oc_alpha <- as.numeric(eNetModel_gm12878_100kb_RUS_oc$bestTune[1])
eNetModel_gm12878_100kb_RUS_oc_lambda <- as.numeric(eNetModel_gm12878_100kb_RUS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_100kb_RUS_OP <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/OP/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_RUS_OP_alpha <- as.numeric(eNetModel_gm12878_100kb_RUS_OP$bestTune[1])
eNetModel_gm12878_100kb_RUS_OP_lambda <- as.numeric(eNetModel_gm12878_100kb_RUS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_100kb_RUS_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/RUS/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_RUS_Distance_alpha <- as.numeric(eNetModel_gm12878_100kb_RUS_Distance$bestTune[1])
eNetModel_gm12878_100kb_RUS_Distance_lambda <- as.numeric(eNetModel_gm12878_100kb_RUS_Distance$bestTune[2])
```

#### SMOTE

##### OC

```{r}
eNetModel_gm12878_100kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OC/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_SMOTE_oc_alpha <- as.numeric(eNetModel_gm12878_100kb_SMOTE_oc$bestTune[1])
eNetModel_gm12878_100kb_SMOTE_oc_lambda <- as.numeric(eNetModel_gm12878_100kb_SMOTE_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_gm12878_100kb_SMOTE_OP <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/OP/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_SMOTE_OP_alpha <- as.numeric(eNetModel_gm12878_100kb_SMOTE_OP$bestTune[1])
eNetModel_gm12878_100kb_SMOTE_OP_lambda <- as.numeric(eNetModel_gm12878_100kb_SMOTE_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_gm12878_100kb_SMOTE_Distance <- readRDS("Z:/TAD_data_analysis/GM12878/100kb/Ensemble/SMOTE/Distance/ENET/eNetModel.rds")
eNetModel_gm12878_100kb_SMOTE_Distance_alpha <- as.numeric(eNetModel_gm12878_100kb_SMOTE_Distance$bestTune[1])
eNetModel_gm12878_100kb_SMOTE_Distance_lambda <- as.numeric(eNetModel_gm12878_100kb_SMOTE_Distance$bestTune[2])
```

## Creating Table

```{r}
eNetModel_gm12878_parameter_tab <- data.frame(Alpha=c(eNetModel_gm12878_10kb_none_oc_alpha,
                                                      eNetModel_gm12878_10kb_none_OP_alpha,
                                                      eNetModel_gm12878_10kb_none_Distance_alpha,

                                                      eNetModel_gm12878_10kb_ROS_oc_alpha,
                                                      eNetModel_gm12878_10kb_ROS_OP_alpha,
                                                      eNetModel_gm12878_10kb_ROS_Distance_alpha,

                                                      eNetModel_gm12878_10kb_RUS_oc_alpha,
                                                      eNetModel_gm12878_10kb_RUS_OP_alpha,
                                                      eNetModel_gm12878_10kb_RUS_Distance_alpha,

                                                      eNetModel_gm12878_10kb_SMOTE_oc_alpha,
                                                      eNetModel_gm12878_10kb_SMOTE_OP_alpha,
                                                      eNetModel_gm12878_10kb_SMOTE_Distance_alpha,

                                                      eNetModel_gm12878_25kb_none_oc_alpha,
                                                      eNetModel_gm12878_25kb_none_OP_alpha,
                                                      eNetModel_gm12878_25kb_none_Distance_alpha,

                                                      eNetModel_gm12878_25kb_ROS_oc_alpha,
                                                      eNetModel_gm12878_25kb_ROS_OP_alpha,
                                                      eNetModel_gm12878_25kb_ROS_Distance_alpha,

                                                      eNetModel_gm12878_25kb_RUS_oc_alpha,
                                                      eNetModel_gm12878_25kb_RUS_OP_alpha,
                                                      eNetModel_gm12878_25kb_RUS_Distance_alpha,

                                                      eNetModel_gm12878_25kb_SMOTE_oc_alpha,
                                                      eNetModel_gm12878_25kb_SMOTE_OP_alpha,
                                                      eNetModel_gm12878_25kb_SMOTE_Distance_alpha,

                                                      eNetModel_gm12878_50kb_none_oc_alpha,
                                                      eNetModel_gm12878_50kb_none_OP_alpha,
                                                      eNetModel_gm12878_50kb_none_Distance_alpha,

                                                      eNetModel_gm12878_50kb_ROS_oc_alpha,
                                                      eNetModel_gm12878_50kb_ROS_OP_alpha,
                                                      eNetModel_gm12878_50kb_ROS_Distance_alpha,

                                                      eNetModel_gm12878_50kb_RUS_oc_alpha,
                                                      eNetModel_gm12878_50kb_RUS_OP_alpha,
                                                      eNetModel_gm12878_50kb_RUS_Distance_alpha,

                                                      eNetModel_gm12878_50kb_SMOTE_oc_alpha,
                                                      eNetModel_gm12878_50kb_SMOTE_OP_alpha,
                                                      eNetModel_gm12878_50kb_SMOTE_Distance_alpha,

                                                      eNetModel_gm12878_100kb_none_oc_alpha,
                                                      eNetModel_gm12878_100kb_none_OP_alpha,
                                                      eNetModel_gm12878_100kb_none_Distance_alpha,

                                                      eNetModel_gm12878_100kb_ROS_oc_alpha,
                                                      eNetModel_gm12878_100kb_ROS_OP_alpha,
                                                      eNetModel_gm12878_100kb_ROS_Distance_alpha,

                                                      eNetModel_gm12878_100kb_RUS_oc_alpha,
                                                      eNetModel_gm12878_100kb_RUS_OP_alpha,
                                                      eNetModel_gm12878_100kb_RUS_Distance_alpha,

                                                      eNetModel_gm12878_100kb_SMOTE_oc_alpha,
                                                      eNetModel_gm12878_100kb_SMOTE_OP_alpha,
                                                      eNetModel_gm12878_100kb_SMOTE_Distance_alpha),
                                              Lambda=c(eNetModel_gm12878_10kb_none_oc_lambda,
                                                      eNetModel_gm12878_10kb_none_OP_lambda,
                                                      eNetModel_gm12878_10kb_none_Distance_lambda,

                                                      eNetModel_gm12878_10kb_ROS_oc_lambda,
                                                      eNetModel_gm12878_10kb_ROS_OP_lambda,
                                                      eNetModel_gm12878_10kb_ROS_Distance_lambda,

                                                      eNetModel_gm12878_10kb_RUS_oc_lambda,
                                                      eNetModel_gm12878_10kb_RUS_OP_lambda,
                                                      eNetModel_gm12878_10kb_RUS_Distance_lambda,

                                                      eNetModel_gm12878_10kb_SMOTE_oc_lambda,
                                                      eNetModel_gm12878_10kb_SMOTE_OP_lambda,
                                                      eNetModel_gm12878_10kb_SMOTE_Distance_lambda,

                                                      eNetModel_gm12878_25kb_none_oc_lambda,
                                                      eNetModel_gm12878_25kb_none_OP_lambda,
                                                      eNetModel_gm12878_25kb_none_Distance_lambda,

                                                      eNetModel_gm12878_25kb_ROS_oc_lambda,
                                                      eNetModel_gm12878_25kb_ROS_OP_lambda,
                                                      eNetModel_gm12878_25kb_ROS_Distance_lambda,

                                                      eNetModel_gm12878_25kb_RUS_oc_lambda,
                                                      eNetModel_gm12878_25kb_RUS_OP_lambda,
                                                      eNetModel_gm12878_25kb_RUS_Distance_lambda,

                                                      eNetModel_gm12878_25kb_SMOTE_oc_lambda,
                                                      eNetModel_gm12878_25kb_SMOTE_OP_lambda,
                                                      eNetModel_gm12878_25kb_SMOTE_Distance_lambda,

                                                      eNetModel_gm12878_50kb_none_oc_lambda,
                                                      eNetModel_gm12878_50kb_none_OP_lambda,
                                                      eNetModel_gm12878_50kb_none_Distance_lambda,

                                                      eNetModel_gm12878_50kb_ROS_oc_lambda,
                                                      eNetModel_gm12878_50kb_ROS_OP_lambda,
                                                      eNetModel_gm12878_50kb_ROS_Distance_lambda,

                                                      eNetModel_gm12878_50kb_RUS_oc_lambda,
                                                      eNetModel_gm12878_50kb_RUS_OP_lambda,
                                                      eNetModel_gm12878_50kb_RUS_Distance_lambda,

                                                      eNetModel_gm12878_50kb_SMOTE_oc_lambda,
                                                      eNetModel_gm12878_50kb_SMOTE_OP_lambda,
                                                      eNetModel_gm12878_50kb_SMOTE_Distance_lambda,

                                                      eNetModel_gm12878_100kb_none_oc_lambda,
                                                      eNetModel_gm12878_100kb_none_OP_lambda,
                                                      eNetModel_gm12878_100kb_none_Distance_lambda,

                                                      eNetModel_gm12878_100kb_ROS_oc_lambda,
                                                      eNetModel_gm12878_100kb_ROS_OP_lambda,
                                                      eNetModel_gm12878_100kb_ROS_Distance_lambda,

                                                      eNetModel_gm12878_100kb_RUS_oc_lambda,
                                                      eNetModel_gm12878_100kb_RUS_OP_lambda,
                                                      eNetModel_gm12878_100kb_RUS_Distance_lambda,

                                                      eNetModel_gm12878_100kb_SMOTE_oc_lambda,
                                                      eNetModel_gm12878_100kb_SMOTE_OP_lambda,
                                                      eNetModel_gm12878_100kb_SMOTE_Distance_lambda))

eNetModel_gm12878_parameter_tab$CL <- "GM12878"
eNetModel_gm12878_parameter_tab$Resolution <- c(rep("10 kb", 12),
                                                rep("25 kb", 12),
                                                rep("50 kb", 12),
                                                rep("100 kb", 12))
eNetModel_gm12878_parameter_tab$PredictorType <- rep(c("OC","OP","Distance"),16)
eNetModel_gm12878_parameter_tab <- eNetModel_gm12878_parameter_tab[,c(3:5,1:2)]
```

## K562

### 10kb

#### none

##### OC

```{r}
eNetModel_K562_10kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OC/ENET/eNetModel.rds")
eNetModel_K562_10kb_none_oc_alpha <- as.numeric(eNetModel_K562_10kb_none_oc$bestTune[1])
eNetModel_K562_10kb_none_oc_lambda <- as.numeric(eNetModel_K562_10kb_none_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_10kb_none_OP <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/OP/ENET/eNetModel.rds")
eNetModel_K562_10kb_none_OP_alpha <- as.numeric(eNetModel_K562_10kb_none_OP$bestTune[1])
eNetModel_K562_10kb_none_OP_lambda <- as.numeric(eNetModel_K562_10kb_none_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_10kb_none_Distance <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/None/Distance/ENET/eNetModel.rds")
eNetModel_K562_10kb_none_Distance_alpha <- as.numeric(eNetModel_K562_10kb_none_Distance$bestTune[1])
eNetModel_K562_10kb_none_Distance_lambda <- as.numeric(eNetModel_K562_10kb_none_Distance$bestTune[2])
```

#### ROS

##### OC

```{r}
eNetModel_K562_10kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/OC/ENET/eNetModel.rds")
eNetModel_K562_10kb_ROS_oc_alpha <- as.numeric(eNetModel_K562_10kb_ROS_oc$bestTune[1])
eNetModel_K562_10kb_ROS_oc_lambda <- as.numeric(eNetModel_K562_10kb_ROS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_10kb_ROS_OP <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/OP/ENET/eNetModel.rds")
eNetModel_K562_10kb_ROS_OP_alpha <- as.numeric(eNetModel_K562_10kb_ROS_OP$bestTune[1])
eNetModel_K562_10kb_ROS_OP_lambda <- as.numeric(eNetModel_K562_10kb_ROS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_10kb_ROS_Distance <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/ROS/Distance/ENET/eNetModel.rds")
eNetModel_K562_10kb_ROS_Distance_alpha <- as.numeric(eNetModel_K562_10kb_ROS_Distance$bestTune[1])
eNetModel_K562_10kb_ROS_Distance_lambda <- as.numeric(eNetModel_K562_10kb_ROS_Distance$bestTune[2])
```

#### RUS

##### OC

```{r}
eNetModel_K562_10kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/OC/ENET/eNetModel.rds")
eNetModel_K562_10kb_RUS_oc_alpha <- as.numeric(eNetModel_K562_10kb_RUS_oc$bestTune[1])
eNetModel_K562_10kb_RUS_oc_lambda <- as.numeric(eNetModel_K562_10kb_RUS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_10kb_RUS_OP <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/OP/ENET/eNetModel.rds")
eNetModel_K562_10kb_RUS_OP_alpha <- as.numeric(eNetModel_K562_10kb_RUS_OP$bestTune[1])
eNetModel_K562_10kb_RUS_OP_lambda <- as.numeric(eNetModel_K562_10kb_RUS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_10kb_RUS_Distance <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/RUS/Distance/ENET/eNetModel.rds")
eNetModel_K562_10kb_RUS_Distance_alpha <- as.numeric(eNetModel_K562_10kb_RUS_Distance$bestTune[1])
eNetModel_K562_10kb_RUS_Distance_lambda <- as.numeric(eNetModel_K562_10kb_RUS_Distance$bestTune[2])
```

#### SMOTE

##### OC

```{r}
eNetModel_K562_10kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/OC/ENET/eNetModel.rds")
eNetModel_K562_10kb_SMOTE_oc_alpha <- as.numeric(eNetModel_K562_10kb_SMOTE_oc$bestTune[1])
eNetModel_K562_10kb_SMOTE_oc_lambda <- as.numeric(eNetModel_K562_10kb_SMOTE_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_10kb_SMOTE_OP <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/OP/ENET/eNetModel.rds")
eNetModel_K562_10kb_SMOTE_OP_alpha <- as.numeric(eNetModel_K562_10kb_SMOTE_OP$bestTune[1])
eNetModel_K562_10kb_SMOTE_OP_lambda <- as.numeric(eNetModel_K562_10kb_SMOTE_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_10kb_SMOTE_Distance <- readRDS("Z:/TAD_data_analysis/K562/10kb/Ensemble/SMOTE/Distance/ENET/eNetModel.rds")
eNetModel_K562_10kb_SMOTE_Distance_alpha <- as.numeric(eNetModel_K562_10kb_SMOTE_Distance$bestTune[1])
eNetModel_K562_10kb_SMOTE_Distance_lambda <- as.numeric(eNetModel_K562_10kb_SMOTE_Distance$bestTune[2])
```

### 25kb

#### none

##### OC

```{r}
eNetModel_K562_25kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OC/ENET/eNetModel.rds")
eNetModel_K562_25kb_none_oc_alpha <- as.numeric(eNetModel_K562_25kb_none_oc$bestTune[1])
eNetModel_K562_25kb_none_oc_lambda <- as.numeric(eNetModel_K562_25kb_none_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_25kb_none_OP <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/OP/ENET/eNetModel.rds")
eNetModel_K562_25kb_none_OP_alpha <- as.numeric(eNetModel_K562_25kb_none_OP$bestTune[1])
eNetModel_K562_25kb_none_OP_lambda <- as.numeric(eNetModel_K562_25kb_none_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_25kb_none_Distance <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/None/Distance/ENET/eNetModel.rds")
eNetModel_K562_25kb_none_Distance_alpha <- as.numeric(eNetModel_K562_25kb_none_Distance$bestTune[1])
eNetModel_K562_25kb_none_Distance_lambda <- as.numeric(eNetModel_K562_25kb_none_Distance$bestTune[2])
```

#### ROS

##### OC

```{r}
eNetModel_K562_25kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/OC/ENET/eNetModel.rds")
eNetModel_K562_25kb_ROS_oc_alpha <- as.numeric(eNetModel_K562_25kb_ROS_oc$bestTune[1])
eNetModel_K562_25kb_ROS_oc_lambda <- as.numeric(eNetModel_K562_25kb_ROS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_25kb_ROS_OP <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/OP/ENET/eNetModel.rds")
eNetModel_K562_25kb_ROS_OP_alpha <- as.numeric(eNetModel_K562_25kb_ROS_OP$bestTune[1])
eNetModel_K562_25kb_ROS_OP_lambda <- as.numeric(eNetModel_K562_25kb_ROS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_25kb_ROS_Distance <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/ROS/Distance/ENET/eNetModel.rds")
eNetModel_K562_25kb_ROS_Distance_alpha <- as.numeric(eNetModel_K562_25kb_ROS_Distance$bestTune[1])
eNetModel_K562_25kb_ROS_Distance_lambda <- as.numeric(eNetModel_K562_25kb_ROS_Distance$bestTune[2])
```

#### RUS

##### OC

```{r}
eNetModel_K562_25kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/OC/ENET/eNetModel.rds")
eNetModel_K562_25kb_RUS_oc_alpha <- as.numeric(eNetModel_K562_25kb_RUS_oc$bestTune[1])
eNetModel_K562_25kb_RUS_oc_lambda <- as.numeric(eNetModel_K562_25kb_RUS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_25kb_RUS_OP <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/OP/ENET/eNetModel.rds")
eNetModel_K562_25kb_RUS_OP_alpha <- as.numeric(eNetModel_K562_25kb_RUS_OP$bestTune[1])
eNetModel_K562_25kb_RUS_OP_lambda <- as.numeric(eNetModel_K562_25kb_RUS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_25kb_RUS_Distance <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/RUS/Distance/ENET/eNetModel.rds")
eNetModel_K562_25kb_RUS_Distance_alpha <- as.numeric(eNetModel_K562_25kb_RUS_Distance$bestTune[1])
eNetModel_K562_25kb_RUS_Distance_lambda <- as.numeric(eNetModel_K562_25kb_RUS_Distance$bestTune[2])
```

#### SMOTE

##### OC

```{r}
eNetModel_K562_25kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/OC/ENET/eNetModel.rds")
eNetModel_K562_25kb_SMOTE_oc_alpha <- as.numeric(eNetModel_K562_25kb_SMOTE_oc$bestTune[1])
eNetModel_K562_25kb_SMOTE_oc_lambda <- as.numeric(eNetModel_K562_25kb_SMOTE_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_25kb_SMOTE_OP <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/OP/ENET/eNetModel.rds")
eNetModel_K562_25kb_SMOTE_OP_alpha <- as.numeric(eNetModel_K562_25kb_SMOTE_OP$bestTune[1])
eNetModel_K562_25kb_SMOTE_OP_lambda <- as.numeric(eNetModel_K562_25kb_SMOTE_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_25kb_SMOTE_Distance <- readRDS("Z:/TAD_data_analysis/K562/25kb/Ensemble/SMOTE/Distance/ENET/eNetModel.rds")
eNetModel_K562_25kb_SMOTE_Distance_alpha <- as.numeric(eNetModel_K562_25kb_SMOTE_Distance$bestTune[1])
eNetModel_K562_25kb_SMOTE_Distance_lambda <- as.numeric(eNetModel_K562_25kb_SMOTE_Distance$bestTune[2])
```

### 50kb

#### none

##### OC

```{r}
eNetModel_K562_50kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OC/ENET/eNetModel.rds")
eNetModel_K562_50kb_none_oc_alpha <- as.numeric(eNetModel_K562_50kb_none_oc$bestTune[1])
eNetModel_K562_50kb_none_oc_lambda <- as.numeric(eNetModel_K562_50kb_none_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_50kb_none_OP <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/OP/ENET/eNetModel.rds")
eNetModel_K562_50kb_none_OP_alpha <- as.numeric(eNetModel_K562_50kb_none_OP$bestTune[1])
eNetModel_K562_50kb_none_OP_lambda <- as.numeric(eNetModel_K562_50kb_none_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_50kb_none_Distance <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/None/Distance/ENET/eNetModel.rds")
eNetModel_K562_50kb_none_Distance_alpha <- as.numeric(eNetModel_K562_50kb_none_Distance$bestTune[1])
eNetModel_K562_50kb_none_Distance_lambda <- as.numeric(eNetModel_K562_50kb_none_Distance$bestTune[2])
```

#### ROS

##### OC

```{r}
eNetModel_K562_50kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/OC/ENET/eNetModel.rds")
eNetModel_K562_50kb_ROS_oc_alpha <- as.numeric(eNetModel_K562_50kb_ROS_oc$bestTune[1])
eNetModel_K562_50kb_ROS_oc_lambda <- as.numeric(eNetModel_K562_50kb_ROS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_50kb_ROS_OP <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/OP/ENET/eNetModel.rds")
eNetModel_K562_50kb_ROS_OP_alpha <- as.numeric(eNetModel_K562_50kb_ROS_OP$bestTune[1])
eNetModel_K562_50kb_ROS_OP_lambda <- as.numeric(eNetModel_K562_50kb_ROS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_50kb_ROS_Distance <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/ROS/Distance/ENET/eNetModel.rds")
eNetModel_K562_50kb_ROS_Distance_alpha <- as.numeric(eNetModel_K562_50kb_ROS_Distance$bestTune[1])
eNetModel_K562_50kb_ROS_Distance_lambda <- as.numeric(eNetModel_K562_50kb_ROS_Distance$bestTune[2])
```

#### RUS

##### OC

```{r}
eNetModel_K562_50kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/OC/ENET/eNetModel.rds")
eNetModel_K562_50kb_RUS_oc_alpha <- as.numeric(eNetModel_K562_50kb_RUS_oc$bestTune[1])
eNetModel_K562_50kb_RUS_oc_lambda <- as.numeric(eNetModel_K562_50kb_RUS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_50kb_RUS_OP <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/OP/ENET/eNetModel.rds")
eNetModel_K562_50kb_RUS_OP_alpha <- as.numeric(eNetModel_K562_50kb_RUS_OP$bestTune[1])
eNetModel_K562_50kb_RUS_OP_lambda <- as.numeric(eNetModel_K562_50kb_RUS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_50kb_RUS_Distance <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/RUS/Distance/ENET/eNetModel.rds")
eNetModel_K562_50kb_RUS_Distance_alpha <- as.numeric(eNetModel_K562_50kb_RUS_Distance$bestTune[1])
eNetModel_K562_50kb_RUS_Distance_lambda <- as.numeric(eNetModel_K562_50kb_RUS_Distance$bestTune[2])
```

#### SMOTE

##### OC

```{r}
eNetModel_K562_50kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/OC/ENET/eNetModel.rds")
eNetModel_K562_50kb_SMOTE_oc_alpha <- as.numeric(eNetModel_K562_50kb_SMOTE_oc$bestTune[1])
eNetModel_K562_50kb_SMOTE_oc_lambda <- as.numeric(eNetModel_K562_50kb_SMOTE_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_50kb_SMOTE_OP <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/OP/ENET/eNetModel.rds")
eNetModel_K562_50kb_SMOTE_OP_alpha <- as.numeric(eNetModel_K562_50kb_SMOTE_OP$bestTune[1])
eNetModel_K562_50kb_SMOTE_OP_lambda <- as.numeric(eNetModel_K562_50kb_SMOTE_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_50kb_SMOTE_Distance <- readRDS("Z:/TAD_data_analysis/K562/50kb/Ensemble/SMOTE/Distance/ENET/eNetModel.rds")
eNetModel_K562_50kb_SMOTE_Distance_alpha <- as.numeric(eNetModel_K562_50kb_SMOTE_Distance$bestTune[1])
eNetModel_K562_50kb_SMOTE_Distance_lambda <- as.numeric(eNetModel_K562_50kb_SMOTE_Distance$bestTune[2])
```

### 100kb

#### none

##### OC

```{r}
eNetModel_K562_100kb_none_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OC/ENET/eNetModel.rds")
eNetModel_K562_100kb_none_oc_alpha <- as.numeric(eNetModel_K562_100kb_none_oc$bestTune[1])
eNetModel_K562_100kb_none_oc_lambda <- as.numeric(eNetModel_K562_100kb_none_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_100kb_none_OP <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/OP/ENET/eNetModel.rds")
eNetModel_K562_100kb_none_OP_alpha <- as.numeric(eNetModel_K562_100kb_none_OP$bestTune[1])
eNetModel_K562_100kb_none_OP_lambda <- as.numeric(eNetModel_K562_100kb_none_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_100kb_none_Distance <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/None/Distance/ENET/eNetModel.rds")
eNetModel_K562_100kb_none_Distance_alpha <- as.numeric(eNetModel_K562_100kb_none_Distance$bestTune[1])
eNetModel_K562_100kb_none_Distance_lambda <- as.numeric(eNetModel_K562_100kb_none_Distance$bestTune[2])
```

#### ROS

##### OC

```{r}
eNetModel_K562_100kb_ROS_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/OC/ENET/eNetModel.rds")
eNetModel_K562_100kb_ROS_oc_alpha <- as.numeric(eNetModel_K562_100kb_ROS_oc$bestTune[1])
eNetModel_K562_100kb_ROS_oc_lambda <- as.numeric(eNetModel_K562_100kb_ROS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_100kb_ROS_OP <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/OP/ENET/eNetModel.rds")
eNetModel_K562_100kb_ROS_OP_alpha <- as.numeric(eNetModel_K562_100kb_ROS_OP$bestTune[1])
eNetModel_K562_100kb_ROS_OP_lambda <- as.numeric(eNetModel_K562_100kb_ROS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_100kb_ROS_Distance <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/ROS/Distance/ENET/eNetModel.rds")
eNetModel_K562_100kb_ROS_Distance_alpha <- as.numeric(eNetModel_K562_100kb_ROS_Distance$bestTune[1])
eNetModel_K562_100kb_ROS_Distance_lambda <- as.numeric(eNetModel_K562_100kb_ROS_Distance$bestTune[2])
```

#### RUS

##### OC

```{r}
eNetModel_K562_100kb_RUS_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/OC/ENET/eNetModel.rds")
eNetModel_K562_100kb_RUS_oc_alpha <- as.numeric(eNetModel_K562_100kb_RUS_oc$bestTune[1])
eNetModel_K562_100kb_RUS_oc_lambda <- as.numeric(eNetModel_K562_100kb_RUS_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_100kb_RUS_OP <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/OP/ENET/eNetModel.rds")
eNetModel_K562_100kb_RUS_OP_alpha <- as.numeric(eNetModel_K562_100kb_RUS_OP$bestTune[1])
eNetModel_K562_100kb_RUS_OP_lambda <- as.numeric(eNetModel_K562_100kb_RUS_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_100kb_RUS_Distance <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/RUS/Distance/ENET/eNetModel.rds")
eNetModel_K562_100kb_RUS_Distance_alpha <- as.numeric(eNetModel_K562_100kb_RUS_Distance$bestTune[1])
eNetModel_K562_100kb_RUS_Distance_lambda <- as.numeric(eNetModel_K562_100kb_RUS_Distance$bestTune[2])
```

#### SMOTE

##### OC

```{r}
eNetModel_K562_100kb_SMOTE_oc <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/OC/ENET/eNetModel.rds")
eNetModel_K562_100kb_SMOTE_oc_alpha <- as.numeric(eNetModel_K562_100kb_SMOTE_oc$bestTune[1])
eNetModel_K562_100kb_SMOTE_oc_lambda <- as.numeric(eNetModel_K562_100kb_SMOTE_oc$bestTune[2])
```

##### OP

```{r}
eNetModel_K562_100kb_SMOTE_OP <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/OP/ENET/eNetModel.rds")
eNetModel_K562_100kb_SMOTE_OP_alpha <- as.numeric(eNetModel_K562_100kb_SMOTE_OP$bestTune[1])
eNetModel_K562_100kb_SMOTE_OP_lambda <- as.numeric(eNetModel_K562_100kb_SMOTE_OP$bestTune[2])
```

##### Distance

```{r}
eNetModel_K562_100kb_SMOTE_Distance <- readRDS("Z:/TAD_data_analysis/K562/100kb/Ensemble/SMOTE/Distance/ENET/eNetModel.rds")
eNetModel_K562_100kb_SMOTE_Distance_alpha <- as.numeric(eNetModel_K562_100kb_SMOTE_Distance$bestTune[1])
eNetModel_K562_100kb_SMOTE_Distance_lambda <- as.numeric(eNetModel_K562_100kb_SMOTE_Distance$bestTune[2])
```

## Creating Table

```{r}
eNetModel_K562_parameter_tab <- data.frame(Alpha=c(eNetModel_K562_10kb_none_oc_alpha,
                                                      eNetModel_K562_10kb_none_OP_alpha,
                                                      eNetModel_K562_10kb_none_Distance_alpha,

                                                      eNetModel_K562_10kb_ROS_oc_alpha,
                                                      eNetModel_K562_10kb_ROS_OP_alpha,
                                                      eNetModel_K562_10kb_ROS_Distance_alpha,

                                                      eNetModel_K562_10kb_RUS_oc_alpha,
                                                      eNetModel_K562_10kb_RUS_OP_alpha,
                                                      eNetModel_K562_10kb_RUS_Distance_alpha,

                                                      eNetModel_K562_10kb_SMOTE_oc_alpha,
                                                      eNetModel_K562_10kb_SMOTE_OP_alpha,
                                                      eNetModel_K562_10kb_SMOTE_Distance_alpha,

                                                      eNetModel_K562_25kb_none_oc_alpha,
                                                      eNetModel_K562_25kb_none_OP_alpha,
                                                      eNetModel_K562_25kb_none_Distance_alpha,

                                                      eNetModel_K562_25kb_ROS_oc_alpha,
                                                      eNetModel_K562_25kb_ROS_OP_alpha,
                                                      eNetModel_K562_25kb_ROS_Distance_alpha,

                                                      eNetModel_K562_25kb_RUS_oc_alpha,
                                                      eNetModel_K562_25kb_RUS_OP_alpha,
                                                      eNetModel_K562_25kb_RUS_Distance_alpha,

                                                      eNetModel_K562_25kb_SMOTE_oc_alpha,
                                                      eNetModel_K562_25kb_SMOTE_OP_alpha,
                                                      eNetModel_K562_25kb_SMOTE_Distance_alpha,

                                                      eNetModel_K562_50kb_none_oc_alpha,
                                                      eNetModel_K562_50kb_none_OP_alpha,
                                                      eNetModel_K562_50kb_none_Distance_alpha,

                                                      eNetModel_K562_50kb_ROS_oc_alpha,
                                                      eNetModel_K562_50kb_ROS_OP_alpha,
                                                      eNetModel_K562_50kb_ROS_Distance_alpha,

                                                      eNetModel_K562_50kb_RUS_oc_alpha,
                                                      eNetModel_K562_50kb_RUS_OP_alpha,
                                                      eNetModel_K562_50kb_RUS_Distance_alpha,

                                                      eNetModel_K562_50kb_SMOTE_oc_alpha,
                                                      eNetModel_K562_50kb_SMOTE_OP_alpha,
                                                      eNetModel_K562_50kb_SMOTE_Distance_alpha,

                                                      eNetModel_K562_100kb_none_oc_alpha,
                                                      eNetModel_K562_100kb_none_OP_alpha,
                                                      eNetModel_K562_100kb_none_Distance_alpha,

                                                      eNetModel_K562_100kb_ROS_oc_alpha,
                                                      eNetModel_K562_100kb_ROS_OP_alpha,
                                                      eNetModel_K562_100kb_ROS_Distance_alpha,

                                                      eNetModel_K562_100kb_RUS_oc_alpha,
                                                      eNetModel_K562_100kb_RUS_OP_alpha,
                                                      eNetModel_K562_100kb_RUS_Distance_alpha,

                                                      eNetModel_K562_100kb_SMOTE_oc_alpha,
                                                      eNetModel_K562_100kb_SMOTE_OP_alpha,
                                                      eNetModel_K562_100kb_SMOTE_Distance_alpha),
                                              Lambda=c(eNetModel_K562_10kb_none_oc_lambda,
                                                      eNetModel_K562_10kb_none_OP_lambda,
                                                      eNetModel_K562_10kb_none_Distance_lambda,

                                                      eNetModel_K562_10kb_ROS_oc_lambda,
                                                      eNetModel_K562_10kb_ROS_OP_lambda,
                                                      eNetModel_K562_10kb_ROS_Distance_lambda,

                                                      eNetModel_K562_10kb_RUS_oc_lambda,
                                                      eNetModel_K562_10kb_RUS_OP_lambda,
                                                      eNetModel_K562_10kb_RUS_Distance_lambda,

                                                      eNetModel_K562_10kb_SMOTE_oc_lambda,
                                                      eNetModel_K562_10kb_SMOTE_OP_lambda,
                                                      eNetModel_K562_10kb_SMOTE_Distance_lambda,

                                                      eNetModel_K562_25kb_none_oc_lambda,
                                                      eNetModel_K562_25kb_none_OP_lambda,
                                                      eNetModel_K562_25kb_none_Distance_lambda,

                                                      eNetModel_K562_25kb_ROS_oc_lambda,
                                                      eNetModel_K562_25kb_ROS_OP_lambda,
                                                      eNetModel_K562_25kb_ROS_Distance_lambda,

                                                      eNetModel_K562_25kb_RUS_oc_lambda,
                                                      eNetModel_K562_25kb_RUS_OP_lambda,
                                                      eNetModel_K562_25kb_RUS_Distance_lambda,

                                                      eNetModel_K562_25kb_SMOTE_oc_lambda,
                                                      eNetModel_K562_25kb_SMOTE_OP_lambda,
                                                      eNetModel_K562_25kb_SMOTE_Distance_lambda,

                                                      eNetModel_K562_50kb_none_oc_lambda,
                                                      eNetModel_K562_50kb_none_OP_lambda,
                                                      eNetModel_K562_50kb_none_Distance_lambda,

                                                      eNetModel_K562_50kb_ROS_oc_lambda,
                                                      eNetModel_K562_50kb_ROS_OP_lambda,
                                                      eNetModel_K562_50kb_ROS_Distance_lambda,

                                                      eNetModel_K562_50kb_RUS_oc_lambda,
                                                      eNetModel_K562_50kb_RUS_OP_lambda,
                                                      eNetModel_K562_50kb_RUS_Distance_lambda,

                                                      eNetModel_K562_50kb_SMOTE_oc_lambda,
                                                      eNetModel_K562_50kb_SMOTE_OP_lambda,
                                                      eNetModel_K562_50kb_SMOTE_Distance_lambda,

                                                      eNetModel_K562_100kb_none_oc_lambda,
                                                      eNetModel_K562_100kb_none_OP_lambda,
                                                      eNetModel_K562_100kb_none_Distance_lambda,

                                                      eNetModel_K562_100kb_ROS_oc_lambda,
                                                      eNetModel_K562_100kb_ROS_OP_lambda,
                                                      eNetModel_K562_100kb_ROS_Distance_lambda,

                                                      eNetModel_K562_100kb_RUS_oc_lambda,
                                                      eNetModel_K562_100kb_RUS_OP_lambda,
                                                      eNetModel_K562_100kb_RUS_Distance_lambda,

                                                      eNetModel_K562_100kb_SMOTE_oc_lambda,
                                                      eNetModel_K562_100kb_SMOTE_OP_lambda,
                                                      eNetModel_K562_100kb_SMOTE_Distance_lambda))

eNetModel_K562_parameter_tab$CL <- "K562"
eNetModel_K562_parameter_tab$Resolution <- c(rep("10 kb", 12),
                                                rep("25 kb", 12),
                                                rep("50 kb", 12),
                                                rep("100 kb", 12))
eNetModel_K562_parameter_tab$PredictorType <- rep(c("OC","OP","Distance"),16)
eNetModel_K562_parameter_tab <- eNetModel_K562_parameter_tab[,c(3:5,1:2)]
```

#All cell lines

```{r}
eNetModel_parameters_table <- rbind.data.frame(eNetModel_gm12878_parameter_tab,
                                               eNetModel_K562_parameter_tab)

write.csv(eNetModel_parameters_table, "C:/Users/stili/Downloads/eNetModel_parameters_table.csv")
```