### 2019-04-22 Scripts/comparing_distance_profiles_using_violin_and_shift_plots.Rmd
added script for evaluating distance profiles using violin and shift plots

---
title: "Comparing Distance Profiles Using Violin Plots and Shift Function"
author: "Spiro Stilianoudakis"
date: "April 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(coin)
library(cowplot)
```

# Functions for shift function

```{r}
setwd("C:/Users/stili/Documents/TAD_miscellaneous/shift_function")

source("Rallfun-v30.txt")
source("wilcox_modified.txt")
source("rgar_visualisation.txt")
source("rgar_utils.txt")

```

# GM12878

## 10kb

```{r}
train <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/training_testing/train.rds"))

enetcoefs <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/enetcoefs.rds"))

names(train) <- gsub(paste(c("`", 
                             "Gm12878", 
                             "-", 
                             #"_", 
                             #"dist", "count", "perc",
                             "Haib", 
                             "Sydh",
                             "Uta",
                             "Uw",
                             "Broad",
                             "Uchicago"), collapse = "|"), "", names(train))
train <- train[,c(1,which(names(train) %in% enetcoefs))]

set.seed(123)
train_smote <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

```

### Top Annotations

####SMC

```{r}
annotname <- gsub(paste(c("`", 
             "Gm12878", 
             "-", 
             #"_", 
             #"dist", "count", "perc",
             "Haib", 
             "Sydh",
             "Uta",
             "Uw",
             "Broad",
             "Uchicago"), collapse = "|"), "", "Gm12878-Smc3ab9263-Sydh_dist")

annotdata.smc <- train_smote[, c(1, which(names(train_smote)==annotname))]

#ks.test(annotdata.smc[which(annotdata.smc$y=="No"),2], annotdata.smc[which(annotdata.smc$y=="Yes"),2], alternative = "two.sided")

perm.pval <- pvalue(oneway_test(annotdata.smc[,2] ~ annotdata.smc[,1],
                     data=annotdata.smc,
                     alternative="greater", 
                     distribution=approximate(B=10000)))[1]

perm.pval <- ifelse(perm.pval==0,"< 0.0001", as.character(round(perm.pval,4)))

plottitle <- paste(gsub(paste(c("Gm12878",
                                  "-","_",
                                  "dist","perc","count",
                                  "Haib",
                                  "Sydh",
                                  "Uta",
                                  "Uw",
                                  "Broad",
                                  "Uchicago"), 
                                collapse = "|"), 
                          "",
                          "Gm12878-Smc3ab9263-Sydh_dist"))

violin.gm12878.10kb.smc <- ggplot(annotdata.smc, aes(x=ordered(y, levels=rev(unique(y))), y=Smc3ab9263_dist, color=y)) + 
  geom_violin(trim=FALSE)+
  geom_boxplot(width=0.1, fill="white")+
  coord_flip() +
  scale_color_discrete(name="Class of\nGenomic Bin",
                         labels=c("Did not overlap\nw/ flanked TAD boundary", "Overlapped w/\nflanked TAD boundary"))+
  xlab("Class of Genomic Bin") +
  ylab("Log2 Distance") +
  ggtitle("SMC3") +
  #annotate("text", x=7, y=.2, label= paste("P-Value:", perm.pval, sep = " "), size=5) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15),
    axis.text.y = element_blank(),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    plot.title = element_text(size=20,
                              hjust=.5))

out <- shifthd(annotdata.smc$Smc3ab9263_dist[which(annotdata.smc$y=="No")],
               annotdata.smc$Smc3ab9263_dist[which(annotdata.smc$y=="Yes")],
               nboot=200)

ylim <- max(max(abs(out$ci_upper)),max(abs(out$ci_lower)))
ylim <- c(-ylim,ylim)
xintercept <- as.numeric(out[5,1]) # get median of group 1
#xbreaks <- out[,1] # get deciles of group 1
xplot = names(out)[1]

shift.gm12878.10kb.smc <- ggplot(out, aes_string(x=xplot, y="difference")) +
  geom_hline(yintercept=0,linetype=2,alpha=0.5) + # x=0 reference line
  geom_vline(xintercept=xintercept,linetype=2,alpha=0.5) + # y=median reference line
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper), colour="black", width=.15) +
  geom_line(colour="black", linetype="solid", size=1.5) +
  geom_point(colour="black", size=4, shape=21, fill="white") + #999999
  xlab("Quantiles for bins that\nOverlapped w/flanked TAD boundaries") +
  ylab("Quantile differences between bin class") +
  theme_bw() +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14),
        axis.title.x = element_text(size=16,face="bold"),
        axis.title.y = element_text(size=16,face="bold")) +
  scale_y_continuous(limits = ylim)

ggarrange(violin.gm12878.10kb.smc,
          shift.gm12878.10kb.smc,
          ncol=1,
          nrow=2,
          common.legend = TRUE,
          legend = "top")
```