### 2019-05-12 Manuscript/Tables/table_legends.Rmd
consolidated figure 4 into a table with bin and ir summaries; updated legends

---
title: "Legends for Tables"
author: "Spiro Stilianoudakis"
date: "November 19, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Table 1. Data Summaries.** The average number of genomic bins and corresponding imbalance ratios, across each of the autosomal chromosomes (omitting chromosome 9), stratified by resolution and cell line. The resolution of the Hi-C contact matrix dictated the width of each genomic bin, with lower resolutions corresponding to wider genomic bins.