### 2019-05-12 Manuscript/Tables/table_legends.Rmd
added number of tads called to table1

---
title: "Legends for Tables"
author: "Spiro Stilianoudakis"
date: "November 19, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Table 1. Data Summaries.** The average number of genomic bins, TADs called, and the corresponding boundaries and imbalance ratios, across each of the autosomal chromosomes (omitting chromosome 9), stratified by resolution and cell line. The resolution of the Hi-C contact matrix dictated the width of each genomic bin, with lower resolutions corresponding to wider genomic bins.