### 2019-05-12 Scripts/evaluating_variable_importances.Rmd
editted results script

---
title: "Evaluating Variable Importances"
author: "Spiro Stilianoudakis"
date: "March 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# Variable Importance Heatmaps for models with distance type predictors with SMOTE resampling

##GM12878

```{r}
annots <- list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878")
annots <- gsub(paste(c("Gm12878-", 
                             "-Haib.bed", 
                             "-Sydh.bed",
                             "-Uta.bed",
                             "-Uw.bed",
                             "-Broad.bed",
                             "-Uchicago.bed"), collapse = "|"), "", annots)
annots <- annots[!duplicated(annots)]
```

###10kb

```{r}
varimp_gm12878_10kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_gm12878_10kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars_tfbs.rds"))
	  varimps$Feature <- gsub("_dist", "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_gm12878_10kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_gm12878_10kb <- data.frame(Feature = rownames(varimp_gm12878_10kb),
                                  aveimp10kb = rowMeans(varimp_gm12878_10kb, na.rm = TRUE),
                                  sdimp10kb = apply(varimp_gm12878_10kb,1, sd, na.rm = TRUE))
rownames(varimp_gm12878_10kb) <- NULL
```

###25kb

```{r}
varimp_gm12878_25kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_gm12878_25kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars_tfbs.rds"))
	  varimps$Feature <- gsub("_dist", "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_gm12878_25kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_gm12878_25kb <- data.frame(Feature = rownames(varimp_gm12878_25kb),
                                  aveimp25kb = rowMeans(varimp_gm12878_25kb, na.rm = TRUE),
                                  sdimp25kb = apply(varimp_gm12878_25kb,1, sd, na.rm = TRUE))
rownames(varimp_gm12878_25kb) <- NULL
```

###50kb

```{r}
varimp_gm12878_50kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_gm12878_50kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars_tfbs.rds"))
	  varimps$Feature <- gsub("_dist", "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_gm12878_50kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_gm12878_50kb <- data.frame(Feature = rownames(varimp_gm12878_50kb),
                                  aveimp50kb = rowMeans(varimp_gm12878_50kb, na.rm = TRUE),
                                  sdimp50kb = apply(varimp_gm12878_50kb,1, sd, na.rm = TRUE))
rownames(varimp_gm12878_50kb) <- NULL
```

###100kb

```{r}
varimp_gm12878_100kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_gm12878_100kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars_tfbs.rds"))
	  varimps$Feature <- gsub("_dist", "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_gm12878_100kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_gm12878_100kb <- data.frame(Feature = rownames(varimp_gm12878_100kb),
                                  aveimp100kb = rowMeans(varimp_gm12878_100kb, na.rm = TRUE),
                                  sdimp100kb = apply(varimp_gm12878_100kb,1, sd, na.rm = TRUE))
rownames(varimp_gm12878_100kb) <- NULL
```

##K562

```{r}
annots <- list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/K562")
annots <- gsub(paste(c("K562-", 
                             "-Haib.bed", 
                             "-Sydh.bed",
                             "-Uta.bed",
                             "-Uw.bed",
                             "-Broad.bed",
                             "-Uchicago.bed"), collapse = "|"), "", annots)
annots <- annots[!duplicated(annots)]
```

###10kb

```{r}
varimp_K562_10kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_K562_10kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars_tfbs.rds"))
	  varimps$Feature <- gsub("_dist", "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_K562_10kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_K562_10kb <- data.frame(Feature = rownames(varimp_K562_10kb),
                                  aveimp10kb = rowMeans(varimp_K562_10kb, na.rm = TRUE),
                                  sdimp10kb = apply(varimp_K562_10kb,1, sd, na.rm = TRUE))
rownames(varimp_K562_10kb) <- NULL
```

###25kb

```{r}
varimp_K562_25kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_K562_25kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars_tfbs.rds"))
	  varimps$Feature <- gsub("_dist", "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_K562_25kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_K562_25kb <- data.frame(Feature = rownames(varimp_K562_25kb),
                                  aveimp25kb = rowMeans(varimp_K562_25kb, na.rm = TRUE),
                                  sdimp25kb = apply(varimp_K562_25kb,1, sd, na.rm = TRUE))
rownames(varimp_K562_25kb) <- NULL
```

###50kb

```{r}
varimp_K562_50kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_K562_50kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars_tfbs.rds"))
	  varimps$Feature <- gsub("_dist", "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_K562_50kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_K562_50kb <- data.frame(Feature = rownames(varimp_K562_50kb),
                                  aveimp50kb = rowMeans(varimp_K562_50kb, na.rm = TRUE),
                                  sdimp50kb = apply(varimp_K562_50kb,1, sd, na.rm = TRUE))
rownames(varimp_K562_50kb) <- NULL
```

###100kb

```{r}
varimp_K562_100kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_K562_100kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars_tfbs.rds"))
	  varimps$Feature <- gsub("_dist", "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_K562_100kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_K562_100kb <- data.frame(Feature = rownames(varimp_K562_100kb),
                                  aveimp100kb = rowMeans(varimp_K562_100kb, na.rm = TRUE),
                                  sdimp100kb = apply(varimp_K562_100kb,1, sd, na.rm = TRUE))
rownames(varimp_K562_100kb) <- NULL
```

##Plots

```{r}
#GM12878
all_varimps_gm12878 <- cbind.data.frame(varimp_gm12878_10kb$Feature,
                                        varimp_gm12878_10kb$aveimp10kb,
                                        varimp_gm12878_25kb$aveimp25kb,
                                        varimp_gm12878_50kb$aveimp50kb,
                                        varimp_gm12878_100kb$aveimp100kb)
names(all_varimps_gm12878) <- c("Feature", "ave10kb", "ave25kb", "ave50kb", "ave100kb")
all_varimps_gm12878 <- all_varimps_gm12878[order(all_varimps_gm12878$ave10kb, decreasing = TRUE),]
all_varimps_gm12878 <- all_varimps_gm12878[1:20,]
all_varimps_gm12878 <- melt(all_varimps_gm12878)

ggplot(all_varimps_gm12878, aes(variable, ordered(Feature, levels=rev(unique(Feature))))) + 
  geom_tile(aes(fill = value), colour = "white") + 
  xlab("Resolution") +
  ylab("Transcription Factor") +
  scale_fill_gradient2(low = "white",
                      mid = "pink",
                      midpoint = 50,
                      high = muted("red"),
                      na.value = "grey50",
                      name="Predictive\nImportance") +
  scale_x_discrete(labels=c("10 kb", "25 kb", "50 kb", "100 kb"))+
  theme_bw() +
  theme(axis.text.x = element_text(size=15,
                                   angle = 45,
                                   hjust = 1),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

#K562
all_varimps_K562 <- cbind.data.frame(varimp_K562_10kb$Feature,
                                        varimp_K562_10kb$aveimp10kb,
                                        varimp_K562_25kb$aveimp25kb,
                                        varimp_K562_50kb$aveimp50kb,
                                        varimp_K562_100kb$aveimp100kb)
names(all_varimps_K562) <- c("Feature", "ave10kb", "ave25kb", "ave50kb", "ave100kb")
all_varimps_K562 <- all_varimps_K562[order(all_varimps_K562$ave10kb, decreasing = TRUE),]
all_varimps_K562 <- all_varimps_K562[1:20,]
all_varimps_K562 <- melt(all_varimps_K562)

ggplot(all_varimps_K562, aes(variable, ordered(Feature, levels=rev(unique(Feature))))) + 
  geom_tile(aes(fill = value), colour = "white") + 
  xlab("Resolution") +
  ylab("Transcription Factor") +
  scale_fill_gradient2(low = "white",
                      mid = "pink",
                      midpoint = 50,
                      high = muted("red"),
                      na.value = "grey50",
                      name="Predictive\nImportance") +
  scale_x_discrete(labels=c("10 kb", "25 kb", "50 kb", "100 kb"))+
  theme_bw() +
  theme(axis.text.x = element_text(size=15,
                                   angle = 45,
                                   hjust = 1),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))
```