### 2019-05-14 Manuscript/Tables/table_legends.Rmd
added tables; updated legends

---
title: "Legends for Tables"
author: "Spiro Stilianoudakis"
date: "November 19, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Table 1. Data Summaries and Imbalance Ratios.** For GM12878, average number of genomic bins, TADs called, and the corresponding boundaries and imbalance ratios, across each of the autosomal chromosomes (omitting chromosome 9), stratified by resolution. The resolution of the Hi-C contact matrix dictated the width of each genomic bin, with lower resolutions corresponding to wider genomic bins.

**Table 2. Inter-Resolution ICCs.** For GM12878, two-way intraclass correlation coefficients between resolutions for the following quantities: Number of genomic bins, number of TADs called, number of unique TAD boundaries, and imbalance ratio. The ICC was computed across 21 of the autosomal chromosomes. All quatities, except for imbalance ratio, was correlated with data resolution. Instead, imbalance ratios remained consistent and low across resolutions.

**Table 3. Jaccard Indices.** For GM12878, the median (range) Jaccard index illustrating the pairwise similarity between annotations within specific groups, including chromatin segmentation states, histone modifications, and transcription factor binding sites. The Jaccard Index measures the ratio of the number of intersecting base pairs between any two sets, to the number of base pairs in the union of the two sets. The number of cell line specific annotations included within each group is also provided.