### 2019-05-14 Scripts/calculating_jaccard_index_between_annotations.Rmd
editted results scripts

---
title: "Evaluating Jaccard Index Between Annotations"
author: "Spiro Stilianoudakis"
date: "April 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(cotools)
library(HelloRanges)
```

# Function for calculating jaccard index between any two annotations

```{r}
#jaccard(A,B) = length of intersection between A and B (in base pairs)/(length of union between A and B - length of intersection between A and B)

jaccard_func <- function(A, B){
  #compute length of intersection between both annotations in base pairs (numerator)
  l.i <- sum(width(pintersect(findOverlapPairs(A, B))))

  #compute length of union between both annotations in base pairs
  l.u <- sum(width(A)) + sum(width(B))

  j <- l.i/(l.u-l.i)

  return(j)
}

```

# Iterate through all annotation and calculate paiwise jaccard index

## GM12878

```{r}
setwd("Z:/TAD_data_analysis/annotations/all_annotations/gm12878")
temp = list.files()

jaccard.mat <- matrix(nrow=length(temp), ncol=length(temp))
colnames(jaccard.mat) <- rownames(jaccard.mat) <- gsub(paste(c("Gm12878", "K562", ".bed",
                          "-",
                          "_",
                          "Haib",
                          "Sydh",
                          "Uta",
                          "Uw",
                          "Broad",
                          "Uchicago"), collapse = "|"),"", temp)

for(i in 27:length(temp)){
  for(j in 1:length(temp)){
    print(c(i,j))
    annotA <- read.table(temp[i],header = FALSE, sep="\t")
    annotA.gr <- GRanges(seqnames=annotA$V1,
                     IRanges(start=annotA$V2, 
                             end=annotA$V3))

    annotB <- read.table(temp[j],header = FALSE, sep="\t")
    annotB.gr <- GRanges(seqnames=annotB$V1,
                     IRanges(start=annotB$V2, 
                             end=annotB$V3))

    if(i==j){jaccard.mat[i,j] <- 1}else{
      code <- R_bedtools_jaccard(annotA.gr,annotB.gr)
      eval.code <- eval(code)
      jaccard.mat[i,j] <- eval.code@listData$jaccard
    }
  }
}
```

###Plots

```{r}
jaccard.mat.gm12878 <- readRDS("Z:/TAD_data_analysis/jaccard/jaccard.mat.gm12878.rds")
jaccard.mat.gm12878 <- jaccard.mat.gm12878[!duplicated(rownames(jaccard.mat.gm12878)), !duplicated(colnames(jaccard.mat.gm12878))]

```

####Chomatin States

```{r}
css <- list.files("Z:/TAD_data_analysis/annotations/BroadHMM")
css <- css[grep("Gm12878", css)]
css <- gsub(paste(c("Gm12878-", ".bed"), collapse = "|"), "", css)
jaccard.mat.gm12878.css <- jaccard.mat.gm12878[rownames(jaccard.mat.gm12878) %in% css, colnames(jaccard.mat.gm12878) %in% css]

corrplot(jaccard.mat.gm12878.css, method = "square",
         type="upper", 
         order="hclust", 
         tl.col="black", 
         tl.srt=90,
         #tl.cex = .4,
         col=colorRampPalette(c("white", "red"))(100),
         cl.pos="n",
         mar = c(0, 0, 0, 0))

get_upper_tri <- function(cormat){
     cormat[lower.tri(cormat)]<- NA
     return(cormat)
}

upper_tri_css <- get_upper_tri(jaccard.mat.gm12878.css)
melted_css <- reshape2::melt(upper_tri_css, na.rm = TRUE)
melted_css <- melted_css[-which(melted_css$Var1==melted_css$Var2),]
summary(melted_css$value)
```

####Histone Modifications

```{r}
hm <- list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")
hm <- hm[grep("Gm12878", hm)]
hm <- gsub(paste(c("Gm12878-", ".bed"), collapse = "|"), "", hm)
jaccard.mat.gm12878.hm <- jaccard.mat.gm12878[rownames(jaccard.mat.gm12878) %in% hm, colnames(jaccard.mat.gm12878) %in% hm]

corrplot(jaccard.mat.gm12878.hm, method = "square",
         type="upper", 
         order="hclust", 
         tl.col="black", 
         tl.srt=90,
         #tl.cex = .4,
         col=colorRampPalette(c("white", "red"))(100),
         cl.pos="n",
         mar = c(0, 0, 0, 0))

get_upper_tri <- function(cormat){
     cormat[lower.tri(cormat)]<- NA
     return(cormat)
}

upper_tri_hm <- get_upper_tri(jaccard.mat.gm12878.hm)
melted_hm <- reshape2::melt(upper_tri_hm, na.rm = TRUE)
melted_hm <- melted_hm[-which(melted_hm$Var1==melted_hm$Var2),]
summary(melted_hm$value)
```

####TFBS

```{r}
tfbs <- list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878")
tfbs <- tfbs[grep("Gm12878", tfbs)]
tfbs <- gsub(paste(c("Gm12878-", ".bed","-Haib","-Sydh","-Uta","-Uw","-Broad","-Uchicago"), collapse = "|"), "", tfbs)
jaccard.mat.gm12878.tfbs <- jaccard.mat.gm12878[rownames(jaccard.mat.gm12878) %in% tfbs, colnames(jaccard.mat.gm12878) %in% tfbs]

corrplot(jaccard.mat.gm12878.tfbs, method = "square",
         type="upper", 
         order="hclust", 
         tl.col="black", 
         tl.srt=90,
         tl.cex = .7,
         col=colorRampPalette(c("white", "red"))(100),
         cl.pos="n",
         mar = c(0, 0, 0, 0))

get_upper_tri <- function(cormat){
     cormat[lower.tri(cormat)]<- NA
     return(cormat)
}

upper_tri_tfbs <- get_upper_tri(jaccard.mat.gm12878.tfbs)
melted_tfbs <- reshape2::melt(upper_tri_tfbs, na.rm = TRUE)
melted_tfbs <- melted_tfbs[-which(melted_tfbs$Var1==melted_tfbs$Var2),]
summary(melted_tfbs$value)
```

####All

```{r}
corrplot(jaccard.mat.gm12878, method = "square",
         type="upper", 
         order="hclust", 
         tl.col="black", 
         tl.srt=90,
         tl.cex = .4,
         col=colorRampPalette(c("white", "red"))(100),
         cl.pos="n",
         mar = c(0, 0, 0, 0))

all_melted_gm12878 <- rbind.data.frame(melted_css,
                                       melted_hm,
                                       melted_tfbs)
all_melted_gm12878$annotation <- c(rep("CSS", length(melted_css$Var1)),
                                   rep("HM", length(melted_hm$Var1)),
                                   rep("TFBS", length(melted_tfbs$Var1)))
all_melted_gm12878$annotation <- factor(all_melted_gm12878$annotation, levels=c("CSS", "HM", "TFBS"))

ggplot(all_melted_gm12878[-which(all_melted_gm12878$annotation=="CSS"),], aes(x=value, group=annotation, color=annotation)) + 
  geom_density(size=1) +
  xlim(0,1) +
  xlab("Jaccard Index") +
  ylab("Density") +
  scale_color_discrete(name="Annotation Group") +
  theme_minimal() +
  theme_bw() +
  theme(axis.text.x = element_text(size=15),
     axis.text.y = element_text(size=15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))
```

## K562

```{r}
setwd("Z:/TAD_data_analysis/annotations/all_annotations/k562")
temp = list.files()

jaccard.mat <- matrix(nrow=length(temp), ncol=length(temp))
colnames(jaccard.mat) <- rownames(jaccard.mat) <- gsub(paste(c("Gm12878", "K562", ".bed",
                          "-",
                          "_",
                          "Haib",
                          "Sydh",
                          "Uta",
                          "Uw",
                          "Broad",
                          "Uchicago"), collapse = "|"),"", temp)

for(i in 1:length(temp)){
  for(j in 1:length(temp)){
    print(c(i,j))
    annotA <- read.table(temp[i],header = FALSE, sep="\t")
    annotA.gr <- GRanges(seqnames=annotA$V1,
                     IRanges(start=annotA$V2, 
                             end=annotA$V3))

    annotB <- read.table(temp[j],header = FALSE, sep="\t")
    annotB.gr <- GRanges(seqnames=annotB$V1,
                     IRanges(start=annotB$V2, 
                             end=annotB$V3))

    if(i==j){jaccard.mat[i,j] <- 1}else{
      code <- R_bedtools_jaccard(annotA.gr,annotB.gr)
      eval.code <- eval(code)
      jaccard.mat[i,j] <- eval.code@listData$jaccard
    }
  }
}
```

###Plots

```{r}
jaccard.mat.k562 <- readRDS("Z:/TAD_data_analysis/jaccard/jaccard.mat.k562.rds")
jaccard.mat.k562 <- jaccard.mat.k562[!duplicated(rownames(jaccard.mat.k562)), !duplicated(colnames(jaccard.mat.k562))]

```

####Chomatin States

```{r}
css <- list.files("Z:/TAD_data_analysis/annotations/BroadHMM")
css <- css[grep("K562", css)]
css <- gsub(paste(c("K562-", ".bed"), collapse = "|"), "", css)
jaccard.mat.k562.css <- jaccard.mat.k562[rownames(jaccard.mat.k562) %in% css, colnames(jaccard.mat.k562) %in% css]

corrplot(jaccard.mat.k562.css, method = "square",
         type="upper", 
         order="hclust", 
         tl.col="black", 
         tl.srt=90,
         #tl.cex = .4,
         col=colorRampPalette(c("white", "red"))(100),
         cl.pos="n",
         mar = c(0, 0, 0, 0))

get_upper_tri <- function(cormat){
     cormat[lower.tri(cormat)]<- NA
     return(cormat)
}

upper_tri_css <- get_upper_tri(jaccard.mat.k562.css)
melted_css <- reshape2::melt(upper_tri_css, na.rm = TRUE)
melted_css <- melted_css[-which(melted_css$Var1==melted_css$Var2),]
summary(melted_css$value)
```

####Histone Modifications

```{r}
hm <- list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")
hm <- hm[grep("K562", hm)]
hm <- gsub(paste(c("K562-", ".bed"), collapse = "|"), "", hm)
jaccard.mat.k562.hm <- jaccard.mat.k562[rownames(jaccard.mat.k562) %in% hm, colnames(jaccard.mat.k562) %in% hm]

corrplot(jaccard.mat.k562.hm, method = "square",
         type="upper", 
         order="hclust", 
         tl.col="black", 
         tl.srt=90,
         #tl.cex = .4,
         col=colorRampPalette(c("white", "red"))(100),
         cl.pos="n",
         mar = c(0, 0, 0, 0))

get_upper_tri <- function(cormat){
     cormat[lower.tri(cormat)]<- NA
     return(cormat)
}

upper_tri_hm <- get_upper_tri(jaccard.mat.k562.hm)
melted_hm <- reshape2::melt(upper_tri_hm, na.rm = TRUE)
melted_hm <- melted_hm[-which(melted_hm$Var1==melted_hm$Var2),]
summary(melted_hm$value)
```

####TFBS

```{r}
tfbs <- list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/k562")
tfbs <- tfbs[grep("K562", tfbs)]
tfbs <- gsub(paste(c("K562-", ".bed","-Haib","-Sydh","-Uta","-Uw","-Broad","-Uchicago"), collapse = "|"), "", tfbs)
jaccard.mat.k562.tfbs <- jaccard.mat.k562[rownames(jaccard.mat.k562) %in% tfbs, colnames(jaccard.mat.k562) %in% tfbs]

corrplot(jaccard.mat.k562.tfbs, method = "square",
         type="upper", 
         order="hclust", 
         tl.col="black", 
         tl.srt=90,
         tl.cex = .6,
         col=colorRampPalette(c("white", "red"))(100),
         cl.pos="n",
         mar = c(0, 0, 0, 0))

get_upper_tri <- function(cormat){
     cormat[lower.tri(cormat)]<- NA
     return(cormat)
}

upper_tri_tfbs <- get_upper_tri(jaccard.mat.k562.tfbs)
melted_tfbs <- reshape2::melt(upper_tri_tfbs, na.rm = TRUE)
melted_tfbs <- melted_tfbs[-which(melted_tfbs$Var1==melted_tfbs$Var2),]
summary(melted_tfbs$value)
```

####All

```{r}
corrplot(jaccard.mat.k562, method = "square",
         type="upper", 
         order="hclust", 
         tl.col="black", 
         tl.srt=90,
         tl.cex = .4,
         col=colorRampPalette(c("white", "red"))(100),
         cl.pos="n",
         mar = c(0, 0, 0, 0))

all_melted_k562 <- rbind.data.frame(melted_css,
                                       melted_hm,
                                       melted_tfbs)
all_melted_k562$annotation <- c(rep("CSS", length(melted_css$Var1)),
                                   rep("HM", length(melted_hm$Var1)),
                                   rep("TFBS", length(melted_tfbs$Var1)))
all_melted_k562$annotation <- factor(all_melted_k562$annotation, levels=c("CSS", "HM", "TFBS"))

ggplot(all_melted_k562[-which(all_melted_k562$annotation=="CSS"),], aes(x=value, group=annotation, color=annotation)) + 
  geom_density(size=1) +
  xlim(0,1) +
  xlab("Jaccard Index") +
  ylab("Density") +
  scale_color_discrete(name="Annotation Group") +
  theme_minimal() +
  theme_bw() +
  theme(axis.text.x = element_text(size=15),
     axis.text.y = element_text(size=15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))
```