### 2019-05-16 Manuscript/Supplementary/supplementary_legends.Rmd
editted figure legends

---
title: "Legends for supplementary material"
output: pdf_document
---

**Supplementary Table 1. Hi-C data.** A complete list of the Hi-C data libraries used to obtain the locations of topologically associating domains.

**Supplementary Table 2. Genomic annotations.** A complete list of the genomic annotations used to build the feature space for all downstream models. The GRCh37/hg19 human genome assembly was used. "Genomic Class" - broad category of genomic features, "Element" - names of genomic features, "Available Cell Type" - the name of the cell type in which the annotation was available, "Cell Type-Specific Source" - download URL specific to the cell line (not all annotations were provided by the same institutions).

**Supplementary Figure 1. GM12878 distance densities.** (A-D) Distances (in millions of basepairs) from the center of each genomic bin to the nearest annotation peak, for binwidths (resolutions) of 10 kb, 25 kb, 50 kb, and 100 kb. (E-H) $Log_{2}$ transformed distances for binwidths (resolutions) of 10 kb, 25 kb, 50 kb, and 100 kb.

**Supplementary Figure 2. K562 distance densities.** (A-D) Distances (in millions of basepairs) from the center of each genomic bin to the nearest annotation peak, for binwidths (resolutions) of 10 kb, 25 kb, 50 kb, and 100 kb. (E-H) $Log_{2}$ transformed distances for binwidths (resolutions) of 10 kb, 25 kb, 50 kb, and 100 kb.

**Supplementary Table 3. Data Summaries and Imbalance Ratios.** For K562, average number of genomic bins, TADs called, and the corresponding boundaries and imbalance ratios, across each of the autosomal chromosomes (omitting chromosome 9), stratified by resolution. The resolution of the Hi-C contact matrix dictated the width of each genomic bin, with lower resolutions corresponding to wider genomic bins.

**Supplementary Table 4. Inter-Resolution ICCs.** For K562, two-way intraclass correlation coefficients between resolutions for the following quantities: Number of genomic bins, number of TADs called, number of unique TAD boundaries, and imbalance ratio. The ICC was computed across 21 of the autosomal chromosomes. All quatities, except for imbalance ratio, was correlated with data resolution. Instead, imbalance ratios remained consistent and low across resolutions.

**Table 5. Jaccard Indices.** For K562, the median (range) Jaccard index illustrating the pairwise similarity between annotations within specific groups, including chromatin segmentation states, histone modifications, and transcription factor binding sites. The Jaccard Index measures the ratio of the number of intersecting base pairs between any two sets, to the number of base pairs in the union of the two sets. The number of cell line specific annotations included within each group is also provided.

**Supplementary Table 6. Full list of aggregated performances for GM12878.** Performances were stratified by resolution, resampling technique, and predictor type, and then averaged across chromosomes.

**Supplementary Table 7. Full list of aggregated performances for K562.** Performances were stratified by resolution, resampling technique, and predictor type, and then averaged across chromosomes.

**Supplementary Figure 3. Comparing ensemble system specific model performances for K562.** Performances of random forest classification models built on each of the 3 predictor types: OC (A), OP (B), and Distance (C). Within each predictor type, and across resolution, MCCs metrics were recorded for models subsequently implemented on re-sampled data using the 4 re-sampling techniques: None (red), ROS (green), RUS (blue), and SMOTE (purple). Performances were aggregated using means across each chromosome, with error bars representing 1 standard deviation away from the mean.

**Supplementary Figure 4. Predictive importance is outweighed by transcription factor binding sites.** Stacked barplots illustrating the proportion of total predictive importance in models using optimal data level characteristics determined by the ensemble framework, for each resolution on K562. Proportions are separated by the set of annotations included in the feature space including chromatin state segmentation, histone modification, and transcription factor binding sites.

**Supplementary Figure 5. Comparing annotation-specific random forest models implemented on SMOTE re-sampled data with distance-type predictors for K562.** MCC performances were compared for models using specific groups of annotations as the feature space including: histone modifications (HM; red), chromatin state segmentations (CSS; green), transcription factor binding sites (TFBS; blue), and a feature space containing all 3 types combined (combined; purple). Each bar represents a model ran through the ensemble framework, with re-sampling technique and predictor-type characteristics fixed to SMOTE and distance-types. The error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosome.

**Supplementary Figure 6. Predictive importance** A heatmap illustrating the importance of the top 20 most predictive transcription factors for random forest models implemented on SMOTE resampled data with distance-type predictors for K562. Importances were sorted according to the 10 kb resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red.

<!--
**Supplementary Figure 12. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 25 kb resolution, on GM12878.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.
-->