### 2019-05-16 Scripts/proportion_of_importance_per_annotation.Rmd

---
title: "proportion_of_importance_per_annotation"
author: "Spiro Stilianoudakis"
date: "May 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

#GM12878

```{r}
annots <- list.files("Z:/TAD_data_analysis/annotations/all_annotations/gm12878")
annots <- gsub(paste(c("Gm12878-",".bed", 
                             "-Haib", 
                             "-Sydh",
                             "-Uta",
                             "-Uw",
                             "-Broad",
                             "-Uchicago"), collapse = "|"), "", annots)
annots <- annots[!duplicated(annots)]

chromstates <- list.files("Z:/TAD_data_analysis/annotations/BroadHMM")
chromstates <- chromstates[grep("Gm12878", chromstates)]
chromstates <- gsub(paste0(c("Gm12878-", ".bed"), collapse = "|"),"",chromstates)

histones <- list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")
histones <- histones[grep("Gm12878", histones)]
histones <- gsub(paste0(c("Gm12878-", ".bed"), collapse = "|"),"",histones)

tfbs <- list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878")
tfbs <- gsub(paste0(c("Gm12878-", ".bed", "-Haib", "-Sydh", "-Broad", "-Uw", "-Uta", "-Chicago"), collapse = "|"),"", tfbs)
tfbs <- tfbs[!duplicated(tfbs)]

```

##10kb

```{r}
varimp_gm12878_10kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_gm12878_10kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars.rds"))
	  varimps$Feature <- gsub(paste0(c("`","_dist"), collapse = "|"), "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_gm12878_10kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_gm12878_10kb <- data.frame(Feature = rownames(varimp_gm12878_10kb),
                                  aveimp10kb = rowMeans(varimp_gm12878_10kb, na.rm = TRUE),
                                  sdimp10kb = apply(varimp_gm12878_10kb,1, sd, na.rm = TRUE))
rownames(varimp_gm12878_10kb) <- NULL
```

##25kb

```{r}
varimp_gm12878_25kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_gm12878_25kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars.rds"))
	  varimps$Feature <- gsub(paste0(c("`","_dist"), collapse = "|"), "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_gm12878_25kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_gm12878_25kb <- data.frame(Feature = rownames(varimp_gm12878_25kb),
                                  aveimp25kb = rowMeans(varimp_gm12878_25kb, na.rm = TRUE),
                                  sdimp25kb = apply(varimp_gm12878_25kb,1, sd, na.rm = TRUE))
rownames(varimp_gm12878_25kb) <- NULL
```

##50kb

```{r}
varimp_gm12878_50kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_gm12878_50kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars.rds"))
	  varimps$Feature <- gsub(paste0(c("`","_dist"), collapse = "|"), "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_gm12878_50kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_gm12878_50kb <- data.frame(Feature = rownames(varimp_gm12878_50kb),
                                  aveimp50kb = rowMeans(varimp_gm12878_50kb, na.rm = TRUE),
                                  sdimp50kb = apply(varimp_gm12878_50kb,1, sd, na.rm = TRUE))
rownames(varimp_gm12878_50kb) <- NULL
```

##100kb

```{r}
varimp_gm12878_100kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_gm12878_100kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars.rds"))
	  varimps$Feature <- gsub(paste0(c("`","_dist"), collapse = "|"), "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_gm12878_100kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_gm12878_100kb <- data.frame(Feature = rownames(varimp_gm12878_100kb),
                                  aveimp100kb = rowMeans(varimp_gm12878_100kb, na.rm = TRUE),
                                  sdimp100kb = apply(varimp_gm12878_100kb,1, sd, na.rm = TRUE))
rownames(varimp_gm12878_100kb) <- NULL
```

##Plots

```{r}
all_propimp_gm12878 <- data.frame(Importance=c(varimp_gm12878_10kb$aveimp10kb,
                                                   varimp_gm12878_25kb$aveimp25kb,
                                                   varimp_gm12878_50kb$aveimp50kb,
                                                   varimp_gm12878_100kb$aveimp100kb),
                                  Feature=c(as.character(varimp_gm12878_10kb$Feature),
                                            as.character(varimp_gm12878_10kb$Feature),
                                            as.character(varimp_gm12878_10kb$Feature),
                                            as.character(varimp_gm12878_10kb$Feature)),
                                  Resolution=c(rep("10 kb", 107),
                                               rep("25 kb", 107),
                                               rep("50 kb", 107),
                                               rep("100 kb", 107))
                                  )

all_propimp_gm12878$Annotation <- ifelse(all_propimp_gm12878$Feature %in% chromstates, "CSS",
                                         ifelse(all_propimp_gm12878$Feature %in% histones, "HM", "TFBS"))

all_propimp_gm12878_summ <- data.frame(Prop.VI = c(
                                                   #10kb
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="10 kb" & all_propimp_gm12878$Annotation=="CSS")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="10 kb")]),
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="10 kb" & all_propimp_gm12878$Annotation=="HM")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="10 kb")]),
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="10 kb" & all_propimp_gm12878$Annotation=="TFBS")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="10 kb")]),

                                                   #25kb
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="25 kb" & all_propimp_gm12878$Annotation=="CSS")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="25 kb")]),
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="25 kb" & all_propimp_gm12878$Annotation=="HM")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="25 kb")]),
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="25 kb" & all_propimp_gm12878$Annotation=="TFBS")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="25 kb")]),

                                                   #50kb
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="50 kb" & all_propimp_gm12878$Annotation=="CSS")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="50 kb")]),
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="50 kb" & all_propimp_gm12878$Annotation=="HM")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="50 kb")]),
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="50 kb" & all_propimp_gm12878$Annotation=="TFBS")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="50 kb")]),

                                                   #100kb
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="100 kb" & all_propimp_gm12878$Annotation=="CSS")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="100 kb")]),
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="100 kb" & all_propimp_gm12878$Annotation=="HM")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="100 kb")]),
                                                   sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="100 kb" & all_propimp_gm12878$Annotation=="TFBS")])/sum(all_propimp_gm12878$Importance[which(all_propimp_gm12878$Resolution=="100 kb")])),
                                       Resolution = c(rep("10 kb", 3),
                                                      rep("25 kb", 3),
                                                      rep("50 kb", 3),
                                                      rep("100 kb", 3)),
                                       Annotation = rep(c("CSS","HM","TFBS"), 4))

all_propimp_gm12878_summ$Resolution <- factor(all_propimp_gm12878_summ$Resolution, levels=c("10 kb", "25 kb", "50 kb", "100 kb"))

ggplot(all_propimp_gm12878_summ, aes(fill=Annotation, y=Prop.VI, x=Resolution)) + 
  geom_bar( stat="identity", position="fill") +
  xlab("Resolution") +
  ylab("Proportion of Importance") +
  scale_fill_discrete(name="Annotation\nGroup")+
  theme_bw() +
  theme(axis.text.x = element_text(size=15,
                                   angle = 45,
                                   hjust = 1),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```

#K562

```{r}
annots <- list.files("Z:/TAD_data_analysis/annotations/all_annotations/k562")
annots <- gsub(paste(c("K562-",".bed", 
                             "-Haib", 
                             "-Sydh",
                             "-Uta",
                             "-Uw",
                             "-Broad",
                             "-Uchicago"), collapse = "|"), "", annots)
annots <- annots[!duplicated(annots)]

chromstates <- list.files("Z:/TAD_data_analysis/annotations/BroadHMM")
chromstates <- chromstates[grep("K562", chromstates)]
chromstates <- gsub(paste0(c("K562-", ".bed"), collapse = "|"),"",chromstates)

histones <- list.files("Z:/TAD_data_analysis/annotations/HistoneModifications")
histones <- histones[grep("K562", histones)]
histones <- gsub(paste0(c("K562-", ".bed"), collapse = "|"),"",histones)

tfbs <- list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/K562")
tfbs <- gsub(paste0(c("K562-", ".bed", "-Haib", "-Sydh", "-Broad", "-Uw", "-Uta", "-Chicago"), collapse = "|"),"", tfbs)
tfbs <- tfbs[!duplicated(tfbs)]

```

##10kb

```{r}
varimp_K562_10kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_K562_10kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars.rds"))
	  varimps$Feature <- gsub(paste0(c("`","_dist"), collapse = "|"), "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_K562_10kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_K562_10kb <- data.frame(Feature = rownames(varimp_K562_10kb),
                                  aveimp10kb = rowMeans(varimp_K562_10kb, na.rm = TRUE),
                                  sdimp10kb = apply(varimp_K562_10kb,1, sd, na.rm = TRUE))
rownames(varimp_K562_10kb) <- NULL
```

##25kb

```{r}
varimp_K562_25kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_K562_25kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars.rds"))
	  varimps$Feature <- gsub(paste0(c("`","_dist"), collapse = "|"), "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_K562_25kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_K562_25kb <- data.frame(Feature = rownames(varimp_K562_25kb),
                                  aveimp25kb = rowMeans(varimp_K562_25kb, na.rm = TRUE),
                                  sdimp25kb = apply(varimp_K562_25kb,1, sd, na.rm = TRUE))
rownames(varimp_K562_25kb) <- NULL
```

##50kb

```{r}
varimp_K562_50kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_K562_50kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars.rds"))
	  varimps$Feature <- gsub(paste0(c("`","_dist"), collapse = "|"), "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_K562_50kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_K562_50kb <- data.frame(Feature = rownames(varimp_K562_50kb),
                                  aveimp50kb = rowMeans(varimp_K562_50kb, na.rm = TRUE),
                                  sdimp50kb = apply(varimp_K562_50kb,1, sd, na.rm = TRUE))
rownames(varimp_K562_50kb) <- NULL
```

##100kb

```{r}
varimp_K562_100kb <- matrix(nrow=length(annots), ncol=21)
rownames(varimp_K562_100kb) <- annots

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
	  varimps <- readRDS(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
									   chrs[i], 
									   "/smote/distance/enet/rfimpvars.rds"))
	  varimps$Feature <- gsub(paste0(c("`","_dist"), collapse = "|"), "", varimps$Feature)	
	  varimps <- varimps[order(varimps$Feature),]

		varimp_K562_100kb[match(varimps$Feature,annots),i] <- varimps$Importance
}

varimp_K562_100kb <- data.frame(Feature = rownames(varimp_K562_100kb),
                                  aveimp100kb = rowMeans(varimp_K562_100kb, na.rm = TRUE),
                                  sdimp100kb = apply(varimp_K562_100kb,1, sd, na.rm = TRUE))
rownames(varimp_K562_100kb) <- NULL
```

##Plots

```{r}
all_propimp_K562 <- data.frame(Importance=c(varimp_K562_10kb$aveimp10kb,
                                                   varimp_K562_25kb$aveimp25kb,
                                                   varimp_K562_50kb$aveimp50kb,
                                                   varimp_K562_100kb$aveimp100kb),
                                  Feature=c(as.character(varimp_K562_10kb$Feature),
                                            as.character(varimp_K562_10kb$Feature),
                                            as.character(varimp_K562_10kb$Feature),
                                            as.character(varimp_K562_10kb$Feature)),
                                  Resolution=c(rep("10 kb", 141),
                                               rep("25 kb", 141),
                                               rep("50 kb", 141),
                                               rep("100 kb", 141))
                                  )

all_propimp_K562$Annotation <- ifelse(all_propimp_K562$Feature %in% chromstates, "CSS",
                                         ifelse(all_propimp_K562$Feature %in% histones, "HM", "TFBS"))

all_propimp_K562_summ <- data.frame(Prop.VI = c(
                                                   #10kb
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="10 kb" & all_propimp_K562$Annotation=="CSS")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="10 kb")]),
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="10 kb" & all_propimp_K562$Annotation=="HM")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="10 kb")]),
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="10 kb" & all_propimp_K562$Annotation=="TFBS")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="10 kb")]),

                                                   #25kb
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="25 kb" & all_propimp_K562$Annotation=="CSS")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="25 kb")]),
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="25 kb" & all_propimp_K562$Annotation=="HM")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="25 kb")]),
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="25 kb" & all_propimp_K562$Annotation=="TFBS")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="25 kb")]),

                                                   #50kb
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="50 kb" & all_propimp_K562$Annotation=="CSS")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="50 kb")]),
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="50 kb" & all_propimp_K562$Annotation=="HM")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="50 kb")]),
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="50 kb" & all_propimp_K562$Annotation=="TFBS")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="50 kb")]),

                                                   #100kb
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="100 kb" & all_propimp_K562$Annotation=="CSS")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="100 kb")]),
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="100 kb" & all_propimp_K562$Annotation=="HM")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="100 kb")]),
                                                   sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="100 kb" & all_propimp_K562$Annotation=="TFBS")])/sum(all_propimp_K562$Importance[which(all_propimp_K562$Resolution=="100 kb")])),
                                       Resolution = c(rep("10 kb", 3),
                                                      rep("25 kb", 3),
                                                      rep("50 kb", 3),
                                                      rep("100 kb", 3)),
                                       Annotation = rep(c("CSS","HM","TFBS"), 4))

all_propimp_K562_summ$Resolution <- factor(all_propimp_K562_summ$Resolution, levels=c("10 kb", "25 kb", "50 kb", "100 kb"))

ggplot(all_propimp_K562_summ, aes(fill=Annotation, y=Prop.VI, x=Resolution)) + 
  geom_bar( stat="identity", position="fill") +
  xlab("Resolution") +
  ylab("Proportion of Importance") +
  scale_fill_discrete(name="Annotation\nGroup")+
  theme_bw() +
  theme(axis.text.x = element_text(size=15,
                                   angle = 45,
                                   hjust = 1),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20))

```