### 2019-05-17 Manuscript/manuscript_structured.Rmd
Abstract complete

---
title: Computational Prediction of 3D Genomic Domains using 1D Genome Annotations
  in Class Imbalance Settings
output:
  word_document: default
  pdf_document:
    fig_caption: no
fontsize: 11pt
---

Spiro Stilianoudakis^1,2^ (stilianoudasc@mymail.vcu.edu), Kellen G. Cresswell^1^ (cresswellkg@vcu.edu), Mikhail G. Dozmorov^1,2\*^ (mikhail.dozmorov@vcuhealth.org)

^1^ Dept. of Biostatistics, Virginia Commonwealth University, Richmond, VA, 23298, USA  
^2^ Equal contribution  
^\*^ To whom correspondence should be addressed: Virginia Commonwealth University, Richmond, VA, 23298, 804-827-2055, mikhail.dozmorov@vcuhealth.org

# Abstract

**Background.** Topologically Associated Domains (TADs) represent fundamental building blocks of the three-dimensional (3D) genome organization and regulation. Several epigenetic marks, such as CTCF, are highly enriched at TAD boundaries, suggesting a strong epigenetic component of TAD boundary formation. Several methods have been developed to leverage the growing amount of (epi)genomic annotation data for predicting TAD boundaries; however, they overlooked key characteristics of the data, such as spatial associations between a genomic annotation and a boundary. Furthermore, the number of regions containing TAD boundaries is much smaller than the rest of the genome, creating a class imbalance problem.

**Methods.** To investigate the problem of spatial genomic relationships and class imbalance problem, we developed an ensemble framework for optimizing prediction of TAD boundaries. We utilized several resampling techniques (random over- and undersampling, Synthetic Minority Over-sampling TEchnique (SMOTE)) coupled with various metrics for classifier performance evaluation (Area Under Receiver Operating Characteristics (AUCROC) curve, Precision-Recall curve (AUPRC), F1, Matthew Correlation Coefficient (MCC), among others) to select the best predictive model in class imbalance settings.

**Results.** Using different feature engineering and resampling strategies, coupled with elastic net-based feature selection, we systematically compared the performance of models predicting the location of TAD boundaries (Gm12878 and K562 Hi-C data at 10kb to 100kb resolutions). Precision-Recall curves were found to be best suited for evaluating model performance in class imbalance settings. Random undersampling and SMOTE-resampled data performed best for optimal parameter selection. Transcription Factor Binding Sites (TFBSs) were best type of genomic annotations for predicting TAD boundaries than histone modifications and chromatin states; we identified well-known CTCF, SMC3, RAD21, and several novel TFBSs as best predictors of TAD boundaries. Importantly, spatial relationships (distance between TAD boundaries and genomic annotations) were best predictors of TAD boundaries as compared with commonly used overlap counts or percent overlap measures.

**Conclusions.** Our results outline strategies for predictive modeling of 3D genomic domains using 1D genome annotation data. The accurate modeling of TAD boundaries will improve our understanding of how epigenetics shapes the 3D structure of the genome.

<!--
# Introduction

## Importance

## Problems and limitations

### Hi-C data issues for TAD boundary identification

Although Hi-C data allows for identification of TADs genome-wide, it has limitations. First, Hi-C requires a large amount of sequencing (one Hi-C experiment requires >20X sequencing of a typical RNA-seq experiment). Consequently, the cost of obtaining Hi-C data remains high and the data availability for various cell lines and tissues remain highly uneven. Second, resolution of Hi-C data, or the size of interacting genomic regions, remains relatively coarse (on the order of kilobases) as compared with transcription factor binding motifs (tenths bases) or ChIP-seq peaks (hundredths bases), and inconsistent across cell lines and tissues. Third, Hi-C data suffers from technology- and DNA-driven biases, sparsity (high proportion of missing values), and limited sequencing depth, adding to inconsistency to cell- and tissue type-specific Hi-C data availability. These factors hinder the consistent TAD boundary identification across tissues and cell types.

### Genome annotation data opportunities

In contrast, 1D genome annotation tracks assembled by the ENCODE and Roadmap consortia provide a fine-grained look at the genome in various tissues and cell types. Although the original genome annotation data was highly uneven (rich annotations available for select cell types while remaining sparse for others), recent genome annotation imputation methods fill the cell type-specific sparsity of genome annotations. Consequently, using existing and imputed genome annotations for TAD boundary predictions represent a viable way to identify TAD boundaries in any cell or tissue type.

### Class imbalance

### Prediction types

## Previous works on solving problems

### Resampling approaches

### Models

### Evaluation metrics

## Pitfalls of previous works

### Distance not considered

## Our work on solving problems

### Consider distance
-->