### 2019-05-21 Scripts/evaluating_cv_metric_on_enet_penalization.Rmd

---
title: "evaluating_cv_metric_on_enet_penalization"
author: "Spiro Stilianoudakis"
date: "May 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(ggplot2, lib.loc = "/home/R/Rlib-3.4.1-rh6/")
library(caret, lib.loc = "/home/R/Rlib-3.4.1-rh6/")
library(DMwR)
#library(PRROC)
library(ModelMetrics, lib.loc = "/home/R/Rlib-3.4.1-rh6/")

```

# Defining summary metric for cross validation

```{r}
#defining function to extract AUPRC
#auprcSummary <- function(data, lev = NULL, model = NULL){
#  lvls <- levels(data$obs) #take the probability of good class
#  prob_good <- data[,lvls[2]] 
#  the_curve <- pr.curve(scores.class0 = prob_good,
#                        weights.class0 = as.numeric(data$obs)-1, #provide the class labels as 0/1
#                        curve = FALSE)
#  out <- the_curve$auc.integral
#  names(out) <- "AUPRC"
#  out
#}

#defining function to extract F1
#f1Summary <- function (data, lev = NULL, model = NULL) {
#  precision <- posPredValue(data$pred, data$obs, positive = "Yes")
#  recall  <- sensitivity(data$pred, data$obs, postive = "Yes")
#  f1_val <- (2 * precision * recall) / (precision + recall)
#  names(f1_val) <- c("F1")
#  f1_val
#}

#defining function to extract mcc
#mccSummary <- function (data, lev = NULL, model = NULL) 
#{
#  lvls <- levels(data$obs)
#  #if (length(lvls) > 2) 
#  #  stop(paste("Your outcome has", length(lvls), "levels. The twoClassSummary() function isn't appropriate."))
#  #requireNamespaceQuietStop("ModelMetrics")
#  #if (!all(levels(data[, "pred"]) == lvls)) 
#  #  stop("levels of observed and predicted data do not match")
#  mcc <- ModelMetrics::mcc(ifelse(data$obs == lev[2], 0, 
#                                  1), data[, lvls[1]], cutoff = .5)
#  out <- c(mcc)
#  names(out) <- c("MCC")
#  out
#}

#One summary function for roc,auprc,f1,and mcc
allSummary <- function (data, lev = NULL, model = NULL) {
  lvls <- levels(data$obs)

  #mcc
  mcc <- ModelMetrics::mcc(ifelse(data$obs == lev[2], 0, 1), data[, lvls[1]], cutoff = .5)

  #roc
  b1 <- twoClassSummary(data, lev, model)

  #auprc & f1
  c1 <- prSummary(data, lev, model)

  out <- c(mcc, b1, c1)
  names(out)[1] <- c("MCC")
  out
}

```

# GM12878 with distance type predictors

```{r}
sampling <- c("none", "ros", "rus", "smote")
resolution <- c("10kb", "25kb", "50kb", "100kb")
metric <- c("ROC", "AUC", "F", "MCC")

l=1
modelList <- list()

alpha_vals <- numeric()
lambda_vals <- numeric()
metric_vals <- character()
resample_vals <- character()
resolution_vals <- character()

for(i in 1:length(resolution)){

  #reading in whole genome data
  ##setting directory specific to cell line and resolution 
  wgd_directory <- paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/",
                          resolution[i],
                          "/training_testing/gm12878_",
                          resolution[i],
                          ".rds")
  whole_genome_dat <- readRDS(wgd_directory)

  #reading in bin data on the whole genome that contains chromosome information
  ##setting directory specific to cell line and resolution 
  bd_directory <- paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/",
                         resolution[i],
                         "/training_testing/binslist",
                         gsub("kb","",resolution[i]),
                         "_center.rds")
  binslist_dat <- readRDS(bd_directory)

  varstorm = c(names(whole_genome_dat)[grep("Ctcf-", names(whole_genome_dat))][-c(1:4)],
               names(whole_genome_dat)[grep("Ebf1sc137065-", names(whole_genome_dat))][-c(1:4)],
               names(whole_genome_dat)[grep("P300-", names(whole_genome_dat))][-c(1:4)],
               names(whole_genome_dat)[grep("Pol2-", names(whole_genome_dat))][-c(1:4)],
               names(whole_genome_dat)[grep("Rad21-", names(whole_genome_dat))][-c(1:4)])

  whole_genome_dat <- whole_genome_dat[,-which(names(whole_genome_dat) %in% varstorm)]

  #reducing the whole genome data to only specific chromosome
  reduce_chr <- "chr1"
  chr_specific_dat <- whole_genome_dat[which(as.character(seqnames(binslist_dat))==reduce_chr),]

  #splitting into training and testing
  set.seed(123)
  inTrainingSet <- sample(length(chr_specific_dat$y),floor(length(chr_specific_dat$y)*.7))
  train <- chr_specific_dat[inTrainingSet,]
  test <- chr_specific_dat[-inTrainingSet,]

  #reduce predictor space
  reduce_predictor = "_dist"
  train <- train[,c(1,grep(reduce_predictor, names(train)))]

  for(j in 1:length(sampling)){

    # set number of iterations
    samps = 50

    if(sampling[j]=="ros"){
      #assign sample indeces
      sampids <- matrix(ncol=samps, 
                        nrow=length(train$y[which(train$y=="No")]))

      #filling in the sample ids matrix
      set.seed(123)
      for(s in 1:samps){
        sampids[,s] <- sample(x = which(train$y=="Yes"),
                              size = length(which(train$y=="No")),
                              replace = TRUE)
      }
      train <- rbind.data.frame(train[which(train$y=="No"),],
                                train[sampids[,1],])
      #Randomly shuffle the data
      set.seed(321)
      train <- train[sample(nrow(train)),]
    }else if(sampling[j]=="rus"){
      #assign sample indeces
      sampids <- matrix(ncol=samps, 
                        nrow=length(train$y[which(train$y=="Yes")]))

      #filling in the sample ids matrix
      set.seed(123)
      for(s in 1:samps){
        sampids[,s] <- sample(x = which(train$y=="No"),
                              size = length(which(train$y=="Yes")),
                              replace = FALSE)
      }
      train <- rbind.data.frame(train[which(train$y=="Yes"),],
                                train[sampids[,1],])

      #Randomly shuffle the data
      set.seed(321)
      train <- train[sample(nrow(train)),]
    }else if(sampling[j]=="smote"){
      set.seed(123)
      train <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

      #Randomly shuffle the data
      set.seed(123)
      train <- train[sample(nrow(train)),]
    }else{train=train}

    for(k in 1:length(metric)){
      print(paste(i, j, k))

      #Feature Selection: ENET
      ##set up grid of alpha and lambda values
      lambda.grid = seq(0,10, length.out = 10)
      alpha.grid = seq(0,1,length=10)
      srchGrid <- expand.grid(.alpha=alpha.grid, .lambda=lambda.grid)

      ##set tuning parameters
      ##set seeds for each cross folds
      set.seed(123)
      #length is = (n_repeats*nresampling)+1
      seeds <- vector(mode = "list", length = 11)
      #(1 is the number of tuning parameter lambda
      for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 100)
      #for the last model
      seeds[[11]]<-sample.int(1000, 1)

      ##setting contols for elastic net
      fitControl <- trainControl(seeds = seeds,
                                 method = "cv",
                                 number = 10,
                                 ## Estimate class probabilities
                                 classProbs = TRUE,
                                 ## Evaluate performance using 
                                 ## the following function
                                 summaryFunction = allSummary)

      #running elastic net model
      eNetModel <- train(relevel(y, ref="Yes") ~ ., data=train, 
                         method = "glmnet", 
                         metric= metric[k], 
                         trControl = fitControl, 
                         family="binomial", 
                         tuneGrid=srchGrid,
                         standardize=FALSE)

      enetcoefs <- data.frame(Feature = gsub("`", "", rownames(coefficients(eNetModel$finalModel, s=eNetModel$bestTune$lambda))[-1]), 
                                                Coefficient = coefficients(eNetModel$finalModel, s=eNetModel$bestTune$lambda)[-1])

      #enetcoefs <- enetcoefs[-which(enetcoefs$Coefficient==0),]

      modelList[[l]] <- enetcoefs

      l <- l+1

      resolution_vals <- rbind(resolution_vals, resolution[i])
      resample_vals <- rbind(resample_vals, sampling[j])
      metric_vals <- rbind(metric_vals, metric[k])
      alpha_vals <- rbind(alpha_vals, as.numeric(eNetModel$bestTune[1])) 
      lambda_vals <- rbind(lambda_vals, as.numeric(eNetModel$bestTune[2])) 
    }
  }
}

saveRDS(modelList, "/home/stilianoudakisc/TAD_data_analysis/comparing_elastic_net_results/modelList.rds")

enet_df <- data.frame(Resolution = resolution_vals,
                      Resampling = resample_vals,
                      Metric = metric_vals,
                      Alpha = alpha_vals,
                      Lambda = lambda_vals)

saveRDS(enet_df, "/home/stilianoudakisc/TAD_data_analysis/comparing_elastic_net_results/enet_df.rds")
```

# Plots