### 2019-05-29 Manuscript/Supplementary/supplementary_legends.Rmd
editted manuscript and legends

---
title: "Legends for supplementary material"
output: pdf_document
---

**Supplementary Table 1. Hi-C data.** A complete list of the Hi-C data libraries used to obtain the locations of topologically associating domains.

**Supplementary Table 2. Genomic annotations.** A complete list of the genomic annotations used to build the feature space for all downstream models. The GRCh37/hg19 human genome assembly was used. "Genomic Class" - broad category of genomic features, "Element" - names of genomic features, "Available Cell Type" - the name of the cell type in which the annotation was available, "Cell Type-Specific Source" - download URL specific to the cell line (not all annotations were provided by the same institutions).

**Supplementary Figure 1. Performing $log_{2}$ transformations normalized skewed distance densities across resolution for GM12878.** When measured in basepairs, distances were highly right skewed for all resolutions: (A) 10 kb, (B) 25 kb, (C) 50 kb, and (D) 100 kb. After $log_{2}$ transforming the distances, densities became much more gaussian distributed for all resolutions: (E) 10 kb, (F) 25 kb, (G) 50 kb, and (H) 100 kb. Each density curve represents a cell line specific annotation, and its relationship with all resolution specific genomic bins.

**Supplementary Figure 2. Performing $log_{2}$ transformations normalized skewed distance densities across resolution for K562.** When measured in basepairs, distances were highly right skewed for all resolutions: (A) 10 kb, (B) 25 kb, (C) 50 kb, and (D) 100 kb. After $log_{2}$ transforming the distances, densities became much more gaussian distributed for all resolutions: (E) 10 kb, (F) 25 kb, (G) 50 kb, and (H) 100 kb. Each density curve represents a cell line specific annotation, and its relationship with all resolution specific genomic bins.

**Supplementary Table 3. Data Summaries and Imbalance Ratios.** For K562, average number of genomic bins, TADs called, and the corresponding boundaries and imbalance ratios, across each of the autosomal chromosomes (omitting chromosome 9), stratified by resolution. The resolution of the Hi-C contact matrix dictated the width of each genomic bin, with lower resolutions corresponding to wider genomic bins.

**Supplementary Table 4. Inter-Resolution ICCs.** For K562, two-way intraclass correlation coefficients between resolutions for the following quantities: Number of genomic bins, number of TADs called, number of unique TAD boundaries, and imbalance ratio. The ICC was computed across 21 of the autosomal chromosomes. All quatities, except for imbalance ratio, was correlated with data resolution. Instead, imbalance ratios remained consistent and low across resolutions.

**Table 5. Jaccard Indices.** For K562, the median (range) Jaccard index illustrating the pairwise similarity between annotations within specific groups, including chromatin segmentation states, histone modifications, and transcription factor binding sites. The Jaccard Index measures the ratio of the number of intersecting base pairs between any two sets, to the number of base pairs in the union of the two sets. The number of cell line specific annotations included within each group is also provided.

**Supplementary Table 6. Full list of aggregated performances for GM12878.** Performances were stratified by resolution, resampling technique, and predictor type, and then averaged across chromosomes.

**Supplementary Table 7. Full list of aggregated performances for K562.** Performances were stratified by resolution, resampling technique, and predictor type, and then averaged across chromosomes.

**Supplementary Figure 3. SMOTE re-sampled models built on distance-type predictors yielded optimal MCCs across resolutions for K562.** Performances of random forest classification models built on each of the 3 predictor types: (A) OC, (B) OP, and (C) Distance. Within each predictor type, and across resolution, training data was re-sampled using different techniques including: None (red), ROS (green), RUS (blue), and SMOTE (purple). Performances were aggregated using means across each chromosome, with error bars representing 1 standard deviation away from the mean.

**Supplementary Figure 4. The proportional importance of optimally built predictive models was outweighed by transcription factor binding sites across resolutions for K562.** Stacked barplots illustrate the proportion of total predictive importance in models using optimal data level characteristics determined by the ensemble framework (SMOTE re-sampling; with distance-type predictors), for each resolution on GM12878. Proportions are grouped by the class of cell line specific annotations in the feature space including chromatin state segmentations (CSS), histone modifications (HM), and transcription factor binding sites (TFBS).

**Supplementary Figure 5. Optimally built TFBS-specific models outperformed other annotation-specific models across resolution for K562** Performances were compared for random forest models built using specific groups of annotations as the feature space including: histone modifications (HM; red), chromatin state segmentations (CSS; green), transcription factor binding sites (TFBS; blue), and a feature space containing all 3 types combined (combined; purple). Each bar represents a model ran through the ensemble framework, with re-sampling technique and predictor-type characteristics fixed to SMOTE and distance-types. The error bars represent $\pm$ 1 standard deviation from the mean MCC across each chromosome.

**Supplementary Figure 6. Top 20 most predictive transcription factor binding sites sorted by 10 kb for K562.** A heatmap illustrating the importance of the key cell line specific transcription factors for optimally built random forest models across resolution. Importances were sorted according to the 10 kb resolution and rescaled to values between 0 and 100. Transcription factors with higher values of predictive importance are denoted in dark red.

**Supplementary Figure 9. Top predictive transcription factor binding sites were significantly closer to bins that overlapped with varying units of flanked TAD boundaries on GM12878** Boxplots illustrating the distribution of distances (on $log_{2}$ scale) around (A) 25 kb, (B) 50 kb, and (C) 100 kb regions on the linear genome that either overlapped with a flanked TAD boundary (Yes) or did not (No), for the most influential cell line specific TFBS including SMC3, CTCF, RAD21, ZNF143, and YY1sc281. P-values obtained from Wilcoxon Rank Sum tests are provided assessing the significance [$NS$ = p > 0.05; $*$ = 0.01 < p $\le$ 0.05; $**$ = 0.001 < p $\le$ 0.01; $***$ = 0.0001 < p $\le$ 0.001; $****$ = p $\le$ 0.00001].

**Supplementary Figure 10. Top predictive transcription factor binding sites were significantly closer to bins that overlapped with varying units of flanked TAD boundaries on K562** Boxplots illustrating the distribution of distances (on $log_{2}$ scale) around (A) 10 kb, (B) 25 kb, (C) 50 kb, and (D) 100 kb regions on the linear genome that either overlapped with a flanked TAD boundary (Yes) or did not (No), for the most influential cell line specific TFBS including SMC3, CTCF, RAD21, and ZNF143. P-values obtained from Wilcoxon Rank Sum tests are provided assessing the significance [$NS$ = p > 0.05; $*$ = 0.01 < p $\le$ 0.05; $**$ = 0.001 < p $\le$ 0.01; $***$ = 0.0001 < p $\le$ 0.001; $****$ = p $\le$ 0.00001].

<!--
**Supplementary Figure 12. Comparing $log_{2}$ distance profiles of top predictive transcription factor binding sites by bin class for 25 kb resolution, on GM12878.** Violin plots (top) and plots of differences in quantiles (bottom) between genomic bin class for 4 of the top predictive TFBS determined by the best performing model, including: SMC3 (A), ZNF143 (B), CTCF (C), and RAD21 (D). For the plots of differences in quantiles, the reference group are the genomic bins that did not overlap with a flanked TAD boundary. That is, quantile differences above 0 indicate that bins that overlapped with flanked TAD boundaries exhibited smaller quantiles.
-->