### 2019-06-10 Scripts/investigating_misclassified_TADs.Rmd
script for predicting finer bins

---
title: "Determining what makes some TADs unpredictable"
author: "Spiro Stilianoudakis"
date: "April 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# Evaluating Missclassified TAD boundaries from best performing model

## SMOTE; distance-type predictors

```{r}
misclassified_func <- function(CL, resolution, sampling="smote", predictor="distance", chromosome){
  #reading in whole genome data
	##setting directory specific to cell line and resolution 
	if(CL=="GM12878"){dataset = "gm12878"}else{dataset = "K562"}
	wgd_directory <- paste0("Z:/TAD_data_analysis/",
	                        CL,
							"/",
							resolution,
							"/training_testing/",
							dataset,
							"_",
							resolution,
							".rds")
	whole_genome_dat <- readRDS(wgd_directory)

	#reading in bin data on the whole genome that contains chromosome information
	##setting directory specific to cell line and resolution 
	bd_directory <- paste0("Z:/TAD_data_analysis/",
	                        CL,
							"/",
							resolution,
							"/training_testing/binslist",
							gsub("kb","",resolution),
							"_center.rds")
	binslist_dat <- readRDS(bd_directory)

	#removing duplicate annotations
	if(CL=="GM12878"){
		varstorm = c(names(whole_genome_dat)[grep("Ctcf-", names(whole_genome_dat))][-c(1:4)],
					 names(whole_genome_dat)[grep("Ebf1sc137065-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("P300-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("Pol2-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("Rad21-", names(whole_genome_dat))][-c(1:4)])} else {
		varstorm = c(names(whole_genome_dat)[grep("Atf3-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("Cmyc-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("Ctcf-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("E2f6-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("Max-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("P300-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("Pol2-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("Rad21-", names(whole_genome_dat))][-c(1:4)],
                     names(whole_genome_dat)[grep("Yy1-", names(whole_genome_dat))][-c(1:4)])
					 }

	whole_genome_dat <- whole_genome_dat[,-which(names(whole_genome_dat) %in% varstorm)]

	#reducing the whole genome data to only specific chromosome
	reduce_chr <- tolower(chromosome)
	chr_specific_dat <- whole_genome_dat[which(as.character(seqnames(binslist_dat))==reduce_chr),]

	#splitting into training and testing
	set.seed(123)
    inTrainingSet <- sample(length(chr_specific_dat$y),floor(length(chr_specific_dat$y)*.7))
    train <- chr_specific_dat[inTrainingSet,]
    test <- chr_specific_dat[-inTrainingSet,]

	#reduce predictor space
	if(predictor=="distance"){reduce_predictor = "_dist"}else if(predictor=="oc"){reduce_predictor = "_count"}else{reduce_predictor = "_perc"}

	train <- train[,c(1,grep(reduce_predictor, names(train)))]
	test <- test[,c(1,grep(reduce_predictor, names(test)))]

	#reading in model and making predictions
	model_directory <- paste0("Z:/TAD_data_analysis/",
	                        CL,
							"/",
							resolution,
							"/results_by_chr/",
							chromosome,
							"/",
							"smote/distance/enet/rfModel_noenet.rds")
  rfModel <- readRDS(model_directory)

  predictions <- predict(rfModel, newdata=test)
  preddata <- cbind.data.frame(Prediction = predictions, True=test$y)

  #subset the false negatives (bins that were predicted as not overlapping with flanked TAD boundary but really were)
  FNdata <- test[which(preddata$Prediction=="No" & preddata$True=="Yes"),]
  FNdatamelt <- melt(FNdata[,-1])
  FNdatamelt$Prediction <- "False Negatives"

  #subset the true positives (bins that were predicted as overlapping with flanked TAD boundary and truely were)
  TPdata <- test[which(preddata$Prediction=="Yes" & preddata$True=="Yes"),]
  TPdatamelt <- melt(TPdata[,-1])
  TPdatamelt$Prediction <- "True Positives"

  prediction_data <- list(FNdatamelt,TPdatamelt)
  return(prediction_data)
}
```

## Plots

```{r}
allpredictiondata <- rbind.data.frame(prediction_data[[1]],prediction_data[[2]])
allpredictiondata$Prediction <- factor(allpredictiondata$Prediction, levels=c("False Negatives","True Positives"))

#looking at specific annotations
##Ctcf
ggplot(allpredictiondata[grep("Ctcf-",allpredictiondata$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Density") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

##Rad21
ggplot(allpredictiondata[grep("Rad21-",allpredictiondata$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Density") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```

```{r}
#reading in test data and reducing
test <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/training_testing/test.rds")
names(test) <- gsub(paste(c("`", 
                            "Gm12878", 
                            "-", 
                            #"_", 
                            #"dist", "count", "perc",
                            "Haib", 
                            "Sydh",
                            "Uta",
                            "Uw",
                            "Broad",
                            "Uchicago"), collapse = "|"), "", names(test))
enetcoefs <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/enetcoefs.rds")
test <- test[,c(1,which(names(test) %in% enetcoefs))]

#reading in model and making predictions
rfmodel_gm12878_10kb <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/Ensemble/SMOTE/Distance/ENET/RF/rfModel.rds")
predictions <- predict(rfmodel_gm12878_10kb, newdata=test)
preddata <- cbind.data.frame(Prediction = predictions, True=test$y)

#subset the false negatives (bins that were predicted as not overlapping with flanked TAD boundary but really were)
FNdata <- test[which(preddata$Prediction=="No" & preddata$True=="Yes"),]
FNdatamelt <- melt(FNdata[,-1])
FNdatamelt$Prediction <- "False Negatives"

#subset the true positives (bins that were predicted as overlapping with flanked TAD boundary and truely were)
TPdata <- test[which(preddata$Prediction=="Yes" & preddata$True=="Yes"),]
TPdatamelt <- melt(TPdata[,-1])
TPdatamelt$Prediction <- "True Positives"

#Wilcoxon test on differences in distances between FN and TP
FNdata <- FNdata[,-1]
TPdata <- TPdata[,-1]
wilcoxon.results <- numeric(length(colnames(FNdata)))
for(i in 1:length(colnames(FNdata))){
 wilcoxon.results[i] <- wilcox.test(FNdata[,i],TPdata[,i])$p.value
}
wilcox.table <- cbind.data.frame(Annotation=gsub("_dist","",colnames(FNdata)), Wilcoxon.PValue=wilcoxon.results)
wilcox.table <- wilcox.table[order(wilcox.table$Wilcoxon.PValue, decreasing = FALSE),]
wilcox.table$Wilcoxon.PValue <- round(wilcox.table$Wilcoxon.PValue,7)

predictiondata_gm12878_10kb <- rbind.data.frame(FNdatamelt,TPdatamelt)
predictiondata_gm12878_10kb$Prediction <- factor(predictiondata_gm12878_10kb$Prediction, levels=c("False Negatives","True Positives"))

#Ploting densities
ggplot(predictiondata_gm12878_10kb, aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

#looking at specific annotations
##Ctcf
ggplot(predictiondata_gm12878_10kb[grep("Ctcf_dist",predictiondata_gm12878_10kb$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

##ZNF143
ggplot(predictiondata_gm12878_10kb[grep("Znf143166181ap_dist",predictiondata_gm12878_10kb$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

##SMC3a
ggplot(predictiondata_gm12878_10kb[grep("Smc3ab9263_dist",predictiondata_gm12878_10kb$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

##Rad21
ggplot(predictiondata_gm12878_10kb[grep("Rad21_dist",predictiondata_gm12878_10kb$variable),], aes(x=value, color=Prediction)) +
  geom_density() +
  scale_color_discrete(name="Prediction Class") +
  ylab("Frequency") +
  xlab("Log2 Distance") +
  #ylab("") +
  #xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=15 #,
                                      #angle = 45, 
                                      #margin = ggplot2::margin(t = 35),
                                      #hjust = 1
     ),
     axis.text.y = element_text(size = 15),
     axis.title.x = element_text(size = 20),
     axis.title.y = element_text(size = 20),
     legend.text=element_text(size=15),
     legend.title=element_text(size=20),
     plot.title = element_text(size=20))

```