### 2019-06-10 Scripts/predicting_finer_bins.Rmd
added result script

---
title: "Predicting Finer Bins"
author: "Spiro Stilianoudakis"
date: "June 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# Predict finer bins

```{r}
finer_resolution_predict_func <- function(CL, resolution, sampling, predictor, chromosome){
	#reading in whole genome data
	##setting directory specific to cell line and resolution 
	if(CL=="GM12878"){dataset = "gm12878"}else{dataset = "K562"}
	wgd_directory <- paste0("Z:/TAD_data_analysis/",
	                        CL,
							"/",
							resolution,
							"/training_testing/",
							dataset,
							"_",
							resolution,
							".rds")
	whole_genome_dat <- readRDS(wgd_directory)

	#reading in bin data on the whole genome that contains chromosome information
	##setting directory specific to cell line and resolution 
	bd_directory <- paste0("Z:/TAD_data_analysis/",
	                        CL,
							"/",
							resolution,
							"/training_testing/binslist",
							gsub("kb","",resolution),
							"_center.rds")
	binslist_dat <- readRDS(bd_directory)

	#annotations to keep
	#############################################
	#!!!make this cell line specific; don't include yy1 for k562
	#############################################

	if(CL=="GM12878" & predictor=="distance"){
	  annots = c("Gm12878-Ctcf-Broad_dist",
				   "Gm12878-Znf143166181ap-Sydh_dist",
				   "Gm12878-Rad21-Haib_dist",
				   "Gm12878-Smc3ab9263-Sydh_dist",
				   "Gm12878-Yy1sc281-Haib_dist")
	}else if(CL=="GM12878" & predictor=="count"){
	  annots = c("Gm12878-Ctcf-Broad_count",
				   "Gm12878-Znf143166181ap-Sydh_count",
				   "Gm12878-Rad21-Haib_count",
				   "Gm12878-Smc3ab9263-Sydh_count",
				   "Gm12878-Yy1sc281-Haib_count")
	}else if(CL=="K562" & predictor=="distance"){
	  annots = c("K562-Ctcf-Broad_dist",
				   "K562-Znf143-Sydh_dist",
				   "K562-Rad21-Haib_dist",
				   "K562-Smc3ab9263-Sydh_dist")
	}else{
	   annots = c("K562-Ctcf-Broad_count",
				   "K562-Znf143-Sydh_count",
				   "K562-Rad21-Haib_count",
				   "K562-Smc3ab9263-Sydh_count") 
	  }

	whole_genome_dat <- whole_genome_dat[,c(1,which(names(whole_genome_dat) %in% annots))]

	#reducing the whole genome data to only specific chromosome
	reduce_chr <- tolower(chromosome)
	chr_specific_dat <- whole_genome_dat[which(as.character(seqnames(binslist_dat))==reduce_chr),]

	train <- chr_specific_dat

	# set number of iterations
	samps = 50

	if(sampling=="ros"){
	    #assign sample indeces
		sampids <- matrix(ncol=samps, 
                  nrow=length(train$y[which(train$y=="No")]))

		#filling in the sample ids matrix
		set.seed(123)
		for(j in 1:samps){
			sampids[,j] <- sample(x = which(train$y=="Yes"),
                        size = length(which(train$y=="No")),
                        replace = TRUE)
		}
		train <- rbind.data.frame(train[which(train$y=="No"),],
                          train[sampids[,1],])
		#Randomly shuffle the data
		set.seed(321)
		train <- train[sample(nrow(train)),]
        }else if(sampling=="rus"){
		    #assign sample indeces
			sampids <- matrix(ncol=samps, 
                  nrow=length(train$y[which(train$y=="Yes")]))

			#filling in the sample ids matrix
			set.seed(123)
			for(j in 1:samps){
				sampids[,j] <- sample(x = which(train$y=="No"),
                        size = length(which(train$y=="Yes")),
                        replace = FALSE)
			}
			train <- rbind.data.frame(train[which(train$y=="Yes"),],
                          train[sampids[,1],])

			#Randomly shuffle the data
			set.seed(321)
			train <- train[sample(nrow(train)),]
			}else if(sampling=="smote"){
				set.seed(123)
				train <- SMOTE(y ~ ., 
                     data=train, 
                     perc.over = 100, 
                     perc.under = 200)

				#Randomly shuffle the data
				set.seed(123)
				train <- train[sample(nrow(train)),]
				}else{train=train}

	##set tuning parameters
	##set seeds for each cross folds
	set.seed(123)
	#length is = (n_repeats*nresampling)+1
	seeds <- vector(mode = "list", length = 11)
	#(1 is the number of tuning parameter lambda
	for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 100)
	#for the last model
	seeds[[11]]<-sample.int(1000, 1)

	##defining one summary function for roc,auprc,f1,and mcc
	allSummary <- function (data, lev = NULL, model = NULL) {
		lvls <- levels(data$obs)

		#mcc
		mcc <- ModelMetrics::mcc(ifelse(data$obs == lev[2], 0, 1), data[, lvls[1]], cutoff = .5)

		#roc
		b1 <- twoClassSummary(data, lev, model)

		#auprc & f1
		c1 <- prSummary(data, lev, model)

		out <- c(mcc, b1, c1)
		names(out)[1] <- c("MCC")
		out
	}

	##setting contols for elastic net
	fitControl <- trainControl(#seeds = seeds,
                           method = "cv",
                           number = 5,
                           ## Estimate class probabilities
                           classProbs = TRUE,
                           ## Evaluate performance using 
                           ## the following function
                           summaryFunction = allSummary)

	#set number of predictors to consider at each node
	tunegrid <- expand.grid(mtry=ceiling(sqrt(dim(train)[2] - 1)))

	#performing random forest
	rfModel <- train(y~., data=train, 
                 method="rf", 
                 metric="AUC", 
                 #tuneGrid=tunegrid, 
				 tuneLength=4,
                 trControl=fitControl, 
                 ntree=500) #as.numeric(rownames(results)[1]))

	return(rfModel)

}

```

# Make finer bins

```{r}
#CL="GM12878"; resolution="10kb"; sampling="smote"; predictor="distance"; chromosome="CHR22"

make_finer_test <- function(CL, resolution, sampling, predictor, chromosome){
	#reading in bin data on the whole genome that contains chromosome information
	##setting directory specific to cell line and resolution 
	bd_directory <- paste0("Z:/TAD_data_analysis/",
	                        CL,
							"/",
							resolution,
							"/training_testing/binslist",
							gsub("kb","",resolution),
							"_center.rds")
	binslist_dat <- readRDS(bd_directory)

	#extracting the first and last chromosome specific genomic coordinate
	reduce_chr <- tolower(chromosome)
	chr_binslist_dat <- binslist_dat[which(as.character(seqnames(binslist_dat))==reduce_chr)]
	min_point <- start(chr_binslist_dat)[1]+1
	max_point <- end(chr_binslist_dat)[length(chr_binslist_dat)]+1

	#creating finer bins for new test data
	##100 bases 
	test_bins <- GRanges(seqnames=reduce_chr, 
						 ranges=IRanges(start=seq(from=min_point, to=max_point, by=100), 
						 width=100))

	#reading chromosome and resolution specific boundary data
	bounds_directory <- paste0("Z:/TAD_data_analysis/",
	                        CL,
							"/",
							resolution,
							"/training_testing/bounds.rds")
	bounds <- readRDS(bounds_directory)
	bounds_f <- flank(bounds, width=100, both=TRUE)
	bounds_f <- bounds_f[which(as.character(seqnames(bounds_f))==tolower(chromosome))]

	#Creating Response Vector Y (1 if tad boundary is in bin; 0 if not)
	y <- countOverlaps(test_bins, bounds_f)
	y <- ifelse(y>=1,1,0)

	test_data <- data.frame(y = y)

	#creating a granges object from the center of every range in the new binned test data
	#this will be used to calculate distances from genomic feature to center of genomic bin
	test_bins_center <- resize(test_bins, width = 1, fix = "center")

	#annotating test data 
	##Creating Feature Space
	###calculating distance
	distance_func <- function(binned_data_center_gr, annot_data_center_gr){
		#distance from center of genomic bin to nearest genomic region of interest
		dist <- mcols(distanceToNearest(binned_data_center_gr, annot_data_center_gr))$distance
		return(dist)
	}

	####calculating count overlaps
	count_func <- function(binned_data_gr, annot_data_gr){
		#Finding the total number of overlaps between genomic bins and the specific genomic annotation
		count_total <- countOverlaps(binned_data_gr, annot_data_gr)

		return(count_total)
	}

	setwd("Z:/TAD_data_analysis/annotations/all_annotations/gm12878/")
	temp = list.files()

	#isolate specific annotations
	#############################################
	#!!!make this cell line specific; don't include yy1 for k562
	#############################################
	annots <- c("Gm12878-Ctcf-Broad.bed",
				"Gm12878-Znf143166181ap-Sydh.bed",
				"Gm12878-Rad21-Haib.bed",
				"Gm12878-Smc3ab9263-Sydh.bed",
				"Gm12878-Yy1sc281-Haib.bed")
	temp <- temp[which(temp %in% annots)]

	for(i in 1:length(temp)){
		#print(temp[i])
		dat <- read.table(temp[i],header = FALSE, sep="\t")
		dat_gr <- GRanges(seqnames=dat$V1,
						IRanges(start=dat$V2, 
								end=dat$V3))
		dat_gr_center <- resize(dat_gr, width = 1, fix = "center")

		####calculating distance
		d <- distance_func(test_bins_center, dat_gr_center)
		test_data <- cbind.data.frame(test_data,d)
		names(test_data)[length(names(test_data))] <- paste(sub(".bed", "", temp[i]), "_dist", sep="")

		####calculating count overlaps
		#co <- count_func(test_bins_center, dat_gr)
		#test_data <- cbind.data.frame(test_data,co)
		#names(test_data)[length(names(test_data))] <- paste(sub(".bed", "", temp[i]), "_count", sep="")

	}

	##Taking log2 transform of distance variables
	cols <- grep("dist",colnames(test_data))
	test_data[,cols] <- apply(test_data[,cols],2,function(x){log(x + 1, base=2)})

	##Changing levels of response (y) to yes no
	test_data$y <- factor(test_data$y)
	levels(test_data$y) <- c("No", "Yes")

	return(list(test_data,test_bins,test_bins_center))

}

test_list <- make_finer_test(CL="GM12878", resolution="10kb", sampling="smote", predictor="distance", chromosome="CHR22")

test_bins <- test_list[[2]]
test_bins_center <- test_list[[3]]
test_data <- test_list[[1]]

```

# Plots

```{r}
pred.rfModel2 <- predict(rfModel,newdata=test_data[,-1],type="prob")[,"Yes"]

	test_data$PredProb <- pred.rfModel2

mcols(test_bins_center)$prediction <- pred.rfModel2

test_data[which(test_data$y=="Yes"),][1:100,c(1,7)]

x <- cbind.data.frame(bin=seq(51224999+1, 51234999, 100),
                      PredProb=test_data$PredProb)

ggplot(x, aes(x=bin, y=PredProb)) + 
    geom_point() +
    geom_smooth(method="loess",size=1)+
    theme_minimal() +
    theme_bw()

```