### 2019-07-09 Scripts/evaluating_cell_line_specific_validation_performances.Rmd
initial commit

---
title: "Cell Type Specific Validation"
author: "Spiro Stilianoudakis"
date: "July 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)

```

# Building optimal models on one cell line and validating on another

## GM12878 on K562

### 10kb

```{r}
mccdata_gm12878_on_k562_10kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs_GM12878_on_K562.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs_GM12878_on_K562.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_gm12878_on_k562_10kb <- rbind(mccdata_gm12878_on_k562_10kb,
			                                      perfdata[15,2])
}

```

### 25kb

```{r}
mccdata_gm12878_on_k562_25kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs_GM12878_on_K562.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs_GM12878_on_K562.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_gm12878_on_k562_25kb <- rbind(mccdata_gm12878_on_k562_25kb,
			                                      perfdata[15,2])
}

```

### 50kb

```{r}
mccdata_gm12878_on_k562_50kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs_GM12878_on_K562.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs_GM12878_on_K562.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_gm12878_on_k562_50kb <- rbind(mccdata_gm12878_on_k562_50kb,
			                                      perfdata[15,2])
}

```

### 100kb

```{r}
mccdata_gm12878_on_k562_100kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs_GM12878_on_K562.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs_GM12878_on_K562.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_gm12878_on_k562_100kb <- rbind(mccdata_gm12878_on_k562_100kb,
			                                      perfdata[15,2])
}

```

## K562 on GM12878

### 10kb

```{r}
mccdata_k562_on_gm12878_10kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs_K562_on_GM12878.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs_K562_on_GM12878.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_k562_on_gm12878_10kb <- rbind(mccdata_k562_on_gm12878_10kb,
			                                      perfdata[15,2])
}

```

### 25kb

```{r}
mccdata_k562_on_gm12878_25kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs_K562_on_GM12878.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs_K562_on_GM12878.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_k562_on_gm12878_25kb <- rbind(mccdata_k562_on_gm12878_25kb,
			                                      perfdata[15,2])
}

```

### 50kb

```{r}
mccdata_k562_on_gm12878_50kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs_K562_on_GM12878.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs_K562_on_GM12878.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_k562_on_gm12878_50kb <- rbind(mccdata_k562_on_gm12878_50kb,
			                                      perfdata[15,2])
}

```

### 100kb

```{r}
mccdata_k562_on_gm12878_100kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs_K562_on_GM12878.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs_K562_on_GM12878.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_k562_on_gm12878_100kb <- rbind(mccdata_k562_on_gm12878_100kb,
			                                      perfdata[15,2])
}

```

## GM12878 on GM12878

### 10kb

```{r}
mccdata_gm12878_on_gm12878_10kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_gm12878_on_gm12878_10kb <- rbind(mccdata_gm12878_on_gm12878_10kb,
			                                      perfdata[15,2])
}

```

### 25kb

```{r}
mccdata_gm12878_on_gm12878_25kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_gm12878_on_gm12878_25kb <- rbind(mccdata_gm12878_on_gm12878_25kb,
			                                      perfdata[15,2])
}

```

### 50kb

```{r}
mccdata_gm12878_on_gm12878_50kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_gm12878_on_gm12878_50kb <- rbind(mccdata_gm12878_on_gm12878_50kb,
			                                      perfdata[15,2])
}

```

### 100kb

```{r}
mccdata_gm12878_on_gm12878_100kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_gm12878_on_gm12878_100kb <- rbind(mccdata_gm12878_on_gm12878_100kb,
			                                      perfdata[15,2])
}

```

## K562 on K562

### 10kb

```{r}
mccdata_k562_on_k562_10kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_k562_on_k562_10kb <- rbind(mccdata_k562_on_k562_10kb,
			                                      perfdata[15,2])
}

```

### 25kb

```{r}
mccdata_k562_on_k562_25kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_k562_on_k562_25kb <- rbind(mccdata_k562_on_k562_25kb,
			                                      perfdata[15,2])
}

```

### 50kb

```{r}
mccdata_k562_on_k562_50kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_k562_on_k562_50kb <- rbind(mccdata_k562_on_k562_50kb,
			                                      perfdata[15,2])
}

```

### 100kb

```{r}
mccdata_k562_on_k562_100kb <- numeric()

for(i in paste0("CHR", c(1:8,10:22))){
			files <- list.files(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/"))
		  if("rfperf_tfbs.rds" %in% files){
		    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
									   i, 
									   "/smote/distance/enet/rfperf_tfbs.rds"))
		  }else{perfdata=data.frame(Metric = c("TN",
                                "FN",
                                "FP",
                                "TP",
                                "Total",
                                "Sensitivity",
                                "Specificity",
                                "Kappa",
                                "Accuracy",
                                "Precision",
                                "FPR",
                                "FNR",
                                "FOR",
                                "NPV",
                                "MCC",
                                "F1",
                                "AUC",
                                "Youden",
                                "AUPRC"), 
                     Performance=NA)}

			mccdata_k562_on_k562_100kb <- rbind(mccdata_k562_on_k562_100kb,
			                                      perfdata[15,2])
}

```

## Plots

```{r}
all_rfperfs_cl_specific_valid <- rbind.data.frame(mccdata_gm12878_on_k562_10kb,
                                                  mccdata_gm12878_on_k562_25kb,
                                                  mccdata_gm12878_on_k562_50kb,
                                                  mccdata_gm12878_on_k562_100kb,
                                                  mccdata_k562_on_gm12878_10kb,
                                                  mccdata_k562_on_gm12878_25kb,
                                                  mccdata_k562_on_gm12878_50kb,
                                                  mccdata_k562_on_gm12878_100kb,
                                                  mccdata_gm12878_on_gm12878_10kb,
                                                  mccdata_gm12878_on_gm12878_25kb,
                                                  mccdata_gm12878_on_gm12878_50kb,
                                                  mccdata_gm12878_on_gm12878_100kb,
                                                  mccdata_k562_on_k562_10kb,
                                                  mccdata_k562_on_k562_25kb,
                                                  mccdata_k562_on_k562_50kb,
                                                  mccdata_k562_on_k562_100kb)

all_rfperfs_cl_specific_valid$Chromosome <- rep(c("CHR1",
                                                  "CHR2",
                                                  "CHR3",
                                                  "CHR4",
                                                  "CHR5",
                                                  "CHR6",
                                                  "CHR7",
                                                  "CHR8",
                                                  "CHR10",
                                                  "CHR11",
                                                  "CHR12",
                                                  "CHR13",
                                                  "CHR14",
                                                  "CHR15",
                                                  "CHR16",
                                                  "CHR17",
                                                  "CHR18",
                                                  "CHR19",
                                                  "CHR20",
                                                  "CHR21",
                                                  "CHR22"),16)
all_rfperfs_cl_specific_valid$Resolution <- rep(c(rep("10 kb", 21),
                                              rep("25 kb", 21),
                                              rep("50 kb", 21),
                                              rep("100 kb", 21)), 4)
all_rfperfs_cl_specific_valid$Validation <- c(rep("GM12878 on K562", 21*4),
                                              rep("K562 on GM12878", 21*4),
                                              rep("GM12878 on GM12878", 21*4),
                                              rep("K562 on K562", 21*4))

all_rfperfs_cl_specific_valid$Validation <- factor(all_rfperfs_cl_specific_valid$Validation, levels = c("GM12878 on GM12878", "K562 on K562", "GM12878 on K562", "K562 on GM12878"))
all_rfperfs_cl_specific_valid$Resolution <- factor(all_rfperfs_cl_specific_valid$Resolution, levels = c("10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs_cl_specific_valid$Chromosome <- factor(all_rfperfs_cl_specific_valid$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_cl_specific_valid_mcc <- all_rfperfs_cl_specific_valid %>%
  dplyr::group_by(Validation, Resolution) %>%
  dplyr::summarise(MCC = mean(V1, na.rm = TRUE),
            MCCSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_cl_specific_valid_mcc, aes(x=Resolution, y=MCC, group=Validation, fill=Validation)) +
  geom_bar(stat="identity", position = position_dodge()) +
  geom_errorbar(aes(ymin=MCC-MCCSD, ymax=MCC+MCCSD), width=.5,
                 position=position_dodge(.9), alpha=0.75) +
  theme_minimal() +
  theme_bw()+
  #scale_fill_discrete(name="Cell Line\nSpecific\nValidation")+
  scale_fill_manual(name="Cell Line\nSpecific\nValidation", 
                    values = c("GM12878 on GM12878" = "red", 
                               "K562 on K562" = "blue", 
                               "GM12878 on K562" = "pink",
                               "K562 on GM12878" = "purple")) +
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                  angle = 45, 
                                  #margin = ggplot2::margin(t = 35),
                                  hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position = "bottom")
```