### 2019-08-17 Scripts/visualizing_example_predicted_tad_bounds.Rmd

---
title: "Visualizing Example of Clustered Predicted TAD Boundary Regions"
author: "Spiro Stilianoudakis"
date: "August 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(scales)
library(GGally)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggpubr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)
library(factoextra)
```

# Steps
##1.	Build chromosome specific test data
###a.	Bin chromosome specific data at bp resolution (i.e. 1 row -> 1 bp)
###b.	Annotate data matrix with distances to top TFBS
##2.	Validate best performing model on test data
##3.	Extract probability vector
##4.	Determine threshold for identifying predicted TAD boundary
###a.	Consider the base pairs where the probability vector is given to be 1 
###b.	Consider the midpoints of all groups of base pairs with probability 1
###c.	Obtain distance matrix of all pairwise distances between midpoints
###d.	Perform PAM clustering on the distance matrix
###i.	Obtain optimal number of clusters using silhouette method
###e.	Assign predicted TAD boundary as the new midpoint (bp) of clustered distances

# GM12878

## CHR 22

###17390000-18640000

```{r}
called_and_pred_bound_regions <- readRDS("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/CHR22/smote/distance/enet/called_and_pred_bound_regions.rds")

predprob_df_1 <- called_and_pred_bound_regions[[1]]
predprob_df <- called_and_pred_bound_regions[[2]]
trueBound_gr_region <- called_and_pred_bound_regions[[3]]
predBound_gr <- called_and_pred_bound_regions[[4]]

predprob_df$baseNum[which(predprob_df$trueBound=="Yes")][1:10]
#17390000 17540000 17550000 17650000 17820000 18010000 18030000 18210000 18280000 18640000

predprob_df_temp <- predprob_df[which(predprob_df$baseNum>=17390000 & predprob_df$baseNum<=18640000),]
table(predprob_df_temp$predBound)

dat1 <- predprob_df_temp[which(predprob_df_temp$PredProb==1),]

ggplot() + 
  geom_line(data=predprob_df_temp, aes(x=baseNum, y=PredProb), 
              color='black',
              size=.5)+
  geom_vline(data=dat1, aes(xintercept=baseNum, 
             color=as.factor(PredProb)),
             size=.75,
             show.legend = TRUE)+
  xlab("Base Pair") +
  ylab("Predicted Probability") +
  scale_color_manual(name = '', 
         values =c("red"),
         labels = c('Predicted Region'))+
  theme_minimal() +
  theme_bw() +
  theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")

x <- dist(predprob_df_temp$baseNum[which(predprob_df_temp$predBound=="Yes")], method = "euclidean")

xlabels <- c("1", "2", "3", "9", "10", "4", "5", "8", "6", "7")
fviz_dist(x, gradient = list(low = "white", mid = "pink", high = muted("red")))+
  xlab("Predicted Boundary Regions")+
  ylab("Predicted Boundary Regions")+
  #scale_fill_continuous("Distance")+
  scale_fill_gradient2("Distance\nBetween\nMidpoints",
                       low = "white", 
                       mid = "pink",
                       midpoint = 450000,
                       high = muted("red"))+
  scale_x_discrete(labels=xlabels)+
  scale_y_discrete(labels=xlabels)+
  theme_minimal()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        #panel.grid.major = element_blank(),
        #panel.grid.minor = element_blank(),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20))

hc1 <- hclust(x, method = "complete", size=2)
plot(hc1)

z <- sapply(2:(dim(as.matrix(x))[1]-1), function(i) { 
    mean(silhouette(cutree(hc1, i), dist=x)[,"sil_width"]) })
plot(2:(dim(as.matrix(x))[1]-1), z)

k = which.max(z)+1
c <- clara(predprob_df_temp$baseNum[which(predprob_df_temp$predBound=="Yes")], k=k, metric = "euclidean",
           stand = FALSE, 
           trace = 2, medoids.x = TRUE)
clustdf <- data.frame(mid=predprob_df_temp$baseNum[which(predprob_df_temp$predBound=="Yes")], clust=c$clustering, medoid=ifelse(predprob_df_temp$baseNum[which(predprob_df_temp$predBound=="Yes")] %in% c$medoids, "Yes", "No"))

dend <- as.dendrogram(hc1)
dend %>% set("labels_col", c("red","green","red","red","green","red","green","red","red","green")) %>% set("labels_cex", 2) %>% plot(main = "Clustered Distances")
rect.hclust(hc1, k = k)

#prob1intervals <- unique(clustdf$clust)
#  mid2 <- numeric()
#  for(i in prob1intervals){
#    dat <- clustdf[which(clustdf$clust==prob1intervals[i]),]
#    n=dim(dat)[1]
#    mid2[i] <- ceiling((dat$mid[1]+dat$mid[n])/2)
#  }

  predprob_df_temp$predBound2 <- ifelse(predprob_df_temp$baseNum %in% c$medoids, "Yes", "No")

  dat2 <- predprob_df_temp[which(predprob_df_temp$predBound2=="Yes"),]

ggplot(predprob_df_temp, aes(x=baseNum, y=PredProb)) +
     geom_line(size=.5)+
     #geom_smooth(method = "loess", size = 5, color="blue")
	   geom_vline(data=dat1, aes(xintercept=baseNum, 
             color=as.factor(PredProb)),
             size=.75,
             show.legend = TRUE)+
     geom_vline(data=dat2, aes(xintercept=baseNum, 
             color=predBound2),
             size=.75,
             show.legend = TRUE)+
     scale_color_manual(name = '', 
         values =c("red", "green"),
         labels = c('Predicted Region',
                    'Predicted Boundary'))+
     xlab("Base Pair") +
     ylab("Predicted Probability") +
     theme_minimal() +
     theme_bw() +
     theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")

#dat3 <- predprob_df_temp[which(predprob_df_temp$predBound2=="Yes"),]
#dat3$predBound2 <- "T"

#ggplot() + 
#  geom_line(data=predprob_df_temp, aes(x=baseNum, y=PredProb), 
#              color='black',
#              size=.5)+
#  geom_vline(data=dat1, aes(xintercept=baseNum, 
#             color=as.factor(PredProb)),
#             size=.75,
#             show.legend = TRUE)+
#  geom_vline(data=dat2, aes(xintercept=baseNum, 
#             color=predBound),
#             size=.75,
#             show.legend = TRUE)+
#  geom_vline(data=dat3, aes(xintercept=baseNum, 
#             color=predBound2),
#             size=.75,
#             show.legend = TRUE)+
#  xlab("Base Pair") +
#  ylab("Predicted Probability") +
#  scale_color_manual(name = '', 
#         values =c("red", "blue", "green"),
#         labels = c('Predicted Region',
#                    'New Predicted Boundary',
#                    'Midpoint'))+
#  theme_minimal() +
#  theme_bw() +
#  theme(axis.text.x = element_text(size=15),
#        axis.text.y = element_text(size = 15),
#        axis.title.x = element_text(size = 20),
#        axis.title.y = element_text(size = 20),
#        #strip.text.x = element_text(size = 15),
#        #panel.spacing = unit(2, "lines"),
#        legend.text=element_text(size=15),
#        legend.title=element_text(size=20),
#        plot.title = element_text(size=20),
#        legend.position = "bottom")

```