### 2019-08-23 Scripts/comparing_predicted_boundaries_vs_called_boundaries.Rmd
edits

---
title: "Comparing Predicted Boundaries vs Called Boundaries"
author: "Spiro Stilianoudakis"
date: "August 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(scales)
library(GGally)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggpubr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)
```

# GM12878

## Create granges from top tfbs

```{r}
source("C:/Users/stili/Documents/TAD_predictive_modeling/Scripts/functions_for_R_package/annots_to_granges_func.r")
#file_path="Z:/TAD_data_analysis/annotations/all_common_annotations/gm12878/chr22_tfbs_gm12878"
file_path="Z:/TAD_data_analysis/annotations/all_common_annotations/gm12878/topTFBS"
annots.GR <- annots_to_granges_func(filepath=file_path)

```

## Overlap between Called Regions and Predicted Boundaroes

```{r}

pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
chrs <- paste0("CHR", c(2:8,10:15,17:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                    chrs[i],
                                    "/smote/distance/enet/called_and_pred_bounds.rds"))

  true_bound_list[[i]] <- called_and_pred[[1]]
  pred_bound_list[[i]] <- called_and_pred[[2]]

}

all_true_bounds <- do.call("c", true_bound_list)
all_pred_bounds <- do.call("c", pred_bound_list)

length(all_true_bounds) #12388
length(all_pred_bounds) #7708

all_pred_bounds_r <- flank(all_pred_bounds, width=10000, both=TRUE)

venn_cnt2venn <- function(venn_cnt){
  n <- which(colnames(venn_cnt)=="Counts") - 1
  SetNames=colnames(venn_cnt)[1:n]
  Weight=venn_cnt[,"Counts"]
  names(Weight) <- apply(venn_cnt[,1:n], 1, paste, collapse="")
  Venn(SetNames=SetNames, Weight=Weight)
}

res <- makeVennDiagram(Peaks=list(all_true_bounds, all_pred_bounds_r),
                       NameOfPeaks=c("Called", "Predicted"),
                       connectedPeaks = "keepAll")
venn_gm12878 <- venn_cnt2venn(res$vennCounts)
venn_gm12878 <- compute.Venn(venn_gm12878)
gp <- VennThemes(venn_gm12878)
gp[["Face"]][["11"]]$fill <-  "purple"
gp[["Face"]][["01"]]$fill <-  "lightblue"
gp[["Face"]][["10"]]$fill <-  "red"
gp[["FaceText"]][["10"]]$cex <- 2
gp[["FaceText"]][["11"]]$cex <- 2
gp[["FaceText"]][["01"]]$cex <- 2
gp[["SetText"]][["Set1"]]$cex <- 2
gp[["SetText"]][["Set2"]]$cex <- 2
gridExtra::grid.arrange(grid::grid.grabExpr(plot(venn_gm12878, 
                                                 gp = gp, 
                                                 show=list(Universe=FALSE))), 
                        top=textGrob("Boundary Region", 
                                     gp=gpar(fontsize=50)))

```

## Enriched Heatmaps

```{r}
annots.GR.f <- lapply(annots.GR, function(x){x[which(as.character(seqnames(x)) %in% tolower(chrs))]})

all_true_bounds_r <- resize(all_true_bounds, width=1,fix = "center")

e = 10000
axis_names <- c("-10kb", "Center", "10kb")
```

### CTCF

#### Predicted

```{r}

mat_pred_ctcf = normalizeToMatrix(annots.GR.f[[1]], 
                         all_pred_bounds, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_pred_ctcf, 
                col = c("black",colorRampPalette(c("white", "pink", "red"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Predicted", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col ="red", 
                                                                                   lwd = 2), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

#### Called

```{r}

mat_called_ctcf = normalizeToMatrix(annots.GR.f[[1]], 
                         all_true_bounds_r, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_called_ctcf, 
                col = c("black",colorRampPalette(c("white", "pink", "red"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Called", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "red", 
                                                                                   lwd = 2,
                                                                                   lty=3), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

### RAD21

#### Predicted

```{r}

mat_pred_rad21 = normalizeToMatrix(annots.GR.f[[2]], 
                         all_pred_bounds, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_pred_rad21, 
                col = c("black",colorRampPalette(c("white", "pink", "red"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Predicted", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "red", 
                                                                                   lwd = 2), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

#### Called

```{r}

mat_called_rad21 = normalizeToMatrix(annots.GR.f[[2]], 
                         all_true_bounds_r, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_called_rad21, 
                col = c("black",colorRampPalette(c("white", "pink", "red"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Called", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "red", 
                                                                                   lwd = 2,
                                                                                   lty = 3), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

### SMC3

#### Predicted

```{r}

mat_pred_smc3 = normalizeToMatrix(annots.GR.f[[3]], 
                         all_pred_bounds, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_pred_smc3, 
                col = c("black",colorRampPalette(c("white", "pink", "red"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Predicted", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "red", 
                                                                                   lwd = 2), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

#### Called

```{r}

mat_called_smc3 = normalizeToMatrix(annots.GR.f[[3]], 
                         all_true_bounds_r, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_called_smc3, 
                col = c("black",colorRampPalette(c("white", "pink", "red"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Called", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "red", 
                                                                                   lwd = 2,
                                                                                   lty = 3), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

### ZNF143

#### Predicted

```{r}

mat_pred_znf143 = normalizeToMatrix(annots.GR.f[[4]], 
                         all_pred_bounds, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_pred_znf143, 
                col = c("black",colorRampPalette(c("white", "pink", "red"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Predicted", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "red", 
                                                                                   lwd = 2), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

#### Called

```{r}

mat_called_znf143 = normalizeToMatrix(annots.GR.f[[4]], 
                         all_true_bounds_r, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_called_znf143, 
                col = c("black",colorRampPalette(c("white", "pink", "red"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Called", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "red", 
                                                                                   lwd = 2,
                                                                                   lty = 3), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

## Distance Plots

<!--
### Predicted

```{r}
chromslength <- numeric()
for(i in 1:length(chrs)){
  chromslength[i] <- max(end(all_pred_bounds[which(as.character(seqnames(all_pred_bounds))==tolower(chrs[i]))]),
                         end(annots.GR.f[[1]][which(as.character(seqnames(annots.GR.f[[1]]))==tolower(chrs[i]))]))
}
names(chromslength) <- tolower(chrs)

p_to_ctcf <- GenometriCorrelation(all_pred_bounds,
                                     annots.GR.f[[1]], 
                                     chromosomes.length = chromslength,
                                     chromosomes.to.proceed = tolower(chrs), 
                                     permut.number = 100,
                                     ecdf.area.permut.number = 100,
                                     mean.distance.permut.number = 100,
                                     jaccard.measure.permut.number = 100,
                                     keep.distributions = TRUE, 
                                     showProgressBar = TRUE)

chromslength <- numeric()
for(i in 1:length(chrs)){
  chromslength[i] <- max(end(all_pred_bounds[which(as.character(seqnames(all_pred_bounds))==tolower(chrs[i]))]),
                         end(annots.GR.f[[2]][which(as.character(seqnames(annots.GR.f[[2]]))==tolower(chrs[i]))]))
}
names(chromslength) <- tolower(chrs)

p_to_rad21 <- GenometriCorrelation(all_pred_bounds,
                                      annots.GR.f[[2]], 
                                      chromosomes.length = chromslength,
                                      chromosomes.to.proceed = tolower(chrs),
                                      permut.number = 100,
                                      ecdf.area.permut.number = 100,
                                      mean.distance.permut.number = 100,
                                      jaccard.measure.permut.number = 100,
                                      keep.distributions = TRUE, 
                                      showProgressBar = TRUE)

chromslength <- numeric()
for(i in 1:length(chrs)){
  chromslength[i] <- max(end(all_pred_bounds[which(as.character(seqnames(all_pred_bounds))==tolower(chrs[i]))]),
                         end(annots.GR.f[[3]][which(as.character(seqnames(annots.GR.f[[3]]))==tolower(chrs[i]))]))
}
names(chromslength) <- tolower(chrs)

p_to_smc3a <- GenometriCorrelation(all_pred_bounds,
                                      annots.GR.f[[3]], 
                                      chromosomes.length = chromslength,
                                      chromosomes.to.proceed = tolower(chrs),
                                      permut.number = 100,
                                      ecdf.area.permut.number = 100,
                                      mean.distance.permut.number = 100,
                                      jaccard.measure.permut.number = 100,
                                      keep.distributions = TRUE, 
                                      showProgressBar = TRUE)

chromslength <- numeric()
for(i in 1:length(chrs)){
  chromslength[i] <- max(end(all_pred_bounds[which(as.character(seqnames(all_pred_bounds))==tolower(chrs[i]))]),
                         end(annots.GR.f[[4]][which(as.character(seqnames(annots.GR.f[[4]]))==tolower(chrs[i]))]))
}
names(chromslength) <- tolower(chrs)

p_to_znf143 <- GenometriCorrelation(all_pred_bounds,
                                       annots.GR.f[[4]], 
                                       chromosomes.length = chromslength,
                                       chromosomes.to.proceed = tolower(chrs), 
                                       permut.number = 100,
                                       ecdf.area.permut.number = 100,
                                       mean.distance.permut.number = 100,
                                       jaccard.measure.permut.number = 100,
                                       keep.distributions = TRUE, 
                                       showProgressBar = TRUE)

```

### Called

```{r}
all_true_bounds <- resize(all_true_bounds, width=1,fix = "center")

chromslength <- numeric()
for(i in 1:length(chrs)){
  chromslength[i] <- max(end(all_true_bounds[which(as.character(seqnames(all_true_bounds))==tolower(chrs[i]))]),
                         end(annots.GR.f[[1]][which(as.character(seqnames(annots.GR.f[[1]]))==tolower(chrs[i]))]))
}
names(chromslength) <- tolower(chrs)

t_to_ctcf <- GenometriCorrelation(all_true_bounds,
                                     annots.GR.f[[1]], 
                                     chromosomes.length = chromslength,
                                     chromosomes.to.proceed = tolower(chrs), 
                                     permut.number = 100,
                                     ecdf.area.permut.number = 100,
                                     mean.distance.permut.number = 100,
                                     jaccard.measure.permut.number = 100,
                                     keep.distributions = TRUE, 
                                     showProgressBar = TRUE)

chromslength <- numeric()
for(i in 1:length(chrs)){
  chromslength[i] <- max(end(all_true_bounds[which(as.character(seqnames(all_true_bounds))==tolower(chrs[i]))]),
                         end(annots.GR.f[[2]][which(as.character(seqnames(annots.GR.f[[2]]))==tolower(chrs[i]))]))
}
names(chromslength) <- tolower(chrs)

t_to_rad21 <- GenometriCorrelation(all_true_bounds,
                                      annots.GR.f[[2]], 
                                      chromosomes.length = chromslength,
                                      chromosomes.to.proceed = tolower(chrs),
                                      permut.number = 100,
                                      ecdf.area.permut.number = 100,
                                      mean.distance.permut.number = 100,
                                      jaccard.measure.permut.number = 100,
                                      keep.distributions = TRUE, 
                                      showProgressBar = TRUE)

chromslength <- numeric()
for(i in 1:length(chrs)){
  chromslength[i] <- max(end(all_true_bounds[which(as.character(seqnames(all_true_bounds))==tolower(chrs[i]))]),
                         end(annots.GR.f[[3]][which(as.character(seqnames(annots.GR.f[[3]]))==tolower(chrs[i]))]))
}
names(chromslength) <- tolower(chrs)

t_to_smc3a <- GenometriCorrelation(all_true_bounds,
                                      annots.GR.f[[3]], 
                                      chromosomes.length = chromslength,
                                      chromosomes.to.proceed = tolower(chrs),
                                      permut.number = 100,
                                      ecdf.area.permut.number = 100,
                                      mean.distance.permut.number = 100,
                                      jaccard.measure.permut.number = 100,
                                      keep.distributions = TRUE, 
                                      showProgressBar = TRUE)

chromslength <- numeric()
for(i in 1:length(chrs)){
  chromslength[i] <- max(end(all_true_bounds[which(as.character(seqnames(all_true_bounds))==tolower(chrs[i]))]),
                         end(annots.GR.f[[4]][which(as.character(seqnames(annots.GR.f[[4]]))==tolower(chrs[i]))]))
}
names(chromslength) <- tolower(chrs)

t_to_znf143 <- GenometriCorrelation(all_true_bounds,
                                       annots.GR.f[[4]], 
                                       chromosomes.length = chromslength,
                                       chromosomes.to.proceed = tolower(chrs), 
                                       permut.number = 100,
                                       ecdf.area.permut.number = 100,
                                       mean.distance.permut.number = 100,
                                       jaccard.measure.permut.number = 100,
                                       keep.distributions = TRUE, 
                                       showProgressBar = TRUE)

```

### Plots

```{r}
pred_v_called_10_df <- data.frame(AbsoluteDist = c(#Predicted
  p_to_ctcf@.Data[[1]]$absolute.min.distance.data,
  p_to_rad21@.Data[[1]]$absolute.min.distance.data,
  p_to_smc3a@.Data[[1]]$absolute.min.distance.data,
  p_to_znf143@.Data[[1]]$absolute.min.distance.data,

  #Called
  t_to_ctcf@.Data[[1]]$absolute.min.distance.data,
  t_to_rad21@.Data[[1]]$absolute.min.distance.data,
  t_to_smc3a@.Data[[1]]$absolute.min.distance.data,
  t_to_znf143@.Data[[1]]$absolute.min.distance.data
),
Annotation = c(#Predicted
    rep("CTCF",length(p_to_ctcf@.Data[[1]]$absolute.min.distance.data)),
    rep("RAD21",length(p_to_rad21@.Data[[1]]$absolute.min.distance.data)),
    rep("SMC3a",length(p_to_smc3a@.Data[[1]]$absolute.min.distance.data)),
    rep("ZNF143",length(p_to_znf143@.Data[[1]]$absolute.min.distance.data)),

    #Called
    rep("CTCF",length(t_to_ctcf@.Data[[1]]$absolute.min.distance.data)),
    rep("RAD21",length(t_to_rad21@.Data[[1]]$absolute.min.distance.data)),
    rep("SMC3a",length(t_to_smc3a@.Data[[1]]$absolute.min.distance.data)),
    rep("ZNF143",length(t_to_znf143@.Data[[1]]$absolute.min.distance.data))),
BoundReg = c(#Predicted
    rep("Predicted", length(p_to_ctcf@.Data[[1]]$absolute.min.distance.data)),
    rep("Predicted", length(p_to_rad21@.Data[[1]]$absolute.min.distance.data)),
    rep("Predicted", length(p_to_smc3a@.Data[[1]]$absolute.min.distance.data)),
    rep("Predicted", length(p_to_znf143@.Data[[1]]$absolute.min.distance.data)),

    #Called
    rep("Called", length(t_to_ctcf@.Data[[1]]$absolute.min.distance.data)),
    rep("Called", length(t_to_rad21@.Data[[1]]$absolute.min.distance.data)),
    rep("Called", length(t_to_smc3a@.Data[[1]]$absolute.min.distance.data)),
    rep("Called", length(t_to_znf143@.Data[[1]]$absolute.min.distance.data))))

pred_v_called_10_df$BoundReg <- factor(pred_v_called_10_df$BoundReg, levels=c("Predicted", "Called"))
pred_v_called_10_df$log2AbsoluteDist <- log(pred_v_called_10_df$AbsoluteDist + 1, base = 2)

ggplot(pred_v_called_10_df, aes(x=BoundReg, y = log2AbsoluteDist, fill=BoundReg))  + 
  facet_grid(~ Annotation) +    
  stat_boxplot(geom ='errorbar', width = 0.2) + 
  geom_boxplot(outlier.shape = NA) +
  geom_signif(test = "wilcox.test", 
              comparisons = list( c("Predicted", "Called")),
              vjust = 0,
              textsize = 4,
              size = .5,
              step_increase = .08) +
  theme_minimal()+
  theme_bw()+
  ylab("Log2 Distance to Annotation")+
  xlab("") +
  scale_fill_discrete(name="Boundary Region")+
  #guides(color=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45,
                                   hjust = 1),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")

```
-->

### Plots v2

```{r}

pred_v_called_10_df <- data.frame(LogDist = c(#Predicted
    log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1),
    log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[2]]))$distance+1),
    log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[3]]))$distance+1),
    log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[4]]))$distance+1),

    #Called
    log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1),
    log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[2]]))$distance+1),
    log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[3]]))$distance+1),
    log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[4]]))$distance+1)
),
Annotation = c(#Predicted
    rep("CTCF",length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("RAD21",length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("SMC3",length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("ZNF143",length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),

    #Called
    rep("CTCF",length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("RAD21",length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("SMC3",length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("ZNF143",length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1)))),
BoundReg = c(#Predicted
    rep("Predicted", length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("Predicted", length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("Predicted", length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("Predicted", length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),

    #Called
    rep("Called", length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("Called", length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("Called", length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("Called", length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1)))))

pred_v_called_10_df$BoundReg <- factor(pred_v_called_10_df$BoundReg, levels=c("Predicted", "Called"))

ggplot(pred_v_called_10_df, aes(x=BoundReg, y = LogDist, fill=BoundReg, color=BoundReg))  + 
    facet_grid(~ Annotation) +    
    stat_boxplot(geom ='errorbar', width = 0.2) + 
    geom_boxplot(outlier.shape = NA) +
    geom_signif(test = "wilcox.test", 
                comparisons = list( c("Predicted", "Called")),
                vjust = 0,
                textsize = 4,
                size = .5,
                step_increase = .08,
                color="black") +
    theme_minimal()+
    theme_bw()+
    ylab("Log2 Distance to Annotation")+
    xlab("") +
    scale_fill_manual(name="Boundary Region",
                      values=c("red",
                               "blue"))+
    scale_color_manual(values=c("red",
                                "blue")) +
    guides(color=FALSE, linetype=FALSE)+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45,
                                     hjust = 1),
          axis.text.y = element_text(size = 15),
          axis.title.x = element_text(size = 20),
          axis.title.y = element_text(size = 20),
          strip.text.x = element_text(size = 15),
          #panel.spacing = unit(2, "lines"),
          legend.text=element_text(size=15),
          legend.title=element_text(size=20),
          plot.title = element_text(size=20),
          legend.position = "bottom")
```

# K562

## Create granges from top tfbs

```{r}
source("C:/Users/stili/Documents/TAD_predictive_modeling/Scripts/functions_for_R_package/annots_to_granges_func.r")
#file_path="Z:/TAD_data_analysis/annotations/all_common_annotations/k562/chr22_tfbs_gm12878"
file_path="Z:/TAD_data_analysis/annotations/all_common_annotations/k562/topTFBS"
annots.GR <- annots_to_granges_func(filepath=file_path)

```

## Overlap between Called Regions and Predicted Boundaroes

```{r}

pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
chrs <- paste0("CHR", c(2:8,10:15,17:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                    chrs[i],
                                    "/smote/distance/enet/called_and_pred_bounds.rds"))

  true_bound_list[[i]] <- called_and_pred[[1]]
  pred_bound_list[[i]] <- called_and_pred[[2]]

}

all_true_bounds <- do.call("c", true_bound_list)
all_pred_bounds <- do.call("c", pred_bound_list)

length(all_true_bounds) #9301
length(all_pred_bounds) #9656

all_pred_bounds_r <- flank(all_pred_bounds, width=10000, both=TRUE)

venn_cnt2venn <- function(venn_cnt){
  n <- which(colnames(venn_cnt)=="Counts") - 1
  SetNames=colnames(venn_cnt)[1:n]
  Weight=venn_cnt[,"Counts"]
  names(Weight) <- apply(venn_cnt[,1:n], 1, paste, collapse="")
  Venn(SetNames=SetNames, Weight=Weight)
}

res <- makeVennDiagram(Peaks=list(all_true_bounds, all_pred_bounds_r),
                       NameOfPeaks=c("Called", "Predicted"),
                       connectedPeaks = "keepAll")
venn_k562 <- venn_cnt2venn(res$vennCounts)
venn_k562 <- compute.Venn(venn_k562)
gp <- VennThemes(venn_k562)
gp[["Face"]][["11"]]$fill <-  "purple"
gp[["Face"]][["01"]]$fill <-  "lightblue"
gp[["Face"]][["10"]]$fill <-  "red"
gp[["FaceText"]][["10"]]$cex <- 2
gp[["FaceText"]][["11"]]$cex <- 2
gp[["FaceText"]][["01"]]$cex <- 2
gp[["SetText"]][["Set1"]]$cex <- 2
gp[["SetText"]][["Set2"]]$cex <- 2
gridExtra::grid.arrange(grid::grid.grabExpr(plot(venn_k562, 
                                                 gp = gp, 
                                                 show=list(Universe=FALSE))), 
                        top=textGrob("Boundary Region", 
                                     gp=gpar(fontsize=50)))

```

## Enriched Heatmaps

```{r}
annots.GR.f <- lapply(annots.GR, function(x){x[which(as.character(seqnames(x)) %in% tolower(chrs))]})

all_true_bounds_r <- resize(all_true_bounds, width=1,fix = "center")

e = 10000
axis_names <- c("-10kb", "Center", "10kb")
```

### CTCF

#### Predicted

```{r}

mat_pred_ctcf = normalizeToMatrix(annots.GR.f[[1]], 
                         all_pred_bounds, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_pred_ctcf, 
                col = c("black",colorRampPalette(c("white", "lightblue", "blue"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Predicted", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col ="blue", 
                                                                                   lwd = 2), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

#### Called

```{r}

mat_called_ctcf = normalizeToMatrix(annots.GR.f[[1]], 
                         all_true_bounds_r, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_called_ctcf, 
                col = c("black",colorRampPalette(c("white", "lightblue", "blue"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Called", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "blue", 
                                                                                   lwd = 2,
                                                                                   lty=3), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

### RAD21

#### Predicted

```{r}

mat_pred_rad21 = normalizeToMatrix(annots.GR.f[[2]], 
                         all_pred_bounds, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_pred_rad21, 
                col = c("black",colorRampPalette(c("white", "lightblue", "blue"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Predicted", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "blue", 
                                                                                   lwd = 2), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

#### Called

```{r}

mat_called_rad21 = normalizeToMatrix(annots.GR.f[[2]], 
                         all_true_bounds_r, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_called_rad21, 
                col = c("black",colorRampPalette(c("white", "lightblue", "blue"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Called", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "blue", 
                                                                                   lwd = 2,
                                                                                   lty = 3), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

### SMC3

#### Predicted

```{r}

mat_pred_smc3 = normalizeToMatrix(annots.GR.f[[3]], 
                         all_pred_bounds, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_pred_smc3, 
                col = c("black",colorRampPalette(c("white", "lightblue", "blue"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Predicted", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "blue", 
                                                                                   lwd = 2), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

#### Called

```{r}

mat_called_smc3 = normalizeToMatrix(annots.GR.f[[3]], 
                         all_true_bounds_r, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_called_smc3, 
                col = c("black",colorRampPalette(c("white", "lightblue", "blue"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Called", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "blue", 
                                                                                   lwd = 2,
                                                                                   lty = 3), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

### ZNF143

#### Predicted

```{r}

mat_pred_znf143 = normalizeToMatrix(annots.GR.f[[4]], 
                         all_pred_bounds, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_pred_znf143, 
                col = c("black",colorRampPalette(c("white", "lightblue", "blue"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Predicted", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "blue", 
                                                                                   lwd = 2), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

#### Called

```{r}

mat_called_znf143 = normalizeToMatrix(annots.GR.f[[4]], 
                         all_true_bounds_r, 
                         value_column = "coverage", 
                         extend = e, 
                         mean_mode = "w0",
                         keep = c(0,.99))

EnrichedHeatmap(mat_called_znf143, 
                col = c("black",colorRampPalette(c("white", "lightblue", "blue"))(99)), 
                #heatmap_legend_param = list(show_heatmap_legend = FALSE),
                row_title="Chromosomes",show_heatmap_legend = FALSE,
                row_title_side="left", 
                column_title="Called", 
                axis_name = axis_names,
                top_annotation = HeatmapAnnotation(lines = anno_enriched(gp = gpar(col = "blue", 
                                                                                   lwd = 2,
                                                                                   lty = 3), 
                                                                         yaxis_side = "right", 
                                                                         yaxis_facing = "left")))
```

### Plots v2

```{r}

pred_v_called_10_df <- data.frame(LogDist = c(#Predicted
    log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1),
    log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[2]]))$distance+1),
    log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[3]]))$distance+1),
    log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[4]]))$distance+1),

    #Called
    log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1),
    log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[2]]))$distance+1),
    log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[3]]))$distance+1),
    log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[4]]))$distance+1)
),
Annotation = c(#Predicted
    rep("CTCF",length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("RAD21",length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("SMC3",length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("ZNF143",length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),

    #Called
    rep("CTCF",length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("RAD21",length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("SMC3",length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("ZNF143",length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1)))),
BoundReg = c(#Predicted
    rep("Predicted", length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("Predicted", length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("Predicted", length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),
    rep("Predicted", length(log2(mcols(distanceToNearest(all_pred_bounds, annots.GR.f[[1]]))$distance+1))),

    #Called
    rep("Called", length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("Called", length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("Called", length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1))),
    rep("Called", length(log2(mcols(distanceToNearest(all_true_bounds_r, annots.GR.f[[1]]))$distance+1)))))

pred_v_called_10_df$BoundReg <- factor(pred_v_called_10_df$BoundReg, levels=c("Predicted", "Called"))

ggplot(pred_v_called_10_df, aes(x=BoundReg, y = LogDist, fill=BoundReg, color=BoundReg))  + 
    facet_grid(~ Annotation) +    
    stat_boxplot(geom ='errorbar', width = 0.2) + 
    geom_boxplot(outlier.shape = NA) +
    geom_signif(test = "wilcox.test", 
                comparisons = list( c("Predicted", "Called")),
                vjust = 0,
                textsize = 4,
                size = .5,
                step_increase = .08,
                color="black") +
    theme_minimal()+
    theme_bw()+
    ylab("Log2 Distance to Annotation")+
    xlab("") +
    scale_fill_manual(name="Boundary Region",
                      values=c("red",
                               "blue"))+
    scale_color_manual(values=c("red",
                                "blue")) +
    guides(color=FALSE, linetype=FALSE)+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45,
                                     hjust = 1),
          axis.text.y = element_text(size = 15),
          axis.title.x = element_text(size = 20),
          axis.title.y = element_text(size = 20),
          strip.text.x = element_text(size = 15),
          #panel.spacing = unit(2, "lines"),
          legend.text=element_text(size=15),
          legend.title=element_text(size=20),
          plot.title = element_text(size=20),
          legend.position = "bottom")
```