### 2019-08-23 Scripts/evaluating_rfe_across_chromosomes.Rmd

---
title: "Evaluating RFE across chromosomes"
author: "Spiro Stilianoudakis"
date: "July 30, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(scales)
library(GGally)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggpubr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
```

# Optimal parameters: 10 kb, TFBS, distance type predictors, SMOTE resampling

## Determining number of annotations

### Line plot

```{r}

rfeperchr <- data.frame()

chrs <- paste0("chr", c(1:8,10:22))
for(i in c("GM12878","K562")){
  for(k in 1:length(chrs)){
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/smote/distance/enet/rfeModelResultsList.rds"))
    rfeModelResults <- rfeModelResultsList[[1]]
    rfeModelResults <- rfeModelResults[,1:2]
    rfeModelResults$CL <- i
    rfeModelResults$Chromosome <- chrs[k]
    rfeperchr <- rbind.data.frame(rfeperchr, rfeModelResults)
  }
}

rfeperchr_summ <- rfeperchr %>%
  group_by(CL, Variables) %>%
  dplyr::summarise(AveMCC = mean(MCC, na.rm = TRUE),
            AveMCCSD = sd(MCC, na.rm = TRUE))

ggplot(rfeperchr_summ, aes(x=Variables, y=AveMCC, color=CL)) + 
    geom_errorbar(aes(ymin=AveMCC-AveMCCSD, ymax=AveMCC+AveMCCSD), 
                  width=2, 
                  position=position_dodge(2),
                  size=1) +
    geom_line(position=position_dodge(2), size=1) +
    geom_point(position=position_dodge(2), size=2) +
    xlab("Number of top ranked annotations") +
    ylab("Cross-Validated MCC") +
    #ylim(.3,.8)+
    scale_color_discrete(name="Cell line")+
    scale_x_continuous(breaks = c(2,4,8,16,32,52),
                     labels = c(2,4,8,16,32,52))+
    theme_minimal() +
    theme_bw()+
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")
```

### Importance heatmap

```{r}
rfeperchr <- data.frame()

set.size = 8
chrs <- paste0("chr", c(1:8,10:22))
for(i in c("GM12878","K562")){
  for(k in 1:length(chrs)){
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/smote/distance/enet/rfeModelResultsList.rds"))
    rfeModelResults <- rfeModelResultsList[[2]]
    rfe.set <- rfeModelResults[rfeModelResults$Variables==set.size,]
    rfe.set <- aggregate(rfe.set[, c("Overall")], list(rfe.set$var), mean)
    rfe.order <- order(rfe.set[, c("x")], decreasing = TRUE)[1:set.size]
    rfe.set <- rfe.set[rfe.order, ]
    rfe.set$CL <- i
    rfe.set$Chromosome <- chrs[k]

    rfeperchr <- rbind.data.frame(rfeperchr, rfe.set)
  }
}

rfeperchr$Group.1 <- gsub(paste0(c("Gm12878-", 
                                   "K562-", 
                                   "-Haib", 
                                   "-Sydh",
                                   "-Uta",
                                   "-Uw",
                                   "-Broad",
                                   "-Uchicago"), collapse = "|"), "", rfeperchr$Group.1)
names(rfeperchr)[1:2] <- c("Feature", "Importance")

rfeperchr_summ <- rfeperchr %>%
  group_by(CL, Feature) %>%
  dplyr::summarise(AveImp = mean(Importance, na.rm = TRUE),
            AveImpSD = sd(Importance, na.rm = TRUE)) %>%
  dplyr::arrange(CL, desc(AveImp))

```

#### GM12878

```{r}

rfegm12878 <- rfeperchr[which(rfeperchr$CL=="GM12878"),]
rfegm12878 <- reshape(rfegm12878[,c(1,2,4)], idvar = "Feature", timevar = "Chromosome", direction = "wide")
names(rfegm12878)[-1] <- paste0("CHR", c(1:8,10:22)) 
rfegm12878.mat <- as.matrix(rfegm12878[,-1])
rownames(rfegm12878.mat) <- rfegm12878$Feature
rfegm12878.mat[is.na(rfegm12878.mat)] <- 0
rfegm12878.mat <- apply(rfegm12878.mat,2,rev)

#using heatmap.2
distance.col = dist(rfegm12878.mat, method = "euclidean")
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "pink", "red"))(ncol(rfegm12878.mat))
heatmap.2(t(rfegm12878.mat),
          #labRow=rownames(rfegm12878.mat),
          #labCol=colnames(rfegm12878.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(12,8),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="column",     # only draw a row dendrogram,
          Rowv = FALSE,
          Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=.5,
          cexCol=1,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          rowsep=0:nrow(t(rfegm12878.mat)),
          colsep=0:ncol(t(rfegm12878.mat)),
          sepwidth=c(0.001,0.001),
          sepcolor="black"
) 
```

#### K562

```{r}
rfek562 <- rfeperchr[which(rfeperchr$CL=="K562"),]
rfek562 <- reshape(rfek562[,c(1,2,4)], idvar = "Feature", timevar = "Chromosome", direction = "wide")
names(rfek562)[-1] <- paste0("CHR", c(1:8,10:22)) 
rfek562.mat <- as.matrix(rfek562[,-1])
rownames(rfek562.mat) <- rfek562$Feature
rfek562.mat[is.na(rfek562.mat)] <- 0
#rfek562.mat <- apply(rfek562.mat,2,rev)

#using heatmap.2
distance.col = dist(rfek562.mat, method = "euclidean")
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "lightblue", "blue"))(ncol(rfek562.mat))
heatmap.2(t(rfek562.mat),
          #labRow=rownames(rfek562.mat),
          #labCol=colnames(rfek562.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(12,8),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="column",     # only draw a row dendrogram,
          Rowv = FALSE,
          Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=.5,
          cexCol=1,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          rowsep=0:nrow(t(rfek562.mat)),
          colsep=0:ncol(t(rfek562.mat)),
          sepwidth=c(0.001,0.001),
          sepcolor="black"
) 
```