### 2019-08-28 Scripts/evaluating_jaccard_indices.Rmd
initial commit

---
title: "Jaccard Index Heatmaps for TFBS"
author: "Spiro Stilianoudakis"
date: "August 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
library(GenomicRanges)
library(caret)
library(data.table)
library(gbm)
library(randomForest)
library(glmnet)
library(pROC)
library(plyr)
library(dplyr)
library(ggplot2)
library(DMwR)
library(gridExtra)
library(pROC)
library(ROCR)
#library(leaps)
library(cluster)
library(scales)
library(GGally)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggpubr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)
```

# GM12878

```{r}
jaccard.mat.gm12878_tfbs <-  readRDS("Z:/TAD_data_analysis/jaccard/jaccard.mat.gm12878_tfbs.rds")

diag(jaccard.mat.gm12878_tfbs) <- NA

distance.col = as.dist(jaccard.mat.gm12878_tfbs)
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "pink", "red"))(ncol(jaccard.mat.gm12878_tfbs))
heatmap.2(t(jaccard.mat.gm12878_tfbs),
          #labRow=rownames(rfegm12878.mat),
          #labCol=colnames(rfegm12878.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(12,8),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="none",     # only draw a row dendrogram,
          #Rowv = FALSE,
          #Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=1,
          cexCol=1)
```

# K562

```{r}
jaccard.mat.k562_tfbs <-  readRDS("Z:/TAD_data_analysis/jaccard/jaccard.mat.k562_tfbs.rds")

diag(jaccard.mat.k562_tfbs) <- NA

distance.col = as.dist(jaccard.mat.k562_tfbs)
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "pink", "red"))(ncol(jaccard.mat.k562_tfbs))
heatmap.2(t(jaccard.mat.k562_tfbs),
          #labRow=rownames(rfegm12878.mat),
          #labCol=colnames(rfegm12878.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(12,8),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="none",     # only draw a row dendrogram,
          #Rowv = FALSE,
          #Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=1,
          cexCol=1)
```