### 2020-01-09 Manuscript/Supplementary/supplementary_legends.Rmd
edits

---
title: "Legends for supplementary material"
output: pdf_document
---

**Supplementary Table 1. Hi-C data.** A complete list of the Hi-C data libraries obtained from GEO GSE63525 [@rao20143d] used to obtain the locations of topologically associating domains.

**Supplementary Table 2. Genomic annotations.** A complete list of the genomic annotations used to build the feature space for all downstream models. The GRCh37/hg19 human genome assembly was used. "Genomic Class" - broad category of genomic features, "Element" - names of genomic features, "Cell Type-Specific Source" - download URL specific to the cell line (not all annotations were provided by the same institutions).

**Supplementary Table 3.** Domain data and class imbalance summaries across resolutions for GM12878.

**Supplementary Figure 1. Performing $log_{2}$ transformations normalized skewed genomic distance densities across all resolutions.** When measured in basepairs, distances were highly right skewed for all resolutions for both the (A) GM12878 and (B) K562 cell lines. After $log_{2}$ transforming the distances, densities became much more gaussian distributed for all resolutions.