### 2020-01-16 Manuscript/Figures/figure_legends.Rmd

---
title: "Legends for Figures"
author: "Spiro Stilianoudakis"
date: "November 19, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Figure 1. Illustrating the resolution-specific data construction and feature engineering used for downstream modeling.** Unique boundaries are denoted by the blue dots. The linear genome was subsequently binned into resolution-specific intervals. The response vector **Y** was defined as a genomic bin that either overlapped with a TAD boundary ($Y = 1$) or did not ($Y = 0$). Three types of associations between bins and genomic annotations (green shapes) were then used to build the feature space of random forest classification models include _Overlap Counts_ (OC), _Overlap Percent_ (OP), and ($log_{2}$) _Distance_.

**Figure 2. A diagram of the ensemble framework detailing the steps for combining different inputs used in the random forest model building pipeline.** Step 1 is to establish the rowise (genomic bins) and colwise (genomic annotations) dimensions of the data matrix, **A**, by combining a resolution (5 kb, 10 kb, 25 kb, 50 kb, or 100 kb) and feature engineering procedure (overlap count (OC), overlap percent (OP), $log_{2}$ distance (Distance)). Step 2 is to split **A** into a training and testing set according to a 7:3 ratio. Step 3 is to apply a re-sampling technique to the training set (no re-sampling, random over-sampling, random under-sampling, or SMOTE re-sampling). Step 4 is to implement the random forest algorithm and perform 5-fold cross-validation to tune for the *mtry* parameter. Finally, step 5 is to validate the model on the separate test data for unbiased model evaluation.

**Figure 3. Models built using distance-type predictors combined with random under-sampling yielded optimal predictive performance.** Line plots comparing the performances of random forest classifiers built using different combinations of data level characteristics. Performances are compared across resolution, within each predictor-type: (A) OC, (B) OP, and (C) Distance, and between each re-sampling technique: no re-sampling (None; red), random over-sampling (ROS; green), random under-sampling (RUS; blue), and synthetic minority over-sampling (SMOTE; purple). Performances were evaluated using MCCs and aggregated by averaging across each chromosome, with error bars representing 1 standard deviation away from the mean.