### 2020-04-29 Scripts_for_results/02_creating_lineplot_of_balanced_accuracies.Rmd

---
title: "Creating line plot of balanced accuracies"
author: "Spiro Stilianoudakis"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and functions

```{r}
library(S4Vectors)
library(GenomicRanges)
library(ggplot2)
library(caret)
library(randomForest)
library(glmnet)
library(gbm)
library(DMwR)
library(rtracklayer)
library(Biostrings)
library(BSgenome)
library(cluster)
library(pROC)
library(SuperLearner)
library(data.table)
library(PRROC)
library(bigmemory)
library(ranger)
library(dplyr)
library(parallel)
library(doSNOW)
library(foreach)
library(pbapply)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)

source("Z:/TAD_data_analysis/functions_for_R_package/preciseTAD.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADRF.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADrfe.R")
source("Z:/TAD_data_analysis/functions_for_R_package/distance_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/percent_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/count_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/binary_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/signal_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/annots_to_granges_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/extract_boundaries_func.R")

```

## ARROWHEAD

### GM12878

####5kb

```{r}
badata_gm12878_5kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_5kb <- rbind(badata_gm12878_5kb,
                                   perfdata[10,2])
    }
  }
}
```

####10kb

```{r}
badata_gm12878_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_10kb <- rbind(badata_gm12878_10kb,
                                   perfdata[10,2])
    }
  }
}

```

####25kb

```{r}
badata_gm12878_25kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_25kb <- rbind(badata_gm12878_25kb,
                                   perfdata[10,2])
    }
  }
}

```

####50kb

```{r}
badata_gm12878_50kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_50kb <- rbind(badata_gm12878_50kb,
                                   perfdata[10,2])
    }
  }
}

```

####100kb

```{r}
badata_gm12878_100kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_100kb <- rbind(badata_gm12878_100kb,
                                    perfdata[10,2])
    }
  }
}

```

#### Plotting

```{r}
all_rfperfs <- rbind.data.frame(badata_gm12878_5kb,
                                badata_gm12878_10kb,
                                badata_gm12878_25kb,
                                badata_gm12878_50kb,
                                badata_gm12878_100kb)
all_rfperfs$CL <- c(rep("GM12878", length(badata_gm12878_5kb)*5))
all_rfperfs$Resolution <- c(rep("5 kb", length(badata_gm12878_5kb)),
                            rep("10 kb", length(badata_gm12878_10kb)),
                            rep("25 kb", length(badata_gm12878_25kb)),
                            rep("50 kb", length(badata_gm12878_50kb)),
                            rep("100 kb", length(badata_gm12878_100kb)))
all_rfperfs$Resampling <- c(rep("None", 4),
                            rep("ROS", 4),
                            rep("RUS", 4),
                            rep("SMOTE", 4))
all_rfperfs$Predictor <- c("Distance", "OC", "OP", "Signal")
all_rfperfs$Chromosome <- c(rep("CHR1", 16),
                            rep("CHR2", 16),
                            rep("CHR3", 16),
                            rep("CHR4", 16),
                            rep("CHR5", 16),
                            rep("CHR6", 16),
                            rep("CHR7", 16),
                            rep("CHR8", 16),
                            rep("CHR10", 16),
                            rep("CHR11", 16),
                            rep("CHR12", 16),
                            rep("CHR13", 16),
                            rep("CHR14", 16),
                            rep("CHR15", 16),
                            rep("CHR16", 16),
                            rep("CHR17", 16),
                            rep("CHR18", 16),
                            rep("CHR19", 16),
                            rep("CHR20", 16),
                            rep("CHR21", 16),
                            rep("CHR22", 16))

all_rfperfs$Resolution <- factor(all_rfperfs$Resolution, levels = c("5 kb", "10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs$Resampling <- factor(all_rfperfs$Resampling, levels = c("None", "ROS", "RUS", "SMOTE"))
all_rfperfs$Predictor <- factor(all_rfperfs$Predictor, levels = c("Signal", "OC", "OP", "Distance"))
all_rfperfs$Chromosome <- factor(all_rfperfs$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_ba <- all_rfperfs %>%
  dplyr::group_by(Resolution, Resampling, Predictor) %>%
  dplyr::summarise(Performance = mean(V1, na.rm = TRUE),
                   PerformanceSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_ba, aes(x=Resolution, y=Performance, color=Resampling, group=Resampling)) +
  geom_line(size=2, position=position_dodge(0.5))+
  geom_point(size=5, position=position_dodge(0.5))+
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=1,
                position=position_dodge(0.5)) +
  facet_grid(. ~ Predictor)+
  theme_minimal() +
  theme_bw()+
  xlab("Resolution") + 
  ylab("Balanced Accuracy") + #ylim(-.1,.5) +
  scale_color_discrete(name="Resampling Technique")+
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  strip.text.y = element_text(size = 15),
  #panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position="bottom")

```

### K562

####5kb

```{r}
badata_k562_5kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/5kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/5kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_5kb <- rbind(badata_k562_5kb,
                               perfdata[10,2])
    }
  }
}

```

####10kb

```{r}
badata_k562_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_10kb <- rbind(badata_k562_10kb,
                                perfdata[10,2])
    }
  }
}

```

####25kb

```{r}
badata_k562_25kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_25kb <- rbind(badata_k562_25kb,
                                perfdata[10,2])
    }
  }
}

```

####50kb

```{r}
badata_k562_50kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_50kb <- rbind(badata_k562_50kb,
                                perfdata[10,2])
    }
  }
}

```

####100kb

```{r}
badata_k562_100kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_100kb <- rbind(badata_k562_100kb,
                                 perfdata[10,2])
    }
  }
}

```

#### Plotting

```{r}
all_rfperfs <- rbind.data.frame(badata_k562_5kb,
                                badata_k562_10kb,
                                badata_k562_25kb,
                                badata_k562_50kb,
                                badata_k562_100kb)
all_rfperfs$CL <- c(rep("K562", length(badata_k562_5kb)*5))
all_rfperfs$Resolution <- c(rep("5 kb", length(badata_k562_5kb)),
                            rep("10 kb", length(badata_k562_10kb)),
                            rep("25 kb", length(badata_k562_25kb)),
                            rep("50 kb", length(badata_k562_50kb)),
                            rep("100 kb", length(badata_k562_100kb)))
all_rfperfs$Resampling <- c(rep("None", 4),
                            rep("ROS", 4),
                            rep("RUS", 4),
                            rep("SMOTE", 4))
all_rfperfs$Predictor <- c("Distance", "OC", "OP", "Signal")
all_rfperfs$Chromosome <- c(rep("CHR1", 16),
                            rep("CHR2", 16),
                            rep("CHR3", 16),
                            rep("CHR4", 16),
                            rep("CHR5", 16),
                            rep("CHR6", 16),
                            rep("CHR7", 16),
                            rep("CHR8", 16),
                            rep("CHR10", 16),
                            rep("CHR11", 16),
                            rep("CHR12", 16),
                            rep("CHR13", 16),
                            rep("CHR14", 16),
                            rep("CHR15", 16),
                            rep("CHR16", 16),
                            rep("CHR17", 16),
                            rep("CHR18", 16),
                            rep("CHR19", 16),
                            rep("CHR20", 16),
                            rep("CHR21", 16),
                            rep("CHR22", 16))

all_rfperfs$Resolution <- factor(all_rfperfs$Resolution, levels = c("5 kb", "10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs$Resampling <- factor(all_rfperfs$Resampling, levels = c("None", "ROS", "RUS", "SMOTE"))
all_rfperfs$Predictor <- factor(all_rfperfs$Predictor, levels = c("Signal", "OC", "OP", "Distance"))
all_rfperfs$Chromosome <- factor(all_rfperfs$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_ba <- all_rfperfs %>%
  dplyr::group_by(Resolution, Resampling, Predictor) %>%
  dplyr::summarise(Performance = mean(V1, na.rm = TRUE),
                   PerformanceSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_ba, aes(x=Resolution, y=Performance, color=Resampling, group=Resampling)) +
  geom_line(size=2, position=position_dodge(0.5))+
  geom_point(size=5, position=position_dodge(0.5))+
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=1,
                position=position_dodge(0.5)) +
  facet_grid(. ~ Predictor)+
  theme_minimal() +
  theme_bw()+
  xlab("Resolution") + 
  ylab("Balanced Accuracy") + #ylim(-.1,.5) +
  scale_color_discrete(name="Resampling Technique")+
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  strip.text.y = element_text(size = 15),
  #panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position="bottom")

```

## SpectralTAD

### GM12878

####5kb

```{r}
badata_gm12878_5kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_5kb <- rbind(badata_gm12878_5kb,
                                   perfdata[10,2])
    }
  }
}
```

####10kb

```{r}
badata_gm12878_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_10kb <- rbind(badata_gm12878_10kb,
                                   perfdata[10,2])
    }
  }
}

```

####25kb

```{r}
badata_gm12878_25kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_25kb <- rbind(badata_gm12878_25kb,
                                   perfdata[10,2])
    }
  }
}

```

####50kb

```{r}
badata_gm12878_50kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/50kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_50kb <- rbind(badata_gm12878_50kb,
                                   perfdata[10,2])
    }
  }
}

```

####100kb

```{r}
badata_gm12878_100kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/100kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_100kb <- rbind(badata_gm12878_100kb,
                                    perfdata[10,2])
    }
  }
}

```

#### Plotting

```{r}
all_rfperfs <- rbind.data.frame(badata_gm12878_5kb,
                                badata_gm12878_10kb,
                                badata_gm12878_25kb,
                                badata_gm12878_50kb,
                                badata_gm12878_100kb)
all_rfperfs$CL <- c(rep("GM12878", length(badata_gm12878_5kb)*5))
all_rfperfs$Resolution <- c(rep("5 kb", length(badata_gm12878_5kb)),
                            rep("10 kb", length(badata_gm12878_10kb)),
                            rep("25 kb", length(badata_gm12878_25kb)),
                            rep("50 kb", length(badata_gm12878_50kb)),
                            rep("100 kb", length(badata_gm12878_100kb)))
all_rfperfs$Resampling <- c(rep("None", 4),
                            rep("ROS", 4),
                            rep("RUS", 4),
                            rep("SMOTE", 4))
all_rfperfs$Predictor <- c("Distance", "OC", "OP", "Signal")
all_rfperfs$Chromosome <- c(rep("CHR1", 16),
                            rep("CHR2", 16),
                            rep("CHR3", 16),
                            rep("CHR4", 16),
                            rep("CHR5", 16),
                            rep("CHR6", 16),
                            rep("CHR7", 16),
                            rep("CHR8", 16),
                            rep("CHR10", 16),
                            rep("CHR11", 16),
                            rep("CHR12", 16),
                            rep("CHR13", 16),
                            rep("CHR14", 16),
                            rep("CHR15", 16),
                            rep("CHR16", 16),
                            rep("CHR17", 16),
                            rep("CHR18", 16),
                            rep("CHR19", 16),
                            rep("CHR20", 16),
                            rep("CHR21", 16),
                            rep("CHR22", 16))

all_rfperfs$Resolution <- factor(all_rfperfs$Resolution, levels = c("5 kb", "10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs$Resampling <- factor(all_rfperfs$Resampling, levels = c("None", "ROS", "RUS", "SMOTE"))
all_rfperfs$Predictor <- factor(all_rfperfs$Predictor, levels = c("Signal", "OC", "OP", "Distance"))
all_rfperfs$Chromosome <- factor(all_rfperfs$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_ba <- all_rfperfs %>%
  dplyr::group_by(Resolution, Resampling, Predictor) %>%
  dplyr::summarise(Performance = mean(V1, na.rm = TRUE),
                   PerformanceSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_ba, aes(x=Resolution, y=Performance, color=Resampling, group=Resampling)) +
  geom_line(size=2, position=position_dodge(0.5))+
  geom_point(size=5, position=position_dodge(0.5))+
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=1,
                position=position_dodge(0.5)) +
  facet_grid(. ~ Predictor)+
  theme_minimal() +
  theme_bw()+
  xlab("Resolution") + 
  ylab("Balanced Accuracy") + #ylim(-.1,.5) +
  scale_color_discrete(name="Resampling Technique")+
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  strip.text.y = element_text(size = 15),
  #panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position="bottom")

```

### K562

####5kb

```{r}
badata_k562_5kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/5kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/5kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_5kb <- rbind(badata_k562_5kb,
                               perfdata[10,2])
    }
  }
}

```

####10kb

```{r}
badata_k562_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_10kb <- rbind(badata_k562_10kb,
                                perfdata[10,2])
    }
  }
}

```

####25kb

```{r}
badata_k562_25kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_25kb <- rbind(badata_k562_25kb,
                                perfdata[10,2])
    }
  }
}

```

####50kb

```{r}
badata_k562_50kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/50kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_50kb <- rbind(badata_k562_50kb,
                                perfdata[10,2])
    }
  }
}

```

####100kb

```{r}
badata_k562_100kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/100kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_spectralTAD.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_100kb <- rbind(badata_k562_100kb,
                                 perfdata[10,2])
    }
  }
}

```

#### Plotting

```{r}
all_rfperfs <- rbind.data.frame(badata_k562_5kb,
                                badata_k562_10kb,
                                badata_k562_25kb,
                                badata_k562_50kb,
                                badata_k562_100kb)
all_rfperfs$CL <- c(rep("K562", length(badata_k562_5kb)*5))
all_rfperfs$Resolution <- c(rep("5 kb", length(badata_k562_5kb)),
                            rep("10 kb", length(badata_k562_10kb)),
                            rep("25 kb", length(badata_k562_25kb)),
                            rep("50 kb", length(badata_k562_50kb)),
                            rep("100 kb", length(badata_k562_100kb)))
all_rfperfs$Resampling <- c(rep("None", 4),
                            rep("ROS", 4),
                            rep("RUS", 4),
                            rep("SMOTE", 4))
all_rfperfs$Predictor <- c("Distance", "OC", "OP", "Signal")
all_rfperfs$Chromosome <- c(rep("CHR1", 16),
                            rep("CHR2", 16),
                            rep("CHR3", 16),
                            rep("CHR4", 16),
                            rep("CHR5", 16),
                            rep("CHR6", 16),
                            rep("CHR7", 16),
                            rep("CHR8", 16),
                            rep("CHR10", 16),
                            rep("CHR11", 16),
                            rep("CHR12", 16),
                            rep("CHR13", 16),
                            rep("CHR14", 16),
                            rep("CHR15", 16),
                            rep("CHR16", 16),
                            rep("CHR17", 16),
                            rep("CHR18", 16),
                            rep("CHR19", 16),
                            rep("CHR20", 16),
                            rep("CHR21", 16),
                            rep("CHR22", 16))

all_rfperfs$Resolution <- factor(all_rfperfs$Resolution, levels = c("5 kb", "10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs$Resampling <- factor(all_rfperfs$Resampling, levels = c("None", "ROS", "RUS", "SMOTE"))
all_rfperfs$Predictor <- factor(all_rfperfs$Predictor, levels = c("Signal", "OC", "OP", "Distance"))
all_rfperfs$Chromosome <- factor(all_rfperfs$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_ba <- all_rfperfs %>%
  dplyr::group_by(Resolution, Resampling, Predictor) %>%
  dplyr::summarise(Performance = mean(V1, na.rm = TRUE),
                   PerformanceSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_ba, aes(x=Resolution, y=Performance, color=Resampling, group=Resampling)) +
  geom_line(size=2, position=position_dodge(0.5))+
  geom_point(size=5, position=position_dodge(0.5))+
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=1,
                position=position_dodge(0.5)) +
  facet_grid(. ~ Predictor)+
  theme_minimal() +
  theme_bw()+
  xlab("Resolution") + 
  ylab("Balanced Accuracy") + #ylim(-.1,.5) +
  scale_color_discrete(name="Resampling Technique")+
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  strip.text.y = element_text(size = 15),
  #panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position="bottom")

```

## PEAKACHU

### GM12878

#### 10kb

```{r}
badata_gm12878_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_peakachu.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_10kb <- rbind(badata_gm12878_10kb,
                                   perfdata[10,2])
    }
  }
}

```

#### Plotting

```{r}
all_rfperfs <- rbind.data.frame(badata_gm12878_10kb)
all_rfperfs$CL <- rep("GM12878", length(badata_gm12878_5kb))
all_rfperfs$Resolution <- rep("10 kb", length(badata_gm12878_10kb))
all_rfperfs$Resampling <- c(rep("None", 4),
                            rep("ROS", 4),
                            rep("RUS", 4),
                            rep("SMOTE", 4))
all_rfperfs$Predictor <- c("Distance", "OC", "OP", "Signal")
all_rfperfs$Chromosome <- c(rep("CHR1", 16),
                            rep("CHR2", 16),
                            rep("CHR3", 16),
                            rep("CHR4", 16),
                            rep("CHR5", 16),
                            rep("CHR6", 16),
                            rep("CHR7", 16),
                            rep("CHR8", 16),
                            rep("CHR10", 16),
                            rep("CHR11", 16),
                            rep("CHR12", 16),
                            rep("CHR13", 16),
                            rep("CHR14", 16),
                            rep("CHR15", 16),
                            rep("CHR16", 16),
                            rep("CHR17", 16),
                            rep("CHR18", 16),
                            rep("CHR19", 16),
                            rep("CHR20", 16),
                            rep("CHR21", 16),
                            rep("CHR22", 16))

all_rfperfs$Resolution <- factor(all_rfperfs$Resolution, levels = c("5 kb", "10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs$Resampling <- factor(all_rfperfs$Resampling, levels = c("None", "ROS", "RUS", "SMOTE"))
all_rfperfs$Predictor <- factor(all_rfperfs$Predictor, levels = c("Signal", "OC", "OP", "Distance"))
all_rfperfs$Chromosome <- factor(all_rfperfs$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_ba <- all_rfperfs %>%
  dplyr::group_by(Resolution, Resampling, Predictor) %>%
  dplyr::summarise(Performance = mean(V1, na.rm = TRUE),
                   PerformanceSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_ba, aes(x=Predictor, y=Performance, fill=Resampling)) +
  geom_bar(stat = "identity", position_dodge(0.9))+
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=1,
                position=position_dodge(0.9)) +
  facet_grid(. ~ Predictor)+
  theme_minimal() +
  theme_bw()+
  xlab("Predictor Type") + 
  ylab("Balanced Accuracy") + #ylim(-.1,.5) +
  scale_color_discrete(name="Resampling Technique")+
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  strip.text.y = element_text(size = 15),
  #panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position="bottom")
```

### K562

#### 10kb

```{r}
badata_k562_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_peakachu.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_10kb <- rbind(badata_k562_10kb,
                                   perfdata[10,2])
    }
  }
}

```

#### Plotting

```{r}
all_rfperfs <- rbind.data.frame(badata_k562_10kb)
all_rfperfs$CL <- rep("K562", length(badata_k562_5kb))
all_rfperfs$Resolution <- rep("10 kb", length(badata_k562_10kb))
all_rfperfs$Resampling <- c(rep("None", 4),
                            rep("ROS", 4),
                            rep("RUS", 4),
                            rep("SMOTE", 4))
all_rfperfs$Predictor <- c("Distance", "OC", "OP", "Signal")
all_rfperfs$Chromosome <- c(rep("CHR1", 16),
                            rep("CHR2", 16),
                            rep("CHR3", 16),
                            rep("CHR4", 16),
                            rep("CHR5", 16),
                            rep("CHR6", 16),
                            rep("CHR7", 16),
                            rep("CHR8", 16),
                            rep("CHR10", 16),
                            rep("CHR11", 16),
                            rep("CHR12", 16),
                            rep("CHR13", 16),
                            rep("CHR14", 16),
                            rep("CHR15", 16),
                            rep("CHR16", 16),
                            rep("CHR17", 16),
                            rep("CHR18", 16),
                            rep("CHR19", 16),
                            rep("CHR20", 16),
                            rep("CHR21", 16),
                            rep("CHR22", 16))

all_rfperfs$Resolution <- factor(all_rfperfs$Resolution, levels = c("5 kb", "10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs$Resampling <- factor(all_rfperfs$Resampling, levels = c("None", "ROS", "RUS", "SMOTE"))
all_rfperfs$Predictor <- factor(all_rfperfs$Predictor, levels = c("Signal", "OC", "OP", "Distance"))
all_rfperfs$Chromosome <- factor(all_rfperfs$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_ba <- all_rfperfs %>%
  dplyr::group_by(Resolution, Resampling, Predictor) %>%
  dplyr::summarise(Performance = mean(V1, na.rm = TRUE),
                   PerformanceSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_ba, aes(x=Predictor, y=Performance, fill=Resampling)) +
  geom_bar(stat = "identity", position_dodge(0.9))+
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=1,
                position=position_dodge(0.9)) +
  facet_grid(. ~ Predictor)+
  theme_minimal() +
  theme_bw()+
  xlab("Predictor Type") + 
  ylab("Balanced Accuracy") + #ylim(-.1,.5) +
  scale_color_discrete(name="Resampling Technique")+
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  strip.text.y = element_text(size = 15),
  #panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position="bottom")
```

## LOLLIPOP

### GM12878

#### 10kb

```{r}
badata_gm12878_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_lollipop.rds"))
      perfdata <- perfdata[[3]]

      badata_gm12878_10kb <- rbind(badata_gm12878_10kb,
                                   perfdata[10,2])
    }
  }
}

```

#### Plotting

```{r}
all_rfperfs <- rbind.data.frame(badata_gm12878_10kb)
all_rfperfs$CL <- rep("GM12878", length(badata_gm12878_5kb))
all_rfperfs$Resolution <- rep("10 kb", length(badata_gm12878_10kb))
all_rfperfs$Resampling <- c(rep("None", 4),
                            rep("ROS", 4),
                            rep("RUS", 4),
                            rep("SMOTE", 4))
all_rfperfs$Predictor <- c("Distance", "OC", "OP", "Signal")
all_rfperfs$Chromosome <- c(rep("CHR1", 16),
                            rep("CHR2", 16),
                            rep("CHR3", 16),
                            rep("CHR4", 16),
                            rep("CHR5", 16),
                            rep("CHR6", 16),
                            rep("CHR7", 16),
                            rep("CHR8", 16),
                            rep("CHR10", 16),
                            rep("CHR11", 16),
                            rep("CHR12", 16),
                            rep("CHR13", 16),
                            rep("CHR14", 16),
                            rep("CHR15", 16),
                            rep("CHR16", 16),
                            rep("CHR17", 16),
                            rep("CHR18", 16),
                            rep("CHR19", 16),
                            rep("CHR20", 16),
                            rep("CHR21", 16),
                            rep("CHR22", 16))

all_rfperfs$Resolution <- factor(all_rfperfs$Resolution, levels = c("5 kb", "10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs$Resampling <- factor(all_rfperfs$Resampling, levels = c("None", "ROS", "RUS", "SMOTE"))
all_rfperfs$Predictor <- factor(all_rfperfs$Predictor, levels = c("Signal", "OC", "OP", "Distance"))
all_rfperfs$Chromosome <- factor(all_rfperfs$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_ba <- all_rfperfs %>%
  dplyr::group_by(Resolution, Resampling, Predictor) %>%
  dplyr::summarise(Performance = mean(V1, na.rm = TRUE),
                   PerformanceSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_ba, aes(x=Predictor, y=Performance, fill=Resampling)) +
  geom_bar(stat = "identity", position_dodge(0.9))+
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=1,
                position=position_dodge(0.9)) +
  facet_grid(. ~ Predictor)+
  theme_minimal() +
  theme_bw()+
  xlab("Predictor Type") + 
  ylab("Balanced Accuracy") + #ylim(-.1,.5) +
  scale_color_discrete(name="Resampling Technique")+
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  strip.text.y = element_text(size = 15),
  #panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position="bottom")
```

### K562

#### 10kb

```{r}
badata_k562_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
  for(j in c("none","ros","rus","smote")){
    for(k in c("distance", "oc", "op", "signal")){
      print(c(i,j,k))
      files <- list.files(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/"))

      perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                 i, 
                                 "/",
                                 j, 
                                 "/", 
                                 k, 
                                 "/TADRF_lollipop.rds"))
      perfdata <- perfdata[[3]]

      badata_k562_10kb <- rbind(badata_k562_10kb,
                                   perfdata[10,2])
    }
  }
}

```

#### Plotting

```{r}
all_rfperfs <- rbind.data.frame(badata_k562_10kb)
all_rfperfs$CL <- rep("K562", length(badata_k562_5kb))
all_rfperfs$Resolution <- rep("10 kb", length(badata_k562_10kb))
all_rfperfs$Resampling <- c(rep("None", 4),
                            rep("ROS", 4),
                            rep("RUS", 4),
                            rep("SMOTE", 4))
all_rfperfs$Predictor <- c("Distance", "OC", "OP", "Signal")
all_rfperfs$Chromosome <- c(rep("CHR1", 16),
                            rep("CHR2", 16),
                            rep("CHR3", 16),
                            rep("CHR4", 16),
                            rep("CHR5", 16),
                            rep("CHR6", 16),
                            rep("CHR7", 16),
                            rep("CHR8", 16),
                            rep("CHR10", 16),
                            rep("CHR11", 16),
                            rep("CHR12", 16),
                            rep("CHR13", 16),
                            rep("CHR14", 16),
                            rep("CHR15", 16),
                            rep("CHR16", 16),
                            rep("CHR17", 16),
                            rep("CHR18", 16),
                            rep("CHR19", 16),
                            rep("CHR20", 16),
                            rep("CHR21", 16),
                            rep("CHR22", 16))

all_rfperfs$Resolution <- factor(all_rfperfs$Resolution, levels = c("5 kb", "10 kb", "25 kb", "50 kb", "100 kb"))
all_rfperfs$Resampling <- factor(all_rfperfs$Resampling, levels = c("None", "ROS", "RUS", "SMOTE"))
all_rfperfs$Predictor <- factor(all_rfperfs$Predictor, levels = c("Signal", "OC", "OP", "Distance"))
all_rfperfs$Chromosome <- factor(all_rfperfs$Chromosome, levels = paste0("CHR", c(1:8,10:22)))

all_rfperfs_ba <- all_rfperfs %>%
  dplyr::group_by(Resolution, Resampling, Predictor) %>%
  dplyr::summarise(Performance = mean(V1, na.rm = TRUE),
                   PerformanceSD = sd(V1, na.rm = TRUE))

ggplot(all_rfperfs_ba, aes(x=Predictor, y=Performance, fill=Resampling)) +
  geom_bar(stat = "identity", position_dodge(0.9))+
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=1,
                position=position_dodge(0.9)) +
  facet_grid(. ~ Predictor)+
  theme_minimal() +
  theme_bw()+
  xlab("Predictor Type") + 
  ylab("Balanced Accuracy") + #ylim(-.1,.5) +
  scale_color_discrete(name="Resampling Technique")+
  guides(shape=FALSE, linetype=FALSE)+
  theme(axis.text.x = element_text(size=15,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1
  ),
  axis.text.y = element_text(size = 15),
  axis.title.x = element_text(size = 20),
  axis.title.y = element_text(size = 20),
  strip.text.x = element_text(size = 15),
  strip.text.y = element_text(size = 15),
  #panel.spacing = unit(2, "lines"),
  legend.text=element_text(size=15),
  legend.title=element_text(size=20),
  plot.title = element_text(size=20),
  legend.position="bottom")
```