### 2020-04-29 Scripts_for_results/04_creating_rfe_lineplot.Rmd

---
title: "Creating RFE line plot"
author: "Spiro Stilianoudakis"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and functions

```{r}
library(S4Vectors)
library(GenomicRanges)
library(ggplot2)
library(caret)
library(randomForest)
library(glmnet)
library(gbm)
library(DMwR)
library(rtracklayer)
library(Biostrings)
library(BSgenome)
library(cluster)
library(pROC)
library(SuperLearner)
library(data.table)
library(PRROC)
library(bigmemory)
library(ranger)
library(dplyr)
library(parallel)
library(doSNOW)
library(foreach)
library(pbapply)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)

source("Z:/TAD_data_analysis/functions_for_R_package/preciseTAD.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADRF.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADrfe.R")
source("Z:/TAD_data_analysis/functions_for_R_package/distance_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/percent_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/count_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/binary_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/signal_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/annots_to_granges_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/extract_boundaries_func.R")

```

## ARROWHEAD

### GM12878

#### Optimal parameters: 5 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- data.frame()
for(i in c("GM12878")){
  for(k in 1:length(chrs)){print(c(i,k));
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "5kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe.rds"))
    rfeModelResults <- rfeModelResultsList[[1]]
    rfeModelResults <- rfeModelResults[,c(1,11)]
    rfeModelResults$CL <- i
    rfeModelResults$Chromosome <- chrs[k]
    rfeperchr <- rbind.data.frame(rfeperchr, rfeModelResults)
  }
}

rfeperchr_summ <- rfeperchr %>%
  dplyr::group_by(Variables) %>%
  dplyr::summarise(Performance = mean(Accuracy, na.rm = TRUE),
                   PerformanceSD = sd(Accuracy, na.rm = TRUE))

ggplot(rfeperchr_summ, aes(x=Variables, y=Performance)) + 
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), 
                width=2, 
                size=1, 
                color="red") +
  geom_line(size=2, color="red") +
  geom_point(size=4, color="red") +
  xlab("Number of top ranked annotations") +
  ylab("Cross-Validated Accuracy") +
  scale_color_discrete(name="Cell line")+
  scale_x_continuous(breaks = c(2,4,8,16,32,52),
                     labels = c(2,4,8,16,32,52))+
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")
```

### K562

#### Optimal parameters: 5 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- data.frame()
for(i in c("K562")){
  for(k in 1:length(chrs)){print(c(i,k));
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "5kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe.rds"))
    rfeModelResults <- rfeModelResultsList[[1]]
    rfeModelResults <- rfeModelResults[,c(1,11)]
    rfeModelResults$CL <- i
    rfeModelResults$Chromosome <- chrs[k]
    rfeperchr <- rbind.data.frame(rfeperchr, rfeModelResults)
  }
}

rfeperchr_summ <- rfeperchr %>%
  dplyr::group_by(Variables) %>%
  dplyr::summarise(Performance = mean(Accuracy, na.rm = TRUE),
                   PerformanceSD = sd(Accuracy, na.rm = TRUE))

ggplot(rfeperchr_summ, aes(x=Variables, y=Performance)) + 
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), 
                width=2, 
                size=1, 
                color="red") +
  geom_line(size=2, color="red") +
  geom_point(size=4, color="red") +
  xlab("Number of top ranked annotations") +
  ylab("Cross-Validated Accuracy") +
  scale_color_discrete(name="Cell line")+
  scale_x_continuous(breaks = c(2,4,8,16,32,52),
                     labels = c(2,4,8,16,32,52))+
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")
```

## PEAKACHU

### GM12878

#### Optimal parameters: 10 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- data.frame()
for(i in c("GM12878")){
  for(k in 1:length(chrs)){print(c(i,k));
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe_peakachu.rds"))
    rfeModelResults <- rfeModelResultsList[[1]]
    rfeModelResults <- rfeModelResults[,c(1,11)]
    rfeModelResults$CL <- i
    rfeModelResults$Chromosome <- chrs[k]
    rfeperchr <- rbind.data.frame(rfeperchr, rfeModelResults)
  }
}

rfeperchr_summ <- rfeperchr %>%
  dplyr::group_by(Variables) %>%
  dplyr::summarise(Performance = mean(Accuracy, na.rm = TRUE),
                   PerformanceSD = sd(Accuracy, na.rm = TRUE))

ggplot(rfeperchr_summ, aes(x=Variables, y=Performance)) + 
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), 
                width=2, 
                size=1, 
                color="red") +
  geom_line(size=2, color="red") +
  geom_point(size=4, color="red") +
  xlab("Number of top ranked annotations") +
  ylab("Cross-Validated Accuracy") +
  scale_color_discrete(name="Cell line")+
  scale_x_continuous(breaks = c(2,4,8,16,32,52),
                     labels = c(2,4,8,16,32,52))+
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")
```

### K562

#### Optimal parameters: 10 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- data.frame()
for(i in c("K562")){
  for(k in 1:length(chrs)){print(c(i,k));
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe_peakachu.rds"))
    rfeModelResults <- rfeModelResultsList[[1]]
    rfeModelResults <- rfeModelResults[,c(1,11)]
    rfeModelResults$CL <- i
    rfeModelResults$Chromosome <- chrs[k]
    rfeperchr <- rbind.data.frame(rfeperchr, rfeModelResults)
  }
}

rfeperchr_summ <- rfeperchr %>%
  dplyr::group_by(Variables) %>%
  dplyr::summarise(Performance = mean(Accuracy, na.rm = TRUE),
                   PerformanceSD = sd(Accuracy, na.rm = TRUE))

ggplot(rfeperchr_summ, aes(x=Variables, y=Performance)) + 
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), 
                width=2, 
                size=1, 
                color="red") +
  geom_line(size=2, color="red") +
  geom_point(size=4, color="red") +
  xlab("Number of top ranked annotations") +
  ylab("Cross-Validated Accuracy") +
  scale_color_discrete(name="Cell line")+
  scale_x_continuous(breaks = c(2,4,8,16,32,52),
                     labels = c(2,4,8,16,32,52))+
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")
```

## LOLLIPOP

### GM12878

#### Optimal parameters: 10 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- data.frame()
for(i in c("GM12878")){
  for(k in 1:length(chrs)){print(c(i,k));
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe_lollipop.rds"))
    rfeModelResults <- rfeModelResultsList[[1]]
    rfeModelResults <- rfeModelResults[,c(1,11)]
    rfeModelResults$CL <- i
    rfeModelResults$Chromosome <- chrs[k]
    rfeperchr <- rbind.data.frame(rfeperchr, rfeModelResults)
  }
}

rfeperchr_summ <- rfeperchr %>%
  dplyr::group_by(Variables) %>%
  dplyr::summarise(Performance = mean(Accuracy, na.rm = TRUE),
                   PerformanceSD = sd(Accuracy, na.rm = TRUE))

ggplot(rfeperchr_summ, aes(x=Variables, y=Performance)) + 
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), 
                width=2, 
                size=1, 
                color="red") +
  geom_line(size=2, color="red") +
  geom_point(size=4, color="red") +
  xlab("Number of top ranked annotations") +
  ylab("Cross-Validated Accuracy") +
  scale_color_discrete(name="Cell line")+
  scale_x_continuous(breaks = c(2,4,8,16,32,52),
                     labels = c(2,4,8,16,32,52))+
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")
```

### K562

#### Optimal parameters: 10 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- data.frame()
for(i in c("K562")){
  for(k in 1:length(chrs)){print(c(i,k));
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe_lollipop.rds"))
    rfeModelResults <- rfeModelResultsList[[1]]
    rfeModelResults <- rfeModelResults[,c(1,11)]
    rfeModelResults$CL <- i
    rfeModelResults$Chromosome <- chrs[k]
    rfeperchr <- rbind.data.frame(rfeperchr, rfeModelResults)
  }
}

rfeperchr_summ <- rfeperchr %>%
  dplyr::group_by(Variables) %>%
  dplyr::summarise(Performance = mean(Accuracy, na.rm = TRUE),
                   PerformanceSD = sd(Accuracy, na.rm = TRUE))

ggplot(rfeperchr_summ, aes(x=Variables, y=Performance)) + 
  geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), 
                width=2, 
                size=1, 
                color="red") +
  geom_line(size=2, color="red") +
  geom_point(size=4, color="red") +
  xlab("Number of top ranked annotations") +
  ylab("Cross-Validated Accuracy") +
  scale_color_discrete(name="Cell line")+
  scale_x_continuous(breaks = c(2,4,8,16,32,52),
                     labels = c(2,4,8,16,32,52))+
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")
```