### 2020-05-01 Scripts_for_results/05_creating_importance_plot.Rmd
edits to results scripts

---
title: "Creating importance plot"
author: "Spiro Stilianoudakis"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and functions

```{r}
library(S4Vectors)
library(GenomicRanges)
library(ggplot2)
library(caret)
library(randomForest)
library(glmnet)
library(gbm)
library(DMwR)
library(rtracklayer)
library(Biostrings)
library(BSgenome)
library(cluster)
library(pROC)
library(SuperLearner)
library(data.table)
library(PRROC)
library(bigmemory)
library(ranger)
library(dplyr)
library(parallel)
library(doSNOW)
library(foreach)
library(pbapply)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)
library(gplots)

source("Z:/TAD_data_analysis/functions_for_R_package/preciseTAD.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADRF.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADrfe.R")
source("Z:/TAD_data_analysis/functions_for_R_package/distance_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/percent_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/count_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/binary_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/signal_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/annots_to_granges_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/extract_boundaries_func.R")

```

## ARROWHEAD

### GM12878

#### Optimal parameters: 5 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- list()
for(i in c("GM12878")){
  for(k in 1:length(chrs)){
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "5kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe.rds"))

    set.size = 8
    rfeModelResults <- rfeModelResultsList[[2]]
    rfe.set <- rfeModelResults[rfeModelResults$Variables==set.size & rfeModelResults$Resample=="Fold1",]
    rfe.set <- aggregate(rfe.set[, c("Overall")], list(rfe.set$var), mean)
    rfe.order <- order(rfe.set[, c("x")], decreasing = TRUE)
    rfe.set <- rfe.set[rfe.order, ]
    rfe.set$CHR <- chrs[k]
    rfeperchr[[k]] <- rfe.set
  }
}

rfeperchr <- do.call(rbind.data.frame, rfeperchr)
rfeperchr <- reshape(rfeperchr, idvar = "Group.1", timevar = "CHR", direction = "wide")
names(rfeperchr) <- c("TFBS", chrs)
rfeperchr[is.na(rfeperchr)] <- 0
rfeperchr <- rfeperchr[order(rowMeans(rfeperchr[,-1], na.rm = TRUE),decreasing = TRUE),]

rfe.set.mat <- as.matrix(rfeperchr[,-1])
rownames(rfe.set.mat) <- rfeperchr$TFBS

rfe.set.mat <- t(apply(rfe.set.mat,1,rev))

distance.col = dist(rfe.set.mat, method = "euclidean")
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "pink", "red"))(ncol(rfe.set.mat))

heatmap.2(t(rfe.set.mat),
          #labRow=rownames(rfe.set.mat),
          #labCol=colnames(rfe.set.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(10,4),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="column",     # only draw a row dendrogram,
          Rowv = FALSE,
          Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=.5,
          cexCol=1,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          rowsep=0:nrow(t(rfe.set.mat)),
          colsep=0:ncol(t(rfe.set.mat)),
          sepwidth=c(0.00001,0.00001),
          sepcolor="black"
)

```

### K562

#### Optimal parameters: 5 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- list()
for(i in c("K562")){
  for(k in 1:length(chrs)){
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "5kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe.rds"))

    set.size = 8
    rfeModelResults <- rfeModelResultsList[[2]]
    rfe.set <- rfeModelResults[rfeModelResults$Variables==set.size & rfeModelResults$Resample=="Fold1",]
    rfe.set <- aggregate(rfe.set[, c("Overall")], list(rfe.set$var), mean)
    rfe.order <- order(rfe.set[, c("x")], decreasing = TRUE)
    rfe.set <- rfe.set[rfe.order, ]
    rfe.set$CHR <- chrs[k]
    rfeperchr[[k]] <- rfe.set
  }
}

rfeperchr <- do.call(rbind.data.frame, rfeperchr)
rfeperchr <- reshape(rfeperchr, idvar = "Group.1", timevar = "CHR", direction = "wide")
names(rfeperchr) <- c("TFBS", chrs)
rfeperchr[is.na(rfeperchr)] <- 0
rfeperchr <- rfeperchr[order(rowMeans(rfeperchr[,-1], na.rm = TRUE),decreasing = TRUE),]

rfe.set.mat <- as.matrix(rfeperchr[,-1])
rownames(rfe.set.mat) <- rfeperchr$TFBS

rfe.set.mat <- t(apply(rfe.set.mat,1,rev))

distance.col = dist(rfe.set.mat, method = "euclidean")
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "lightblue", "blue"))(ncol(rfe.set.mat))

heatmap.2(t(rfe.set.mat),
          #labRow=rownames(rfe.set.mat),
          #labCol=colnames(rfe.set.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(10,4),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="column",     # only draw a row dendrogram,
          Rowv = FALSE,
          Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=.5,
          cexCol=1,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          rowsep=0:nrow(t(rfe.set.mat)),
          colsep=0:ncol(t(rfe.set.mat)),
          sepwidth=c(0.00001,0.00001),
          sepcolor="black"
)

```

## PEAKACHU

### GM12878

#### Optimal parameters: 5 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- list()
for(i in c("GM12878")){
  for(k in 1:length(chrs)){
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe_peakachu.rds"))

    set.size = 8
    rfeModelResults <- rfeModelResultsList[[2]]
    rfe.set <- rfeModelResults[rfeModelResults$Variables==set.size & rfeModelResults$Resample=="Fold1",]
    rfe.set <- aggregate(rfe.set[, c("Overall")], list(rfe.set$var), mean)
    rfe.order <- order(rfe.set[, c("x")], decreasing = TRUE)
    rfe.set <- rfe.set[rfe.order, ]
    rfe.set$CHR <- chrs[k]
    rfeperchr[[k]] <- rfe.set
  }
}

rfeperchr <- do.call(rbind.data.frame, rfeperchr)
rfeperchr <- reshape(rfeperchr, idvar = "Group.1", timevar = "CHR", direction = "wide")
names(rfeperchr) <- c("TFBS", chrs)
rfeperchr[is.na(rfeperchr)] <- 0
rfeperchr <- rfeperchr[order(rowMeans(rfeperchr[,-1], na.rm = TRUE),decreasing = TRUE),]

rfe.set.mat <- as.matrix(rfeperchr[,-1])
rownames(rfe.set.mat) <- rfeperchr$TFBS

rfe.set.mat <- t(apply(rfe.set.mat,1,rev))

distance.col = dist(rfe.set.mat, method = "euclidean")
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "pink", "red"))(ncol(rfe.set.mat))

heatmap.2(t(rfe.set.mat),
          #labRow=rownames(rfe.set.mat),
          #labCol=colnames(rfe.set.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(10,4),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="column",     # only draw a row dendrogram,
          Rowv = FALSE,
          Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=.5,
          cexCol=1,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          rowsep=0:nrow(t(rfe.set.mat)),
          colsep=0:ncol(t(rfe.set.mat)),
          sepwidth=c(0.00001,0.00001),
          sepcolor="black"
)

```

### K562

#### Optimal parameters: 5 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- list()
for(i in c("K562")){
  for(k in 1:length(chrs)){
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe_peakachu.rds"))

    set.size = 8
    rfeModelResults <- rfeModelResultsList[[2]]
    rfe.set <- rfeModelResults[rfeModelResults$Variables==set.size & rfeModelResults$Resample=="Fold1",]
    rfe.set <- aggregate(rfe.set[, c("Overall")], list(rfe.set$var), mean)
    rfe.order <- order(rfe.set[, c("x")], decreasing = TRUE)
    rfe.set <- rfe.set[rfe.order, ]
    rfe.set$CHR <- chrs[k]
    rfeperchr[[k]] <- rfe.set
  }
}

rfeperchr <- do.call(rbind.data.frame, rfeperchr)
rfeperchr <- reshape(rfeperchr, idvar = "Group.1", timevar = "CHR", direction = "wide")
names(rfeperchr) <- c("TFBS", chrs)
rfeperchr[is.na(rfeperchr)] <- 0
rfeperchr <- rfeperchr[order(rowMeans(rfeperchr[,-1], na.rm = TRUE),decreasing = TRUE),]

rfe.set.mat <- as.matrix(rfeperchr[,-1])
rownames(rfe.set.mat) <- rfeperchr$TFBS

rfe.set.mat <- t(apply(rfe.set.mat,1,rev))

distance.col = dist(rfe.set.mat, method = "euclidean")
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "lightblue", "blue"))(ncol(rfe.set.mat))

heatmap.2(t(rfe.set.mat),
          #labRow=rownames(rfe.set.mat),
          #labCol=colnames(rfe.set.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(10,4),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="column",     # only draw a row dendrogram,
          Rowv = FALSE,
          Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=.5,
          cexCol=1,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          rowsep=0:nrow(t(rfe.set.mat)),
          colsep=0:ncol(t(rfe.set.mat)),
          sepwidth=c(0.00001,0.00001),
          sepcolor="black"
)

```

## LOLLIPOP

### GM12878

#### Optimal parameters: 5 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- list()
for(i in c("GM12878")){
  for(k in 1:length(chrs)){
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe_lollipop.rds"))

    set.size = 8
    rfeModelResults <- rfeModelResultsList[[2]]
    rfe.set <- rfeModelResults[rfeModelResults$Variables==set.size & rfeModelResults$Resample=="Fold1",]
    rfe.set <- aggregate(rfe.set[, c("Overall")], list(rfe.set$var), mean)
    rfe.order <- order(rfe.set[, c("x")], decreasing = TRUE)
    rfe.set <- rfe.set[rfe.order, ]
    rfe.set$CHR <- chrs[k]
    rfeperchr[[k]] <- rfe.set
  }
}

rfeperchr <- do.call(rbind.data.frame, rfeperchr)
rfeperchr <- reshape(rfeperchr, idvar = "Group.1", timevar = "CHR", direction = "wide")
names(rfeperchr) <- c("TFBS", chrs)
rfeperchr[is.na(rfeperchr)] <- 0
rfeperchr <- rfeperchr[order(rowMeans(rfeperchr[,-1], na.rm = TRUE),decreasing = TRUE),]

rfe.set.mat <- as.matrix(rfeperchr[,-1])
rownames(rfe.set.mat) <- rfeperchr$TFBS

rfe.set.mat <- t(apply(rfe.set.mat,1,rev))

distance.col = dist(rfe.set.mat, method = "euclidean")
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "pink", "red"))(ncol(rfe.set.mat))

heatmap.2(t(rfe.set.mat),
          #labRow=rownames(rfe.set.mat),
          #labCol=colnames(rfe.set.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(10,4),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="column",     # only draw a row dendrogram,
          Rowv = FALSE,
          Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=.5,
          cexCol=1,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          rowsep=0:nrow(t(rfe.set.mat)),
          colsep=0:ncol(t(rfe.set.mat)),
          sepwidth=c(0.00001,0.00001),
          sepcolor="black"
)

```

### K562

#### Optimal parameters: 5 kb, TFBS, distance type predictors, RUS

```{r}
chrs <- paste0("CHR",c(1:8,10:22))
rfeperchr <- list()
for(i in c("K562")){
  for(k in 1:length(chrs)){
    rfeModelResultsList <- readRDS(paste0("Z:/TAD_data_analysis/",
                                          i,
                                          "/",
                                          "10kb/results_by_chr/",
                                          chrs[k],
                                          "/rus/distance/TADrfe_lollipop.rds"))

    set.size = 8
    rfeModelResults <- rfeModelResultsList[[2]]
    rfe.set <- rfeModelResults[rfeModelResults$Variables==set.size & rfeModelResults$Resample=="Fold1",]
    rfe.set <- aggregate(rfe.set[, c("Overall")], list(rfe.set$var), mean)
    rfe.order <- order(rfe.set[, c("x")], decreasing = TRUE)
    rfe.set <- rfe.set[rfe.order, ]
    rfe.set$CHR <- chrs[k]
    rfeperchr[[k]] <- rfe.set
  }
}

rfeperchr <- do.call(rbind.data.frame, rfeperchr)
rfeperchr <- reshape(rfeperchr, idvar = "Group.1", timevar = "CHR", direction = "wide")
names(rfeperchr) <- c("TFBS", chrs)
rfeperchr[is.na(rfeperchr)] <- 0
rfeperchr <- rfeperchr[order(rowMeans(rfeperchr[,-1], na.rm = TRUE),decreasing = TRUE),]

rfe.set.mat <- as.matrix(rfeperchr[,-1])
rownames(rfe.set.mat) <- rfeperchr$TFBS

rfe.set.mat <- t(apply(rfe.set.mat,1,rev))

distance.col = dist(rfe.set.mat, method = "euclidean")
cluster.col = hclust(distance.col, method = "average")
my_palette <- colorRampPalette(c("white", "lightblue", "blue"))(ncol(rfe.set.mat))

heatmap.2(t(rfe.set.mat),
          #labRow=rownames(rfe.set.mat),
          #labCol=colnames(rfe.set.mat),
          #cellnote = ,  # same data set for cell labels
          #main = "", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins=c(10,4),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier
          na.color = "gray",
          #breaks=col_breaks,    # enable color transition at specified limits
          key=FALSE,
          #keysize = 1.5,
          #key.title = "Elastic-Net Coefficient",
          dendrogram="column",     # only draw a row dendrogram,
          Rowv = FALSE,
          Colv = reorder(as.dendrogram(cluster.col), 37:1),
          cexRow=.5,
          cexCol=1,
          #labRow = FALSE,
          #lmat=rbind( c(0, 3), c(2,1), c(0,4) ), 
          #lhei=c(.5,6,0),
          rowsep=0:nrow(t(rfe.set.mat)),
          colsep=0:ncol(t(rfe.set.mat)),
          sepwidth=c(0.00001,0.00001),
          sepcolor="black"
)

```