### 2020-05-15 Scripts_for_results/10_evaluating_jaccard_indices.Rmd
initial commit

---
title: "Evaluating Jaccard Indices"
author: "Spiro Stilianoudakis"
date: "5/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and functions

```{r}
library(S4Vectors)
library(GenomicRanges)
library(ggplot2)
library(caret)
library(randomForest)
library(glmnet)
library(gbm)
library(DMwR)
library(rtracklayer)
library(Biostrings)
library(BSgenome)
library(cluster)
library(pROC)
library(SuperLearner)
library(data.table)
library(PRROC)
library(bigmemory)
library(ranger)
library(dplyr)
library(parallel)
library(doSNOW)
library(foreach)
library(pbapply)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)

source("Z:/TAD_data_analysis/functions_for_R_package/preciseTAD.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADRF.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADrfe.R")
source("Z:/TAD_data_analysis/functions_for_R_package/distance_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/percent_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/count_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/binary_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/signal_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/annots_to_granges_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/extract_boundaries_func.R")

```

# ARROWHEAD

## GM12878

```{r}
pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
chrs <- paste0("CHR", c(21:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                                    chrs[i],
                                    "/rus/distance/preciseTAD.rds"))

  true_bound_list[[i]] <- called_and_pred[[3]]
  pred_bound_list[[i]] <- called_and_pred[[2]]

}

all_true_bounds_gm12878 <- do.call("c", true_bound_list)
all_pred_bounds_gm12878 <- do.call("c", pred_bound_list)

all_true_bounds_gm12878_arrowhead <- flank(all_true_bounds_gm12878, 5000, both=TRUE)
all_pred_bounds_gm12878_arrowhead <- flank(all_pred_bounds_gm12878, 5000, both=TRUE)

```

## K562

```{r}
pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
chrs <- paste0("CHR", c(21:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/K562/5kb/results_by_chr/",
                                    chrs[i],
                                    "/rus/distance/preciseTAD.rds"))

  true_bound_list[[i]] <- called_and_pred[[3]]
  pred_bound_list[[i]] <- called_and_pred[[2]]

}

all_true_bounds_k562 <- do.call("c", true_bound_list)
all_pred_bounds_k562 <- do.call("c", pred_bound_list)

all_true_bounds_k562_arrowhead <- flank(all_true_bounds_k562, 5000, both=TRUE)
all_pred_bounds_k562_arrowhead <- flank(all_pred_bounds_k562, 5000, both=TRUE)

```

## Jaccard Indices

```{r}
#original
jaccard_o <- numeric()
for(i in 1:length(chrs)){
  g1 <- all_true_bounds_gm12878_arrowhead[which(as.character(seqnames(all_true_bounds_gm12878_arrowhead))==tolower(chrs[i]))]
  g2 <- all_true_bounds_k562_arrowhead[which(as.character(seqnames(all_true_bounds_k562_arrowhead))==tolower(chrs[i]))]

  co <- countOverlaps(g1,g2)

  int <- length(co[which(co>0)])

  uni <- length(g1) + length(g2) - int

  jaccard_o[i] <- int/uni
}
summary(jaccard_o)

#preciseTAD
jaccard_p <- numeric()
for(i in 1:length(chrs)){
  g1 <- all_pred_bounds_gm12878_arrowhead[which(as.character(seqnames(all_pred_bounds_gm12878_arrowhead))==tolower(chrs[i]))]
  g2 <- all_pred_bounds_k562_arrowhead[which(as.character(seqnames(all_pred_bounds_k562_arrowhead))==tolower(chrs[i]))]

  co <- countOverlaps(g1,g2)

  int <- length(co[which(co>0)])

  uni <- length(g1) + length(g2) - int

  jaccard_p[i] <- int/uni
}
summary(jaccard_p)

wilcox.test(jaccard_o,jaccard_p)
```

# PEAKACHU

## GM12878

```{r}
pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
chrs <- paste0("CHR", c(21:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                    chrs[i],
                                    "/rus/distance/preciseTAD_peakachu.rds"))

  true_bound_list[[i]] <- called_and_pred[[3]]
  pred_bound_list[[i]] <- called_and_pred[[2]]

}

all_true_bounds_gm12878 <- do.call("c", true_bound_list)
all_pred_bounds_gm12878 <- do.call("c", pred_bound_list)

all_true_bounds_gm12878_peakachu <- flank(all_true_bounds_gm12878, 10000, both=TRUE)
all_pred_bounds_gm12878_peakachu <- flank(all_pred_bounds_gm12878, 10000, both=TRUE)

```

## K562

```{r}
pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
chrs <- paste0("CHR", c(21:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                    chrs[i],
                                    "/rus/distance/preciseTAD_peakachu.rds"))

  true_bound_list[[i]] <- called_and_pred[[3]]
  pred_bound_list[[i]] <- called_and_pred[[2]]

}

all_true_bounds_k562 <- do.call("c", true_bound_list)
all_pred_bounds_k562 <- do.call("c", pred_bound_list)

all_true_bounds_k562_peakachu <- flank(all_true_bounds_k562, 10000, both=TRUE)
all_pred_bounds_k562_peakachu <- flank(all_pred_bounds_k562, 10000, both=TRUE)

```

## Jaccard Indices

```{r}
#original
jaccard_o <- numeric()
for(i in 1:length(chrs)){
  g1 <- all_true_bounds_gm12878_peakachu[which(as.character(seqnames(all_true_bounds_gm12878_peakachu))==tolower(chrs[i]))]
  g2 <- all_true_bounds_k562_peakachu[which(as.character(seqnames(all_true_bounds_k562_peakachu))==tolower(chrs[i]))]

  co <- countOverlaps(g1,g2)

  int <- length(co[which(co>0)])

  uni <- length(g1) + length(g2) - int

  jaccard_o[i] <- int/uni
}
summary(jaccard_o)

#preciseTAD
jaccard_p <- numeric()
for(i in 1:length(chrs)){
  g1 <- all_pred_bounds_gm12878_peakachu[which(as.character(seqnames(all_pred_bounds_gm12878_peakachu))==tolower(chrs[i]))]
  g2 <- all_pred_bounds_k562_peakachu[which(as.character(seqnames(all_pred_bounds_k562_peakachu))==tolower(chrs[i]))]

  co <- countOverlaps(g1,g2)

  int <- length(co[which(co>0)])

  uni <- length(g1) + length(g2) - int

  jaccard_p[i] <- int/uni
}
summary(jaccard_p)

wilcox.test(jaccard_o,jaccard_p)
```