### 2020-07-31 Scripts_for_results/12_cross_cell_line_performances.Rmd
initial commit

---
title: "Cross cell line performance"
author: "Spiro Stilianoudakis"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages and functions

```{r}
library(S4Vectors)
library(GenomicRanges)
library(ggplot2)
library(caret)
library(randomForest)
library(glmnet)
library(gbm)
library(DMwR)
library(rtracklayer)
library(Biostrings)
library(BSgenome)
library(cluster)
library(pROC)
library(SuperLearner)
library(data.table)
library(PRROC)
library(bigmemory)
library(ranger)
library(dplyr)
library(parallel)
library(doSNOW)
library(foreach)
library(pbapply)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)
library(ComplexHeatmap)
library(circlize)
library(cotools)
library(ggcorrplot)
library(circlize)

source("Z:/TAD_data_analysis/functions_for_R_package/preciseTAD.R")
source("Z:/TAD_data_analysis/functions_for_R_package/createTADdata.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADRF.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADrfe.R")
source("Z:/TAD_data_analysis/functions_for_R_package/distance_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/percent_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/count_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/binary_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/signal_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/annots_to_granges_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/extract_boundaries_func.R")

```

# ARROWHEAD

```{r}
arrowhead_g_on_g <- list()
arrowhead_g_on_g_auc <- numeric()
arrowhead_g_on_k <- list()
arrowhead_g_on_k_auc <- numeric()
arrowhead_k_on_k <- list()
arrowhead_k_on_k_auc <- numeric()
arrowhead_k_on_g <- list()
arrowhead_k_on_g_auc <- numeric()

peakachu_g_on_g <- list()
peakachu_g_on_g_auc <- numeric()
peakachu_g_on_k <- list()
peakachu_g_on_k_auc <- numeric()
peakachu_k_on_k <- list()
peakachu_k_on_k_auc <- numeric()
peakachu_k_on_g <- list()
peakachu_k_on_g_auc <- numeric()

genomicElements.GR_gm12878 <- annots_to_granges_func(filepath = "Z:/TAD_data_analysis/annotations/all_common_annotations/gm12878/topTFBS_nonames",
                                             pattern="*.bed",
                                             signal=4)
genomicElements.GR_k562 <- annots_to_granges_func(filepath = "Z:/TAD_data_analysis/annotations/all_common_annotations/k562/topTFBS_nonames",
                                             pattern="*.bed",
                                             signal=4)

domains_arrowhead_gm12878 <- read.table("Z:/TAD_data_analysis/GSE63525_data/arrowhead_output/GM12878/GM12878_domain_data_5000.b.txt", header=F)
bounds.GR_arrowhead_gm12878 <- extract_boundaries_func(domains.mat=domains_arrowhead_gm12878, 
                                     preprocess=FALSE, 
                                     CHR=paste0("CHR", c(1:8,10:22)), 
                                     resolution=5000)
domains_arrowhead_k562 <- read.table("Z:/TAD_data_analysis/GSE63525_data/arrowhead_output/K562/K562_domain_data_5000.b.txt", header=F)
bounds.GR_arrowhead_k562 <- extract_boundaries_func(domains.mat=domains_arrowhead_k562, 
                                     preprocess=FALSE, 
                                     CHR=paste0("CHR", c(1:8,10:22)), 
                                     resolution=5000)

domains_peakachu_gm12878 <- readRDS("Z:/TAD_data_analysis/peakachu/peakachu_GM12878.rds")
bounds.GR_peakachu_gm12878 <- extract_boundaries_func(domains.mat=domains_peakachu_gm12878, 
                                     preprocess=FALSE, 
                                     CHR=paste0("CHR", c(1:8,10:22)), 
                                     resolution=5000)
domains_peakachu_k562 <- readRDS("Z:/TAD_data_analysis/peakachu/peakachu_K562.rds")
bounds.GR_peakachu_k562 <- extract_boundaries_func(domains.mat=domains_peakachu_k562, 
                                     preprocess=FALSE, 
                                     CHR=paste0("CHR", c(1:8,10:22)), 
                                     resolution=5000)

chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
  print(chrs[i])

  #ARROWHEAD

  TADRF_arrowhead_g_on_g <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                                           chrs[i],
                                           "/rus/distance/TADRF_holdout_gm12878_on_gm12878.rds"))

  arrowhead_g_on_g_auc[i] <- TADRF_arrowhead_g_on_g[[3]][18,2]

  TADRF_arrowhead_g_on_k <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                                           chrs[i],
                                           "/rus/distance/TADRF_holdout_gm12878_on_k562.rds"))

  arrowhead_g_on_k_auc[i] <- TADRF_arrowhead_g_on_k[[3]][18,2]

  TADRF_arrowhead_k_on_k <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/K562/5kb/results_by_chr/",
                                           chrs[i],
                                           "/rus/distance/TADRF_holdout_k562_on_k562.rds"))

  arrowhead_k_on_k_auc[i] <- TADRF_arrowhead_k_on_k[[3]][18,2]

  TADRF_arrowhead_k_on_g <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/K562/5kb/results_by_chr/",
                                           chrs[i],
                                           "/rus/distance/TADRF_holdout_k562_on_gm12878.rds"))

  arrowhead_k_on_g_auc[i] <- TADRF_arrowhead_k_on_g[[3]][18,2]

  tadData_arrowhead_gm12878 <- createTADdata(bounds.GR=bounds.GR_arrowhead_gm12878,
                         resolution=5000,
                         genomicElements.GR=genomicElements.GR_gm12878,
                         featureType="distance",
                         resampling="rus",
                         trainCHR=paste0("CHR",c(1:8,10:22))[-which(paste0("CHR",c(1:8,10:22))==chrs[i])],
                         predictCHR=chrs[i],
                         seed=123)
  tadData_arrowhead_k562 <- createTADdata(bounds.GR=bounds.GR_arrowhead_k562,
                         resolution=5000,
                         genomicElements.GR=genomicElements.GR_k562,
                         featureType="distance",
                         resampling="rus",
                         trainCHR=paste0("CHR",c(1:8,10:22))[-which(paste0("CHR",c(1:8,10:22))==chrs[i])],
                         predictCHR=chrs[i],
                         seed=123)

  pred <- as.vector(predict(TADRF_arrowhead_g_on_g[[1]], 
                                       newdata=tadData_arrowhead_gm12878[[2]][,-1], 
                                       type="prob")[,"Yes"])
  ROC <- pROC::roc(tadData_arrowhead_gm12878[[2]]$y, pred, quiet = TRUE)
  arrowhead_g_on_g[[i]] <- data.frame(specificity=ROC$specificities,sensitivity=ROC$sensitivities,chr=chrs[i])

  pred <- as.vector(predict(TADRF_arrowhead_g_on_k[[1]], 
                                       newdata=tadData_arrowhead_k562[[2]][,-1], 
                                       type="prob")[,"Yes"])
  ROC <- pROC::roc(tadData_arrowhead_k562[[2]]$y, pred, quiet = TRUE)
  arrowhead_g_on_k[[i]] <- data.frame(specificity=ROC$specificities,sensitivity=ROC$sensitivities,chr=chrs[i])

  pred <- as.vector(predict(TADRF_arrowhead_k_on_k[[1]], 
                                       newdata=tadData_arrowhead_k562[[2]][,-1], 
                                       type="prob")[,"Yes"])
  ROC <- pROC::roc(tadData_arrowhead_k562[[2]]$y, pred, quiet = TRUE)
  arrowhead_k_on_k[[i]] <- data.frame(specificity=ROC$specificities,sensitivity=ROC$sensitivities,chr=chrs[i])

  pred <- as.vector(predict(TADRF_arrowhead_k_on_g[[1]], 
                                       newdata=tadData_arrowhead_gm12878[[2]][,-1], 
                                       type="prob")[,"Yes"])
  ROC <- pROC::roc(tadData_arrowhead_gm12878[[2]]$y, pred, quiet = TRUE)
  arrowhead_k_on_g[[i]] <- data.frame(specificity=ROC$specificities,sensitivity=ROC$sensitivities,chr=chrs[i])

  #########################################################################################################

  #PEAKACHU

  TADRF_peakachu_g_on_g <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                           chrs[i],
                                           "/rus/distance/TADRF_holdout_gm12878_on_gm12878_peakachu.rds"))

  peakachu_g_on_g_auc[i] <- TADRF_peakachu_g_on_g[[3]][18,2]

  TADRF_peakachu_g_on_k <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                           chrs[i],
                                           "/rus/distance/TADRF_holdout_gm12878_on_k562_peakachu.rds"))

  peakachu_g_on_k_auc[i] <- TADRF_peakachu_g_on_k[[3]][18,2]

  TADRF_peakachu_k_on_k <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/K562/10kb/results_by_chr/",
                                           chrs[i],
                                           "/rus/distance/TADRF_holdout_k562_on_k562_peakachu.rds"))

  peakachu_k_on_k_auc[i] <- TADRF_peakachu_k_on_k[[3]][18,2]

  TADRF_peakachu_k_on_g <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/K562/10kb/results_by_chr/",
                                           chrs[i],
                                           "/rus/distance/TADRF_holdout_k562_on_gm12878_peakachu.rds"))

  peakachu_k_on_g_auc[i] <- TADRF_peakachu_k_on_g[[3]][18,2]

  tadData_peakachu_gm12878 <- createTADdata(bounds.GR=bounds.GR_peakachu_gm12878,
                         resolution=5000,
                         genomicElements.GR=genomicElements.GR_gm12878,
                         featureType="distance",
                         resampling="rus",
                         trainCHR=paste0("CHR",c(1:8,10:22))[-which(paste0("CHR",c(1:8,10:22))==chrs[i])],
                         predictCHR=chrs[i],
                         seed=123)
  tadData_peakachu_k562 <- createTADdata(bounds.GR=bounds.GR_peakachu_k562,
                         resolution=5000,
                         genomicElements.GR=genomicElements.GR_k562,
                         featureType="distance",
                         resampling="rus",
                         trainCHR=paste0("CHR",c(1:8,10:22))[-which(paste0("CHR",c(1:8,10:22))==chrs[i])],
                         predictCHR=chrs[i],
                         seed=123)

  pred <- as.vector(predict(TADRF_peakachu_g_on_g[[1]], 
                                       newdata=tadData_peakachu_gm12878[[2]][,-1], 
                                       type="prob")[,"Yes"])
  ROC <- pROC::roc(tadData_peakachu_gm12878[[2]]$y, pred, quiet = TRUE)
  peakachu_g_on_g[[i]] <- data.frame(specificity=ROC$specificities,sensitivity=ROC$sensitivities,chr=chrs[i])

  pred <- as.vector(predict(TADRF_peakachu_g_on_k[[1]], 
                                       newdata=tadData_peakachu_k562[[2]][,-1], 
                                       type="prob")[,"Yes"])
  ROC <- pROC::roc(tadData_peakachu_k562[[2]]$y, pred, quiet = TRUE)
  peakachu_g_on_k[[i]] <- data.frame(specificity=ROC$specificities,sensitivity=ROC$sensitivities,chr=chrs[i])

  pred <- as.vector(predict(TADRF_peakachu_k_on_k[[1]], 
                                       newdata=tadData_peakachu_k562[[2]][,-1], 
                                       type="prob")[,"Yes"])
  ROC <- pROC::roc(tadData_peakachu_k562[[2]]$y, pred, quiet = TRUE)
  peakachu_k_on_k[[i]] <- data.frame(specificity=ROC$specificities,sensitivity=ROC$sensitivities,chr=chrs[i])

  pred <- as.vector(predict(TADRF_peakachu_k_on_g[[1]], 
                                       newdata=tadData_peakachu_gm12878[[2]][,-1], 
                                       type="prob")[,"Yes"])
  ROC <- pROC::roc(tadData_peakachu_gm12878[[2]]$y, pred, quiet = TRUE)
  peakachu_k_on_g[[i]] <- data.frame(specificity=ROC$specificities,sensitivity=ROC$sensitivities,chr=chrs[i])

}

saveRDS(arrowhead_g_on_g, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/arrowhead_g_on_g.rds")
saveRDS(arrowhead_g_on_g_auc, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/arrowhead_g_on_g_auc.rds") 
saveRDS(arrowhead_g_on_k, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/arrowhead_g_on_k.rds")
saveRDS(arrowhead_g_on_k_auc, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/arrowhead_g_on_k_auc.rds")
saveRDS(arrowhead_k_on_k, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/arrowhead_k_on_k.rds")
saveRDS(arrowhead_k_on_k_auc, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/arrowhead_k_on_k_auc.rds")
saveRDS(arrowhead_k_on_g, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/arrowhead_k_on_g.rds")
saveRDS(arrowhead_k_on_g_auc, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/arrowhead_k_on_g_auc.rds")

saveRDS(peakachu_g_on_g, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/peakachu_g_on_g.rds")
saveRDS(peakachu_g_on_g_auc, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/peakachu_g_on_g_auc.rds") 
saveRDS(peakachu_g_on_k, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/peakachu_g_on_k.rds")
saveRDS(peakachu_g_on_k_auc, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/peakachu_g_on_k_auc.rds") 
saveRDS(peakachu_k_on_k, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/peakachu_k_on_k.rds")
saveRDS(peakachu_k_on_k_auc, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/peakachu_k_on_k_auc.rds") 
saveRDS(peakachu_k_on_g, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/peakachu_k_on_g.rds")
saveRDS(peakachu_k_on_g_auc, "/home/stilianoudakisc/TAD_data_analysis/miscellaneous/peakachu_k_on_g_auc.rds")

```

# Plotting

```{r}
arrowhead_g_on_g <- readRDS("Z:/TAD_data_analysis/miscellaneous/arrowhead_g_on_k.rds")
arrowhead_g_on_g_auc <- readRDS("Z:/TAD_data_analysis/miscellaneous/arrowhead_g_on_k_auc.rds") 
arrowhead_g_on_k <- readRDS("Z:/TAD_data_analysis/miscellaneous/arrowhead_g_on_g.rds")
arrowhead_g_on_k_auc <- readRDS("Z:/TAD_data_analysis/miscellaneous/arrowhead_g_on_g_auc.rds") 
arrowhead_k_on_g <- readRDS("Z:/TAD_data_analysis/miscellaneous/arrowhead_k_on_k.rds")
arrowhead_k_on_g_auc <- readRDS("Z:/TAD_data_analysis/miscellaneous/arrowhead_k_on_k_auc.rds") 
arrowhead_k_on_k <- readRDS("Z:/TAD_data_analysis/miscellaneous/arrowhead_k_on_g.rds")
arrowhead_k_on_k_auc <- readRDS("Z:/TAD_data_analysis/miscellaneous/arrowhead_k_on_g_auc.rds")

peakachu_g_on_g <- readRDS("Z:/TAD_data_analysis/miscellaneous/peakachu_g_on_g.rds")
peakachu_g_on_g_auc <- readRDS("Z:/TAD_data_analysis/miscellaneous/peakachu_g_on_g_auc.rds") 
peakachu_g_on_k <- readRDS("Z:/TAD_data_analysis/miscellaneous/peakachu_g_on_k.rds")
peakachu_g_on_k_auc <- readRDS("Z:/TAD_data_analysis/miscellaneous/peakachu_g_on_k_auc.rds") 
peakachu_k_on_k <- readRDS("Z:/TAD_data_analysis/miscellaneous/peakachu_k_on_k.rds")
peakachu_k_on_k_auc <- readRDS("Z:/TAD_data_analysis/miscellaneous/peakachu_k_on_k_auc.rds") 
peakachu_k_on_g <- readRDS("Z:/TAD_data_analysis/miscellaneous/peakachu_k_on_g.rds")
peakachu_k_on_g_auc <- readRDS("Z:/TAD_data_analysis/miscellaneous/peakachu_k_on_g_auc.rds")

arrowhead_g_on_k <- do.call("rbind.data.frame", arrowhead_g_on_k)

arrowhead_g_on_g[[12]] <- rbind.data.frame(arrowhead_g_on_g[[12]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR13"),
                                           data.frame(specificity=1,sensitivity=0,chr="CHR13"),
                                           data.frame(specificity=1,sensitivity=0,chr="CHR13"))
arrowhead_g_on_g[[13]] <- rbind.data.frame(arrowhead_g_on_g[[13]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR14"))
arrowhead_g_on_g[[20]] <- rbind.data.frame(arrowhead_g_on_g[[20]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR21"),
                                           data.frame(specificity=1,sensitivity=0,chr="CHR21"),
                                           data.frame(specificity=1,sensitivity=0,chr="CHR21"),
                                           data.frame(specificity=1,sensitivity=0,chr="CHR21"))
arrowhead_g_on_g <- do.call("rbind.data.frame", arrowhead_g_on_g)

arrowhead_k_on_k[[15]] <- rbind.data.frame(arrowhead_k_on_k[[15]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR16"))
arrowhead_k_on_k[[17]] <- rbind.data.frame(arrowhead_k_on_k[[17]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR18"),
                                           data.frame(specificity=1,sensitivity=0,chr="CHR18"))
arrowhead_k_on_k[[20]] <- rbind.data.frame(arrowhead_k_on_k[[20]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR21"),
                                           data.frame(specificity=1,sensitivity=0,chr="CHR21"))
arrowhead_k_on_k <- do.call("rbind.data.frame", arrowhead_k_on_k)

arrowhead_k_on_g[[12]] <- rbind.data.frame(arrowhead_k_on_g[[12]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR13"))
arrowhead_k_on_g <- do.call("rbind.data.frame", arrowhead_k_on_g)

peakachu_g_on_g[[20]] <- rbind.data.frame(peakachu_g_on_g[[20]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR21"))
peakachu_g_on_g <- do.call("rbind.data.frame", peakachu_g_on_g)

peakachu_g_on_k[[20]] <- rbind.data.frame(peakachu_g_on_k[[20]],
                                           data.frame(specificity=1,sensitivity=0,chr="CHR21"))
peakachu_g_on_k <- do.call("rbind.data.frame", peakachu_g_on_k)

peakachu_k_on_g[[20]] <- rbind.data.frame(peakachu_k_on_g[[20]],
                                          data.frame(specificity=1,sensitivity=0,chr="CHR21"),
                                          data.frame(specificity=1,sensitivity=0,chr="CHR21"))
peakachu_k_on_g <- do.call("rbind.data.frame", peakachu_k_on_g)

peakachu_k_on_k <- do.call("rbind.data.frame", peakachu_k_on_k)

# GM12878 
##arrowhead
data <- rbind.data.frame(arrowhead_g_on_g, arrowhead_g_on_k)
data$id <- rep(factor(rep(c(1:502),21)),2)
data$CL <- c(rep("GM12878 on GM12878", nrow(arrowhead_g_on_g)),
                 rep("GM12878 on K562", nrow(arrowhead_g_on_k)))

a1 <- reshape(data[which(data$CL=="GM12878 on GM12878"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "specificity")
a1 <- rowMeans(a1[,-c(1:3)])
b1 <- reshape(data[which(data$CL=="GM12878 on GM12878"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "sensitivity")
b1 <- rowMeans(b1[,-c(1:3)])
ave_sens_spec_gm12878_on_gm12878 <- data.frame(specificity=a1, sensitivity=b1)

a2 <- reshape(data[which(data$CL=="GM12878 on K562"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "specificity")
a2 <- rowMeans(a2[,-c(1:3)])
b2 <- reshape(data[which(data$CL=="GM12878 on K562"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "sensitivity")
b2 <- rowMeans(b2[,-c(1:3)])
ave_sens_spec_gm12878_on_k562 <- data.frame(specificity=a2, sensitivity=b2)

ave_sens_spec_arrow <- rbind.data.frame(ave_sens_spec_gm12878_on_gm12878,ave_sens_spec_gm12878_on_k562)
ave_sens_spec_arrow$Training_Testing <- c(rep("GM12878 on GM12878", nrow(ave_sens_spec_gm12878_on_gm12878)),
                                       rep("GM12878 on K562", nrow(ave_sens_spec_gm12878_on_k562)))
ave_sens_spec_arrow$Tool <- "ARROWHEAD"

##peakachu
data <- rbind.data.frame(peakachu_g_on_g, peakachu_g_on_k)
data$id <- rep(factor(rep(c(1:502),21)),2)
data$CL <- c(rep("GM12878 on GM12878", nrow(peakachu_g_on_g)),
                 rep("GM12878 on K562", nrow(peakachu_g_on_k)))

a1 <- reshape(data[which(data$CL=="GM12878 on GM12878"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "specificity")
a1 <- rowMeans(a1[,-c(1:3)])
b1 <- reshape(data[which(data$CL=="GM12878 on GM12878"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "sensitivity")
b1 <- rowMeans(b1[,-c(1:3)])
ave_sens_spec_gm12878_on_gm12878 <- data.frame(specificity=a1, sensitivity=b1)

a2 <- reshape(data[which(data$CL=="GM12878 on K562"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "specificity")
a2 <- rowMeans(a2[,-c(1:3)])
b2 <- reshape(data[which(data$CL=="GM12878 on K562"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "sensitivity")
b2 <- rowMeans(b2[,-c(1:3)])
ave_sens_spec_gm12878_on_k562 <- data.frame(specificity=a2, sensitivity=b2)

ave_sens_spec_peak <- rbind.data.frame(ave_sens_spec_gm12878_on_gm12878,ave_sens_spec_gm12878_on_k562)
ave_sens_spec_peak$Training_Testing <- c(rep("GM12878 on GM12878", nrow(ave_sens_spec_gm12878_on_gm12878)),
                                       rep("GM12878 on K562", nrow(ave_sens_spec_gm12878_on_k562)))
ave_sens_spec_peak$Tool <- "PEAKACHU"

all_ave_sens_spec_df <- rbind.data.frame(ave_sens_spec_arrow,ave_sens_spec_peak)

ggplot(all_ave_sens_spec_df, aes(x=1-specificity, y=sensitivity, linetype=Training_Testing, color=Tool)) +
  facet_grid(. ~ Tool) +
  geom_line(size=1) +
  geom_abline(slope = 1, intercept = 0) +
  xlab("1-Specificity") + 
  ylab("Sensitivity") + 
  scale_linetype_discrete(name="Training/Testing") +
  scale_color_manual(values = c("blue","red")) +
  guides(color=FALSE, linetype=guide_legend(nrow=2,byrow=TRUE)) +
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x = element_text(size = 15),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=15),
        legend.title=element_text(size=20),
        plot.title = element_text(size=20),
        legend.position = "bottom")

##################################################################################################################

# K562
##arrowhead
data <- rbind.data.frame(arrowhead_k_on_k, arrowhead_k_on_g)
data$id <- rep(factor(rep(c(1:502),21)),2)
data$CL <- c(rep("K562 on K562", nrow(arrowhead_k_on_k)),
                 rep("K562 on GM12878", nrow(arrowhead_k_on_g)))

a1 <- reshape(data[which(data$CL=="K562 on K562"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "specificity")
a1 <- rowMeans(a1[,-c(1:3)])
b1 <- reshape(data[which(data$CL=="K562 on K562"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "sensitivity")
b1 <- rowMeans(b1[,-c(1:3)])
ave_sens_spec_k562_on_k562 <- data.frame(specificity=a1, sensitivity=b1)

a2 <- reshape(data[which(data$CL=="K562 on GM12878"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "specificity")
a2 <- rowMeans(a2[,-c(1:3)])
b2 <- reshape(data[which(data$CL=="K562 on GM12878"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "sensitivity")
b2 <- rowMeans(b2[,-c(1:3)])
ave_sens_spec_k562_on_gm12878 <- data.frame(specificity=a2, sensitivity=b2)

ave_sens_spec_arrow <- rbind.data.frame(ave_sens_spec_k562_on_k562,ave_sens_spec_k562_on_gm12878)
ave_sens_spec_arrow$Training_Testing <- c(rep("K562 on K562", nrow(ave_sens_spec_k562_on_k562)),
                                       rep("K562 on GM12878", nrow(ave_sens_spec_k562_on_gm12878)))
ave_sens_spec_arrow$Tool <- "ARROWHEAD"

##peakachu
data <- rbind.data.frame(peakachu_k_on_k, peakachu_k_on_g)
data$id <- rep(factor(rep(c(1:502),21)),2)
data$CL <- c(rep("K562 on K562", nrow(peakachu_k_on_k)),
                 rep("K562 on GM12878", nrow(peakachu_k_on_g)))

a1 <- reshape(data[which(data$CL=="K562 on K562"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "specificity")
a1 <- rowMeans(a1[,-c(1:3)])
b1 <- reshape(data[which(data$CL=="K562 on K562"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "sensitivity")
b1 <- rowMeans(b1[,-c(1:3)])
ave_sens_spec_k562_on_k562 <- data.frame(specificity=a1, sensitivity=b1)

a2 <- reshape(data[which(data$CL=="K562 on GM12878"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "specificity")
a2 <- rowMeans(a2[,-c(1:3)])
b2 <- reshape(data[which(data$CL=="K562 on GM12878"),], 
             idvar = "id",
             timevar="chr",
             direction="wide",
             v.names = "sensitivity")
b2 <- rowMeans(b2[,-c(1:3)])
ave_sens_spec_k562_on_gm12878 <- data.frame(specificity=a2, sensitivity=b2)

ave_sens_spec_peak <- rbind.data.frame(ave_sens_spec_k562_on_k562,ave_sens_spec_k562_on_gm12878)
ave_sens_spec_peak$Training_Testing <- c(rep("K562 on K562", nrow(ave_sens_spec_k562_on_k562)),
                                       rep("K562 on GM12878", nrow(ave_sens_spec_k562_on_gm12878)))
ave_sens_spec_peak$Tool <- "PEAKACHU"

all_ave_sens_spec_df <- rbind.data.frame(ave_sens_spec_arrow,ave_sens_spec_peak)

ggplot(all_ave_sens_spec_df, aes(x=1-specificity, y=sensitivity, linetype=Training_Testing, color=Tool)) +
  facet_grid(. ~ Tool) +
  geom_line(size=1) +
  geom_abline(slope = 1, intercept = 0) +
  xlab("1-Specificity") + 
  ylab("Sensitivity") + 
  scale_linetype_discrete(name="Training/Testing") +
  scale_color_manual(values = c("blue","red")) +
  guides(color=FALSE, linetype=guide_legend(nrow=2,byrow=TRUE)) +
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x = element_text(size = 10),
        #panel.spacing = unit(2, "lines"),
        legend.text=element_text(size=10),
        legend.title=element_text(size=15),
        plot.title = element_text(size=15),
        legend.position = "bottom")

```