### 2020-08-02 Scripts_for_results/13_saving_boundary_data_for_deeptools.Rmd
initial commit

---
title: "Saving boundary datasets for DeepTools"
author: "Spiro Stilianoudakis"
date: "7/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages and functions

```{r}
library(S4Vectors)
library(GenomicRanges)
library(ggplot2)
library(caret)
library(randomForest)
library(glmnet)
library(gbm)
library(DMwR)
library(rtracklayer)
library(Biostrings)
library(BSgenome)
library(cluster)
library(pROC)
library(SuperLearner)
library(data.table)
library(PRROC)
library(bigmemory)
library(ranger)
library(dplyr)
library(parallel)
library(doSNOW)
library(foreach)
library(pbapply)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)

source("Z:/TAD_data_analysis/functions_for_R_package/preciseTAD.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADRF.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADrfe.R")
source("Z:/TAD_data_analysis/functions_for_R_package/distance_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/percent_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/count_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/binary_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/signal_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/annots_to_granges_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/extract_boundaries_func.R")

```

# ARROWHEAD

## GM12878

```{r}
pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                                    chrs[i],
                                    "/rus/distance/preciseTAD_holdout.rds"))

  pred_bound_list[[i]] <- called_and_pred[[2]]
  true_bound_list[[i]] <- called_and_pred[[3]]

}

all_true_bounds <- do.call("c", true_bound_list)
all_pred_bounds <- do.call("c", pred_bound_list)

length(all_true_bounds) #15468
length(all_pred_bounds) #7846

all_true_bounds_df <- data.frame(chr=as.character(seqnames(all_true_bounds)),
                                 start=start(all_true_bounds),
                                 end=start(all_true_bounds)+1)

all_pred_bounds_df <- data.frame(chr=as.character(seqnames(all_pred_bounds)),
                                 start=start(all_pred_bounds),
                                 end=start(all_pred_bounds)+1)

write.table(all_true_bounds_df, "C:/Users/stili/Documents/TAD_miscellaneous/data/bedfiles_for_deeptools/arrowhead_gm12878_5kb.bed",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE,
            sep = "\t")

write.table(all_pred_bounds_df, "C:/Users/stili/Documents/TAD_miscellaneous/data/bedfiles_for_deeptools/preciseTAD_arrowhead_gm12878_5kb.bed",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE,
            sep = "\t")
```

## K562

```{r}
pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/K562/5kb/results_by_chr/",
                                    chrs[i],
                                    "/rus/distance/preciseTAD_holdout.rds"))

  pred_bound_list[[i]] <- called_and_pred[[2]]
  true_bound_list[[i]] <- called_and_pred[[3]]

}

all_true_bounds <- do.call("c", true_bound_list)
all_pred_bounds <- do.call("c", pred_bound_list)

length(all_true_bounds) #9316
length(all_pred_bounds) #8883

all_true_bounds_df <- data.frame(chr=as.character(seqnames(all_true_bounds)),
                                 start=start(all_true_bounds),
                                 end=start(all_true_bounds)+1)

all_pred_bounds_df <- data.frame(chr=as.character(seqnames(all_pred_bounds)),
                                 start=start(all_pred_bounds),
                                 end=start(all_pred_bounds)+1)

write.table(all_true_bounds_df, "C:/Users/stili/Documents/TAD_miscellaneous/data/bedfiles_for_deeptools/arrowhead_k562_5kb.bed",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE,
            sep = "\t")

write.table(all_pred_bounds_df, "C:/Users/stili/Documents/TAD_miscellaneous/data/bedfiles_for_deeptools/preciseTAD_arrowhead_k562_5kb.bed",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE,
            sep = "\t")
```

# PEAKACHU

## GM12878

```{r}
pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10:22))
for(i in 1:length(chrs)){
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                                    chrs[i],
                                    "/rus/distance/preciseTAD_holdout_peakachu.rds"))

  pred_bound_list[[i]] <- called_and_pred[[2]]
  true_bound_list[[i]] <- called_and_pred[[3]]

}

all_true_bounds <- do.call("c", true_bound_list)
all_pred_bounds <- do.call("c", pred_bound_list)

length(all_true_bounds) #21421
length(all_pred_bounds) #15838

all_true_bounds_df <- data.frame(chr=as.character(seqnames(all_true_bounds)),
                                 start=start(all_true_bounds),
                                 end=start(all_true_bounds)+1)

all_pred_bounds_df <- data.frame(chr=as.character(seqnames(all_pred_bounds)),
                                 start=start(all_pred_bounds),
                                 end=start(all_pred_bounds)+1)

write.table(all_true_bounds_df, "C:/Users/stili/Documents/TAD_miscellaneous/data/bedfiles_for_deeptools/peakachu_gm12878_10kb.bed",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE,
            sep = "\t")

write.table(all_pred_bounds_df, "C:/Users/stili/Documents/TAD_miscellaneous/data/bedfiles_for_deeptools/preciseTAD_peakachu_gm12878_10kb.bed",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE,
            sep = "\t")
```

## K562

```{r}
pred_bound_list <- GRangesList()
true_bound_list <- GRangesList()
chrs <- paste0("CHR", c(1:8,10,11,13:22))
for(i in 1:length(chrs)){
  print(i)
  called_and_pred <- readRDS(paste0("Z:/TAD_data_analysis/K562/10kb/results_by_chr/",
                                    chrs[i],
                                    "/rus/distance/preciseTAD_holdout_peakachu.rds"))

  pred_bound_list[[i]] <- called_and_pred[[2]]
  true_bound_list[[i]] <- called_and_pred[[3]]

}

all_true_bounds <- do.call("c", true_bound_list)
all_pred_bounds <- do.call("c", pred_bound_list)

length(all_true_bounds) #20773
length(all_pred_bounds) #10379

all_true_bounds_df <- data.frame(chr=as.character(seqnames(all_true_bounds)),
                                 start=start(all_true_bounds),
                                 end=start(all_true_bounds)+1)

all_pred_bounds_df <- data.frame(chr=as.character(seqnames(all_pred_bounds)),
                                 start=start(all_pred_bounds),
                                 end=start(all_pred_bounds)+1)

write.table(all_true_bounds_df, "C:/Users/stili/Documents/TAD_miscellaneous/data/bedfiles_for_deeptools/peakachu_k562_10kb.bed",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE,
            sep = "\t")

write.table(all_pred_bounds_df, "C:/Users/stili/Documents/TAD_miscellaneous/data/bedfiles_for_deeptools/preciseTAD_peakachu_k562_10kb.bed",
            col.names = FALSE,
            row.names = FALSE,
            quote = FALSE,
            sep = "\t")
```