### 2020-08-06 Scripts_for_results/14_evaluating_preciseTAD_parameters.Rmd
initial commit

---
title: "Evaluating preciseTAD parameters"
author: "Spiro Stilianoudakis"
date: "8/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(GenomicRanges)
library(ggplot2)

source("Z:/TAD_data_analysis/functions_for_R_package/preciseTAD.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADRF.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADrfe.R")
source("Z:/TAD_data_analysis/functions_for_R_package/distance_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/percent_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/count_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/binary_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/signal_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/annots_to_granges_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/extract_boundaries_func.R")

```

# Extracting boundaries for juicer

```{r}
juicer_func <- function(grdat){
  #n <- length(unique(as.character(seqnames(grdat))))

  chrs <- unique(as.character(seqnames(grdat)))

  mat_list <- list()

  for(i in 1:length(chrs)){
    grdat_chr <- grdat[which(as.character(seqnames(grdat))==chrs[i])]

    if(length(grdat_chr)>1){
      mymat <- matrix(nrow=(length(grdat_chr)-1),
                      ncol=6)
      mymat[,1] <- mymat[,4] <- chrs[i]
      for(j in 1:(length(grdat_chr)-1)){
        mymat[j,2] <- mymat[j,5] <- start(grdat_chr)[j]
        mymat[j,3] <- mymat[j,6] <- start(grdat_chr)[j+1]
      }

      mat_list[[i]] <- mymat

    }

  }

  mat_list <- do.call("rbind",mat_list)
  return(mat_list)

}

```

## Chromosome 22

### t=1

#### e=1000

```{r}
pt_1_1000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_1_1000_22.rds")

length(pt_1_1000_22[[2]]) #447
length(pt_1_1000_22[[3]]) #306

pt_1_1000_22_juicer <- juicer_func(pt_1_1000_22[[2]])
write.table(pt_1_1000_22_juicer, 
            "C:/Users/stili/Downloads/pt_1_1000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")

arrowhead_22_juicer <- juicer_func(pt_1_1000_22[[3]])
write.table(arrowhead_22_juicer, 
            "C:/Users/stili/Downloads/arrowhead_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=2500

```{r}
pt_1_2500_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_1_2500_22.rds")

length(pt_1_2500_22[[2]]) #351
length(pt_1_2500_22[[3]]) #306

pt_1_2500_22_juicer <- juicer_func(pt_1_2500_22[[2]])
write.table(pt_1_2500_22_juicer, 
            "C:/Users/stili/Downloads/pt_1_2500_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=5000

```{r}
pt_1_5000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_1_5000_22.rds")
#pt_1_5000_22_o <- readRDS("Z:/TAD_data_analysis/GM12878/5kb/results_by_chr/CHR22/rus/distance/preciseTAD_holdout.rds")

length(pt_1_5000_22[[2]]) #247
#length(pt_1_5000_22_o[[2]]) #119
length(pt_1_5000_22[[3]]) #306

pt_1_5000_22_juicer <- juicer_func(pt_1_5000_22[[2]])
write.table(pt_1_5000_22_juicer, 
            "C:/Users/stili/Downloads/pt_1_5000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")

#pt_1_5000_22_juicer_o <- juicer_func(pt_1_5000_22_o[[2]])
#write.table(pt_1_5000_22_juicer_o, 
#            "C:/Users/stili/Downloads/pt_1_5000_22_juicer_o.bed",
#            quote = FALSE,
#            col.names = FALSE,
#            row.names = FALSE,
#            sep = "\t")
```

#### e=10000

```{r}
pt_1_10000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_1_10000_22.rds")

length(pt_1_10000_22[[2]]) #226
length(pt_1_10000_22[[3]]) #306

pt_1_10000_22_juicer <- juicer_func(pt_1_10000_22[[2]])
write.table(pt_1_10000_22_juicer, 
            "C:/Users/stili/Downloads/pt_1_10000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

### t=.99

#### e=1000

```{r}
pt_99_1000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_99_1000_22.rds")

length(pt_99_1000_22[[2]]) #
length(pt_99_1000_22[[3]]) #306

pt_99_1000_22_juicer <- juicer_func(pt_99_1000_22[[2]])
write.table(pt_99_1000_22_juicer, 
            "C:/Users/stili/Downloads/pt_99_1000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=2500

```{r}
pt_99_2500_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_99_2500_22.rds")

length(pt_99_2500_22[[2]]) #486
length(pt_99_2500_22[[3]]) #306

pt_99_2500_22_juicer <- juicer_func(pt_99_2500_22[[2]])
write.table(pt_99_2500_22_juicer, 
            "C:/Users/stili/Downloads/pt_99_2500_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=5000

```{r}
pt_99_5000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_99_5000_22.rds")

length(pt_99_5000_22[[2]]) #368
length(pt_99_5000_22[[3]]) #306

pt_99_5000_22_juicer <- juicer_func(pt_99_5000_22[[2]])
write.table(pt_99_5000_22_juicer, 
            "C:/Users/stili/Downloads/pt_99_5000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=10000

```{r}
pt_99_10000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_99_10000_22.rds")

length(pt_99_10000_22[[2]]) #299
length(pt_99_10000_22[[3]]) #306

pt_99_10000_22_juicer <- juicer_func(pt_99_10000_22[[2]])
write.table(pt_99_10000_22_juicer, 
            "C:/Users/stili/Downloads/pt_99_10000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

### t=.975

#### e=1000

```{r}
pt_975_1000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_975_1000_22.rds")

length(pt_975_1000_22[[2]]) #
length(pt_975_1000_22[[3]]) #306

pt_975_1000_22_juicer <- juicer_func(pt_975_1000_22[[2]])
write.table(pt_975_1000_22_juicer, 
            "C:/Users/stili/Downloads/pt_975_1000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=2500

```{r}
pt_975_2500_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_975_2500_22.rds")

length(pt_975_2500_22[[2]]) #599
length(pt_975_2500_22[[3]]) #306

pt_975_2500_22_juicer <- juicer_func(pt_975_2500_22[[2]])
write.table(pt_975_2500_22_juicer, 
            "C:/Users/stili/Downloads/pt_975_2500_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=5000

```{r}
pt_975_5000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_975_5000_22.rds")

length(pt_975_5000_22[[2]]) #418
length(pt_975_5000_22[[3]]) #306

pt_975_5000_22_juicer <- juicer_func(pt_975_5000_22[[2]])
write.table(pt_975_5000_22_juicer, 
            "C:/Users/stili/Downloads/pt_975_5000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=10000

```{r}
pt_975_10000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_975_10000_22.rds")

length(pt_975_10000_22[[2]]) #312
length(pt_975_10000_22[[3]]) #306

pt_975_10000_22_juicer <- juicer_func(pt_975_10000_22[[2]])
write.table(pt_975_10000_22_juicer, 
            "C:/Users/stili/Downloads/pt_975_10000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

### t=.95

#### e=1000

```{r}
#pt_95_1000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/t_95_1000_22.rds")

#length(pt_95_1000_22[[2]]) #
#length(pt_95_1000_22[[3]]) #306

#pt_95_1000_22_juicer <- juicer_func(pt_95_1000_22[[2]])
#write.table(pt_95_1000_22_juicer, 
#            "C:/Users/stili/Downloads/pt_95_1000_22_juicer.bed",
#            quote = FALSE,
#            col.names = FALSE,
#            row.names = FALSE,
#            sep = "\t")
```

#### e=2500

```{r}
pt_95_2500_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_95_2500_22.rds")

length(pt_95_2500_22[[2]]) #636
length(pt_95_2500_22[[3]]) #306

pt_95_2500_22_juicer <- juicer_func(pt_95_2500_22[[2]])
write.table(pt_95_2500_22_juicer, 
            "C:/Users/stili/Downloads/pt_95_2500_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=5000

```{r}
pt_95_5000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_95_5000_22.rds")

length(pt_95_5000_22[[2]]) #396
length(pt_95_5000_22[[3]]) #306

pt_95_5000_22_juicer <- juicer_func(pt_95_5000_22[[2]])
write.table(pt_95_5000_22_juicer, 
            "C:/Users/stili/Downloads/pt_95_5000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

#### e=10000

```{r}
pt_95_10000_22 <- readRDS("Z:/TAD_data_analysis/miscellaneous/evaluating_preciseTAD_parameters/pt_95_10000_22.rds")

length(pt_95_10000_22[[2]]) #298
length(pt_95_10000_22[[3]]) #306

pt_95_10000_22_juicer <- juicer_func(pt_95_10000_22[[2]])
write.table(pt_95_10000_22_juicer, 
            "C:/Users/stili/Downloads/pt_95_10000_22_juicer.bed",
            quote = FALSE,
            col.names = FALSE,
            row.names = FALSE,
            sep = "\t")
```

# Plotting density of distances between boundaries

```{r}
dist_dens <- data.frame(Distance = c(diff(start(pt_1_1000_22[[2]]))[-which(diff(start(pt_1_1000_22[[2]]))>2000000)],
                                     diff(start(pt_1_2500_22[[2]]))[-which(diff(start(pt_1_2500_22[[2]]))>2000000)],
                                     diff(start(pt_1_5000_22[[2]]))[-which(diff(start(pt_1_5000_22[[2]]))>2000000)],
                                     diff(start(pt_1_10000_22[[2]]))[-which(diff(start(pt_1_10000_22[[2]]))>2000000)],
                                     diff(start(pt_99_1000_22[[2]])),
                                     diff(start(pt_99_2500_22[[2]])),
                                     diff(start(pt_99_5000_22[[2]])),
                                     diff(start(pt_99_10000_22[[2]])),
                                     diff(start(pt_975_1000_22[[2]])),
                                     diff(start(pt_975_2500_22[[2]])),
                                     diff(start(pt_975_5000_22[[2]]))[-which(diff(start(pt_975_5000_22[[2]]))>2000000)],
                                     diff(start(pt_975_10000_22[[2]]))[-which(diff(start(pt_975_10000_22[[2]]))>2000000)],
                                     #diff(start(pt_95_1000_22[[2]]))[-which(diff(start(pt_95_1000_22[[2]]))>2000000)],
                                     diff(start(pt_95_2500_22[[2]])),
                                     diff(start(pt_95_5000_22[[2]])),
                                     diff(start(pt_95_10000_22[[2]]))),
                        Threshold = c(rep("t=1.0", length(diff(start(pt_1_1000_22[[2]]))[-which(diff(start(pt_1_1000_22[[2]]))>2000000)])+
                                        length(diff(start(pt_1_2500_22[[2]]))[-which(diff(start(pt_1_2500_22[[2]]))>2000000)])+
                                        length(diff(start(pt_1_5000_22[[2]]))[-which(diff(start(pt_1_5000_22[[2]]))>2000000)])+
                                        length(diff(start(pt_1_10000_22[[2]]))[-which(diff(start(pt_1_10000_22[[2]]))>2000000)])),
                                      rep("t=0.99", length(diff(start(pt_99_1000_22[[2]])))+
                                        length(diff(start(pt_99_2500_22[[2]])))+
                                        length(diff(start(pt_99_5000_22[[2]])))+
                                        length(diff(start(pt_99_10000_22[[2]])))),
                                      rep("t=0.975", length(diff(start(pt_975_1000_22[[2]])))+
                                        length(diff(start(pt_975_2500_22[[2]])))+
                                        length(diff(start(pt_975_5000_22[[2]]))[-which(diff(start(pt_975_5000_22[[2]]))>2000000)])+
                                        length(diff(start(pt_975_10000_22[[2]]))[-which(diff(start(pt_975_10000_22[[2]]))>2000000)])),
                                      rep("t=0.95", #length(diff(start(pt_95_1000_22[[2]]))[-which(diff(start(pt_95_1000_22[[2]]))>2000000)])+
                                        length(diff(start(pt_95_2500_22[[2]])))+
                                        length(diff(start(pt_95_5000_22[[2]])))+
                                        length(diff(start(pt_95_10000_22[[2]]))))),
                        Epsilon = c(rep("eps=0.2r", length(diff(start(pt_1_1000_22[[2]]))[-which(diff(start(pt_1_1000_22[[2]]))>2000000)])),
                                    rep("eps=0.5r", length(diff(start(pt_1_2500_22[[2]]))[-which(diff(start(pt_1_2500_22[[2]]))>2000000)])),
                                    rep("eps=r", length(diff(start(pt_1_5000_22[[2]]))[-which(diff(start(pt_1_5000_22[[2]]))>2000000)])),
                                    rep("eps=2r", length(diff(start(pt_1_10000_22[[2]]))[-which(diff(start(pt_1_10000_22[[2]]))>2000000)])),
                                    rep("eps=0.2r", length(diff(start(pt_99_1000_22[[2]])))),
                                    rep("eps=0.5r", length(diff(start(pt_99_2500_22[[2]])))),
                                    rep("eps=r", length(diff(start(pt_99_5000_22[[2]])))),
                                    rep("eps=2r", length(diff(start(pt_99_10000_22[[2]])))),
                                    rep("eps=0.2r", length(diff(start(pt_975_1000_22[[2]])))),
                                    rep("eps=0.5r", length(diff(start(pt_975_2500_22[[2]])))),
                                    rep("eps=r", length(diff(start(pt_975_5000_22[[2]]))[-which(diff(start(pt_975_5000_22[[2]]))>2000000)])),
                                    rep("eps=2r", length(diff(start(pt_975_10000_22[[2]]))[-which(diff(start(pt_975_10000_22[[2]]))>2000000)])),
                                    #rep("eps=0.2r", length(diff(start(pt_95_1000_22[[2]]))[-which(diff(start(pt_95_1000_22[[2]]))>2000000)])),
                                    rep("eps=0.5r", length(diff(start(pt_95_2500_22[[2]])))),
                                    rep("eps=r", length(diff(start(pt_95_5000_22[[2]])))),
                                    rep("eps=2r", length(diff(start(pt_95_10000_22[[2]]))))))

dist_dens$Epsilon <- factor(dist_dens$Epsilon, levels=c("eps=0.2r",
                                                        "eps=0.5r",
                                                        "eps=r",
                                                        "eps=2r"))

ggplot(data=dist_dens, aes(x=Distance)) +
  facet_grid(Epsilon ~ Threshold) +
  geom_density(fill="gray", color="black", alpha=0.8) +
  ylab("Density") +
  xlab("Distance between predicted TAD boundaries") +
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size=10,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.background.x = element_blank(),
        #strip.background.y = element_blank(),
        legend.text=element_text(size=10),
        legend.title=element_text(size=15),
        plot.title = element_text(size=15))
```

# Overlaps per boundaries

```{r}
genomicElements.GR <- annots_to_granges_func(filepath = "Z:/TAD_data_analysis/annotations/all_common_annotations/gm12878/topTFBS",
                                             pattern="*.bed",
											 signal=4)

df <- data.frame(Overlap = c(length(unique(subjectHits(findOverlaps(flank(pt_1_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_1_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_1_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_1_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_1_10000_22[[2]])-.1,

length(unique(subjectHits(findOverlaps(flank(pt_99_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_99_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_99_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_99_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_99_10000_22[[2]]),

length(unique(subjectHits(findOverlaps(flank(pt_975_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_975_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_975_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_975_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_975_10000_22[[2]]),

NA,
length(unique(subjectHits(findOverlaps(flank(pt_95_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_95_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_95_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_95_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_95_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[1]]))))/length(pt_95_10000_22[[2]]),

length(unique(subjectHits(findOverlaps(flank(pt_1_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_1_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_1_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_1_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_1_10000_22[[2]])-.1,

length(unique(subjectHits(findOverlaps(flank(pt_99_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_99_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_99_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_99_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_99_10000_22[[2]]),

length(unique(subjectHits(findOverlaps(flank(pt_975_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_975_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_975_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_975_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_975_10000_22[[2]]),

NA,
length(unique(subjectHits(findOverlaps(flank(pt_95_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_95_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_95_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_95_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_95_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[2]]))))/length(pt_95_10000_22[[2]]),

length(unique(subjectHits(findOverlaps(flank(pt_1_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_1_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_1_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_1_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_1_10000_22[[2]])-.1,

length(unique(subjectHits(findOverlaps(flank(pt_99_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_99_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_99_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_99_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_99_10000_22[[2]]),

length(unique(subjectHits(findOverlaps(flank(pt_975_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_975_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_975_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_975_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_975_10000_22[[2]]),

NA,
length(unique(subjectHits(findOverlaps(flank(pt_95_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_95_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_95_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_95_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_95_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[3]]))))/length(pt_95_10000_22[[2]]),

length(unique(subjectHits(findOverlaps(flank(pt_1_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_1_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_1_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_1_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_1_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_1_10000_22[[2]])-.1,

length(unique(subjectHits(findOverlaps(flank(pt_99_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_99_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_99_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_99_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_99_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_99_10000_22[[2]]),

length(unique(subjectHits(findOverlaps(flank(pt_975_1000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_975_1000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_975_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_975_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_975_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_975_10000_22[[2]]),

NA,
length(unique(subjectHits(findOverlaps(flank(pt_95_2500_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_95_2500_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_95_5000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_95_5000_22[[2]]),
length(unique(subjectHits(findOverlaps(flank(pt_95_10000_22[[2]], width = 5000, both = TRUE),genomicElements.GR[[4]]))))/length(pt_95_10000_22[[2]])),

Threshold = c(rep(c(rep("t=1.0",4),
              rep("t=0.99",4),
              rep("t=0.975",4),
              rep("t=0.95",4)),4)),

Epsilon = rep(c("eps=1000",
              "eps=2500",
			  "eps=5000",
			  "eps=10000"),16),

Annotation = c(rep("CTCF", 16),
               rep("RAD21", 16),
			   rep("SMC3", 16),
			   rep("ZNF143", 16))
)

df$Epsilon <- factor(df$Epsilon, levels=c("eps=1000", "eps=2500", "eps=5000", "eps=10000"))

ggplot(data=df, aes(x=Annotation, y=Overlap, fill=Annotation)) +
  facet_grid(Epsilon ~ Threshold) +
  geom_bar(stat='identity') +
  ylim(0,1.5)+
  ylab("Normalized Overlap") +
  xlab("Annotation") +
  guides(fill=FALSE) +
  theme_minimal() +
  theme_bw()+
  theme(axis.text.x = element_text(size=10,
                                   angle = 45, 
                                   #margin = ggplot2::margin(t = 35),
                                   hjust = 1),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #strip.background.x = element_blank(),
        #strip.background.y = element_blank(),
        legend.text=element_text(size=10),
        legend.title=element_text(size=15),
        plot.title = element_text(size=15))
```