### 2020-08-14 Scripts_for_results/03_creating_annotation_specifc_barplots.Rmd
updated figures/tables

---
title: "Creating annotation specific barplots"
author: "Spiro Stilianoudakis"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages and functions

```{r}
library(S4Vectors)
library(GenomicRanges)
library(ggplot2)
library(caret)
library(randomForest)
library(glmnet)
library(gbm)
library(DMwR)
library(rtracklayer)
library(Biostrings)
library(BSgenome)
library(cluster)
library(pROC)
library(SuperLearner)
library(data.table)
library(PRROC)
library(bigmemory)
library(ranger)
library(dplyr)
library(parallel)
library(doSNOW)
library(foreach)
library(pbapply)
library(ggpubr)
library(scales)
library(GGally)
library(VennDiagram)
library(network)
library(sna)
library(gtable)
library(grid)
library(magrittr) 
library(DT)
library(tidyverse)
library(reshape)
library(ROSE)
library(lattice)
library(corrplot)
library(cluster)
library(RColorBrewer)
library(GenometriCorr)
library(ggsignif)
library(Vennerable)
library(VennDiagram)
library(ChIPpeakAnno)
library(EnrichedHeatmap)

source("Z:/TAD_data_analysis/functions_for_R_package/preciseTAD.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADRF.R")
source("Z:/TAD_data_analysis/functions_for_R_package/TADrfe.R")
source("Z:/TAD_data_analysis/functions_for_R_package/distance_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/percent_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/count_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/binary_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/signal_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/annots_to_granges_func.R")
source("Z:/TAD_data_analysis/functions_for_R_package/extract_boundaries_func.R")

```

## ARROWHEAD

### GM12878 (rus; distance types)

#### 5kb

```{r}
badata_gm12878_5kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
 for(j in c("BroadHMM", "HistoneModifications", "TFBS", "all")){
   print(c(i,j))
  if(j %in% c("BroadHMM", "HistoneModifications", "TFBS")){
    perfdata <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_",
                               j, 
                               ".rds"))
    perfdata <- perfdata[[3]]
  }else{
    perfdata <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/5kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF.rds"))
    perfdata <- perfdata[[3]]
  }

  badata_gm12878_5kb <- rbind(badata_gm12878_5kb,
                               perfdata[10,2])

 }
}

```

#### Plotting

```{r}
all_rfperfs <- data.frame(Perf=badata_gm12878_5kb)
all_rfperfs$Resolution <- rep("5 kb", length(badata_gm12878_5kb))
all_rfperfs$Annotation <- c("BroadHMM","HM","TFBS","ALL")
all_rfperfs$Annotation <- factor(all_rfperfs$Annotation, levels = c("HM", "BroadHMM", "TFBS", "ALL"))
all_rfperfs$Chromosome <- c(rep("CHR1", 4),
                            rep("CHR2", 4),
                            rep("CHR3", 4),
                            rep("CHR4", 4),
                            rep("CHR5", 4),
                            rep("CHR6", 4),
                            rep("CHR7", 4),
                            rep("CHR8", 4),
                            rep("CHR10", 4),
                            rep("CHR11", 4),
                            rep("CHR12", 4),
                            rep("CHR13", 4),
                            rep("CHR14", 4),
                            rep("CHR15", 4),
                            rep("CHR16", 4),
                            rep("CHR17", 4),
                            rep("CHR18", 4),
                            rep("CHR19", 4),
                            rep("CHR20", 4),
                            rep("CHR21", 4),
                            rep("CHR22", 4))

all_rfperfs <- all_rfperfs %>% 
  group_by(Resolution, Annotation) %>% 
  dplyr::summarise(Performance = mean(Perf, na.rm = TRUE),
                   PerformanceSD = sd(Perf, na.rm = TRUE))

ggplot(all_rfperfs, aes(x=Annotation, y=Performance, fill=Annotation)) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=.75) +
    theme_minimal() +
    theme_bw()+
    xlab("Resolution") + 
    ylab("Balanced Accuracy") + 
    scale_fill_discrete(name="Genomic Class")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    legend.position = "bottom",
    plot.title = element_text(size=20))
```

### K562 (rus; distance types)

#### 5kb

```{r}
badata_k562_5kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
 for(j in c("BroadHMM", "HistoneModifications", "TFBS", "all")){
   print(c(i,j))
  if(j %in% c("BroadHMM", "HistoneModifications", "TFBS")){
    perfdata <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/K562/5kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_",
                               j, 
                               ".rds"))
    perfdata <- perfdata[[3]]
  }else{
    perfdata <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/K562/5kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF.rds"))
    perfdata <- perfdata[[3]]
  }

  badata_k562_5kb <- rbind(badata_k562_5kb,
                               perfdata[10,2])

 }
}

```

#### Plotting

```{r}
all_rfperfs <- data.frame(Perf=badata_k562_5kb)
all_rfperfs$Resolution <- rep("5 kb", length(badata_k562_5kb))
all_rfperfs$Annotation <- c("BroadHMM","HM","TFBS","ALL")
all_rfperfs$Annotation <- factor(all_rfperfs$Annotation, levels = c("HM", "BroadHMM", "TFBS", "ALL"))
all_rfperfs$Chromosome <- c(rep("CHR1", 4),
                            rep("CHR2", 4),
                            rep("CHR3", 4),
                            rep("CHR4", 4),
                            rep("CHR5", 4),
                            rep("CHR6", 4),
                            rep("CHR7", 4),
                            rep("CHR8", 4),
                            rep("CHR10", 4),
                            rep("CHR11", 4),
                            rep("CHR12", 4),
                            rep("CHR13", 4),
                            rep("CHR14", 4),
                            rep("CHR15", 4),
                            rep("CHR16", 4),
                            rep("CHR17", 4),
                            rep("CHR18", 4),
                            rep("CHR19", 4),
                            rep("CHR20", 4),
                            rep("CHR21", 4),
                            rep("CHR22", 4))

all_rfperfs <- all_rfperfs %>% 
  group_by(Resolution, Annotation) %>% 
  dplyr::summarise(Performance = mean(Perf, na.rm = TRUE),
                   PerformanceSD = sd(Perf, na.rm = TRUE))

ggplot(all_rfperfs, aes(x=Annotation, y=Performance, fill=Annotation)) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=.75) +
    theme_minimal() +
    theme_bw()+
    xlab("Resolution") + 
    ylab("Balanced Accuracy") + 
    scale_fill_discrete(name="Genomic Class")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    legend.position = "bottom",
    plot.title = element_text(size=20))
```

## PEAKACHU

### GM12878 (rus; distance types)

#### 10kb

```{r}
badata_gm12878_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
 for(j in c("BroadHMM", "HistoneModifications", "TFBS", "all")){
   print(c(i,j))
  if(j %in% c("BroadHMM", "HistoneModifications", "TFBS")){
    perfdata <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_peakachu_",
                               j, 
                               ".rds"))
    perfdata <- perfdata[[3]]
  }else{
    perfdata <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/GM12878/10kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_peakachu.rds"))
    perfdata <- perfdata[[3]]
  }

  badata_gm12878_10kb <- rbind(badata_gm12878_10kb,
                               perfdata[10,2])

 }
}

```

#### Plotting

```{r}
all_rfperfs <- data.frame(Perf=badata_gm12878_10kb)
all_rfperfs$Resolution <- rep("10 kb", length(badata_gm12878_10kb))
all_rfperfs$Annotation <- c("BroadHMM","HM","TFBS","ALL")
all_rfperfs$Annotation <- factor(all_rfperfs$Annotation, levels = c("HM", "BroadHMM", "TFBS", "ALL"))
all_rfperfs$Chromosome <- c(rep("CHR1", 4),
                            rep("CHR2", 4),
                            rep("CHR3", 4),
                            rep("CHR4", 4),
                            rep("CHR5", 4),
                            rep("CHR6", 4),
                            rep("CHR7", 4),
                            rep("CHR8", 4),
                            rep("CHR10", 4),
                            rep("CHR11", 4),
                            rep("CHR12", 4),
                            rep("CHR13", 4),
                            rep("CHR14", 4),
                            rep("CHR15", 4),
                            rep("CHR16", 4),
                            rep("CHR17", 4),
                            rep("CHR18", 4),
                            rep("CHR19", 4),
                            rep("CHR20", 4),
                            rep("CHR21", 4),
                            rep("CHR22", 4))

all_rfperfs <- all_rfperfs %>% 
  group_by(Resolution, Annotation) %>% 
  dplyr::summarise(Performance = mean(Perf, na.rm = TRUE),
                   PerformanceSD = sd(Perf, na.rm = TRUE))

ggplot(all_rfperfs, aes(x=Annotation, y=Performance, fill=Annotation)) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=.75) +
    theme_minimal() +
    theme_bw()+
    xlab("Resolution") + 
    ylab("Balanced Accuracy") + 
    scale_fill_discrete(name="Genomic Class")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    legend.position = "bottom",
    plot.title = element_text(size=20))
```

### K562 (rus; distance types)

#### 10kb

```{r}
badata_k562_10kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
 for(j in c("BroadHMM", "HistoneModifications", "TFBS", "all")){
   print(c(i,j))
  if(j %in% c("BroadHMM", "HistoneModifications", "TFBS")){
    perfdata <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/K562/10kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_peakachu_",
                               j, 
                               ".rds"))
    perfdata <- perfdata[[3]]
  }else{
    perfdata <- readRDS(paste0("/home/stilianoudakisc/TAD_data_analysis/K562/10kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_peakachu.rds"))
    perfdata <- perfdata[[3]]
  }

  badata_k562_10kb <- rbind(badata_k562_10kb,
                               perfdata[10,2])

 }
}

```

#### Plotting

```{r}
all_rfperfs <- data.frame(Perf=badata_k562_10kb)
all_rfperfs$Resolution <- rep("10 kb", length(badata_k562_10kb))
all_rfperfs$Annotation <- c("BroadHMM","HM","TFBS","ALL")
all_rfperfs$Annotation <- factor(all_rfperfs$Annotation, levels = c("HM", "BroadHMM", "TFBS", "ALL"))
all_rfperfs$Chromosome <- c(rep("CHR1", 4),
                            rep("CHR2", 4),
                            rep("CHR3", 4),
                            rep("CHR4", 4),
                            rep("CHR5", 4),
                            rep("CHR6", 4),
                            rep("CHR7", 4),
                            rep("CHR8", 4),
                            rep("CHR10", 4),
                            rep("CHR11", 4),
                            rep("CHR12", 4),
                            rep("CHR13", 4),
                            rep("CHR14", 4),
                            rep("CHR15", 4),
                            rep("CHR16", 4),
                            rep("CHR17", 4),
                            rep("CHR18", 4),
                            rep("CHR19", 4),
                            rep("CHR20", 4),
                            rep("CHR21", 4),
                            rep("CHR22", 4))

all_rfperfs <- all_rfperfs %>% 
  group_by(Resolution, Annotation) %>% 
  dplyr::summarise(Performance = mean(Perf, na.rm = TRUE),
                   PerformanceSD = sd(Perf, na.rm = TRUE))

ggplot(all_rfperfs, aes(x=Annotation, y=Performance, fill=Annotation)) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=.75) +
    theme_minimal() +
    theme_bw()+
    xlab("Resolution") + 
    ylab("Balanced Accuracy") + 
    scale_fill_discrete(name="Genomic Class")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    legend.position = "bottom",
    plot.title = element_text(size=20))
```

## SpectralTAD

### GM12878 (rus; distance types)

#### 25kb

```{r}
badata_gm12878_25kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
 for(j in c("BroadHMM", "HistoneModifications", "TFBS", "all")){
   print(c(i,j))
  if(j %in% c("BroadHMM", "HistoneModifications", "TFBS")){
    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_spectralTAD_",
                               j, 
                               ".rds"))
    perfdata <- perfdata[[3]]
  }else{
    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/GM12878/25kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_spectralTAD.rds"))
    perfdata <- perfdata[[3]]
  }

  badata_gm12878_25kb <- rbind(badata_gm12878_25kb,
                               perfdata[10,2])

 }
}

```

#### Plotting

```{r}
all_rfperfs <- data.frame(Perf=badata_gm12878_25kb)
all_rfperfs$Resolution <- rep("25 kb", length(badata_gm12878_25kb))
all_rfperfs$Annotation <- c("BroadHMM","HM","TFBS","ALL")
all_rfperfs$Annotation <- factor(all_rfperfs$Annotation, levels = c("HM", "BroadHMM", "TFBS", "ALL"))
all_rfperfs$Chromosome <- c(rep("CHR1", 4),
                            rep("CHR2", 4),
                            rep("CHR3", 4),
                            rep("CHR4", 4),
                            rep("CHR5", 4),
                            rep("CHR6", 4),
                            rep("CHR7", 4),
                            rep("CHR8", 4),
                            rep("CHR10", 4),
                            rep("CHR11", 4),
                            rep("CHR12", 4),
                            rep("CHR13", 4),
                            rep("CHR14", 4),
                            rep("CHR15", 4),
                            rep("CHR16", 4),
                            rep("CHR17", 4),
                            rep("CHR18", 4),
                            rep("CHR19", 4),
                            rep("CHR20", 4),
                            rep("CHR21", 4),
                            rep("CHR22", 4))

all_rfperfs <- all_rfperfs %>% 
  group_by(Resolution, Annotation) %>% 
  dplyr::summarise(Performance = mean(Perf, na.rm = TRUE),
                   PerformanceSD = sd(Perf, na.rm = TRUE))

ggplot(all_rfperfs, aes(x=Annotation, y=Performance, fill=Annotation)) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=.75) +
    theme_minimal() +
    theme_bw()+
    xlab("Resolution") + 
    ylab("Balanced Accuracy") + 
    scale_fill_discrete(name="Genomic Class")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    legend.position = "bottom",
    plot.title = element_text(size=20))
```

### K562 (rus; distance types)

#### 25kb

```{r}
badata_k562_25kb <- numeric()
for(i in paste0("CHR", c(1:8,10:22))){
 for(j in c("BroadHMM", "HistoneModifications", "TFBS", "all")){
   print(c(i,j))
  if(j %in% c("BroadHMM", "HistoneModifications", "TFBS")){
    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_spectralTAD_",
                               j, 
                               ".rds"))
    perfdata <- perfdata[[3]]
  }else{
    perfdata <- readRDS(paste0("Z:/TAD_data_analysis/K562/25kb/results_by_chr/",
                               i,
                               "/rus/distance/TADRF_spectralTAD.rds"))
    perfdata <- perfdata[[3]]
  }

  badata_k562_25kb <- rbind(badata_k562_25kb,
                               perfdata[10,2])

 }
}

```

#### Plotting

```{r}
all_rfperfs <- data.frame(Perf=badata_k562_25kb)
all_rfperfs$Resolution <- rep("10 kb", length(badata_k562_25kb))
all_rfperfs$Annotation <- c("BroadHMM","HM","TFBS","ALL")
all_rfperfs$Annotation <- factor(all_rfperfs$Annotation, levels = c("HM", "BroadHMM", "TFBS", "ALL"))
all_rfperfs$Chromosome <- c(rep("CHR1", 4),
                            rep("CHR2", 4),
                            rep("CHR3", 4),
                            rep("CHR4", 4),
                            rep("CHR5", 4),
                            rep("CHR6", 4),
                            rep("CHR7", 4),
                            rep("CHR8", 4),
                            rep("CHR10", 4),
                            rep("CHR11", 4),
                            rep("CHR12", 4),
                            rep("CHR13", 4),
                            rep("CHR14", 4),
                            rep("CHR15", 4),
                            rep("CHR16", 4),
                            rep("CHR17", 4),
                            rep("CHR18", 4),
                            rep("CHR19", 4),
                            rep("CHR20", 4),
                            rep("CHR21", 4),
                            rep("CHR22", 4))

all_rfperfs <- all_rfperfs %>% 
  group_by(Resolution, Annotation) %>% 
  dplyr::summarise(Performance = mean(Perf, na.rm = TRUE),
                   PerformanceSD = sd(Perf, na.rm = TRUE))

ggplot(all_rfperfs, aes(x=Annotation, y=Performance, fill=Annotation)) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=.75) +
    theme_minimal() +
    theme_bw()+
    xlab("Resolution") + 
    ylab("Balanced Accuracy") + 
    scale_fill_discrete(name="Genomic Class")+
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    legend.position = "bottom",
    plot.title = element_text(size=20))
```

#############################################

#Plotting

```{r}
ggplot(all_rfperfs_a_p, aes(x=Annotation, y=Performance, fill=Tool)) +
    facet_grid(. ~ Tool) +
    geom_bar(stat="identity") +
    geom_errorbar(aes(ymin=Performance-PerformanceSD, ymax=Performance+PerformanceSD), width=.75, size=1) +
    guides(fill=FALSE) +
    scale_fill_manual(values = c("blue", "red"))+
    theme_minimal() +
    theme_bw()+
    xlab("Genomic Class") + 
    ylab("Balanced Accuracy") + 
    theme(axis.text.x = element_text(size=15,
                                     angle = 45, 
                                     #margin = ggplot2::margin(t = 35),
                                     hjust = 1
    ),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    strip.text.x = element_text(size = 20),
    legend.text=element_text(size=15),
    legend.title=element_text(size=20),
    legend.position = "bottom",
    plot.title = element_text(size=20))
```