### 2020-09-11 Scripts_for_results/16_psuedocode.Rmd
initial commit

---
title: "psuedocode"
author: "Spiro Stilianoudakis"
date: "9/6/2020"
output: pdf_document
header-includes:
- \usepackage{stackengine}
- \usepackage{amsmath}
- \usepackage{algorithm2e}
- \usepackage[noend]{algpseudocode}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# preciseTAD (CV)

\begin{algorithm}[H]
\DontPrintSemicolon
\SetAlgoLined
1. Consider an optimized RF model (M) built on the set of autosomal chromosomes $\{ k |i \notin k \}$ binned at some resolution $r$
\BlankLine
2. \For{each chr $i$ do}{
\BlankLine
3.    Construct the base-level resolution predictor space $A_{n \times p}$ where $n$ is the length of chr $i$ and $p$ is the number of predictors\;
\BlankLine
4.    Assign threshold $\{t | 0 \le t \le 1 \}$ and $\{\epsilon | \epsilon > 0 \}$\;
\BlankLine
5.    \eIf{$|t| > 1$ or $|\epsilon| > 1$}{
\BlankLine
6.          Assign grid $(g)$ of combinations of $t$ and $\epsilon$ \;
\BlankLine
7.               \For{each combination $l$ in g}{
\BlankLine
8.                  Evaluate M on $A_{n \times p}$ to get the probability of each genomic coordinate as being a domain boundary $\pi_{n}$\;
\BlankLine
9.                  Subset $\{\pi_{n} | \pi_{n} \ge t_{l} \}$\;
\BlankLine
10.                 Construct the pairwise distance matrix $D$ between genomic coordinates where $\pi_{n} \ge t_{l}$\;
\BlankLine
11.                 Apply DBSCAN on $D$ with $MinPts=3$ and $\epsilon=\epsilon_{l}$\;
\BlankLine
12.                 \For{each cluster $k$ identified by DBSCAN}{
\BlankLine
13.                     Assign $w_{k}$ as the number of coordinates that span each cluster of bases in $k$ (PTBR)\;
\BlankLine
14.                     Perform PAM on the sub-distance matrix $D_{k}$ to extract the cluster medoid $b_{k}$ (PTBP)\;
\BlankLine
15.                     \For{each predictor p}{
\BlankLine
16.                         Calculate the normalized enrichment (NE) over all predictors
                            $$NE = \dfrac{1}{p} \left[ \Sigma_{s=1}^{p} \left[ \dfrac{1}{b} \Sigma_{k=1}^{b} e_{ks} \times c_{k} \right] \right]$$
                            where $e_{ks} = \textbf{I}\{ r_{s} \in (b_{k}-f, b_{k}+f)\}$ is the number of elemental regions $r$ of predictor $p$ that overlap with each flanked boundary and $c_{k} = \dfrac{ \textbf{I}\{ \pi_{n} \ge t_{l} \in w_{k} \} }{w_{k}}$ is the coverage probability of each PTBR\;
\BlankLine
17.                         Determine where $NE$ converges as optimal $\{t,\epsilon\}$ combination\;
\BlankLine
18.                         Repeat steps 12-18 on $A_{nxp}$ with optimal $\{t,\epsilon\}$\;
\BlankLine
                     }
\BlankLine
                 }
\BlankLine
              }
\BlankLine
   }{
\BlankLine
19.  Perform steps 12-18 on $A_{nxp}$ such that $t=t_{0}$ and $\epsilon=\epsilon_{0}$\;
\BlankLine
   }
\BlankLine
}
\caption{Psuedocode for $preciseTAD$ implementation.}
\end{algorithm}