### 2020-09-26 BMC_Genome_Biology_submission/additional_files/additional_file_legends.Rmd
submission folder initial commit

---
title: "preciseTAD: A machine learning framework for precise 3D domain boundary prediction at base-level resolution"
subtitle: "Additional File Legends"
output:
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Additional File 1

An example script for applying Arrowhead to in situ Hi-C data (HIC001-HIC018) to obtain chromosome-specific TAD boundaries on the GM12878 cell line at 5kb, 10kb, 25kb, 50kb, 100kb resolutions. An example of deepTools commands to generate plotProfiles and plotHeatmaps is also provided.